```{r setup_r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls())
library(tidyverse)
library(patchwork)

cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
```

## Visualise pSer2 vs total from western blot

mAC-POLR2A cells treated for 4 h with 5-Ph-IAA at different doses (or vehicle). Visualise changes over dose and correspondence between total and pSer2 from western blot data.

```{r}
wb_csv <- "data/20260610_degron_titration.csv"

wb_data <- read_csv(wb_csv) %>%
  select(Replicate,concentration=`5-Ph-IAA concentration`,D8L4Y, S2, vinculin) %>%
  filter(!is.na(Replicate)) %>%
  mutate(total_D8L4Y = D8L4Y / vinculin,
         pSer2_3E10 = S2 / vinculin,
         concentration_numeric = case_when(
           concentration=="DMSO" ~ 1e-9,
           concentration=="10 nM" ~ 10e-9,
           concentration=="100 nM" ~ 100e-9,
           concentration=="1 µM" ~ 1e-6,
           concentration=="10 µM" ~ 10e-6,
           TRUE ~ NA_real_)) 

wb_data <- wb_data %>%
  left_join(filter(wb_data,concentration=="DMSO") %>% 
              select(Replicate,
                     total_D8L4Y_untreated=total_D8L4Y,
                     pSer2_3E10_untreated=pSer2_3E10)) %>%
  mutate(total_D8L4Y_normalised = total_D8L4Y / total_D8L4Y_untreated,
         pSer2_3E10_normalised = pSer2_3E10 / pSer2_3E10_untreated) 

wb_data_long <- wb_data %>%
  pivot_longer(cols=c(total_D8L4Y_normalised,pSer2_3E10_normalised),
               names_to = "antibody",
               values_to = "intensity")

wb_data_long_summary <- group_by(wb_data_long,antibody,concentration_numeric) %>%
              summarise(sd_intensity=sd(intensity),
                        intensity=mean(intensity))

wb_data_long %>%
  ggplot(aes(x=concentration_numeric,y=intensity,col=antibody)) +
    geom_point(pch='x') + 
  scale_x_log10(breaks=c(1e-9,1e-8,1e-7,1e-6,1e-5),
                labels=c("Vehicle","10 nM","100 nM","1 µM","10 µM")) +
  geom_line(data=wb_data_long_summary) +
  geom_point(data=wb_data_long_summary) +
  geom_errorbar(data=wb_data_long_summary,aes(ymin=intensity-sd_intensity,ymax=intensity+sd_intensity),width=0.1) +
  scale_y_continuous(limits = c(0,NA)) +
  theme_bw()

wb_data %>%
  ggplot(aes(x=total_D8L4Y_normalised,pSer2_3E10_normalised,col=log10(concentration_numeric))) +
  geom_point() +
  geom_abline() +
  scale_x_continuous("Total Pol II (D8L4Y western blot)",
                     breaks=c(0,0.5,1)) +
  scale_y_continuous("pSer2 Pol II (3E10 western blot)",
                     breaks=c(0,0.5,1)) +
  coord_cartesian(xlim=c(0,NA),ylim = c(0,NA)) +
  theme_bw(base_size = 7) +
  scale_color_viridis_c(
    name="5-Ph-IAA\nconcentration",
    begin=0,
    end=0.8,
    breaks=c(-9,-8,-7,-6,-5),
    labels=c("Vehicle","10 nM","100 nM","1 µM","10 µM")) +
  theme(panel.grid = element_blank(),
        aspect.ratio = 1,
        legend.key.size = unit(0.2,"cm"))
ggsave(filename = "activity_abundance_plot_WB_Ben.pdf",units="cm",width=8,height=5)

```

## Compare immunofluorescence with direct GFP measurements

Experimental data with mAC-POLR2A at different doses and times. GFP in 488 channel, F12 (mouse) in 561 channel, D8L4Y in 647 channel. Check quantitative correspondence of different measurements.

```{r}
read_quantification <- function(dir_name) {
  csv_files <- Sys.glob(file.path(dir_name,"*.csv"))
  dt_list <- lapply(csv_files, data.table::fread)
  df <- tibble(path = csv_files,
               feature_values = dt_list) %>%
    mutate(dir_name = dir_name,
           file_name = basename(path),
           feature_values = purrr::map(feature_values,tibble)) %>%
    separate(file_name,into = c("nuclei","well_name","ch1","ch2","ch3","ch4","sample","id"),sep="[,_.]", extra = "drop") %>%
    mutate(well_name = str_replace(well_name, "^Well", ""),
           field_id = as.integer(id))
  return(df)
}
```


### Load data

```{r}
data_dir_images <- "/srv/scratch/berrylab/z3536241/NikonSpinningDisk/260522_mACPOLR2A_Kinetic_pSer2/20260526_181658_086/"
data_dir_quant <- "/srv/scratch/berrylab/z3536241/NikonSpinningDisk/260522_mACPOLR2A_Kinetic_pSer2/20260526_181658_086/"
plate_layout_file <- "/srv/scratch/z3532965/src/blana/Alex/260522_mACPOLR2A_Kinetic_pSer2/260522_mACPOLR2A_Kinetic_pSer2.csv"

# Read plate layout
plate_layout <- read_csv(plate_layout_file, show_col_types = FALSE) %>%
  rename(well_name=WELL)

# Read and process quantification data
f <- read_quantification(file.path(data_dir_quant, "QUANTIFICATION")) %>%
     dplyr::mutate(data_dir = dirname(dirname(path))) %>%
     dplyr::select(-c(ch1, ch2, ch3, ch4, sample))

dat <- left_join(f,plate_layout,by="well_name") %>%
  unnest(feature_values) %>%
  mutate(unique_id = paste(well_name, label, sep = "_"))

```

### Check plate layout

```{r}

source("/srv/scratch/z3532965/src/blana/Scott/plot_plate_layout.R")

plate_layout %>% group_by(CELL,DRUG_CONCENTRATION,TIME_MINUTES,PRIMARY_1,SECONDARY_1,PRIMARY_2,SECONDARY_2) %>%
  count()

plot_plate_layout(plate_layout, color_by = "CELL", title = "Cell type")
plot_plate_layout(plate_layout, color_by = "DRUG", title = "Drug")
plot_plate_layout(plate_layout, color_by = "DRUG_CONCENTRATION", title = "Drug (µM)", log10_scale = TRUE)
plot_plate_layout(plate_layout, color_by = "TIME_MINUTES", title = "Time (min)")
plot_plate_layout(plate_layout, color_by = "PRIMARY_1", title = "Primary antibody 1")
plot_plate_layout(plate_layout, color_by = "SECONDARY_1", title = "Secondary antibody 1")
plot_plate_layout(plate_layout, color_by = "PRIMARY_2", title = "Primary antibody 2")
plot_plate_layout(plate_layout, color_by = "SECONDARY_2", title = "Secondary antibody 2")

```
```{r}
dat <- dat %>%
  rename_with(~ str_replace(., "Channel:0:0_", "Nuclei_")) %>%
  mutate(
    well_row = str_extract(well_name, "^[A-Z]"),
    well_col = as.integer(str_extract(well_name, "[0-9]+$"))
  ) %>%
  filter(Nuclei_is_border==F)

names(dat)

```
### QC

```{r,fig.height=8,fig.width=8}

bkgrd_primary_1 <- filter(dat,PRIMARY_1=="No primary") %>%
  group_by(well_name) %>%
  summarise(mean_bkgrd_561 = mean(Nuclei_intensity_mean_561)) %>%
  summarise(mean_bkgrd_561 = mean(mean_bkgrd_561)) %>%
  pull(mean_bkgrd_561)

dat %>% 
  filter(DRUG=="Vehicle") %>%
  group_by(well_name,CELL,DRUG_CONCENTRATION,TIME_MINUTES,PRIMARY_1,SECONDARY_1,PRIMARY_2,SECONDARY_2) %>%
  summarise(across(Nuclei_intensity_mean_561,mean)) %>%
  ggplot(aes(x=PRIMARY_1,y=Nuclei_intensity_mean_561,col=CELL)) +
  geom_boxplot() +
  geom_hline(yintercept=bkgrd_primary_1)

bkgrd_primary_2 <- filter(dat,PRIMARY_2=="No primary") %>%
  group_by(well_name) %>%
  summarise(mean_bkgrd_647 = mean(Nuclei_intensity_mean_647)) %>%
  summarise(mean_bkgrd_647 = mean(mean_bkgrd_647)) %>%
  pull(mean_bkgrd_647)

dat %>% 
  filter(DRUG=="Vehicle") %>%
  group_by(well_name,CELL,DRUG_CONCENTRATION,TIME_MINUTES,PRIMARY_1,SECONDARY_1,PRIMARY_2,SECONDARY_2) %>%
  summarise(across(Nuclei_intensity_mean_647,mean)) %>%
  ggplot(aes(x=PRIMARY_2,y=Nuclei_intensity_mean_647,col=CELL)) +
  geom_boxplot() +
  geom_hline(yintercept=bkgrd_primary_2)

bkgrd_gfp <- filter(dat,CELL=="HCT116 OsTIRF74G") %>%
  group_by(well_name) %>%
  summarise(mean_bkgrd_488 = mean(Nuclei_intensity_mean_488)) %>%
  summarise(mean_bkgrd_488 = mean(mean_bkgrd_488)) %>%
  pull(mean_bkgrd_488)

dat %>% 
  #filter(DRUG=="Vehicle") %>%
  group_by(well_name,CELL,DRUG,DRUG_CONCENTRATION,TIME_MINUTES,PRIMARY_1,SECONDARY_1,PRIMARY_2,SECONDARY_2) %>%
  summarise(across(Nuclei_intensity_mean_488,mean)) %>%
  ggplot(aes(x=CELL,y=Nuclei_intensity_mean_488,col=interaction(CELL,DRUG,DRUG_CONCENTRATION))) +
  geom_boxplot() +
  geom_jitter(position = position_jitterdodge()) +
  geom_hline(yintercept=bkgrd_gfp) +
  theme(axis.text.x = element_text(angle=90,hjust=1,vjust=0.5))

```
Replicate-to-replicate variability due to row/plate effects. Focus on replicate wells in upper half of plate.

```{r}
dat %>%
  filter(CELL=="HCT116 OsTIRF74G mAC-POLR2A Clone 5" & PRIMARY_1=="Total Pol II (F12)") %>%
  ggplot(aes(x=Nuclei_intensity_mean_561,col=well_row,group=well_name)) +
  geom_density() +
  facet_wrap(~interaction(DRUG,DRUG_CONCENTRATION))
```

Make comparisons

```{r}

f12_vs_gfp_vs_d8l4y <- dat %>%
  filter(well_row %in% c("D","E","F","G","H")) %>% 
  filter(CELL=="HCT116 OsTIRF74G mAC-POLR2A Clone 5" & PRIMARY_1=="Total Pol II (F12)") %>%
  group_by(well_name,CELL,DRUG,DRUG_CONCENTRATION,TIME_MINUTES,PRIMARY_1,SECONDARY_1,PRIMARY_2,SECONDARY_2) %>%
  summarise(mean_total_F12 = mean(Nuclei_intensity_mean_561),
            mean_total_D8L4Y = mean(Nuclei_intensity_mean_647),
            mean_GFP = mean(Nuclei_intensity_mean_488),
            .groups = 'drop') %>%
  group_by(CELL,DRUG,DRUG_CONCENTRATION,TIME_MINUTES,PRIMARY_1,SECONDARY_1,PRIMARY_2,SECONDARY_2) %>%
  summarise(sd_total_F12 = sd(mean_total_F12),
            sd_total_D8L4Y = sd(mean_total_D8L4Y),
            sd_GFP = sd(mean_GFP),
            mean_total_F12 = mean(mean_total_F12),
            mean_total_D8L4Y = mean(mean_total_D8L4Y),
            mean_GFP = mean(mean_GFP),
            .groups = 'drop') %>%
  mutate(DRUG_CONCENTRATION = if_else(DRUG_CONCENTRATION == 0, 1, DRUG_CONCENTRATION))

f12_vs_gfp_vs_d8l4y_norm <- f12_vs_gfp_vs_d8l4y %>%
  filter(DRUG=="Vehicle") %>%
  summarise(normaliser_total_F12 = mean(mean_total_F12),
            normaliser_total_D8L4Y = mean(mean_total_D8L4Y),
            normaliser_GFP = mean(mean_GFP)) %>%
  bind_cols(f12_vs_gfp_vs_d8l4y) %>%
  mutate(mean_total_F12 = mean_total_F12 / normaliser_total_F12,
         sd_total_F12 = sd_total_F12 / normaliser_total_F12,
         mean_total_D8L4Y = mean_total_D8L4Y / normaliser_total_D8L4Y,
         sd_total_D8L4Y = sd_total_D8L4Y / normaliser_total_D8L4Y,
         mean_GFP = mean_GFP / normaliser_GFP,
         sd_GFP = sd_GFP / normaliser_GFP)

# Calculate correlations for all three plot combinations
cor_gfp_f12 <- cor.test(f12_vs_gfp_vs_d8l4y_norm$mean_GFP, 
                         f12_vs_gfp_vs_d8l4y_norm$mean_total_F12)
cor_gfp_d8l4y <- cor.test(f12_vs_gfp_vs_d8l4y_norm$mean_GFP, 
                            f12_vs_gfp_vs_d8l4y_norm$mean_total_D8L4Y)
cor_f12_d8l4y <- cor.test(f12_vs_gfp_vs_d8l4y_norm$mean_total_F12, 
                            f12_vs_gfp_vs_d8l4y_norm$mean_total_D8L4Y)

make_cor_label <- function(cor_test) {
  r_str <- sprintf("%.3f", cor_test$estimate)
  paste0("italic(r)=='", r_str, "'")
}

f12_vs_gfp_plot <- f12_vs_gfp_vs_d8l4y_norm %>%
  ggplot(aes(x=mean_GFP,y=mean_total_F12,col=log10(DRUG_CONCENTRATION))) +
  geom_smooth(aes(x=mean_GFP,y=mean_total_F12),method = "lm",se=F, inherit.aes = F,col="black",linewidth = 0.5,fullrange = T,  linetype="dashed") +
  geom_errorbar(aes(ymin=mean_total_F12-sd_total_F12,ymax=mean_total_F12+sd_total_F12)) +
  geom_errorbarh(aes(xmin=mean_GFP-sd_GFP,xmax=mean_GFP+sd_GFP)) +
  geom_point() +
  coord_cartesian(xlim=c(0,NA),ylim=c(0,NA)) +
  geom_hline(yintercept = bkgrd_primary_1/f12_vs_gfp_vs_d8l4y_norm$normaliser_total_F12) +
  geom_vline(xintercept = bkgrd_gfp/f12_vs_gfp_vs_d8l4y_norm$normaliser_GFP) +
  annotate("label", x=0.01, y=1.05, label=make_cor_label(cor_gfp_f12),
         parse=TRUE, hjust=0, size=2.5,
         fill="white", colour="black", label.size=0.3) +
  scale_color_viridis_c(
    name="5-Ph-IAA\nconcentration",
    begin=0,
    end=0.8,
    breaks=c(0,1,2,3,4),
    labels=c("Vehicle","10 nM","100 nM","1 µM","10 µM")) +
  scale_x_continuous(limits = c(0,1.25),breaks=c(0,0.5,1.0)) +
  scale_x_continuous(name="POLR2A (mClover)",limits = c(0,1.1),breaks=c(0,0.5,1.0)) +
  scale_y_continuous(name="POLR2A IF (F12)",limits = c(0,1.1),breaks=c(0,0.5,1.0))

d8l4y_vs_gfp_plot <- f12_vs_gfp_vs_d8l4y_norm %>%
  ggplot(aes(x=mean_GFP,y=mean_total_D8L4Y,col=log10(DRUG_CONCENTRATION))) +
  geom_smooth(aes(x=mean_GFP,y=mean_total_D8L4Y),method = "lm",se=F, inherit.aes = F,col="black",linewidth = 0.5,fullrange = T,  linetype="dashed") +
  geom_errorbar(aes(ymin=mean_total_D8L4Y-sd_total_D8L4Y,ymax=mean_total_D8L4Y+sd_total_D8L4Y)) +
  geom_errorbarh(aes(xmin=mean_GFP-sd_GFP,xmax=mean_GFP+sd_GFP)) +
  geom_point() +
  coord_cartesian(xlim=c(0,NA),ylim=c(0,NA)) +
  geom_hline(yintercept = bkgrd_primary_2/f12_vs_gfp_vs_d8l4y_norm$normaliser_total_D8L4Y) +
  geom_vline(xintercept = bkgrd_gfp/f12_vs_gfp_vs_d8l4y_norm$normaliser_GFP) +
  annotate("label", x=0.01, y=1.05, label=make_cor_label(cor_gfp_d8l4y),
         parse=TRUE, hjust=0, size=2.5,
         fill="white", colour="black", label.size=0.3) +
  scale_color_viridis_c(
    name="5-Ph-IAA\nconcentration",
    begin=0,
    end=0.8,
    breaks=c(0,1,2,3,4),
    labels=c("Vehicle","10 nM","100 nM","1 µM","10 µM")) +
  scale_x_continuous(name="POLR2A (mClover)",limits = c(0,1.1),breaks=c(0,0.5,1.0)) +
  scale_y_continuous(name="POLR2A IF (D8L4Y)",limits = c(0,1.1),breaks=c(0,0.5,1.0))

d8l4y_vs_f12_plot <- f12_vs_gfp_vs_d8l4y_norm %>%
  ggplot(aes(x=mean_total_F12,y=mean_total_D8L4Y,col=log10(DRUG_CONCENTRATION))) +
  geom_smooth( aes(x=mean_total_F12,y=mean_total_D8L4Y),method = "lm",se=F, inherit.aes = F,col="black",linewidth = 0.5,fullrange = T, linetype="dashed") +
  geom_errorbar(aes(ymin=mean_total_D8L4Y-sd_total_D8L4Y,ymax=mean_total_D8L4Y+sd_total_D8L4Y)) +
  geom_errorbarh(aes(xmin=mean_total_F12-sd_total_F12,xmax=mean_total_F12+sd_total_F12)) +
  geom_point() +
  coord_cartesian(xlim=c(0,NA),ylim=c(0,NA)) +
  geom_hline(yintercept = bkgrd_primary_2/f12_vs_gfp_vs_d8l4y_norm$normaliser_total_D8L4Y) +
  geom_vline(xintercept = bkgrd_primary_1/f12_vs_gfp_vs_d8l4y_norm$normaliser_total_F12) +
  annotate("label", x=0.01, y=1.05, label=make_cor_label(cor_f12_d8l4y),
         parse=TRUE, hjust=0, size=2.5,
         fill="white", colour="black", label.size=0.3) +
  scale_color_viridis_c(
    name="5-Ph-IAA\nconcentration",
    begin=0,
    end=0.8,
    breaks=c(0,1,2,3,4),
    labels=c("Vehicle","10 nM","100 nM","1 µM","10 µM")) +
  scale_x_continuous(name="POLR2A IF (F12)",limits = c(0,1.1),breaks=c(0,0.5,1.0)) +
  scale_y_continuous(name="POLR2A IF (D8L4Y)",limits = c(0,1.1),breaks=c(0,0.5,1.0))

d8l4y_vs_f12_plot

design <-
"
A#
BC
"

f12_vs_gfp_plot + d8l4y_vs_gfp_plot + d8l4y_vs_f12_plot + plot_layout(design = design) & 
  theme_bw(base_size = 8) & 
  theme(legend.position = "none",
        panel.grid = element_blank())

ggsave("f12_vs_gfp_vs_d8l4y.pdf",width=10,height=8,units="cm")

l <- cowplot::get_legend(f12_vs_gfp_plot + theme(legend.key.size = unit(0.4,"cm")))
ggsave("f12_vs_gfp_vs_d8l4y_legend.pdf", plot=l, width = 5, height = 5,units="cm")

```

## Compare WB and IF

We can now also match the 4 h data with the western blots performed at this timepoint to see how they compare

```{r wb_to_IF}

wb_data_summary <- wb_data %>%
  group_by(concentration,concentration_numeric) %>%
  summarise(across(c(total_D8L4Y_normalised,pSer2_3E10_normalised),c(mean=~mean(.),sd=~sd(.))),.groups = 'drop') %>%
  mutate(DRUG=if_else(concentration=="DMSO","Vehicle","5PhIAA"),
         DRUG_CONCENTRATION=concentration_numeric*10^9) %>%
  select(DRUG,
         DRUG_CONCENTRATION,
         D8L4Y_WB_mean=total_D8L4Y_normalised_mean,
         D8L4Y_WB_sd=total_D8L4Y_normalised_sd)

if_data_summary <- f12_vs_gfp_vs_d8l4y %>%
  filter(TIME_MINUTES==240) %>%
  mutate(mean_total_D8L4Y = mean_total_D8L4Y - bkgrd_primary_2) %>%
  mutate(mean_total_F12 = mean_total_F12 - bkgrd_primary_1) %>%
  select(DRUG,DRUG_CONCENTRATION,
         D8L4Y_IF_mean=mean_total_D8L4Y,
         D8L4Y_IF_sd=sd_total_D8L4Y,
         F12_IF_mean=mean_total_F12,
         F12_IF_sd=sd_total_F12)

if_norm <- if_data_summary %>%
  filter(DRUG=="Vehicle") %>%
  summarise(D8L4Y_IF_mean_normaliser = mean(D8L4Y_IF_mean),
            F12_IF_mean_normaliser = mean(F12_IF_mean)) %>%
  bind_cols(if_data_summary) %>%
    mutate(
      D8L4Y_IF_mean = D8L4Y_IF_mean / D8L4Y_IF_mean_normaliser,
      D8L4Y_IF_sd = D8L4Y_IF_sd / D8L4Y_IF_mean_normaliser,
      F12_IF_mean = F12_IF_mean / F12_IF_mean_normaliser,
      F12_IF_sd = F12_IF_sd / F12_IF_mean_normaliser) %>%
  pivot_longer(
    cols = c(D8L4Y_IF_mean, D8L4Y_IF_sd, F12_IF_mean, F12_IF_sd,
             D8L4Y_IF_mean_normaliser, F12_IF_mean_normaliser),
    names_to = c("antibody", ".value"),
    names_pattern = "^(D8L4Y|F12)_(.+)$"
  )

plot_data <- left_join(if_norm, wb_data_summary)

# Calculate correlation per antibody
cor_labels <- plot_data %>%
  group_by(antibody) %>%
  summarise(
    r = cor(IF_mean, D8L4Y_WB_mean, use = "complete.obs"),
    p = cor.test(IF_mean, D8L4Y_WB_mean)$p.value,
    label = paste0("italic(r)==", round(r, 3)),
    x = 0.05,
    y = ifelse(antibody == "D8L4Y", 1.05, 0.95)
  )

plot_data %>%
  ggplot(aes(x = IF_mean, y = D8L4Y_WB_mean, colour = antibody)) +
  geom_smooth(aes(x = IF_mean, y = D8L4Y_WB_mean, group = antibody, colour = antibody),
              method = "lm", se = F, inherit.aes = F, linewidth = 0.5,
              fullrange = T, linetype = "dashed", show.legend = FALSE) +
  geom_errorbarh(aes(xmin = IF_mean - IF_sd, xmax = IF_mean + IF_sd),show.legend = FALSE) +
  geom_errorbar(aes(ymin = D8L4Y_WB_mean - D8L4Y_WB_sd, ymax = D8L4Y_WB_mean + D8L4Y_WB_sd),show.legend = FALSE) +
    geom_point() +
  geom_text(data = cor_labels,
          aes(x = x, y = y, label = label, colour = antibody),
          hjust = 0, size = 2.5, inherit.aes = F, parse = TRUE, show.legend = FALSE) +
  scale_x_continuous(name = "POLR2A immunofluorescence",
                     limits = c(0, 1.1), breaks = c(0, 0.5, 1.0)) +
  scale_y_continuous(name = "POLR2A western\nblot (D8L4Y)",
                     limits = c(0, 1.1), breaks = c(0, 0.5, 1.0)) +
  theme_bw(base_size = 8) +
  theme(legend.position = "right",
        legend.title = element_blank(),
        panel.grid = element_blank(),
        aspect.ratio = 1) +
  scale_colour_manual(values = c("#0072B2", "#D55E00"))
ggsave("if_vs_wb_d8l4y.pdf",width=8,height=5,units="cm")

```


## Single-cell

Let's look at variability between single-cells in pSer2 and D84LY, in the same cells.

```{r}
dat_costain_D8L4Y_pSer2 <- filter(dat,well_row=="N") %>%
  filter(PRIMARY_1 %in% c("pSer2 Pol II (3E10)") & PRIMARY_2 %in% c("Total Pol II (D8L4Y)"))

single_cell_variability_D8L4Y_pSer2 <- dat_costain_D8L4Y_pSer2 %>%
  mutate(pSer2 = Nuclei_intensity_mean_561 - bkgrd_primary_1,
         total_D8L4Y = Nuclei_intensity_mean_647 - bkgrd_primary_2) %>%
  group_by(well_name) %>%
  summarise(mean_pSer2 = mean(pSer2),
            sd_pSer2 = sd(pSer2),
            cv_pSer2 = sd_pSer2 / mean_pSer2,
            mean_total_D8L4Y = mean(total_D8L4Y),
            sd_total_D8L4Y = sd(total_D8L4Y),
            cv_total_D8L4Y = sd_total_D8L4Y / mean_total_D8L4Y) 

single_cell_variability_D8L4Y_pSer2 %>%
  pivot_longer(cols=c(cv_pSer2,cv_total_D8L4Y),values_to = "CV",names_to = "antibody",names_prefix = "cv_") %>%
  ggplot(aes(x=antibody,y=CV)) + geom_boxplot() + ylim(0,NA)

```

```{r}

dat_costain_D8L4Y_pSer2_normalised <- dat_costain_D8L4Y_pSer2 %>%
  mutate(pSer2 = Nuclei_intensity_mean_561 - bkgrd_primary_1,
         total_D8L4Y = Nuclei_intensity_mean_647 - bkgrd_primary_2) %>%
  left_join(dplyr::select(single_cell_variability_D8L4Y_pSer2,well_name,mean_pSer2,mean_total_D8L4Y),by = join_by(well_name)) %>%
  mutate(total_D8L4Y_norm=total_D8L4Y/mean_total_D8L4Y,pSer2_norm=pSer2/mean_pSer2)

dat_costain_D8L4Y_pSer2_normalised %>%
  dplyr::select(total_D8L4Y_norm,pSer2_norm,well_name) %>%
  pivot_longer(cols = c(total_D8L4Y_norm,pSer2_norm), values_to = "intensity", names_to = "antibody") %>%
  ggplot(aes(x=intensity,col=antibody,group = interaction(well_name,antibody))) + geom_density()

perc_limits <- dat_costain_D8L4Y_pSer2_normalised %>%
  summarise(total_low = quantile(total_D8L4Y_norm,0.01),
            total_high = quantile(total_D8L4Y_norm,0.99),
            pSer2_low = quantile(pSer2_norm,0.01),
            pSer2_high = quantile(pSer2_norm,0.99))

library(ggpointdensity)

dat_costain_D8L4Y_pSer2_normalised %>%
  filter(Nuclei_centroid_1 > 250 & Nuclei_centroid_0 < 2000 & Nuclei_centroid_1 > 250 & Nuclei_centroid_1 < 2000) %>%
  ggplot(aes(x=Nuclei_intensity_sum_405,col=well_name)) +
  geom_density()

dat_costain_D8L4Y_pSer2_normalised %>%
  filter(Nuclei_centroid_1 > 250 & Nuclei_centroid_0 < 2000 & Nuclei_centroid_1 > 250 & Nuclei_centroid_1 < 2000) %>%
  ggplot(aes(x=Nuclei_area)) +
  geom_density() +
  geom_vline(xintercept = 750) +
  geom_vline(xintercept = 5000)

dat_costain_D8L4Y_pSer2_normalised_clean <- dat_costain_D8L4Y_pSer2_normalised %>%
  filter(Nuclei_area > 750 & Nuclei_intensity_sum_405 > 5000) %>%
  filter(Nuclei_intensity_sum_405 > 2e5 & Nuclei_intensity_sum_405 < 1.2e6) %>%
  filter(Nuclei_centroid_1 > 250 & Nuclei_centroid_0 < 2000 & Nuclei_centroid_1 > 250 & Nuclei_centroid_1 < 2000)

 dat_costain_D8L4Y_pSer2_normalised_clean %>%
  ggplot(aes(x = total_D8L4Y_norm, y = pSer2_norm)) + 
  geom_pointdensity() +
  scale_color_viridis_c() +
  geom_abline() + 
  geom_smooth(data = filter(dat_costain_D8L4Y_pSer2_normalised_clean,
                            total_D8L4Y_norm > perc_limits$total_low & total_D8L4Y_norm < perc_limits$total_high),
              colour="red",
              method = "lm") +
  coord_fixed()

```

```{r}
dat_costain_D8L4Y_pSer2_normalised %>%
  ggplot(aes(x=Nuclei_centroid_1,y=Nuclei_centroid_0,col=log10(Nuclei_intensity_mean_647))) + 
  geom_point() +
  scale_color_viridis_c()

dat_costain_D8L4Y_pSer2_normalised %>%
  ggplot(aes(x=Nuclei_centroid_1,y=Nuclei_centroid_0,col=log10(Nuclei_intensity_mean_561))) + 
  geom_point() +
  scale_color_viridis_c()
```
```{r}
mac_F12_D8L4Y <- filter(dat, CELL == "HCT116 OsTIRF74G mAC-POLR2A Clone 5" & PRIMARY_1=="Total Pol II (F12)" & DRUG=="Vehicle")
mac_Ser2_D8L4Y <- filter(dat, CELL == "HCT116 OsTIRF74G mAC-POLR2A Clone 5" & PRIMARY_1=="pSer2 Pol II (3E10)" & DRUG=="Vehicle")

mac_F12_D8L4Y_positive <- mac_F12_D8L4Y %>%
  mutate(total_F12 = Nuclei_intensity_mean_561 - bkgrd_primary_1,
         total_D8L4Y = Nuclei_intensity_mean_647 - bkgrd_primary_2,
         GFP = Nuclei_intensity_mean_488 - bkgrd_gfp) %>%
  filter(total_F12 > 0 & total_D8L4Y > 0) %>%
  mutate(log_total_F12 = log(total_F12),
         log_total_D8L4Y = log(total_D8L4Y),
         log_GFP = log(GFP))

n_dropped_F12 <- nrow(mac_F12_D8L4Y_positive) - nrow(mac_F12_D8L4Y)

mac_Ser2_D8L4Y_positive <- mac_Ser2_D8L4Y %>%
  mutate(pSer2 = Nuclei_intensity_mean_561 - bkgrd_primary_1,
         total_D8L4Y = Nuclei_intensity_mean_647 - bkgrd_primary_2,
         GFP = Nuclei_intensity_mean_488 - bkgrd_gfp) %>%
  filter(pSer2 > 0 & total_D8L4Y > 0) %>%
  mutate(log_pSer2 = log(pSer2),
         log_total_D8L4Y = log(total_D8L4Y),
         log_GFP = log(GFP))
  
n_dropped_Ser2 <- nrow(mac_Ser2_D8L4Y_positive) - nrow(mac_Ser2_D8L4Y)

```


```{r}

tidy_rlm <- function(d) {
  broom::tidy(MASS::rlm(log_y ~ log_GFP, data = d, method = "MM", maxit = 200))
}
tidy_rlm_safe <- purrr::possibly(
  tidy_rlm,
  otherwise = tibble::tibble(term = "log_GFP", estimate = NA_real_, std.error = NA_real_)
)

slopes_Ser2_D8L4Y_vsGFP <- mac_Ser2_D8L4Y_positive %>%
  group_by(well_name) %>%
  filter(log_GFP >= quantile(log_GFP, 0.01),
         log_GFP <= quantile(log_GFP, 0.99)) %>%
  ungroup() %>%
  pivot_longer(c(log_pSer2, log_total_D8L4Y),
               names_to = "readout", values_to = "log_y") %>%
  group_by(well_name, readout) %>%
  nest() %>%
  mutate(
    n_cells = map_int(data, nrow),
    coefs   = map(data, tidy_rlm_safe)
  ) %>%
  dplyr::select(well_name, readout, n_cells, coefs) %>%
  unnest(coefs) %>%
  filter(term == "log_GFP") %>%
  transmute(well_name, readout, n_cells, slope = estimate, se = std.error) %>%
  ungroup()

slopes_F12_D8L4Y_vsGFP <- mac_F12_D8L4Y_positive %>%
  group_by(well_name) %>%
  filter(log_GFP >= quantile(log_GFP, 0.01),
         log_GFP <= quantile(log_GFP, 0.99)) %>%
  ungroup() %>%
  pivot_longer(c(log_total_F12, log_total_D8L4Y),
               names_to = "readout", values_to = "log_y") %>%
  group_by(well_name, readout) %>%
  nest() %>%
  mutate(
    n_cells = map_int(data, nrow),
    coefs   = map(data, tidy_rlm_safe)
  ) %>%
  dplyr::select(well_name, readout, n_cells, coefs) %>%
  unnest(coefs) %>%
  filter(term == "log_GFP") %>%
  transmute(well_name, readout, n_cells, slope = estimate, se = std.error) %>%
  ungroup()

slopes_Ser2_D8L4Y_wide <- slopes_Ser2_D8L4Y_vsGFP %>%
  dplyr::select(well_name, readout, slope) %>%
  pivot_wider(names_from = readout, values_from = slope) %>%
  mutate(beta = log_pSer2 / log_total_D8L4Y)   # per-well attenuation-corrected elasticity

slopes_F12_D8L4Y_wide <- slopes_F12_D8L4Y_vsGFP %>%
  dplyr::select(well_name, readout, slope) %>%
  pivot_wider(names_from = readout, values_from = slope) %>%
  mutate(beta = log_total_F12 / log_total_D8L4Y)   # per-well attenuation-corrected elasticity

# significance: is the pSer2 slope below the D8L4Y (proportional-total) slope?
t.test(slopes_Ser2_D8L4Y_wide$log_pSer2, slopes_Ser2_D8L4Y_wide$log_total_D8L4Y, paired = TRUE)

# effect size: the true elasticity, GFP noise cancelled
mean(slopes_Ser2_D8L4Y_wide$beta)

```


```{r}

slope_Ser2_D8L4Y_means <- slopes_Ser2_D8L4Y_vsGFP %>%
  group_by(readout) %>%
  summarise(mean_slope = mean(slope), .groups = "drop")

Ser2_slope_plot <- ggplot(slopes_Ser2_D8L4Y_vsGFP, aes(readout, slope, group = well_name)) +
  geom_hline(yintercept = 1, linetype = "dotted", colour = "grey50") +
  geom_line(alpha = 0.35) +
  geom_point(size = 2, alpha = 0.6) +
  # well-averaged slope for each readout
  geom_line(data = slope_Ser2_D8L4Y_means, aes(readout, mean_slope, group = 1),
            inherit.aes = FALSE, colour = "firebrick", linewidth = 1) +
  geom_point(data = slope_Ser2_D8L4Y_means, aes(readout, mean_slope),
             inherit.aes = FALSE, colour = "firebrick", size = 4) +
  scale_x_discrete(limits = c("log_total_D8L4Y", "log_pSer2"),
                   labels = c(log_total_D8L4Y = "total (D8L4Y)\nvs GFP",
                              log_pSer2 = "pSer2\nvs GFP")) +
  scale_y_continuous(limits = c(0,1)) +
  labs(x = NULL, y = "log-log slope (regressed on GFP)") +
  theme_minimal() +
  theme_bw()

slope_F12_D8L4Y_means <- slopes_F12_D8L4Y_vsGFP %>%
  group_by(readout) %>%
  summarise(mean_slope = mean(slope), .groups = "drop")

F12_slope_plot <- ggplot(slopes_F12_D8L4Y_vsGFP, aes(readout, slope, group = well_name)) +
  geom_hline(yintercept = 1, linetype = "dotted", colour = "grey50") +
  geom_line(alpha = 0.35) +
  geom_point(size = 2, alpha = 0.6) +
  # well-averaged slope for each readout
  geom_line(data = slope_F12_D8L4Y_means, aes(readout, mean_slope, group = 1),
            inherit.aes = FALSE, colour = "firebrick", linewidth = 1) +
  geom_point(data = slope_F12_D8L4Y_means, aes(readout, mean_slope),
             inherit.aes = FALSE, colour = "firebrick", size = 4) +
  scale_x_discrete(limits = c("log_total_D8L4Y", "log_total_F12"),
                   labels = c(log_total_D8L4Y = "total (D8L4Y)\nvs GFP",
                              log_total_F12 = "total (F12)\nvs GFP")) +
  scale_y_continuous(limits = c(0,1)) +
  labs(x = NULL, y = "log-log slope (regressed on GFP)") +
  theme_bw()

library(patchwork)
F12_slope_plot + Ser2_slope_plot

```
```{r, fig.height=8,fig.width=8}


plot_df_all <- mac_Ser2_D8L4Y_positive %>%
  pivot_longer(c(log_pSer2, log_total_D8L4Y), names_to = "readout", values_to = "log_y") %>%
  group_by(well_name) %>%
  mutate(
    log_GFP_c = log_GFP - mean(log_GFP),
    in_range  = dplyr::between(log_GFP,
                               quantile(log_GFP, 0.01),
                               quantile(log_GFP, 0.99))
  ) %>%
  group_by(well_name, readout) %>%
  mutate(log_y_c = log_y - mean(log_y)) %>%
  ungroup()

ggplot(plot_df_all, aes(log_GFP_c, log_y_c, colour = readout)) +
  geom_point(alpha = 0.05, size = 0.4) +                       # all cells
  geom_abline(slope = 1, intercept = 0, linetype = "dotted", colour = "grey60") +
  geom_smooth(data = ~ dplyr::filter(.x, in_range),            # fit central 90% only
              method = MASS::rlm, method.args = list(method = "MM", maxit = 200),
              formula = y ~ x, se = FALSE, linewidth = 0.9) +
  facet_wrap(~ well_name) +
  scale_colour_manual(values = c(log_total_D8L4Y = "steelblue", log_pSer2 = "firebrick"),
                      labels = c(log_total_D8L4Y = "total (D8L4Y)", log_pSer2 = "pSer2")) +
  labs(x = "centred log(GFP)", y = "centred log(readout)",
       colour = NULL) +
  coord_cartesian(xlim = c(-1,1),ylim = c(-1,1)) +
  theme_minimal() +
  theme(legend.position = "top")


```

## Well-average analysis across mAC-POLR2A data

Load data and check for plate effects

```{r}
F12_D8L4Y_all <- filter(dat, PRIMARY_1=="Total Pol II (F12)")
Ser2_D8L4Y_all <- filter(dat, PRIMARY_1=="pSer2 Pol II (3E10)")

F12_D8L4Y_all_summary <- F12_D8L4Y_all %>%
  mutate(total_F12 = Nuclei_intensity_mean_561 - bkgrd_primary_1,
         total_D8L4Y = Nuclei_intensity_mean_647 - bkgrd_primary_2,
         GFP = Nuclei_intensity_mean_488 - bkgrd_gfp) %>%
  group_by(well_name,CELL,DRUG_CONCENTRATION,TIME_MINUTES,PRIMARY_1,SECONDARY_1,PRIMARY_2,SECONDARY_2,well_row,well_col) %>%
  summarise(across(c(total_F12,total_D8L4Y,GFP),c(mean=~mean(.))),.groups = "drop")

Ser2_D8L4Y_all_summary <- Ser2_D8L4Y_all %>%
  mutate(pSer2 = Nuclei_intensity_mean_561 - bkgrd_primary_1,
         total_D8L4Y = Nuclei_intensity_mean_647 - bkgrd_primary_2,
         GFP = Nuclei_intensity_mean_488 - bkgrd_gfp) %>%
  group_by(well_name,CELL,DRUG_CONCENTRATION,TIME_MINUTES,PRIMARY_1,SECONDARY_1,PRIMARY_2,SECONDARY_2,well_row,well_col) %>%
  summarise(across(c(pSer2,total_D8L4Y,GFP),c(mean=~mean(.))),.groups = "drop")

# row effects for F12
F12_D8L4Y_all_summary %>% 
  filter(CELL=="HCT116 OsTIRF74G") %>%
  ggplot(aes(x=well_row,y=total_F12_mean,col=well_col)) + geom_point()

# row effects for D8L4Y (also correct per left/right block of the plate)
bind_rows(F12_D8L4Y_all_summary,Ser2_D8L4Y_all_summary) %>% 
  filter(CELL=="HCT116 OsTIRF74G") %>%
  ggplot(aes(x=well_row,y=total_D8L4Y_mean,col=well_col)) + geom_point()

# row effects for Ser2
Ser2_D8L4Y_all_summary %>% 
  filter(CELL=="HCT116 OsTIRF74G") %>%
  ggplot(aes(x=well_row,y=pSer2_mean,col=well_col)) + geom_point()

# GFP intensity is stable during acquisition
bind_rows(F12_D8L4Y_all_summary,Ser2_D8L4Y_all_summary) %>% 
  filter(CELL=="HCT116 OsTIRF74G mAC-POLR2A Clone 5" & DRUG_CONCENTRATION==0) %>%
  ggplot(aes(x=well_row,y=GFP_mean,col=well_col)) + 
  geom_boxplot() +
  geom_jitter()

```
```{r}
# Correct row effects by normalising to the control cells in each row 
F12_D8L4Y_normalisers <- F12_D8L4Y_all_summary %>%
  group_by(well_row) %>%
  summarise(across(c(total_F12_mean,total_D8L4Y_mean),c(row_mean=~mean(.))),.groups = "drop")

F12_D8L4Y_all_summary_normalised <- F12_D8L4Y_all_summary %>% 
  inner_join(F12_D8L4Y_normalisers,by = join_by(well_row)) %>%
  mutate(total_F12_mean_normalised = total_F12_mean/total_F12_mean_row_mean,
         total_D8L4Y_mean_normalised = total_D8L4Y_mean/total_D8L4Y_mean_row_mean)

Ser2_D8L4Y_normalisers <- Ser2_D8L4Y_all_summary %>%
  group_by(well_row) %>%
  summarise(across(c(pSer2_mean,total_D8L4Y_mean),c(row_mean=~mean(.))),.groups = "drop")

Ser2_D8L4Y_all_summary_normalised <- Ser2_D8L4Y_all_summary %>% 
  inner_join(Ser2_D8L4Y_normalisers,by = join_by(well_row)) %>%
  mutate(pSer2_mean_normalised = pSer2_mean/pSer2_mean_row_mean,
         total_D8L4Y_mean_normalised = total_D8L4Y_mean/total_D8L4Y_mean_row_mean)


```



```{r}

vehicle_treated <- F12_D8L4Y_all_summary_normalised %>%
  filter(CELL=="HCT116 OsTIRF74G mAC-POLR2A Clone 5") %>%
  filter(DRUG_CONCENTRATION==0) %>%
  summarise(total_F12_mean_normalised = mean(total_F12_mean_normalised),
            total_D8L4Y_mean_normalised = mean(total_D8L4Y_mean_normalised),
            GFP_mean = mean(GFP_mean))

D8L4Y_vs_F12 <- F12_D8L4Y_all_summary_normalised %>%
  filter(CELL=="HCT116 OsTIRF74G mAC-POLR2A Clone 5") %>%
  ggplot(aes(x=total_F12_mean_normalised/vehicle_treated$total_F12_mean_normalised,
             y=total_D8L4Y_mean_normalised/vehicle_treated$total_D8L4Y_mean_normalised)) + 
  geom_point() +
  geom_abline() +
  coord_cartesian(xlim = c(0,NA),ylim = c(0,NA))

D8L4Y_vs_GFP <- F12_D8L4Y_all_summary_normalised %>%
  filter(CELL=="HCT116 OsTIRF74G mAC-POLR2A Clone 5") %>%
  ggplot(aes(x=GFP_mean/vehicle_treated$GFP_mean,
             y=total_D8L4Y_mean_normalised/vehicle_treated$total_D8L4Y_mean_normalised)) + 
  geom_point() +
  geom_abline() +
  coord_cartesian(xlim = c(0,NA),ylim = c(0,NA))

F12_vs_GFP <- F12_D8L4Y_all_summary_normalised %>%
  filter(CELL=="HCT116 OsTIRF74G mAC-POLR2A Clone 5") %>%
  ggplot(aes(x=GFP_mean/vehicle_treated$GFP_mean,
             y=total_F12_mean_normalised/vehicle_treated$total_F12_mean_normalised)) + 
  geom_point() +
  geom_abline() +
  coord_cartesian(xlim = c(0,NA),ylim = c(0,NA))

F12_vs_GFP

```