```{r setup_r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls())
library(tidyverse)
library(patchwork)

cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
```

# Population-averaged

## Visualise pSer2 vs total from western blot

mAC-POLR2A cells treated for 4 h with 5-Ph-IAA at different doses (or vehicle). Visualise changes over dose and correspondence between total and pSer2 from western blot data.

```{r}
wb_csv <- "data/20260610_degron_titration.csv"

wb_data <- read_csv(wb_csv) %>%
  select(Replicate,concentration=`5-Ph-IAA concentration`,D8L4Y, S2, vinculin) %>%
  filter(!is.na(Replicate)) %>%
  mutate(total_D8L4Y = D8L4Y / vinculin,
         pSer2_3E10 = S2 / vinculin,
         concentration_numeric = case_when(
           concentration=="DMSO" ~ 1e-9,
           concentration=="10 nM" ~ 10e-9,
           concentration=="100 nM" ~ 100e-9,
           concentration=="1 µM" ~ 1e-6,
           concentration=="10 µM" ~ 10e-6,
           TRUE ~ NA_real_)) 

wb_data <- wb_data %>%
  left_join(filter(wb_data,concentration=="DMSO") %>% 
              select(Replicate,
                     total_D8L4Y_untreated=total_D8L4Y,
                     pSer2_3E10_untreated=pSer2_3E10)) %>%
  mutate(total_D8L4Y_normalised = total_D8L4Y / total_D8L4Y_untreated,
         pSer2_3E10_normalised = pSer2_3E10 / pSer2_3E10_untreated) 

wb_data_long <- wb_data %>%
  pivot_longer(cols=c(total_D8L4Y_normalised,pSer2_3E10_normalised),
               names_to = "antibody",
               values_to = "intensity")

wb_data_long_summary <- group_by(wb_data_long,antibody,concentration_numeric) %>%
              summarise(sd_intensity=sd(intensity),
                        intensity=mean(intensity))

wb_data_long %>%
  ggplot(aes(x=concentration_numeric,y=intensity,col=antibody)) +
    geom_point(pch='x') + 
  scale_x_log10(breaks=c(1e-9,1e-8,1e-7,1e-6,1e-5),
                labels=c("Vehicle","10 nM","100 nM","1 µM","10 µM")) +
  geom_line(data=wb_data_long_summary) +
  geom_point(data=wb_data_long_summary) +
  geom_errorbar(data=wb_data_long_summary,aes(ymin=intensity-sd_intensity,ymax=intensity+sd_intensity),width=0.1) +
  scale_y_continuous(limits = c(0,NA)) +
  theme_bw()

wb_data %>%
  ggplot(aes(x=total_D8L4Y_normalised,pSer2_3E10_normalised,col=log10(concentration_numeric))) +
  geom_point(size=1.5) +
  geom_abline() +
  scale_x_continuous("POLR2A western\nblot (D8L4Y )",
                     breaks=c(0,0.5,1)) +
  scale_y_continuous("pSer2-POLR2A\nwestern blot (3E10)",
                     breaks=c(0,0.5,1)) +
  coord_cartesian(xlim=c(0,NA),ylim = c(0,NA)) +
  theme_bw(base_size = 8) +
  scale_color_viridis_c(
    name="5-Ph-IAA\nconcentration",
    begin=0,
    end=0.8,
    breaks=c(-9,-8,-7,-6,-5),
    labels=c("Vehicle","10 nM","100 nM","1 µM","10 µM")) +
  theme(panel.grid = element_blank(),
        aspect.ratio = 1,
        legend.key.size = unit(0.3,"cm"),
        axis.text=element_text(size=8),
        axis.title=element_text(size=8),
        legend.title=element_text(size=8),
        legend.text=element_text(size=8),
        panel.border = element_rect(linewidth = 1))
ggsave(filename = "activity_abundance_plot_WB_Ben.pdf",units="cm",width=7,height=5)

```

## Compare immunofluorescence with direct GFP measurements

Experimental data with mAC-POLR2A at different doses and times. GFP in 488 channel, F12 (mouse) in 561 channel, D8L4Y in 647 channel. Check quantitative correspondence of different measurements.

```{r}
read_quantification <- function(dir_name) {
  csv_files <- Sys.glob(file.path(dir_name,"*.csv"))
  dt_list <- lapply(csv_files, data.table::fread)
  df <- tibble(path = csv_files,
               feature_values = dt_list) %>%
    mutate(dir_name = dir_name,
           file_name = basename(path),
           feature_values = purrr::map(feature_values,tibble)) %>%
    separate(file_name,into = c("nuclei","well_name","ch1","ch2","ch3","ch4","sample","id"),sep="[,_.]", extra = "drop") %>%
    mutate(well_name = str_replace(well_name, "^Well", ""),
           field_id = as.integer(id))
  return(df)
}
```


### Load data

```{r}
data_dir_images <- "/srv/scratch/berrylab/z3536241/NikonSpinningDisk/260522_mACPOLR2A_Kinetic_pSer2/20260526_181658_086/"
data_dir_quant <- "/srv/scratch/berrylab/z3532965/NikonSpinningDisk/260522_mACPOLR2A_Kinetic_pSer2/20260526_181658_086/"
plate_layout_file <- "/srv/scratch/z3532965/src/blana/Alex/260522_mACPOLR2A_Kinetic_pSer2/260522_mACPOLR2A_Kinetic_pSer2.csv"

# Read plate layout
plate_layout <- read_csv(plate_layout_file, show_col_types = FALSE) %>%
  rename(well_name=WELL)

# Read and process quantification data
f <- read_quantification(file.path(data_dir_quant, "QUANTIFICATION")) %>%
     dplyr::mutate(data_dir = dirname(dirname(path))) %>%
     dplyr::select(-c(ch1, ch2, ch3, ch4, sample))

dat <- left_join(f,plate_layout,by="well_name") %>%
  unnest(feature_values) %>%
  mutate(unique_id = paste(well_name, label, sep = "_"))

```

### Check plate layout

```{r}

source("/srv/scratch/z3532965/src/blana/Scott/plot_plate_layout.R")

plate_layout %>% group_by(CELL,DRUG_CONCENTRATION,TIME_MINUTES,PRIMARY_1,SECONDARY_1,PRIMARY_2,SECONDARY_2) %>%
  count()

plot_plate_layout(plate_layout, color_by = "CELL", title = "Cell type")
plot_plate_layout(plate_layout, color_by = "DRUG", title = "Drug")
plot_plate_layout(plate_layout, color_by = "DRUG_CONCENTRATION", title = "Drug (µM)", log10_scale = TRUE)
plot_plate_layout(plate_layout, color_by = "TIME_MINUTES", title = "Time (min)")
plot_plate_layout(plate_layout, color_by = "PRIMARY_1", title = "Primary antibody 1")
plot_plate_layout(plate_layout, color_by = "SECONDARY_1", title = "Secondary antibody 1")
plot_plate_layout(plate_layout, color_by = "PRIMARY_2", title = "Primary antibody 2")
plot_plate_layout(plate_layout, color_by = "SECONDARY_2", title = "Secondary antibody 2")

```
```{r}
dat <- dat %>%
  rename_with(~ str_replace(., "Channel:0:0_", "Nuclei_")) %>%
  mutate(
    well_row = str_extract(well_name, "^[A-Z]"),
    well_col = as.integer(str_extract(well_name, "[0-9]+$"))
  ) %>%
  filter(Nuclei_is_border==F)

names(dat)

```
### QC

```{r,fig.height=8,fig.width=8}

bkgrd_primary_1 <- filter(dat,PRIMARY_1=="No primary") %>%
  group_by(well_name) %>%
  summarise(mean_bkgrd_561 = mean(Nuclei_intensity_mean_561)) %>%
  summarise(mean_bkgrd_561 = mean(mean_bkgrd_561)) %>%
  pull(mean_bkgrd_561)

dat %>% 
  filter(DRUG=="Vehicle") %>%
  group_by(well_name,CELL,DRUG_CONCENTRATION,TIME_MINUTES,PRIMARY_1,SECONDARY_1,PRIMARY_2,SECONDARY_2) %>%
  summarise(across(Nuclei_intensity_mean_561,mean)) %>%
  ggplot(aes(x=PRIMARY_1,y=Nuclei_intensity_mean_561,col=CELL)) +
  geom_boxplot() +
  geom_hline(yintercept=bkgrd_primary_1)

bkgrd_primary_2 <- filter(dat,PRIMARY_2=="No primary") %>%
  group_by(well_name) %>%
  summarise(mean_bkgrd_647 = mean(Nuclei_intensity_mean_647)) %>%
  summarise(mean_bkgrd_647 = mean(mean_bkgrd_647)) %>%
  pull(mean_bkgrd_647)

dat %>% 
  filter(DRUG=="Vehicle") %>%
  group_by(well_name,CELL,DRUG_CONCENTRATION,TIME_MINUTES,PRIMARY_1,SECONDARY_1,PRIMARY_2,SECONDARY_2) %>%
  summarise(across(Nuclei_intensity_mean_647,mean)) %>%
  ggplot(aes(x=PRIMARY_2,y=Nuclei_intensity_mean_647,col=CELL)) +
  geom_boxplot() +
  geom_hline(yintercept=bkgrd_primary_2)

bkgrd_gfp <- filter(dat,CELL=="HCT116 OsTIRF74G") %>%
  group_by(well_name) %>%
  summarise(mean_bkgrd_488 = mean(Nuclei_intensity_mean_488)) %>%
  summarise(mean_bkgrd_488 = mean(mean_bkgrd_488)) %>%
  pull(mean_bkgrd_488)

dat %>% 
  #filter(DRUG=="Vehicle") %>%
  group_by(well_name,CELL,DRUG,DRUG_CONCENTRATION,TIME_MINUTES,PRIMARY_1,SECONDARY_1,PRIMARY_2,SECONDARY_2) %>%
  summarise(across(Nuclei_intensity_mean_488,mean)) %>%
  ggplot(aes(x=CELL,y=Nuclei_intensity_mean_488,col=interaction(CELL,DRUG,DRUG_CONCENTRATION))) +
  geom_boxplot() +
  geom_jitter(position = position_jitterdodge()) +
  geom_hline(yintercept=bkgrd_gfp) +
  theme(axis.text.x = element_text(angle=90,hjust=1,vjust=0.5))

```
Replicate-to-replicate variability due to row/plate effects. Focus on replicate wells in upper half of plate.

```{r}
dat %>%
  filter(CELL=="HCT116 OsTIRF74G mAC-POLR2A Clone 5" & PRIMARY_1=="Total Pol II (F12)") %>%
  ggplot(aes(x=Nuclei_intensity_mean_561,col=well_row,group=well_name)) +
  geom_density() +
  facet_wrap(~interaction(DRUG,DRUG_CONCENTRATION))
```

# Quantitative agreement between mClover, F12 and D8L4Y

Each channel is an affine function of true abundance $T$:

$$\text{signal}_c = \text{bg}_c + a_c T + n_c$$

`bg_c` is what the no-primary (561, 647) and parental-untagged (488) controls
measure and what we subtract. `n_c` is any signal the antibody still produces at
zero POLR2A, e.g. non-specific binding of the primary, which a no-primary control
does not remove. After subtracting `bg_c` and normalising to the vehicle
mean, writing $\nu_c = n_c / a_c$ as a fraction of the vehicle-specific signal:

$$y_c = \frac{T + \nu_c}{1 + \nu_c} = \underbrace{\frac{1}{1+\nu_c}}_{\text{slope}} T + \underbrace{\frac{\nu_c}{1+\nu_c}}_{\text{intercept}}$$

so slope + intercept = 1, and normalisation forces the line through (1,1).
Fitting one channel against another, the slope $\beta$ is then exactly the
conversion factor between reported percentage reductions:

$$R_y = \beta \, R_x$$

## Well-level table, background subtracted *before* normalisation

Subtract the plate-wide background first, then normalise. 

```{r agreement_welltable}
CH <- c("mC", "F12", "D8L4Y")

wt <- dat %>%
  filter(well_row %in% c("D","E","F","G","H"),
         CELL == "HCT116 OsTIRF74G mAC-POLR2A Clone 5",
         PRIMARY_1 == "Total Pol II (F12)") %>%
  group_by(well_name, well_row, DRUG, DRUG_CONCENTRATION, TIME_MINUTES) %>%
  summarise(F12     = mean(Nuclei_intensity_mean_561) - bkgrd_primary_1,
            D8L4Y   = mean(Nuclei_intensity_mean_647) - bkgrd_primary_2,
            mC      = mean(Nuclei_intensity_mean_488) - bkgrd_gfp,
            n_cells = n(), .groups = "drop")

# normalise each channel by its own vehicle mean
normalisers <- wt %>% filter(DRUG == "Vehicle") %>%
  summarise(across(all_of(CH), mean))

wt <- wt %>%
  mutate(across(all_of(CH), ~ .x / normalisers[[cur_column()]])) %>%
  mutate(cond = interaction(DRUG, DRUG_CONCENTRATION, TIME_MINUTES, drop = TRUE)) %>%
  as.data.frame()

cat("wells:", nrow(wt), " conditions:", nlevels(wt$cond),
    " wells/condition:", paste(range(table(wt$cond)), collapse = "-"), "\n")
cat("normalisers (raw units, bg-subtracted):\n"); print(round(unlist(normalisers), 1))
cat("\nrange of normalised signal per channel:\n")
print(round(sapply(wt[CH], range), 3))
```

## Estimators

Both channels equal 1 in vehicle by construction, so the calibration line is
anchored at (1,1). The conversion factor is therefore estimated directly as the
least-squares slope of the fractional reduction in one channel on the fractional
reduction in the other, forced through the origin. The intercept it implies is
exactly $1-\beta$ — which is what makes $R_y = \beta R_x$ hold.

Measurement error is estimated from replicate wells, but only as a
sensitivity check (the correction is Warton *et al.*, 2006, *Biol. Rev.* **81**,
259–291, eqn 6). The unconstrained two-parameter fit is also reported, as the
test of whether the (1,1) anchor is justified: slope + intercept should be 1.

```{r agreement_helpers}
err_cov_wells <- function(wt, chs, cond) {
  cond <- as.factor(cond)
  resid <- do.call(cbind, lapply(chs, function(ch) wt[[ch]] - ave(wt[[ch]], cond)))
  colnames(resid) <- chs
  df <- nrow(wt) - nlevels(droplevels(cond))
  if (df < 1) return(matrix(0, length(chs), length(chs), dimnames = list(chs, chs)))
  crossprod(resid) / df
}

# free 2-parameter fit, corrected for measurement error
slope_free <- function(wt, xch, ych, Sig) {
  b <- (cov(wt[[xch]], wt[[ych]]) - Sig[xch, ych]) /
       (var(wt[[xch]])            - Sig[xch, xch])
  c(slope = b, intercept = mean(wt[[ych]]) - b * mean(wt[[xch]]))
}

# HEADLINE ESTIMATOR. beta = least-squares slope of fractional reduction on
# fractional reduction, through the origin. Equivalent to fitting
# y = beta*x + (1 - beta), so that R_y = beta * R_x exactly.
beta_reduction <- function(wt, xch, ych) {
  u <- 1 - wt[[xch]]; v <- 1 - wt[[ych]]
  unname(sum(u * v) / sum(u * u))
}

# the same estimator with a measurement-error correction from replicate wells.
# Sensitivity check only; not the reported value.
beta_reduction_mom <- function(wt, xch, ych, Sig) {
  u <- 1 - wt[[xch]]; v <- 1 - wt[[ych]]; n <- nrow(wt)
  unname((sum(u * v) - n * Sig[xch, ych]) / (sum(u * u) - n * Sig[xch, xch]))
}

slope_ols <- function(wt, xch, ych) unname(coef(lm(wt[[ych]] ~ wt[[xch]]))[2])

slope_deming <- function(wt, xch, ych, Sig) {
  x <- wt[[xch]]; y <- wt[[ych]]
  lambda <- Sig[ych, ych] / Sig[xch, xch]
  Sxx <- var(x); Syy <- var(y); Sxy <- cov(x, y)
  unname(((Syy - lambda * Sxx) +
          sqrt((Syy - lambda * Sxx)^2 + 4 * lambda * Sxy^2)) / (2 * Sxy))
}

# bootstrap wells, stratified by condition, re-estimating the error covariance
boot_pair <- function(wt, xch, ych, chs, cond, B = 4000) {
  cond <- as.factor(cond); idx <- split(seq_len(nrow(wt)), cond)
  out <- matrix(NA_real_, B, 3,
                dimnames = list(NULL, c("free_slope", "free_int", "beta11")))
  for (b in seq_len(B)) {
    s <- unlist(lapply(idx, function(i) sample(i, length(i), TRUE)))
    wb <- wt[s, , drop = FALSE]; cb <- cond[s]
    Sb <- err_cov_wells(wb, chs, cb)
    if (var(wb[[xch]]) - Sb[xch, xch] <= 0) next
    f <- slope_free(wb, xch, ych, Sb)
    out[b, ] <- c(f[["slope"]], f[["intercept"]], beta_reduction(wb, xch, ych))
  }
  out[complete.cases(out), , drop = FALSE]
}

bland_altman <- function(wt, xch, ych) {
  m <- (wt[[xch]] + wt[[ych]]) / 2; d <- wt[[ych]] - wt[[xch]]
  list(avg = m, diff = d, bias = mean(d),
       loa = mean(d) + c(-1.96, 1.96) * sd(d),
       prop_p = summary(lm(d ~ m))$coefficients[2, 4])
}
```

## Slopes and conversion factors

```{r agreement_slopes}
Sigma <- err_cov_wells(wt, CH, wt$cond)

cat("within-condition error CV per channel:\n")
print(round(100 * sqrt(diag(Sigma)) / 1, 1))
cat("\nerror correlation between channels (shared well-level nuisance):\n")
print(round(cov2cor(Sigma), 2))

PAIRS <- list(c("mC","F12"), c("mC","D8L4Y"), c("F12","D8L4Y"))
res <- lapply(PAIRS, function(p) {
  bs <- boot_pair(wt, p[1], p[2], CH, wt$cond)
  f  <- slope_free(wt, p[1], p[2], Sigma)
  list(pair = p, boot = bs,
       ols = slope_ols(wt, p[1], p[2]),
       deming = slope_deming(wt, p[1], p[2], Sigma),
       free = f,
       beta11 = beta_reduction(wt, p[1], p[2]),
       ci11 = quantile(bs[, "beta11"], c(.025, .975)),
       ci_int = quantile(bs[, "free_int"], c(.025, .975)))
})
names(res) <- sapply(PAIRS, paste, collapse = "_vs_")

for (r in res) {
  cat(sprintf("\n%s (x) vs %s (y)\n", r$pair[1], r$pair[2]))
  cat(sprintf("  naive OLS slope        %.3f   (what geom_smooth draws)\n", r$ols))
  cat(sprintf("  Deming slope           %.3f\n", r$deming))
  cat(sprintf("  moment-corrected       slope %.3f  intercept %+.3f [%+.3f, %+.3f]\n",
              r$free[["slope"]], r$free[["intercept"]], r$ci_int[1], r$ci_int[2]))
  cat(sprintf("  slope + intercept      %.3f   (1.00 if the affine model holds)\n",
              r$free[["slope"]] + r$free[["intercept"]]))
  b_mom <- beta_reduction_mom(wt, r$pair[1], r$pair[2], Sigma)
  cat(sprintf("  CONVERSION FACTOR beta %.3f [%.3f, %.3f]  (reduction on reduction, through origin)\n",
              r$beta11, r$ci11[1], r$ci11[2]))
  cat(sprintf("    sensitivity: measurement-error corrected %.3f (%+.2f%%)\n",
              b_mom, 100 * (b_mom / r$beta11 - 1)))
  if (r$pair[1] == "mC") {
    nu <- 1 / r$beta11 - 1
    cat(sprintf("  implied nu (residual non-specific, fraction of vehicle) %.3f\n", nu))
  }
}
```

## Does it matter which measurement you use?

`beta` converts directly: a reduction of `R` by channel x reads as `beta * R` by
channel y. Pre-specify a tolerance and check whether the interval sits inside it,
rather than reading a non-significant difference as agreement.

```{r agreement_answer}

for (r in res) {
  cat(sprintf("A 50%% reduction measured by %s reads as %.1f%% [%.1f, %.1f] by %s.\n",
              r$pair[1], 50 * r$beta11, 50 * r$ci11[1], 50 * r$ci11[2], r$pair[2]))
}
cat(sprintf("\nBased on %d wells across %d dose x time conditions.\n",
            nrow(wt), nlevels(wt$cond)))
```

## Bland-Altman

Standard method-comparison view: mean of the two measurements against their
difference, with 95% limits of agreement. A slope in this plot is proportional
bias, which is what a non-specific offset produces.

```{r agreement_bland_altman, fig.height=3, fig.width=9}
ba_df <- do.call(rbind, lapply(PAIRS, function(p) {
  ba <- bland_altman(wt, p[1], p[2])
  data.frame(pair = paste(p[2], "-", p[1]), avg = ba$avg, diff = ba$diff,
             bias = ba$bias, lo = ba$loa[1], hi = ba$loa[2], prop_p = ba$prop_p)
}))

for (p in PAIRS) {
  ba <- bland_altman(wt, p[1], p[2])
  cat(sprintf("%6s - %-6s : bias %+.3f  LoA [%+.3f, %+.3f]  proportional-bias p = %.3g\n",
              p[2], p[1], ba$bias, ba$loa[1], ba$loa[2], ba$prop_p))
}

ggplot(ba_df, aes(avg, diff)) +
  geom_hline(yintercept = 0, colour = "grey70") +
  geom_hline(aes(yintercept = bias), linetype = "solid", linewidth = 0.3) +
  geom_hline(aes(yintercept = lo), linetype = "21", linewidth = 0.3) +
  geom_hline(aes(yintercept = hi), linetype = "21", linewidth = 0.3) +
  geom_point(size = 1.2) +
  facet_wrap(~pair) +
  labs(x = "mean of the two measurements", y = "difference") +
  theme_bw(base_size = 8) + theme(panel.grid = element_blank())
```

## Corrected panels


```{r agreement_panels, fig.height=8, fig.width=10}
cond_summary <- wt %>%
  group_by(DRUG, DRUG_CONCENTRATION, TIME_MINUTES) %>%
  summarise(across(all_of(CH), list(mean = mean, sd = sd)), .groups = "drop") %>%
  mutate(conc_plot = if_else(DRUG_CONCENTRATION == 0, 1, DRUG_CONCENTRATION))

panel_df <- function(xv, yv) {
  data.frame(x    = cond_summary[[paste0(xv, "_mean")]],
             x_sd = cond_summary[[paste0(xv, "_sd")]],
             y    = cond_summary[[paste0(yv, "_mean")]],
             y_sd = cond_summary[[paste0(yv, "_sd")]],
             conc = cond_summary$conc_plot)
}

mk_panel <- function(xv, yv, xlab, ylab, r) {
  d <- panel_df(xv, yv)
  # normalisation puts both channels at 1 in vehicle, so the line is anchored
  # at (1,1): y = beta*x + (1 - beta). Drawn line and quoted values are the same
  # fit, and alpha = 1 - beta by construction rather than fitted freely 
  # The unconstrained fit and its intercept CI are in the agreement_slopes output.
  b <- r$beta11
  a <- 1 - b
  lab <- sprintf("beta==%.3f*','~alpha==%.3f", b, a)
  ggplot(d, aes(x = x, y = y, col = log10(conc))) +
    geom_abline(slope = b, intercept = a,
                col = "black", linewidth = 0.5, linetype = "21") +
    geom_errorbar(aes(ymin = y - y_sd, ymax = y + y_sd)) +
    geom_errorbarh(aes(xmin = x - x_sd, xmax = x + x_sd)) +
    geom_point(size = 1.2) +
    coord_cartesian(xlim = c(0, NA), ylim = c(0, NA)) +
    annotate("label", x = 0.01, y = 1.05, label = lab, parse = TRUE, hjust = 0,
             size = 2.5, fill = "white", colour = "black", label.size = 0.3) +
    scale_color_viridis_c(
      name = "5-Ph-IAA\nconcentration",
      begin = 0, end = 0.8,
      breaks = c(0, 1, 2, 3, 4),
      labels = c("Vehicle", "10 nM", "100 nM", "1 \u00b5M", "10 \u00b5M")) +
    scale_x_continuous(name = xlab, limits = c(0, 1.1), breaks = c(0, 0.5, 1.0)) +
    scale_y_continuous(name = ylab, limits = c(0, 1.1), breaks = c(0, 0.5, 1.0))
}

f12_vs_gfp_plot   <- mk_panel("mC",  "F12",   "POLR2A (mClover)", "POLR2A IF (F12)",
                              res$mC_vs_F12)
d8l4y_vs_gfp_plot <- mk_panel("mC",  "D8L4Y", "POLR2A (mClover)", "POLR2A IF (D8L4Y)",
                              res$mC_vs_D8L4Y)
d8l4y_vs_f12_plot <- mk_panel("F12", "D8L4Y", "POLR2A IF (F12)",  "POLR2A IF (D8L4Y)",
                              res$F12_vs_D8L4Y)

design <-
"
A#
BC
"

f12_vs_gfp_plot + d8l4y_vs_gfp_plot + d8l4y_vs_f12_plot + plot_layout(design = design) &
  theme_bw(base_size = 8) &
  theme(legend.position = "none",
        panel.grid = element_blank(),
        axis.text=element_text(size=8),
        axis.title=element_text(size=8),
        legend.title=element_text(size=8),
        legend.text=element_text(size=8),
        panel.border = element_rect(linewidth = 1))

ggsave("f12_vs_gfp_vs_d8l4y.pdf",width=10,height=8,units="cm")

l <- cowplot::get_legend(f12_vs_gfp_plot + theme(legend.key.size = unit(0.4,"cm"),legend.title=element_text(size=8),
        legend.text=element_text(size=8)))
ggsave("f12_vs_gfp_vs_d8l4y_legend.pdf", plot=l, width = 5, height = 5,units="cm")
```

# F12 as the working readout

Most experiments use F12, so another useful direction is F12 &rarr; the others. 

```{r f12_anchored}
res_f12 <- lapply(list(c("F12","mC"), c("F12","D8L4Y")), function(p) {
  bs <- boot_pair(wt, p[1], p[2], CH, wt$cond)
  f  <- slope_free(wt, p[1], p[2], Sigma)
  list(pair = p, boot = bs, free = f,
       beta11 = beta_reduction(wt, p[1], p[2]),
       ci11 = quantile(bs[, "beta11"], c(.025, .975)))
})
names(res_f12) <- c("F12_to_mC", "F12_to_D8L4Y")

cat("A reduction measured by F12 corresponds to:\n\n")
cat(sprintf("%-12s %-22s %s\n", "R_F12", "by mClover", "by D8L4Y"))
for (R in c(10, 20, 30, 40, 50, 60, 75, 90)) {
  a <- res_f12$F12_to_mC; b <- res_f12$F12_to_D8L4Y
  cat(sprintf("%-12s %5.1f%% [%.1f, %.1f]     %5.1f%% [%.1f, %.1f]\n",
              paste0(R, "%"),
              R * a$beta11, R * a$ci11[1], R * a$ci11[2],
              R * b$beta11, R * b$ci11[1], R * b$ci11[2]))
}
for (r in res_f12)
  cat(sprintf("\nF12 -> %-6s : beta = %.3f [%.3f, %.3f]", r$pair[2],
              r$beta11, r$ci11[1], r$ci11[2]))
cat("\n")
```

## Panels with F12 on the x-axis

Same data and fits, reoriented so the working readout is the predictor.

```{r agreement_panels_f12x, fig.height=8, fig.width=10}
f12_vs_gfp_plot_f12x   <- mk_panel("F12", "mC",    "POLR2A IF (F12)", "POLR2A (mClover)",
                                   res_f12$F12_to_mC)
d8l4y_vs_f12_plot_f12x <- mk_panel("F12", "D8L4Y", "POLR2A IF (F12)", "POLR2A IF (D8L4Y)",
                                   res_f12$F12_to_D8L4Y)
d8l4y_vs_gfp_plot_f12x <- mk_panel("mC",  "D8L4Y", "POLR2A (mClover)", "POLR2A IF (D8L4Y)",
                                   res$mC_vs_D8L4Y)

f12_vs_gfp_plot_f12x + d8l4y_vs_f12_plot_f12x + d8l4y_vs_gfp_plot_f12x +
  plot_layout(design = design) &
  theme_bw(base_size = 8) &
  theme(legend.position = "none",
        panel.grid = element_blank(),
        axis.text=element_text(size=8),
        axis.title=element_text(size=8),
        legend.title=element_text(size=8),
        legend.text=element_text(size=8),
        panel.border = element_rect(linewidth = 1))

ggsave("f12_vs_gfp_vs_d8l4y_F12x.pdf",width=10,height=8,units="cm")
```

# Compare WB and IF

Matching the 4 h data with the western blots performed at this timepoint. The IF
side reuses `cond_summary` from the agreement analysis, so it is already
background-subtracted; it is re-normalised within the 4 h subset to match the WB
normalisation to its own DMSO control.

```{r wb_to_IF}
wb_data_summary <- wb_data %>%
  group_by(concentration, concentration_numeric) %>%
  summarise(across(c(total_D8L4Y_normalised, pSer2_3E10_normalised),
                   c(mean = ~mean(.), sd = ~sd(.))),
            WB_n = n(), .groups = 'drop') %>%
  mutate(DRUG = if_else(concentration == "DMSO", "Vehicle", "5PhIAA"),
         # DMSO is coded 1e-9 so it plots on a log axis; the IF side keeps the
         # true zero, so force both to 0 or the vehicle row will not join
         DRUG_CONCENTRATION = if_else(DRUG == "Vehicle", 0,
                                      concentration_numeric * 10^9)) %>%
  select(DRUG, DRUG_CONCENTRATION, WB_n,
         D8L4Y_WB_mean = total_D8L4Y_normalised_mean,
         D8L4Y_WB_sd   = total_D8L4Y_normalised_sd)

if_n <- wt %>%
  filter(TIME_MINUTES == 240) %>%
  count(DRUG, DRUG_CONCENTRATION, name = "IF_n")

if_data_summary <- cond_summary %>%
  filter(TIME_MINUTES == 240) %>%
  select(DRUG, DRUG_CONCENTRATION,
         D8L4Y_IF_mean = D8L4Y_mean, D8L4Y_IF_sd = D8L4Y_sd,
         F12_IF_mean   = F12_mean,   F12_IF_sd   = F12_sd)

if_norm <- if_data_summary %>%
  filter(DRUG == "Vehicle") %>%
  summarise(D8L4Y_IF_mean_normaliser = mean(D8L4Y_IF_mean),
            F12_IF_mean_normaliser   = mean(F12_IF_mean)) %>%
  bind_cols(if_data_summary) %>%
  mutate(D8L4Y_IF_mean = D8L4Y_IF_mean / D8L4Y_IF_mean_normaliser,
         D8L4Y_IF_sd   = D8L4Y_IF_sd   / D8L4Y_IF_mean_normaliser,
         F12_IF_mean   = F12_IF_mean   / F12_IF_mean_normaliser,
         F12_IF_sd     = F12_IF_sd     / F12_IF_mean_normaliser) %>%
  select(-D8L4Y_IF_mean_normaliser, -F12_IF_mean_normaliser) %>%
  pivot_longer(cols = c(D8L4Y_IF_mean, D8L4Y_IF_sd, F12_IF_mean, F12_IF_sd),
               names_to = c("antibody", ".value"),
               names_pattern = "^(D8L4Y|F12)_(.+)$")

plot_data <- left_join(if_norm, wb_data_summary,
                       by = c("DRUG", "DRUG_CONCENTRATION")) %>%
  left_join(if_n, by = c("DRUG", "DRUG_CONCENTRATION"))

# join check: every IF condition should have found a WB partner
cat("rows:", nrow(plot_data),
    " unmatched WB:", sum(is.na(plot_data$D8L4Y_WB_mean)), "\n")
print(plot_data %>%
        select(antibody, DRUG, DRUG_CONCENTRATION, IF_mean, D8L4Y_WB_mean) %>%
        arrange(antibody, DRUG_CONCENTRATION))
```

Both axes are normalised to their own vehicle, so the same anchored estimator
applies: `beta_reduction` on the IF and WB fractional reductions. The
unconstrained fit is reported alongside it, using a diagonal error covariance
built from the standard error of each condition mean — here the two sets of
errors really are independent, since IF and WB are separate assays.

```{r wb_slopes}
fit_wb <- function(d) {
  d <- d[complete.cases(d[c("IF_mean", "D8L4Y_WB_mean")]), ]
  Sig <- diag(c(mean(d$IF_sd^2 / d$IF_n, na.rm = TRUE),
                mean(d$D8L4Y_WB_sd^2 / d$WB_n, na.rm = TRUE)))
  dimnames(Sig) <- list(c("IF", "WB"), c("IF", "WB"))
  dd <- data.frame(IF = d$IF_mean, WB = d$D8L4Y_WB_mean)
  f <- slope_free(dd, "IF", "WB", Sig)
  data.frame(n = nrow(dd),
             ols       = slope_ols(dd, "IF", "WB"),
             slope     = f[["slope"]],
             intercept = f[["intercept"]],
             beta11    = beta_reduction(dd, "IF", "WB"))
}

wb_slopes <- plot_data %>%
  group_by(antibody) %>%
  group_modify(~ fit_wb(.x)) %>%
  ungroup()

print(as.data.frame(wb_slopes), digits = 3)
cat("\nbeta11 is the conversion factor: a reduction of R by IF reads as beta*R by WB.\n")
for (i in seq_len(nrow(wb_slopes)))
  cat(sprintf("  50%% reduction by %-6s IF -> %.1f%% by western blot\n",
              wb_slopes$antibody[i], 50 * wb_slopes$beta11[i]))
```
```{r wb_plot, fig.height=3, fig.width=4}
# both axes are normalised to their own vehicle, so the line is anchored at
# (1,1): y = beta*x + (1 - beta). This is the line beta refers to.
slope_labels <- wb_slopes %>%
  mutate(intercept11 = 1 - beta11,
         label = sprintf("beta==%.3f*','~alpha==%.3f", beta11, intercept11),
         x = 0.05,
         y = ifelse(antibody == "D8L4Y", 1.05, 0.95))

plot_data %>%
  ggplot(aes(x = IF_mean, y = D8L4Y_WB_mean, colour = antibody)) +
  geom_abline(data = slope_labels,
              aes(slope = beta11, intercept = intercept11, colour = antibody),
              linewidth = 0.75, linetype = "21", show.legend = FALSE) +
  geom_errorbarh(aes(xmin = IF_mean - IF_sd, xmax = IF_mean + IF_sd),
                 show.legend = FALSE, linewidth = 0.4, width = 0.05) +
  geom_errorbar(aes(ymin = D8L4Y_WB_mean - D8L4Y_WB_sd,
                    ymax = D8L4Y_WB_mean + D8L4Y_WB_sd),
                show.legend = FALSE, linewidth = 0.4, width = 0.05) +
  geom_point(size = 1.5) +
  geom_text(data = slope_labels,
            aes(x = x, y = y, label = label, colour = antibody),
            hjust = 0, size = 2.5, inherit.aes = F, parse = TRUE, show.legend = FALSE) +
  scale_x_continuous(name = "POLR2A immunofluorescence",
                     limits = c(0, 1.1), breaks = c(0, 0.5, 1.0)) +
  scale_y_continuous(name = "POLR2A western\nblot (D8L4Y)",
                     limits = c(0, 1.1), breaks = c(0, 0.5, 1.0)) +
  theme_bw(base_size = 8) +
  theme(legend.position = "right",
        legend.title = element_blank(),
        panel.grid = element_blank(),
        aspect.ratio = 1,
        axis.text=element_text(size=8),
        axis.title=element_text(size=8),
        legend.text=element_text(size=8),
        panel.border = element_rect(linewidth = 1)) +
  scale_colour_manual(values = c("#0072B2", "#D55E00"))
ggsave("if_vs_wb_d8l4y.pdf",width=7,height=5,units="cm")
```



# Consistency check


```{r consistency_check}
# ---- 1. does each plotted line reproduce its own quoted beta? -------------
# every panel draws y = beta*x + (1 - beta), so at x = 0.5 the implied
# reduction must be exactly beta * 50%.
chk <- function(tag, b) {
  a <- 1 - b
  data.frame(panel = tag, beta = b, alpha = a,
             slope_plus_intercept = b + a,
             pct_from_drawn_line = 100 * (1 - (b * 0.5 + a)),
             pct_from_quoted_beta = 50 * b)
}
panel_chk <- rbind(
  chk("mC -> F12          (fig 1, A)", res$mC_vs_F12$beta11),
  chk("mC -> D8L4Y        (fig 1, B)", res$mC_vs_D8L4Y$beta11),
  chk("F12 -> D8L4Y       (fig 1, C)", res$F12_vs_D8L4Y$beta11),
  chk("F12 -> mC          (fig 2, A)", res_f12$F12_to_mC$beta11),
  chk("F12 -> D8L4Y       (fig 2, B)", res_f12$F12_to_D8L4Y$beta11))
print(format(panel_chk, digits = 4), row.names = FALSE)
stopifnot(all(abs(panel_chk$slope_plus_intercept - 1) < 1e-10),
          all(abs(panel_chk$pct_from_drawn_line -
                  panel_chk$pct_from_quoted_beta) < 1e-8))
cat("OK: every plotted line reproduces its quoted beta exactly.\n")

# ---- 2. what would the other estimators have said? ------------------------
# a general (slope, intercept) pair implies R_y = 100*(1 - a - b) + b*R_x,
# which collapses to b*R_x only when a + b = 1.
red_at   <- function(b, a, Rx = 50) 100 * (1 - a - b) + b * Rx
int_from <- function(b, xch, ych) mean(wt[[ych]]) - b * mean(wt[[xch]])

cat("\n--- reduction implied when channel x shows a 50% reduction ---\n")
for (r in res) {
  x <- r$pair[1]; y <- r$pair[2]
  b_mom <- beta_reduction_mom(wt, x, y, Sigma)
  rows <- c(
    "reduction-OLS (reported)" = red_at(r$beta11, 1 - r$beta11),
    "+ measurement-error corr" = red_at(b_mom, 1 - b_mom),
    "unconstrained affine fit" = red_at(r$free[["slope"]], r$free[["intercept"]]),
    "raw OLS"                  = red_at(r$ols, int_from(r$ols, x, y)),
    "Deming"                   = red_at(r$deming, int_from(r$deming, x, y)))
  cat(sprintf("\n%s -> %s\n", x, y))
  for (nm in names(rows)) cat(sprintf("  %-26s %5.1f%%\n", nm, rows[[nm]]))
  cat(sprintf("  spread across estimators: %.1f pp   (bootstrap CI half-width: %.1f pp)\n",
              diff(range(rows)), 50 * (r$ci11[2] - r$ci11[1]) / 2))
}
cat("\nIf the spread across estimators is small relative to the bootstrap interval,\n",
    "the choice of estimator is not what limits the conclusion.\n")

# ---- 3. internal consistency of the three pairwise betas ------------------
cat("\n--- internal consistency ---\n")
cat(sprintf("multiplicative closure : beta(mC->F12) x beta(F12->D8L4Y) = %.4f  vs  beta(mC->D8L4Y) = %.4f\n",
            res$mC_vs_F12$beta11 * res$F12_vs_D8L4Y$beta11,
            res$mC_vs_D8L4Y$beta11))
cat(sprintf("reciprocity            : beta(F12->mC) = %.4f  vs  1/beta(mC->F12) = %.4f\n",
            res_f12$F12_to_mC$beta11, 1 / res$mC_vs_F12$beta11))
cat("  (these two are NOT expected to match exactly: for least-squares through the\n",
    "  origin, beta(x->y) * beta(y->x) equals the squared correlation, so the\n",
    "  reciprocal is the larger of the two by a factor of about 1/r^2.)\n")
cat(sprintf("  implied r^2 = %.4f\n",
            res_f12$F12_to_mC$beta11 * res$mC_vs_F12$beta11))

# ---- 4. the same check for the western blot panel -------------------------
cat("\n--- western blot panel ---\n")
for (i in seq_len(nrow(wb_slopes))) {
  b <- wb_slopes$beta11[i]; a <- 1 - b
  cat(sprintf("%-6s IF -> WB : drawn line at x=0.5 gives %.1f%%, quoted beta gives %.1f%%",
              wb_slopes$antibody[i], 100 * (1 - (b * 0.5 + a)), 50 * b))
  cat(sprintf("   | unconstrained fit would give %.1f%%\n",
              red_at(wb_slopes$slope[i], wb_slopes$intercept[i])))
}
```
