---
title: "EU in ARMC5 KO cells"
author: "Scott Berry"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup_r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

rm(list=ls())
library(tidyverse)
library(patchwork)
library(lme4)
library(emmeans)

wt_col <- "#184cee"
armc5_col <- "#9d24ec"
wt_trp_col <- "#0095ff"
armc5_trp_col <- "#d682ff"

cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
```

# Experiment

# Data handling

Original data (Nikon `.nd2` files) are stored on UNSW's data archive under Berry Lab project D0419427. Data were maximum-intensity projected and converted to OME-TIFF using the `blimp` tools.

# Analysis

## Image analysis 

Image analysis performed using the following series of python scripts (found on github for [hek_inhibitors_EU](https://github.com/berrygroup/publications/tree/main/2024_ARMC5/hek_inhibitors_EU))

- `illumination_correction_EU.ipynb`
- `save_example_images_EU.ipynb` (used to generate example images and training data for the cellpose model)
- `run_quantification_EU.py` (parallelised using e.g. `PBS/batch_quantify_EU_240304.pbs`)

## Data analysis

Quantitative data analysis performed in this document.

### Load data

Load data from `scratch/berrylab` on UNSW's Katana HPC.

```{r load_data}

ana_dir <- "/srv/scratch/z3532965/src/blana/Alex/202402_ARMC5_Fixed"
data_dir <- "/srv/scratch/berrylab/z3532965/NikonSpinningDisk"
acquisition_dirs <- purrr::map_chr(
  c("240229/240226_ARMC5KO_EU/20240229_172045_717",
    "240304/240301_ARMC5KO_EU/20240304_180009_911",
    "240307/240304_ARMC5KO_EU/20240307_184329_444"),~file.path(data_dir,.))
intensity_image_dirs <- map_chr(acquisition_dirs,~file.path(.,"OME-TIFF"))
intensity_image_mip_dirs <- map_chr(acquisition_dirs,~file.path(.,"OME-TIFF-MIP"))
label_image_dirs <- map_chr(acquisition_dirs,~file.path(.,"SEGMENTATION"))
quantification_dirs <- map_chr(acquisition_dirs,~file.path(.,"QUANTIFICATION"))

read_quantification <- function(dir_name) {
  csv_files <- Sys.glob(file.path(dir_name,"*.csv"))
  dt_list <- lapply(csv_files, data.table::fread)
  df <- tibble(path = csv_files,
               feature_values = dt_list) %>%
    mutate(dir_name = dir_name,
           file_name = basename(path),
           feature_values = purrr::map(feature_values,tibble)) %>%
    separate(file_name,into = c("well_name","ch1","ch2","ch3","ch4","sample","id"),sep="[,_.]", extra = "drop") %>%
    mutate(well_name = str_replace(well_name, "^Well", ""),
           field_id = as.integer(id))
  return(df)
}

pad_well_names <- function(names) {
  gsub("(\\D)(\\d)$", "\\10\\2", names)
}

# features
f <- lapply(quantification_dirs,read_quantification) %>%
  bind_rows() %>%
  mutate(acquisition_dir = dirname(dirname(path)),
         imaging_date = basename(acquisition_dir) %>% str_extract("\\d{8}(?=_)")) %>%
  select(-c(ch1, ch2, ch3, ch4, sample))

# layouts
plate_layouts <- tibble(layout_filename=c("240226_ARMC5KO_EU_longlayout.csv","240301_ARMC5KO_EU_longlayout.csv","240304_ARMC5KO_EU_longlayout.csv"),
                        plating_date=c("240226","240301","240304"),
                        imaging_date=c("20240229","20240304","20240307"),
                        metadata=purrr::map(layout_filename,~read_csv(file.path(ana_dir,"METADATA",.),show_col_types = FALSE, name_repair = ~sub(" ", "_", .)))) %>%
  unnest(metadata) %>%
  mutate(well_name=pad_well_names(Well_name)) %>%
  select(-Well_name) %>%
  left_join(distinct(f,acquisition_dir,well_name,imaging_date),by = join_by(imaging_date, well_name))

```

Check that the data are complete.

```{r check_complete}

# check completeness (count wells)
f %>% 
  distinct(acquisition_dir,well_name) %>%
  group_by(acquisition_dir) %>%
  count()

# check completeness (count sites per well)
f %>% 
  group_by(acquisition_dir,well_name) %>%
  count() %>%
  filter(!(n %in% c(9,16)))

```

Load metadata and join with quantified image data.

```{r load metadata}
# read image metadata and join to plate metadata
m <- tibble(acquisition_dir = acquisition_dirs) %>%
  mutate(image_dir = purrr::map_chr(acquisition_dirs,~file.path(.,"OME-TIFF-MIP"))) %>%
  mutate(metadata_file = purrr::map(image_dir,~list.files(.))) %>%
  unnest(metadata_file) %>%
  filter(grepl("metadata.csv$",metadata_file)) %>%
  mutate(dat = purrr::map(file.path(image_dir,metadata_file),~read_csv(.,show_col_types = FALSE))) %>%
  separate(metadata_file,into = c("well_name","ch1","ch2","ch3","ch4","sample"),sep="[,_.]", extra = "drop") %>%
  unnest(dat) %>%
  mutate(well_name = str_replace(well_name, "^Well", ""),
           field_id = as.integer(field_id)) %>%
  select(-c(ch1, ch2, ch3, ch4, sample)) %>%
  left_join(plate_layouts,by=join_by(acquisition_dir,well_name))

# attach image and plate metadata 
dat <- right_join(m,f,by = join_by(acquisition_dir, well_name, field_id, imaging_date)) %>%
  unnest(feature_values) %>%
  select(-matches("^Nuclei_x"))

```

Some basic QC regarding number of cells per well, etc.

```{r cell_counts, fig.width=8,fig.height=8}

dat %>% 
  distinct(Cell,plating_date,imaging_date,well_name,Drug,Drug_time,EU_concentration,EU_time) %>%
  group_by(Cell,plating_date,imaging_date,Drug,Drug_time,EU_concentration,EU_time) %>%
  count() %>%
  ungroup() %>%
  ggplot(aes(x=Cell,y=n,fill=Drug)) + geom_col(position = "dodge") + facet_grid(interaction(EU_time,EU_concentration)~imaging_date) +
  coord_flip() +
  ggtitle("Number of wells of each condition")

dat %>% 
  group_by(Cell,plating_date,imaging_date,Drug,Drug_time,EU_concentration,EU_time) %>%
  count() %>%
  ungroup() %>%
  ggplot(aes(x=Cell,y=n,fill=Drug)) + geom_col(position = "dodge") + facet_grid(interaction(EU_time,EU_concentration)~imaging_date) +
  coord_flip() +
  ggtitle("Total number of cells in each condition")

```

### Subtract background

Calculate background by maximum intensity projecting, illumination-correcting, and getting the mean intensity in a region containing no cells.


```{r subtract_background}

bg_647 <- mean(c(159.894, 157.667))
bg_DAPI <-mean(c(146.643, 148.027))
bg_568 <-mean(c(125.420, 124.781))
bg_488 <-mean(c(182.254, 180.156))

dat_minus_bg <- mutate(dat,
                       Nuclei_intensity_mean_647 = Nuclei_intensity_mean_647 - bg_647,
                       Nuclei_intensity_mean_DAPI = Nuclei_intensity_mean_DAPI - bg_DAPI,
                       Nuclei_intensity_mean_568 = Nuclei_intensity_mean_568 - bg_568,
                       Nuclei_intensity_mean_488 = Nuclei_intensity_mean_488 - bg_488,
                       
                       Nuclei_intensity_median_647 = Nuclei_intensity_median_647 - bg_647,
                       Nuclei_intensity_median_DAPI = Nuclei_intensity_median_DAPI - bg_DAPI,
                       Nuclei_intensity_median_568 = Nuclei_intensity_median_568 - bg_568,
                       Nuclei_intensity_median_488 = Nuclei_intensity_median_488 - bg_488)

```

### Sum intensities

Compute sum intensities by multiplying the background-subtracted means by the corresponding area

```{r sum_intensities}

dat_minus_bg <- mutate(dat_minus_bg,
                       Nuclei_intensity_sum_647 = Nuclei_intensity_mean_647 * Nuclei_area,
                       Nuclei_intensity_sum_DAPI = Nuclei_intensity_mean_DAPI * Nuclei_area,
                       Nuclei_intensity_sum_568 = Nuclei_intensity_mean_568 * Nuclei_area,
                       Nuclei_intensity_sum_488 = Nuclei_intensity_mean_488 * Nuclei_area)

```

## Data cleanup

Remove border cells 

```{r remove_variables}

dat_non_border <- filter(dat_minus_bg,!Nuclei_is_border)

```

### Check illumination correction

After correction of illumination biases across the field of view, there should be no bias for mean intensities across the field-of-view.

```{r test_illum_corr}

dat_non_border_sample_longer <- dat_non_border %>%
  filter(imaging_date=="20240304") %>%
  filter(well_name=="F06") %>%
  select(Nuclei_centroid_0,Nuclei_centroid_1,Nuclei_intensity_mean_DAPI,Nuclei_intensity_mean_488,Nuclei_intensity_mean_647) %>%
  pivot_longer(-c(Nuclei_centroid_0,Nuclei_centroid_1)) 

rescaling_limits <- dat_non_border_sample_longer %>%
  group_by(name) %>%
  summarise(upper_limit = quantile(value,0.99))

dat_non_border_sample_longer <- left_join(dat_non_border_sample_longer,rescaling_limits,by="name")

channels <- unique(dat_non_border_sample_longer$name)

p <- list(length=length(channels))
for (i in 1:length(channels)) {
  p[[i]] <- dat_non_border_sample_longer %>%
    filter(name==channels[i]) %>%
    ggplot(aes(x=Nuclei_centroid_0,y=Nuclei_centroid_1,col=value/upper_limit)) +
    geom_point(size=0.5) +
    scale_color_viridis_c(limits=c(0,1)) +
    facet_wrap(~name) +
    coord_fixed() +
    theme(legend.position = "none")
}

wrap_plots(p) & ggtitle("Illumination biases")

```

### Check DAPI distributions

```{r check_DAPI}

dat_non_border %>%
  ggplot(aes(x=Nuclei_intensity_sum_DAPI,col=as.factor(well_name))) +
  geom_density() +
  facet_wrap(imaging_date~Cell,scales = "free") +
  theme(legend.position = "none") +
  scale_x_continuous(limits = c(0,6e6)) +
  scale_y_continuous(labels = NULL)

```

Major well-to-well variation in DAPI staining. Looks to be a row- or column-effect based on clustered wells showing similar intensities.

### Nuclear Area

```{r nuclear_area_qc}

dat_non_border %>%
  ggplot(aes(x=Nuclei_area,col=as.factor(well_name))) +
  geom_density() +
  facet_wrap(imaging_date~Cell,scales = "free") +
  theme(legend.position = "none") +
  geom_vline(xintercept = 750) +
  scale_y_continuous(labels = NULL)

dat_non_border <- filter(dat_non_border,Nuclei_area>750)

```

## Quantify stains of interest

### Check EU distributions

```{r EU_by_cell, fig.width=14,fig.height=10}

dat_non_border <- mutate(dat_non_border,Drug=factor(Drug,levels=c("None","Vehicle","Triptolide 300 nM","LDC4297 10 µM","THZ1 1 µM","AZD4573 1 µM","DRB 100 µM"))) 

p15min <- dat_non_border %>%
  filter(Drug!="None") %>% 
  filter(EU_time=="15 min") %>%
  ggplot(aes(x=interaction(Cell,EU_concentration),y=Nuclei_intensity_mean_647,col=Drug,group=interaction(well_name,Drug))) + 
  geom_boxplot(outlier.shape=NA) + facet_grid(EU_time~imaging_date) +
  coord_cartesian(ylim = c(0,600)) +
  theme_bw(base_size = 8) +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.title.y = element_blank())

p30min <- dat_non_border %>% 
  filter(Drug!="None") %>% 
  filter(EU_time=="30 min") %>%
  ggplot(aes(x=interaction(Cell,EU_concentration),y=Nuclei_intensity_mean_647,col=Drug,group=interaction(well_name,Drug))) + 
  geom_boxplot(outlier.shape=NA) + facet_grid(EU_time~imaging_date) +
  coord_cartesian(ylim = c(0,1750)) +
  theme_bw(base_size = 8) +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank())

p60min <- dat_non_border %>% 
  filter(Drug!="None") %>% 
  filter(EU_time=="60 min") %>%
  ggplot(aes(x=interaction(Cell,EU_concentration),y=Nuclei_intensity_mean_647,col=Drug,group=interaction(well_name,Drug))) + 
  geom_boxplot(outlier.shape=NA) + facet_grid(EU_time~imaging_date) +
  coord_cartesian(ylim = c(0,5000)) +
  theme_bw(base_size = 8) +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_text(angle=90,hjust=1,vjust=0.5),
        axis.title.y = element_blank())

p15min / p30min / p60min + plot_layout(guides="collect")

ggsave("PLOTS/EU_boxplot_per_well_by_drug.pdf",width=14,height=10)

```

```{r EU_by_treatment, fig.width=14,fig.height=10}

p15min <- dat_non_border %>%
  filter(Drug!="None") %>% 
  filter(EU_time=="15 min") %>%
  ggplot(aes(x=interaction(Drug,EU_concentration),y=Nuclei_intensity_mean_647,col=Cell,group=interaction(well_name,Cell))) + 
  geom_boxplot(outlier.shape=NA) + facet_grid(EU_time~imaging_date) +
  coord_cartesian(ylim = c(0,600)) +
  theme_bw(base_size = 8) +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.title.y = element_blank())

p30min <- dat_non_border %>% 
  filter(Drug!="None") %>% 
  filter(EU_time=="30 min") %>%
  ggplot(aes(x=interaction(Drug,EU_concentration),y=Nuclei_intensity_mean_647,col=Cell,group=interaction(well_name,Cell))) + 
  geom_boxplot(outlier.shape=NA) + facet_grid(EU_time~imaging_date) +
  coord_cartesian(ylim = c(0,1750)) +
  theme_bw(base_size = 8) +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank())

p60min <- dat_non_border %>% 
  filter(Drug!="None") %>% 
  filter(EU_time=="60 min") %>%
  ggplot(aes(x=interaction(Drug,EU_concentration),y=Nuclei_intensity_mean_647,col=Cell,group=interaction(well_name,Cell))) + 
  geom_boxplot(outlier.shape=NA) + facet_grid(EU_time~imaging_date) +
  coord_cartesian(ylim = c(0,5000)) +
  theme_bw(base_size = 8) +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_text(angle=90,hjust=1,vjust=0.5),
        axis.title.y = element_blank())

p15min / p30min / p60min + plot_layout(guides="collect")

ggsave("PLOTS/EU_boxplot_per_well_by_cell.pdf",width=14,height=10)

```

### Check plate layouts

Visualise plate layouts and examine the well-means across the plates for each experiment.

```{r check_plate_layouts, fig.width=9,fig.height=3}

dat_non_border %>%
  mutate(well_row = factor(str_extract(well_name, "[A-Z]"),levels=LETTERS),
         well_column = as.integer(as.character(str_extract(well_name, "[0-9]+")))) %>%
  distinct(imaging_date,well_name,well_row,well_column,Drug) %>%
  ggplot(aes(x=well_column,y=well_row,fill=Drug)) +
  geom_tile() +
  scale_y_discrete(limits=rev) +
  facet_wrap(~imaging_date) +
  ggtitle("Drug")

dat_non_border %>%
  mutate(well_row = factor(str_extract(well_name, "[A-Z]"),levels=LETTERS),
         well_column = as.integer(as.character(str_extract(well_name, "[0-9]+")))) %>%
  distinct(imaging_date,well_name,well_row,well_column,Cell) %>%
  ggplot(aes(x=well_column,y=well_row,fill=Cell)) +
  geom_tile() +
  scale_y_discrete(limits=rev) +
  facet_wrap(~imaging_date) +
  ggtitle("Cell")

dat_non_border %>%
  mutate(well_row = factor(str_extract(well_name, "[A-Z]"),levels=LETTERS),
         well_column = as.integer(as.character(str_extract(well_name, "[0-9]+")))) %>%
  distinct(imaging_date,well_name,well_row,well_column,EU_concentration) %>%
  ggplot(aes(x=well_column,y=well_row,fill=EU_concentration)) +
  geom_tile() +
  scale_y_discrete(limits=rev) +
  facet_wrap(~imaging_date) +
  ggtitle("EU concentration")

dat_non_border %>%
  mutate(well_row = factor(str_extract(well_name, "[A-Z]"),levels=LETTERS),
         well_column = as.integer(as.character(str_extract(well_name, "[0-9]+")))) %>%
  distinct(imaging_date,well_name,well_row,well_column,EU_time) %>%
  ggplot(aes(x=well_column,y=well_row,fill=EU_time)) +
  geom_tile() +
  scale_y_discrete(limits=rev) +
  facet_wrap(~imaging_date) +
  ggtitle("EU time")


dat_non_border %>%
  mutate(well_row = factor(str_extract(well_name, "[A-Z]"),levels=LETTERS),
         well_column = as.integer(as.character(str_extract(well_name, "[0-9]+")))) %>%
  count(imaging_date,well_name,well_row,well_column,EU_concentration) %>%
  ggplot(aes(x=well_column,y=well_row,fill=n)) +
  geom_tile() +
  scale_y_discrete(limits=rev) +
  facet_wrap(~imaging_date) +
  scale_fill_viridis_c() +
  ggtitle("EU concentration")


# compute fold-change from mean for each well
dat_non_border %>%
  mutate(well_row = factor(str_extract(well_name, "[A-Z]"),levels=LETTERS),
         well_column = as.integer(as.character(str_extract(well_name, "[0-9]+")))) %>%
  group_by(imaging_date,well_name,well_row,well_column,EU_concentration,EU_time,Cell,Drug) %>%
  summarise(well_mean_EU = mean(Nuclei_intensity_mean_647),.groups="drop") %>%
  left_join(dat_non_border %>%
              group_by(imaging_date,well_name,EU_concentration,EU_time,Cell,Drug) %>%
              summarise(well_mean_EU = mean(Nuclei_intensity_mean_647),.groups="drop") %>%
              group_by(imaging_date,EU_concentration,EU_time,Cell,Drug) %>%
              summarise(plate_mean_EU = mean(well_mean_EU),.groups="drop"),
            by = join_by(imaging_date, EU_concentration, EU_time, Cell, Drug)) %>%
  mutate(log2FC_from_mean = log2(well_mean_EU/plate_mean_EU)) %>%
  ggplot(aes(x=well_column,y=well_row,fill=log2FC_from_mean)) +
  geom_tile() +
  scale_fill_distiller(type = "div", palette = "RdBu", limits = c(-2,2)) +
  scale_y_discrete(limits=rev) +
  facet_wrap(~imaging_date) +
  ggtitle("Within plate difference from mean per well")

# compute fold-change from mean for each well
dat_non_border %>%
  mutate(well_row = factor(str_extract(well_name, "[A-Z]"),levels=LETTERS),
         well_column = as.integer(as.character(str_extract(well_name, "[0-9]+")))) %>%
  group_by(imaging_date,well_name,well_row,well_column,EU_concentration,EU_time,Cell,Drug) %>%
  summarise(well_mean_EU = mean(Nuclei_intensity_mean_647),.groups="drop") %>%
  left_join(dat_non_border %>%
              group_by(imaging_date,well_name,EU_concentration,EU_time,Cell,Drug) %>%
              summarise(well_mean_EU = mean(Nuclei_intensity_mean_647),.groups="drop") %>%
              group_by(EU_concentration,EU_time,Cell,Drug) %>%
              summarise(overall_mean_EU = mean(well_mean_EU),.groups="drop"),
            by = join_by(EU_concentration, EU_time, Cell, Drug)) %>%
  mutate(log2FC_from_mean = log2(well_mean_EU/overall_mean_EU)) %>%
  ggplot(aes(x=well_column,y=well_row,fill=log2FC_from_mean)) +
  geom_tile() +
  scale_fill_distiller(type = "div", palette = "RdBu", limits = c(-2,2)) +
  scale_y_discrete(limits=rev) +
  facet_wrap(~imaging_date) +
  ggtitle("Experiment difference from mean per well")

```

Plate overviews indicate that imaging date 20240229 rows M/N are clear outliers. This is likely due to pipetting inaccuracies during click reaction addition (Alex said he added click manually after the other rows). We exclude this data.

```{r exclude_data}

dat_non_border_with_exclusions <- dat_non_border %>%
    mutate(well_row = factor(str_extract(well_name, "[A-Z]"),levels=LETTERS),
           well_column = as.integer(as.character(str_extract(well_name, "[0-9]+")))) %>%
  filter(!(imaging_date=="20240229" & well_row %in% c("M","N")))

```

Correct for plate-to-plate variation in EU staining for plotting (don't use normalisation for statistics, just include plate and well as random effects.)

```{r plate_normalisation}

normalisation_data <- dat_non_border_with_exclusions %>%
  filter(EU_concentration!="None" & Drug=="Vehicle") %>%
  group_by(imaging_date,well_name,EU_time,Cell) %>%
  summarise(well_mean_IF = mean(Nuclei_intensity_mean_647),.groups="drop") %>%
  group_by(imaging_date,EU_time,Cell) %>%
  summarise(plate_mean_IF = mean(well_mean_IF),.groups="drop")

normalisation_data %>%
  pivot_wider(names_from = imaging_date, values_from=plate_mean_IF) %>%
  ggplot(aes(x=`20240304`,y=`20240229`)) + geom_point() + geom_abline() + facet_wrap(~EU_time)

normalisation_data %>%
  pivot_wider(names_from = imaging_date, values_from=plate_mean_IF) %>%
  ggplot(aes(x=`20240304`,y=`20240307`)) + geom_point() + geom_abline() + facet_wrap(~EU_time)

# compute a normalisation factor for plates on imaging_date 20240305 and 20240308

normalisation_factors <- normalisation_data %>%
  filter(imaging_date=="20240304") %>%
  select(-imaging_date) %>%
  rename(normalise_to = plate_mean_IF) %>%
  left_join(normalisation_data,by = join_by(EU_time, Cell)) %>%
  mutate(normalisation_factor=normalise_to/plate_mean_IF) %>%
  group_by(EU_time,imaging_date) %>%
  summarise(plate_normaliser = mean(normalisation_factor),.groups = "drop")

normalisation_factors

```

Now average over replicates, including plate-to-plate normalisation factors.

```{r average_over_replicates,fig.height=4,fig.width=12}

well_summaries <- dat_non_border_with_exclusions %>%
  filter(EU_concentration!="None" & Drug!="None") %>%
  left_join(normalisation_factors, by = join_by(imaging_date, EU_time)) %>%
  mutate(Nuclei_intensity_mean_647_plate_normalised = Nuclei_intensity_mean_647 * plate_normaliser) %>%
  group_by(imaging_date,well_name,EU_time,Cell,Drug) %>%
  summarise(Nuclei_intensity_mean_647 = mean(Nuclei_intensity_mean_647),
            Nuclei_intensity_mean_647_plate_normalised = mean(Nuclei_intensity_mean_647_plate_normalised),
            .groups='drop')

overall_summaries <- well_summaries %>%
  group_by(EU_time,Cell,Drug) %>%
  summarise(mean_Nuclei_intensity_mean_647 = mean(Nuclei_intensity_mean_647),
            sd_Nuclei_intensity_mean_647 = sd(Nuclei_intensity_mean_647),
            mean_Nuclei_intensity_mean_647_plate_normalised = mean(Nuclei_intensity_mean_647_plate_normalised),
            sd_Nuclei_intensity_mean_647_plate_normalised = sd(Nuclei_intensity_mean_647_plate_normalised),
            .groups='drop')

well_summaries %>%
  ggplot(aes(x=Drug,y=Nuclei_intensity_mean_647_plate_normalised,col=Cell)) +
  facet_wrap(~factor(EU_time, levels=c("15 min","30 min","60 min")), nrow=1,scales="free_y") +
  geom_col(data = overall_summaries, aes(y=mean_Nuclei_intensity_mean_647_plate_normalised,fill=Cell),col=NA,alpha=0.4,position=position_dodge(0.8),width=0.8) +
  geom_point(data = overall_summaries, aes(y=mean_Nuclei_intensity_mean_647_plate_normalised),position=position_dodge(0.8)) +
  geom_errorbar(data = overall_summaries, aes(y=mean_Nuclei_intensity_mean_647_plate_normalised,
                                              ymin=mean_Nuclei_intensity_mean_647_plate_normalised-sd_Nuclei_intensity_mean_647_plate_normalised,
                                              ymax=mean_Nuclei_intensity_mean_647_plate_normalised+sd_Nuclei_intensity_mean_647_plate_normalised),
                position=position_dodge(0.8),width=0.1) +
  geom_jitter(position=position_jitterdodge(jitter.width=0.15,dodge.width=0.8),pch=4,alpha=0.5) +
  scale_y_continuous(limits = c(0,NA),name="Nuclear intensity (A.U.)") +
  theme_bw() + 
  theme(axis.text.x = element_text(angle=90,hjust=1,vjust=0.5),
        axis.title.x = element_blank(),
        panel.grid = element_blank())
ggsave("PLOTS/EU_barplots_by_cell.pdf",width=12,height=4)

well_summaries %>%
  ggplot(aes(x=Cell,y=Nuclei_intensity_mean_647_plate_normalised,col=Drug)) +
  facet_wrap(~factor(EU_time, levels=c("15 min","30 min","60 min")), nrow=1,scales="free_y") +
  geom_col(data = overall_summaries, aes(y=mean_Nuclei_intensity_mean_647_plate_normalised,fill=Drug),col=NA,alpha=0.4,position=position_dodge(0.8),width=0.8) +
  geom_point(data = overall_summaries, aes(y=mean_Nuclei_intensity_mean_647_plate_normalised),position=position_dodge(0.8)) +
  geom_errorbar(data = overall_summaries, aes(y=mean_Nuclei_intensity_mean_647_plate_normalised,
                                              ymin=mean_Nuclei_intensity_mean_647_plate_normalised-sd_Nuclei_intensity_mean_647_plate_normalised,
                                              ymax=mean_Nuclei_intensity_mean_647_plate_normalised+sd_Nuclei_intensity_mean_647_plate_normalised),
                position=position_dodge(0.8),width=0.1) +
  scale_y_continuous(limits = c(0,NA),name="Nuclear intensity (A.U.)") +
  theme_bw() + 
  theme(axis.text.x = element_text(angle=90,hjust=1,vjust=0.5),
        axis.title.x = element_blank(),
        panel.grid = element_blank())
ggsave("PLOTS/EU_barplots_by_drug.pdf",width=12,height=4)

```

EU data has more convincing signal-to-background at 30 min than at 15 min incubation time. Particularly for studying the effects of transcriptional inhibitors. Therefore 15 min data are noisier and 60 min data are not preferred (due to longer EU incubation). We therefore focus on 30 min data.


```{r plot_EU_30min_data}

well_summaries %>%
  filter(EU_time=="30 min") %>%
  ggplot(aes(x=Drug,y=Nuclei_intensity_mean_647_plate_normalised,col=Cell)) +
  geom_col(data = filter(overall_summaries,EU_time=="30 min"), aes(y=mean_Nuclei_intensity_mean_647_plate_normalised,fill=Cell),col=NA,alpha=0.4,position=position_dodge(0.8),width=0.8) +
  geom_point(data = filter(overall_summaries,EU_time=="30 min"), aes(y=mean_Nuclei_intensity_mean_647_plate_normalised),position=position_dodge(0.8)) +
  geom_errorbar(data = filter(overall_summaries,EU_time=="30 min"), aes(y=mean_Nuclei_intensity_mean_647_plate_normalised,
                                              ymin=mean_Nuclei_intensity_mean_647_plate_normalised-sd_Nuclei_intensity_mean_647_plate_normalised,
                                              ymax=mean_Nuclei_intensity_mean_647_plate_normalised+sd_Nuclei_intensity_mean_647_plate_normalised),
                position=position_dodge(0.8),width=0.1) +
  geom_jitter(position=position_jitterdodge(jitter.width=0.15,dodge.width=0.8),pch=4,alpha=0.5) +
  scale_y_continuous(limits = c(0,NA),name="Nuclear intensity (A.U.)") +
  theme_bw() + 
  theme(axis.text.x = element_text(angle=90,hjust=1,vjust=0.5),
        axis.title.x = element_blank(),
        panel.grid = element_blank())
ggsave("PLOTS/EU_barplots_by_cell_30min.pdf",width=6,height=4)

```

## Statistical modelling

Can see a stronger effect of Triptolide on parental cells compared to ARMC5 KO, however the effect appears mild when one considers the higher starting level of EU in these clones. Can also see pronounced clone-to-clone variability.

To make this more rigorous, we use linear mixed-effects models on single-cell data.

```{r stats_testing_unperturbed}

to_fit_unperturbed <- dat_non_border_with_exclusions %>%
  filter(Drug=="Vehicle" & EU_concentration=="100 µM" & EU_time=="30 min") %>%
  select(imaging_date,well_name,Cell,Nuclei_intensity_mean_647) %>%
  mutate(Cell=factor(Cell,levels=c("HEK293 parental","HEK293 ARMC5 KO 1A3","HEK293 ARMC5 KO 2B6")),
         imaging_date=factor(imaging_date),
         well_name=factor(well_name)) %>%
  rename(EU_incorporation=Nuclei_intensity_mean_647) %>%
  filter(EU_incorporation > 0) %>%
  mutate(log_EU_incorporation=log(EU_incorporation))
  
unperturbed_model <- nlme::lme(fixed = log_EU_incorporation ~ Cell,
                               random= ~1 | imaging_date/well_name,
                               data = to_fit_unperturbed)

summary(unperturbed_model)

emmeans(unperturbed_model, trt.vs.ctrl1 ~ Cell, adjust="bonferroni")

to_fit_unperturbed %>%
  ggplot(aes(x=Cell,y=log_EU_incorporation,col=imaging_date,group=interaction(imaging_date,well_name,Cell))) +
  geom_boxplot() +
  ggtitle("unperturbed")

to_fit_unperturbed %>%
  ggplot(aes(x=imaging_date,y=log_EU_incorporation,col=Cell,group=interaction(imaging_date,well_name,Cell))) +
  geom_boxplot() +
  ggtitle("unperturbed")

```

One of the two clones has significantly higher EU incorporation at 30 min.

This is interesting, but what we actually want to know is whether the triptolide has a different effect on ARMC5 KO compared to parental cells.

```{r stats_testing_triptolide,fig.height=5,fig.width=8}

# remove the few cells with EU < 0

to_fit_triptolide <- dat_non_border_with_exclusions %>%
  filter(Drug %in% c("Vehicle","Triptolide 300 nM") & EU_concentration=="100 µM" & EU_time=="30 min") %>%
  select(imaging_date,well_name,Cell,Drug,Nuclei_intensity_median_647) %>%
  mutate(Cell=factor(Cell,levels=c("HEK293 parental","HEK293 ARMC5 KO 1A3","HEK293 ARMC5 KO 2B6")),
         Drug=factor(Drug,levels=c("Vehicle","Triptolide 300 nM")),
         imaging_date=factor(imaging_date),
         well_name=factor(well_name)) %>%
  rename(EU_incorporation=Nuclei_intensity_median_647) %>%
  filter(EU_incorporation > 0) %>%
  mutate(log_EU_incorporation=log(EU_incorporation))
  
triptolide_model <- nlme::lme(fixed = log_EU_incorporation ~ Cell*Drug,
                               random= ~1 | imaging_date/well_name,
                               data = to_fit_triptolide)

summary(triptolide_model)

# Use emmeans to look at marginal effects across cell lines. 
# looks like a different effect of Triptolide on the cell lines
emmip(triptolide_model, Cell ~ Drug)

# Test for Drug effect in each cell type (treated vs vehicle)
contrast(emmeans(triptolide_model, ~ Drug|Cell),interaction = "trt.vs.ctrl1", adjust="bonferroni")
# Test for Cell effect in each treatment (KO vs parental)
contrast(emmeans(triptolide_model, ~ Cell|Drug),interaction = "trt.vs.ctrl1", adjust="bonferroni")

# Test for difference in drug effect in each KO cell line
contrast(emmeans(triptolide_model, ~ Drug*Cell),interaction = "trt.vs.ctrl1", adjust="bonferroni")

to_fit_triptolide %>%
  ggplot(aes(x=Drug,y=log_EU_incorporation,col=Cell)) +
  geom_boxplot() +
  facet_wrap(~imaging_date,ncol=3)

```

Not only does triptolide affect EU for all cell lines, but it seems there is a significant interaction between drug and cell (i.e. Triptolide has a different effect on the KO cells compared to the parental cells)

```{r plots}

to_plot_well_summaries <- well_summaries %>%
  filter(EU_time=="30 min" & Drug %in% c("Vehicle","Triptolide 300 nM")) %>%
  mutate(cell_label=case_when(Cell == "HEK293 parental" ~ "WT",
                              Cell == "HEK293 ARMC5 KO 1A3" ~ "K.O. 1",
                              Cell == "HEK293 ARMC5 KO 2B6" ~ "K.O. 2",
                              TRUE ~ NA_character_),
         cell_label=factor(cell_label,levels=c("WT","K.O. 1","K.O. 2"))) %>%
  mutate(condition_label=paste(cell_label,Drug),
         condition_label=factor(condition_label,
                                levels=c("WT Vehicle","K.O. 1 Vehicle","K.O. 2 Vehicle",
                                         "WT Triptolide 300 nM","K.O. 1 Triptolide 300 nM","K.O. 2 Triptolide 300 nM")))

to_plot_overall_summaries <- overall_summaries %>%
  filter(EU_time=="30 min" & Drug %in% c("Vehicle","Triptolide 300 nM")) %>%
    mutate(cell_label=case_when(Cell == "HEK293 parental" ~ "WT",
                              Cell == "HEK293 ARMC5 KO 1A3" ~ "K.O. 1",
                              Cell == "HEK293 ARMC5 KO 2B6" ~ "K.O. 2",
                              TRUE ~ NA_character_),
         cell_label=factor(cell_label,levels=c("WT","K.O. 1","K.O. 2"))) %>%
  mutate(condition_label=paste(cell_label,Drug),
         condition_label=factor(condition_label,
                                levels=c("WT Vehicle","K.O. 1 Vehicle","K.O. 2 Vehicle",
                                         "WT Triptolide 300 nM","K.O. 1 Triptolide 300 nM","K.O. 2 Triptolide 300 nM")))

axis_norm <- to_plot_overall_summaries %>% filter(condition_label=="WT Vehicle") %>% pull(mean_Nuclei_intensity_mean_647_plate_normalised)

to_plot_well_summaries %>%
  ggplot(aes(x=Drug,y=Nuclei_intensity_mean_647_plate_normalised/axis_norm,col=condition_label)) +
  geom_col(data = to_plot_overall_summaries, aes(y=mean_Nuclei_intensity_mean_647_plate_normalised/axis_norm,fill=condition_label),
           col=NA,alpha=0.4,position=position_dodge(0.8),width=0.6) +
  geom_point(data = to_plot_overall_summaries, aes(y=mean_Nuclei_intensity_mean_647_plate_normalised/axis_norm),position=position_dodge(0.8)) +
  geom_errorbar(data = to_plot_overall_summaries,
                aes(y=mean_Nuclei_intensity_mean_647_plate_normalised/axis_norm,
                    ymin=(mean_Nuclei_intensity_mean_647_plate_normalised-sd_Nuclei_intensity_mean_647_plate_normalised)/axis_norm,
                    ymax=(mean_Nuclei_intensity_mean_647_plate_normalised+sd_Nuclei_intensity_mean_647_plate_normalised)/axis_norm),
                position=position_dodge(0.8),width=0.25) +
  geom_jitter(position=position_jitterdodge(jitter.width=0.15,dodge.width=0.8),pch=4,alpha=0.5) +
  scale_y_continuous(limits = c(0,1.55),name="Mean nuclear\nintensity",expand = c(0,0)) +
  scale_fill_manual(values = c(wt_col,armc5_col,armc5_col,wt_trp_col,armc5_trp_col,armc5_trp_col)) +
  scale_colour_manual(values = c(wt_col,armc5_col,armc5_col,wt_trp_col,armc5_trp_col,armc5_trp_col)) +
  theme_bw() + 
  theme(axis.text.x = element_text(angle=90,hjust=1,vjust=0.5),
        axis.title.x = element_blank(),
        panel.grid = element_blank(),
        legend.position = "none")
ggsave("PLOTS/EU_trp_only_30min.pdf",width=2,height=2.5)

```

```{r well_numbers}

dat_non_border %>%
  filter(EU_time=="30 min" & Drug %in% c("Vehicle","Triptolide 300 nM")) %>%
  group_by(imaging_date,Cell,Drug,EU_time,EU_concentration,well_name) %>%
  count() 

to_plot_well_summaries %>%
  group_by(imaging_date,Cell,Drug,EU_time) %>%
  count()

to_plot_well_summaries %>%
  group_by(Cell,Drug,EU_time) %>%
  count()


```

```{r save_summaries}

write_csv(x = well_summaries,file = "SUMMARIES/Drug_EU_well_summaries.csv")
write_csv(x = overall_summaries,file = "SUMMARIES/Drug_EU_overall_summaries.csv")

```


```{r session_info}

sessionInfo()

```
