---
title: "Buffering of POLR2A variation on pSer2 in unperturbed cells"
subtitle: "Single-cell immunofluorescence, using direct protein detection to correct for measurement noise"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: show
    df_print: paged
date: "`r format(Sys.Date())`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

rm(list=ls())
library(tidyverse)
library(patchwork)
library(caret)
library(ranger)

cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

PLOTS_DIR <- "PLOTS"
dir.create(PLOTS_DIR, showWarnings = FALSE, recursive = TRUE)

read_quantification <- function(dir_name) {
  csv_files <- Sys.glob(file.path(dir_name,"*.csv"))
  dt_list <- lapply(csv_files, data.table::fread)
  df <- tibble(path = csv_files,
               feature_values = dt_list) %>%
    mutate(dir_name = dir_name,
           file_name = basename(path),
           feature_values = purrr::map(feature_values,tibble)) %>%
    separate(file_name,into = c("well_name","ch1","ch2","ch3","ch4","ch5","ch6","sample","id"),sep="[,_.]", extra = "drop") %>%
    mutate(well_name = str_replace(well_name, "^Well", ""),
           field_id = as.integer(id))
  return(df)
}

```

## Question and experimental design

### Summary

Having observed limited changes in pSer2 (CTD Ser2 phosphorylation) at the level
of cell populations when total POLR2A levels are perturbed, we now ask whether
such buffering is observed at the level of single cells to prevent propagation
of natural cell-to-cell variability in POLR2A abundance to transcription.

### Experiment and analysis approach

We measured nuclear pSer2 via immunofluorescence with monoclonal 3E10 antibody
and two readouts of total Pol II: one direct, using a mClover-POLR2A fusion, and
secondly with a monoclonal D8L4Y POLR2A antibody, all in the same cells as we
visualised pSer2. These experiments were repeated in mCherry-POLR2A cells and
using different combinations of secondary channels for detecting the Rat and
Rabbit primary antibodies (secondary switch).

Buffering is quantified as the slope, $\beta$, of $\log(\text{pSer2})$ on
$\log(\text{total Pol II})$. $\beta=1$ is proportional (constant phosphorylated
fraction), $\beta<1$ is buffering, $\beta=0$ is perfect homeostasis.

The difficulty in determining this slope is measurement noise, which biases the
slope in two opposing ways. Noise in the quantity on the x-axis (total Pol II)
flattens any fitted slope toward zero (regression-dilution), which may
appear as buffering. In the other direction, a per-cell imaging factor
(confounding factor) shared by some or all fluorescence channels
(immunofluorescence artifact, segmentation inaccuracy, nuclear thickness)
inflates the apparent slope, $\beta$, toward 1 and so tends to mask true
buffering.

To investigat (and correct for) these issues, we make use of the
mClover-POLR2A and mCherry-POLR2A fusions. Because fluorescent proteins carry
no antibody noise, we can use these as a reference to remove correlated
immunofluorescence noise from
the slope, removing the regression-dilution bias. The shared-imaging inflation
is not fully removed, but it only pushes the slope upward, so any slope we find
below 1 is a conservative statement of buffering.

## Load, verify, and clean up data

### Load data

Load data from csvs. Nuclei segmented with cellpose 4 and quantified using 
``blimp``.

```{r}
data_dir_images <- "/srv/scratch/berrylab/z3532965/systems_Ti2/20260626_POLR2A_heterogeneity/20260629_134305_894/OME-TIFF-MIP"
data_dir_quant <- "/srv/scratch/berrylab/z3532965/systems_Ti2/20260626_POLR2A_heterogeneity/20260629_134305_894/QUANTIFICATION"
plate_layout_file <- "METADATA/20260626_single_cell_POLR2A_plate_layout.csv"

# Read plate layout
plate_layout <- read_csv(plate_layout_file, show_col_types = FALSE) %>%
  rename(well_name=well)

# Read and process quantification data
f <- read_quantification(data_dir_quant) %>%
  dplyr::mutate(data_dir = dirname(dirname(path))) %>%
  dplyr::select(-c(ch1, ch2, ch3, ch4, ch5, ch6, sample))

dat <- left_join(f,plate_layout,by="well_name") %>%
  unnest(feature_values) %>%
  mutate(unique_id = paste(well_name, label, sep = "_")) %>%
  mutate(well_row=substr(well_name,1,1),
         well_col=substr(well_name,2,3)) %>%
  ungroup()

```

### Plot plate layouts

Plot plate layouts to verify and get an overview of the experiment.

```{r}

source("/srv/scratch/z3532965/src/blana/Scott/plot_plate_layout.R")

plate_layout %>% group_by(CELLS,PRIMARY_RAT,PRIMARY_RABBIT,SECONDARY_488,SECONDARY_568,SECONDARY_647,DAPI) %>%
  count()

plot_plate_layout(plate_layout, color_by = "CELLS", title = "Cell type")
plot_plate_layout(plate_layout, color_by = "PRIMARY_RAT", title = "Primary (Rat)")
plot_plate_layout(plate_layout, color_by = "PRIMARY_RABBIT", title = "Primary (Rabbit)")
plot_plate_layout(plate_layout, color_by = "SECONDARY_488", title = "Secondary (488)")
plot_plate_layout(plate_layout, color_by = "SECONDARY_568", title = "Secondary (568)")
plot_plate_layout(plate_layout, color_by = "SECONDARY_647", title = "Secondary (647)")

```
### Clean up data

Training data has been manually annotated for several fields of view to identify
mitotic cells and oversegmented cells. We focus on properly segmented interphase
cells (where transcription is ongoing and properly quantified) and therefore
remove these cells.

```{r machine_learning_cleanup}

# define helpers
normalise_stem <- function(x) x %>% str_replace_all(",", "_") %>% str_replace("\\.ome$", "_ome")
path_to_stem   <- function(p) basename(p) %>% str_remove("\\.csv$") %>% normalise_stem()

read_annotations <- function(anno_dir) {
  list.files(anno_dir, pattern = "_annotation\\.csv$", full.names = TRUE) %>%
    set_names() %>%
    map_dfr(~ read_csv(.x, show_col_types = FALSE,
                       col_select = c(label, annotations, annotation_names)),
            .id = "file") %>%
    filter(!is.na(annotation_names)) %>%
    transmute(file,
              stem  = normalise_stem(str_remove(basename(file), "_annotation\\.csv$")),
              label, class = annotation_names)
}

zscore_405 <- function(df, well_col_id = "well_name") {
  cols_405 <- names(df) %>% str_subset("^Nuclei_intensity_.*_405$")
  df %>%
    select(-matches("^Nuclei_intensity_.*_405_z$")) %>%
    group_by(across(all_of(well_col_id))) %>%
    mutate(across(all_of(cols_405), ~ as.numeric(scale(.x)), .names = "{.col}_z")) %>%
    ungroup()
}

fit_classifier <- function(df, anno_dir, feature_cols, positive_class, name,
                           well_col_id = "well_name", corr_cutoff = 0.9, seed = 1) {
  
  df <- zscore_405(df, well_col_id = well_col_id)
  
  # swap any raw 405 intensity features for their z-scored versions
  feature_cols <- feature_cols %>%
    map_chr(\(nm) if (str_detect(nm, "^Nuclei_intensity_.*_405$")) paste0(nm, "_z") else nm) %>%
    intersect(names(df))
  
  missing <- setdiff(feature_cols, names(df))
  if (length(missing)) warning("features not found: ", paste(missing, collapse = ", "))
  
  anno   <- read_annotations(anno_dir)
  joined <- anno %>% left_join(mutate(df, stem = path_to_stem(path)),
                               by = c("stem", "label"))
  n_bad <- sum(is.na(joined$path))
  if (n_bad) warning(sprintf("[%s] %d/%d annotations didn't match df — check keys.",
                             name, n_bad, nrow(anno)))
  
  train <- joined %>%
    transmute(class = factor(class), across(all_of(feature_cols))) %>%
    drop_na()
  stopifnot(nlevels(train$class) >= 2, positive_class %in% levels(train$class))
  train$class <- relevel(train$class, ref = positive_class)
  
  x <- as.data.frame(train[, feature_cols, drop = FALSE])
  y <- train$class
  w <- as.numeric(1 / table(y)[y]); w <- w / mean(w)
  
  grid <- expand.grid(mtry = c(2, 3, 4, 5), splitrule = "gini", min.node.size = c(5, 10))
  
  set.seed(seed)
  fit <- train(
    x = x, y = y, weights = w,
    method = "ranger", metric = "ROC",
    preProcess = c("nzv", "corr"),
    trControl = trainControl(method = "cv", number = 5,
                             classProbs = TRUE, summaryFunction = twoClassSummary,
                             preProcOptions = list(cutoff = corr_cutoff)),
    tuneGrid = grid, num.trees = 1000, importance = "impurity"
  )
  
  # performance from the final model's OOB predictions
  oob_prob <- fit$finalModel$predictions[, positive_class]
  oob_pred <- factor(ifelse(oob_prob > 0.5, positive_class,
                            setdiff(levels(y), positive_class)), levels = levels(y))
  cm  <- confusionMatrix(oob_pred, y, positive = positive_class)
  pos <- oob_prob[y == positive_class]; neg <- oob_prob[y != positive_class]
  auc <- mean(outer(pos, neg, ">") + 0.5 * outer(pos, neg, "=="))
  
  # which features survived the nzv/corr reduction
  kept    <- fit$finalModel$xNames
  omitted <- setdiff(feature_cols, kept)
  
  # assemble summary objects for the notebook to render 
  class_counts <- as_tibble(table(class = y))
  
  metrics <- tibble(
    positive     = positive_class,
    n            = length(y),
    n_pos        = sum(y == positive_class),
    n_feat_in    = length(feature_cols),
    n_feat_kept  = length(kept),
    AUC          = auc,
    Precision    = unname(cm$byClass["Precision"]),
    Recall       = unname(cm$byClass["Recall"]),
    F1           = unname(cm$byClass["F1"]),
    best_mtry    = fit$bestTune$mtry,
    best_node    = fit$bestTune$min.node.size,
    corr_cutoff  = corr_cutoff
  )
  
  confusion <- as.data.frame.matrix(cm$table)
  
  importance <- fit$finalModel$variable.importance %>%
    enframe("feature", "importance") %>%
    arrange(desc(importance))
  
  features <- tibble(
    feature = feature_cols,
    kept    = feature_cols %in% kept
  )
  
  list(model = fit, feature_cols = kept, positive_class = positive_class,
       name = name, well_col_id = well_col_id, anno_dir = anno_dir,
       class_counts = class_counts, metrics = metrics, confusion = confusion,
       importance = importance, features = features, dropped = omitted)
}

apply_classifier <- function(df, bundle, threshold = 0.5) {
  scored <- zscore_405(df, well_col_id = bundle$well_col_id)
  fc <- bundle$feature_cols
  ok <- complete.cases(scored[, fc])
  prob <- rep(NA_real_, nrow(scored))
  prob[ok] <- predict(bundle$model, scored[ok, fc, drop = FALSE],
                      type = "prob")[[bundle$positive_class]]
  df[[paste0(bundle$name, "_prob")]] <- prob
  df[[bundle$name]] <- prob > threshold
  df
}

write_image_predictions <- function(df, bundle, out_dir = NULL) {
  name      <- bundle$name
  prob_col  <- paste0(name, "_prob")
  stopifnot(prob_col %in% names(df))
  
  # map each training image's stem back to its annotation file path
  anno <- read_annotations(bundle$anno_dir)
  file_by_stem <- anno %>% distinct(stem, file)
  
  # ground-truth labels (only exist for annotated cells)
  truth <- anno %>% select(stem, label, annotated_class = class)
  
  df %>%
    mutate(stem = path_to_stem(path)) %>%
    inner_join(file_by_stem, by = "stem") %>%
    left_join(truth, by = c("stem", "label")) %>%
    select(file, stem, label,
           any_of(c("well_name")),
           annotated_class,
           !!prob_col := all_of(prob_col),
           !!name     := all_of(name)) %>%
    group_by(file) %>%
    group_walk(~ {
      out_name <- str_replace(basename(.y$file), "_annotation\\.csv$", "_predictions.csv")
      dir_out  <- out_dir %||% dirname(.y$file)
      write_csv(.x, file.path(dir_out, out_name))
    })
  invisible(df)
}

```

We select features to pass to the classifiers, train and apply classifiers, and
write predictions back to file for manual verification together with images. 

```{r fit_classifiers}
feature_cols <- c(
  paste0("Nuclei_", c("area","area_convex","axis_major_length","axis_minor_length",
                      "eccentricity","extent","feret_diameter_max","solidity",
                      "perimeter","perimeter_crofton","euler_number")),
  names(dat) %>% str_subset("^Nuclei_intensity_.*_405$"),
  names(dat) %>% str_subset("^Nuclei_405_Haralick")
)
feature_cols <- intersect(feature_cols, names(dat))

mito <- fit_classifier(dat, "ANNOTATIONS/MITOTIC", feature_cols, "Mitotic",
                       name = "MITOTIC", well_col_id = "well_name",
                       corr_cutoff = 0.975)

knitr::kable(t(mito$metrics), digits = 3, caption = "MITOTIC: OOB performance & best tune")
knitr::kable(mito$confusion, caption = "Confusion matrix (rows = predicted, cols = actual)")
knitr::kable(head(mito$importance, 10), digits = 3, caption = "Top 10 features (after reduction)")

dat <- apply_classifier(dat, mito)
write_image_predictions(dat, mito)

oversegmented <- fit_classifier(dat, "ANNOTATIONS/OVERSEGMENTED", feature_cols, "Oversegmented",
                                name = "OVERSEGMENTED", well_col_id = "well_name",
                                corr_cutoff = 0.975)

knitr::kable(t(oversegmented$metrics), digits = 3, caption = "OVERSEGMENTED: OOB performance & best tune")
knitr::kable(oversegmented$confusion, caption = "Confusion matrix (rows = predicted, cols = actual)")
knitr::kable(head(oversegmented$importance, 10), digits = 3, caption = "Top 10 features (after reduction)")

dat <- apply_classifier(dat, oversegmented)
write_image_predictions(dat, oversegmented)
```

Confirmed visual inspection of classifier predictions looks appropriate.

We now these classifications, and conventional filtering steps to cleanup the 
single-cell data.

```{r cleanup}

# remove border cells
filtered_dat <- filter(dat,Nuclei_is_border==FALSE) %>%
  filter(DAPI=="DAPI")

area_lower_limit <- 400
area_upper_limit <- 10000

filtered_dat %>%
  ggplot(aes(x=Nuclei_area,group=well_name)) + 
  facet_wrap(~CELLS,ncol=1) + 
  geom_density() +
  coord_cartesian(xlim=c(0,5000)) +
  geom_vline(xintercept=area_lower_limit) +
  geom_vline(xintercept=area_upper_limit)

filtered_dat <- filter(filtered_dat,
                       Nuclei_area > area_lower_limit & 
                         Nuclei_area < area_upper_limit)

# apply classifiers based on DAPI
DAPI_bkgrd <- 200

# subtract background and z-score
filtered_dat <- filtered_dat %>%
  group_by(well_name) %>%
  mutate(DAPI_sum = Nuclei_area*(Nuclei_intensity_mean_405-DAPI_bkgrd),
         DAPI_sum_zscore = as.numeric(scale(DAPI_sum,center = T, scale=T))) %>%
  ungroup()

# count fraction positive
filtered_dat %>% group_by(CELLS,well_name) %>% 
  count(MITOTIC) %>% 
  group_by(CELLS,well_name) %>% 
  mutate(fraction_mitotic = n/sum(n)) %>%
  filter(MITOTIC==TRUE) %>%
  select(-MITOTIC) %>%
  ggplot(aes(y=fraction_mitotic,x=CELLS)) + geom_boxplot() 

# count fraction positive (after removing MITOTIC)
filtered_dat %>% group_by(CELLS,well_name) %>% 
  filter(!MITOTIC) %>%
  count(OVERSEGMENTED) %>% 
  group_by(CELLS,well_name) %>% 
  mutate(fraction_oversegmented = n/sum(n)) %>%
  filter(OVERSEGMENTED==TRUE) %>%
  select(-OVERSEGMENTED) %>%
  ggplot(aes(y=fraction_oversegmented,x=CELLS)) + geom_boxplot()

# check DAPI histograms for MITOTIC classifier
filtered_dat %>%
  ggplot(aes(x=DAPI_sum_zscore,colour=MITOTIC)) +
  facet_wrap(~CELLS,ncol=1) + 
  geom_density()

# check DAPI histograms for OVERSEGMENTED classifier
filtered_dat %>%
  ggplot(aes(x=DAPI_sum_zscore,colour=OVERSEGMENTED)) +
  facet_wrap(~CELLS,ncol=1) + 
  geom_density()

# remove cells labelled as OVERSEGMENTED or MITOTIC
cells_before_classifiers <- filtered_dat %>% group_by(well_name) %>% count(name = "before")
filtered_dat <- filter(filtered_dat,!MITOTIC & !OVERSEGMENTED)
#filtered_dat <- filter(filtered_dat,!MITOTIC)
cells_after_classifiers <- filtered_dat %>% group_by(well_name) %>% count(name = "after")

# repeat subtract background and z-score after this filtering
filtered_dat <- filtered_dat %>%
  select(-DAPI_sum,-DAPI_sum_zscore) %>%
  group_by(well_name) %>%
  mutate(DAPI_sum = Nuclei_area*(Nuclei_intensity_mean_405-DAPI_bkgrd),
         DAPI_sum_zscore = scale(DAPI_sum,center = T, scale=T)) %>%
  ungroup()

dapi_sum_zscore_lower_limit <- -2
dapi_sum_zscore_upper_limit <- 4

filtered_dat %>%
  ggplot(aes(x=DAPI_sum,group=well_name,col=well_row)) +
  facet_wrap(~CELLS,ncol=1) + 
  coord_cartesian(xlim=c(0,2e7)) +
  geom_density()

filtered_dat %>%
  ggplot(aes(x=DAPI_sum_zscore,group=well_name,col=well_row)) +
  facet_wrap(~CELLS,ncol=1) + 
  coord_cartesian(xlim=c(-5,5)) +
  geom_density() +
  geom_vline(xintercept=dapi_sum_zscore_lower_limit) +
  geom_vline(xintercept=dapi_sum_zscore_upper_limit)

filtered_dat <- filter(filtered_dat,
                       DAPI_sum_zscore > dapi_sum_zscore_lower_limit & 
                         DAPI_sum_zscore < dapi_sum_zscore_upper_limit)

```
### Background calculation

GFP and mCherry background come from cells imaged alongside that do not carry these fluorescent proteins and are not stained with secondary antibodies. 

```{r qc_FP}

GFP_bkgrd <- filter(filtered_dat,CELLS=="OsTIR(F74G)" & SECONDARY_488=="None" & SECONDARY_568=="None" & SECONDARY_647=="None") %>%
  ungroup() %>%
  group_by(well_name) %>%
  summarise(bkgrd = median(Nuclei_intensity_mean_488),.groups = 'drop') %>%
  summarise(bkgrd = min(bkgrd)) %>%
  pull(bkgrd)

# GFP intensities
filtered_dat %>%
  filter(CELLS=="mAC-POLR2A" | (CELLS=="OsTIR(F74G)" & SECONDARY_488=="None" & SECONDARY_568=="None" & SECONDARY_647=="None")) %>%
  ggplot(aes(x=Nuclei_intensity_mean_488,group=well_name,col=CELLS)) + 
  geom_density() + 
  geom_vline(xintercept = GFP_bkgrd) +
  facet_wrap(~CELLS,ncol=1,scale="free_y") +
  coord_cartesian(xlim=c(0,800))

mCherry_bkgrd <- filter(filtered_dat,CELLS=="OsTIR(F74G)" & SECONDARY_488=="None" & SECONDARY_568=="None" & SECONDARY_647=="None") %>%
  ungroup() %>%
  group_by(well_name) %>%
  summarise(bkgrd = median(Nuclei_intensity_mean_561),.groups = 'drop') %>%
  summarise(bkgrd = min(bkgrd)) %>%
  pull(bkgrd)

# mCherry intensities
filtered_dat %>%
  filter(CELLS=="mCherry-POLR2A" | (CELLS=="OsTIR(F74G)" & SECONDARY_488=="None" & SECONDARY_568=="None" & SECONDARY_647=="None")) %>%
  ggplot(aes(x=Nuclei_intensity_mean_561,group=well_name,col=CELLS)) + 
  geom_density() + 
  geom_vline(xintercept = mCherry_bkgrd) +
  facet_wrap(~CELLS,ncol=1,scale="free_y") +
  coord_cartesian(xlim=c(0,500))
```

IF background comes from cells stained with secondary antibodies but without primary antibodies. There are several secondary antibodies used that need to be examined separately.

```{r qc_IF}
bkgrd_488 <- filtered_dat %>% 
  filter(CELLS!="mAC-POLR2A") %>%
  filter(PRIMARY_RABBIT=="None" & PRIMARY_RAT=="None") %>%
  group_by(SECONDARY_488,well_name,CELLS) %>%
  summarise(bkgrd=median(Nuclei_intensity_mean_488)) %>%
  group_by(SECONDARY_488,CELLS) %>%
  summarise(bkgrd_488=mean(bkgrd)) %>%
  ungroup()

bkgrd_568 <- filtered_dat %>% 
  filter(CELLS!="mCherry-POLR2A") %>%
  filter(PRIMARY_RABBIT=="None" & PRIMARY_RAT=="None") %>%
  group_by(SECONDARY_568,well_name,CELLS) %>%
  summarise(bkgrd=median(Nuclei_intensity_mean_561)) %>%
  group_by(SECONDARY_568,CELLS) %>%
  summarise(bkgrd_561=mean(bkgrd)) %>%
  ungroup()

bkgrd_647 <- filtered_dat %>% 
  filter(PRIMARY_RABBIT=="None" & PRIMARY_RAT=="None") %>%
  group_by(SECONDARY_647,well_name,CELLS) %>%
  summarise(bkgrd=median(Nuclei_intensity_mean_647)) %>%
  group_by(SECONDARY_647,CELLS) %>%
  summarise(bkgrd_647=mean(bkgrd)) %>%
  ungroup()

# IF over secondary (Rat 488)
filtered_dat %>%
  filter(CELLS!="mAC-POLR2A") %>%
  filter(SECONDARY_488=="Rat" | SECONDARY_488=="None") %>% 
  ggplot(aes(x=SECONDARY_488,y=Nuclei_intensity_mean_488,group=well_name,col=PRIMARY_RAT)) + 
  geom_boxplot(outlier.shape = NA) + 
  facet_wrap(~CELLS,labeller=label_both) + 
  coord_cartesian(ylim=c(0,10000)) +
  geom_hline(yintercept=filter(bkgrd_488,SECONDARY_488=="Rat") %>% pull(bkgrd_488)) +
  ggtitle("Rat 488")

# IF over secondary (Rabbit 488)
filtered_dat %>%
  filter(CELLS!="mAC-POLR2A") %>%
  filter(SECONDARY_488=="Rabbit" | SECONDARY_488=="None") %>% 
  ggplot(aes(x=SECONDARY_488,y=Nuclei_intensity_mean_488,group=well_name,col=PRIMARY_RABBIT)) + 
  geom_boxplot(outlier.shape = NA) + 
  facet_wrap(~CELLS,labeller=label_both) + 
  coord_cartesian(ylim=c(0,5000)) +
  geom_hline(yintercept=filter(bkgrd_488,SECONDARY_488=="Rabbit") %>% pull(bkgrd_488)) +
  ggtitle("Rabbit 488")

# IF over secondary (Rat 568)
filtered_dat %>%
  filter(CELLS!="mCherry-POLR2A") %>%
  filter(SECONDARY_568=="Rat" | SECONDARY_568=="None") %>% 
  ggplot(aes(x=SECONDARY_568,y=Nuclei_intensity_mean_561,group=well_name,col=PRIMARY_RAT)) + 
  geom_boxplot(outlier.shape = NA) + 
  facet_wrap(~CELLS,labeller=label_both) + 
  coord_cartesian(ylim=c(0,10000)) +
  geom_hline(yintercept=filter(bkgrd_568,SECONDARY_568=="Rat") %>% pull(bkgrd_561)) +
  ggtitle("Rat 568")

# IF over secondary (Rat 568)
filtered_dat %>%
  filter(CELLS!="mCherry-POLR2A") %>%
  filter(SECONDARY_568=="Rabbit" | SECONDARY_568=="None") %>% 
  ggplot(aes(x=SECONDARY_568,y=Nuclei_intensity_mean_561,group=well_name,col=PRIMARY_RABBIT)) + 
  geom_boxplot(outlier.shape = NA) + 
  facet_wrap(~CELLS,labeller=label_both) + 
  coord_cartesian(ylim=c(0,5000)) +
  geom_hline(yintercept=filter(bkgrd_568,SECONDARY_568=="Rabbit") %>% pull(bkgrd_561)) +
  ggtitle("Rabbit 568")

# IF over secondary (Rat 647)
filtered_dat %>%
  filter(SECONDARY_647=="Rat" | SECONDARY_647=="None") %>% 
  ggplot(aes(x=SECONDARY_647,y=Nuclei_intensity_mean_647,group=well_name,col=PRIMARY_RAT)) + 
  geom_boxplot(outlier.shape = NA) + 
  facet_wrap(~CELLS,labeller=label_both) + 
  coord_cartesian(ylim=c(0,3000)) +
  geom_hline(yintercept=filter(bkgrd_647,SECONDARY_647=="Rat") %>% pull(bkgrd_647)) +
  ggtitle("Rat 647")

# IF over secondary (Rabbit 647)
filtered_dat %>%
  filter(SECONDARY_647=="Rabbit" | SECONDARY_647=="None") %>% 
  ggplot(aes(x=SECONDARY_647,y=Nuclei_intensity_mean_647,group=well_name,col=PRIMARY_RABBIT)) + 
  geom_boxplot(outlier.shape = NA) + 
  facet_wrap(~CELLS,labeller=label_both) + 
  coord_cartesian(ylim=c(0,3000)) +
  geom_hline(yintercept=filter(bkgrd_647,SECONDARY_647=="Rabbit") %>% pull(bkgrd_647)) +
  ggtitle("Rabbit 647")


```
### Bleed-through

#### mCherry

mCherry and mClover are much weaker than immunofluorescence stains and DAPI. We check bleed-through (channel crosstalk) into these channels by comparing mCherry intensity in the presence and absence of stains in other channels.

```{r mCherry_bleed_through}

# mCherry
mCherry_bleed_rat_488 <- filtered_dat %>%
  filter(CELLS=="mCherry-POLR2A" & SECONDARY_488=="Rat") %>%
  ggplot(aes(x=SECONDARY_488,y=Nuclei_intensity_mean_561,col=PRIMARY_RAT)) + 
  geom_boxplot(outlier.shape = NA) + 
  coord_cartesian(ylim=c(0,500)) +
  ggtitle("Check Rat-488 bleed into mCherry") +
  theme(legend.position = "bottom") +
  geom_hline(yintercept=mCherry_bkgrd)

mCherry_bleed_rabbit_488 <- filtered_dat %>%
  filter(CELLS=="mCherry-POLR2A" & SECONDARY_488=="Rabbit") %>%
  ggplot(aes(x=SECONDARY_488,y=Nuclei_intensity_mean_561,col=PRIMARY_RABBIT)) + 
  geom_boxplot(outlier.shape = NA) + 
  coord_cartesian(ylim=c(0,500)) +
  ggtitle("Check Rabbit-488 bleed into mCherry") +
  theme(legend.position = "bottom") +
  geom_hline(yintercept=mCherry_bkgrd)

mCherry_bleed_rat_488 + mCherry_bleed_rabbit_488

mCherry_bleed_rat_647 <- filtered_dat %>%
  filter(CELLS=="mCherry-POLR2A" & SECONDARY_647=="Rat") %>%
  ggplot(aes(x=SECONDARY_647,y=Nuclei_intensity_mean_561,col=PRIMARY_RAT)) + 
  geom_boxplot(outlier.shape = NA) + 
  facet_wrap(~CELLS,labeller=label_both) +
  coord_cartesian(ylim=c(0,500)) +
  ggtitle("Check Rat-647 bleed into mCherry") +
  theme(legend.position = "bottom") +
  geom_hline(yintercept=mCherry_bkgrd)

mCherry_bleed_rabbit_647 <- filtered_dat %>%
  filter(CELLS=="mCherry-POLR2A" & SECONDARY_647=="Rabbit") %>%
  ggplot(aes(x=SECONDARY_647,y=Nuclei_intensity_mean_561,col=PRIMARY_RABBIT)) + 
  geom_boxplot(outlier.shape = NA) + 
  facet_wrap(~CELLS,labeller=label_both) +
  coord_cartesian(ylim=c(0,500)) +
  ggtitle("Check Rabbit-647 bleed into mCherry") +
  theme(legend.position = "bottom") +
  geom_hline(yintercept=mCherry_bkgrd)

mCherry_bleed_rat_647 + mCherry_bleed_rabbit_647
```

Can see a small amount of bleed-through into mCherry from 647 channel. Quantify this as a fraction of signal.

```{r fraction_bleed}
# calculate percentage bleed-through from rat in 647 channel into mCherry
filtered_dat %>%
  filter(CELLS=="mCherry-POLR2A" & SECONDARY_647=="Rat") %>%
  group_by(well_name,PRIMARY_RAT) %>%
  summarise(Nuclei_intensity_mean_561_well_median = median(Nuclei_intensity_mean_561),.groups = 'drop') %>%
  mutate(Nuclei_intensity_mean_561_well_median_minus_bkgrd = Nuclei_intensity_mean_561_well_median - mCherry_bkgrd) %>%
  group_by(PRIMARY_RAT) %>%
  summarise(m = mean(Nuclei_intensity_mean_561_well_median_minus_bkgrd)) %>%
  pivot_wider(names_from = "PRIMARY_RAT", values_from = "m") %>%
  mutate(fraction_3E8_bleed = (`3E8`-None)/None,
         fraction_3E10_bleed = (`3E10`-None)/None)

# calculate percentage bleed-through from rabbit in 647 channel into mCherry
filtered_dat %>%
  filter(CELLS=="mCherry-POLR2A" & SECONDARY_647=="Rabbit") %>%
  group_by(well_name,PRIMARY_RABBIT) %>%
  summarise(Nuclei_intensity_mean_561_well_median = median(Nuclei_intensity_mean_561),.groups = 'drop') %>%
  mutate(Nuclei_intensity_mean_561_well_median_minus_bkgrd = Nuclei_intensity_mean_561_well_median - mCherry_bkgrd) %>%
  group_by(PRIMARY_RABBIT) %>%
  summarise(m = mean(Nuclei_intensity_mean_561_well_median_minus_bkgrd)) %>%
  pivot_wider(names_from = "PRIMARY_RABBIT", values_from = "m") %>%
  mutate(fraction_D8L4Y_bleed = (`D8L4Y`-None)/None)

```

Negligible for D8L4Y. <5% for 3E10 but almost 10% for 3E8 (the strongest stain). Prefer Rat-488 / Rabbit-647 only for mCherry comparisons.

#### mClover

```{r mClover_bleed_through}

# mClover
mClover_bleed_rat_568 <- filtered_dat %>%
  filter(CELLS=="mAC-POLR2A" & SECONDARY_568=="Rat") %>%
  ggplot(aes(x=SECONDARY_568,y=Nuclei_intensity_mean_488,col=PRIMARY_RAT)) + 
  geom_boxplot(outlier.shape = NA) + 
  coord_cartesian(ylim=c(0,500)) +
  ggtitle("Check Rat-568 bleed into mClover") +
  theme(legend.position = "bottom") +
  geom_hline(yintercept=GFP_bkgrd)

mClover_bleed_rabbit_568 <- filtered_dat %>%
  filter(CELLS=="mAC-POLR2A" & SECONDARY_568=="Rabbit") %>%
  ggplot(aes(x=SECONDARY_568,y=Nuclei_intensity_mean_488,col=PRIMARY_RABBIT)) + 
  geom_boxplot(outlier.shape = NA) + 
  coord_cartesian(ylim=c(0,500)) +
  ggtitle("Check Rabbit-568 bleed into mClover") +
  theme(legend.position = "bottom") +
  geom_hline(yintercept=GFP_bkgrd)

mClover_bleed_rat_568 + mClover_bleed_rabbit_568

```

No bleed-through into mClover.

### Check for antibody cross-reactivity

We also check that rat and rabbit antibodies do not cross-react.

```{r cross_reactivity}

# columns 3 and 16 have combinations without one of the two primary antibodies, compare these to secondary only.
no_primary_cross_reactivity_controls <- filter(filtered_dat,well_col %in% c("03","16")) %>% 
  ungroup() %>%
  distinct(PRIMARY_RAT,PRIMARY_RABBIT,SECONDARY_488,SECONDARY_568,SECONDARY_647) %>%
  filter(!(PRIMARY_RAT=="None" & PRIMARY_RABBIT=="None")) %>%
  inner_join(filtered_dat) %>%
  mutate(primary_combination = paste("Rat",PRIMARY_RAT,"Rabbit",PRIMARY_RABBIT,sep="_")) %>%
  mutate(secondary_combination = paste("488",SECONDARY_488,"568",SECONDARY_568,"647",SECONDARY_647,sep="_"))

secondary_only <- filter(filtered_dat,PRIMARY_RAT=="None" & PRIMARY_RABBIT=="None") %>%
  inner_join(distinct(no_primary_cross_reactivity_controls,SECONDARY_488,SECONDARY_568,SECONDARY_647)) %>%
  mutate(primary_combination = "None") %>%
  mutate(secondary_combination = paste("488",SECONDARY_488,"568",SECONDARY_568,"647",SECONDARY_647,sep="_"))


# 568 rat secondary = same (background) signal when D8L4Y is present or not
# 568 rabbit scondary = same (background) signal when 3E10 or 3E8 is present or not
bind_rows(no_primary_cross_reactivity_controls,secondary_only) %>%
  filter(SECONDARY_568!="None") %>%
  filter(!(SECONDARY_568=="Rat" & PRIMARY_RAT %in% c("3E10","3E8"))) %>%
  filter(!(SECONDARY_568=="Rabbit" & PRIMARY_RABBIT %in% c("D8L4Y"))) %>%
  ggplot(aes(x=secondary_combination,y=Nuclei_intensity_mean_561,col=primary_combination)) +
  geom_boxplot() +
  coord_cartesian(ylim=c(0,300)) +
  theme(axis.text.x=element_text(angle=90,hjust=0.5,vjust=0.5))

# 647 rat secondary = same (background) signal when D8L4Y is present or not
# 647 rabbit scondary = same (background) signal when 3E10 or 3E8 is present or not
bind_rows(no_primary_cross_reactivity_controls,secondary_only) %>%
  filter(SECONDARY_647!="None" & SECONDARY_568=="None") %>%
  filter(!(SECONDARY_647=="Rat" & PRIMARY_RAT %in% c("3E10","3E8"))) %>%
  filter(!(SECONDARY_647=="Rabbit" & PRIMARY_RABBIT %in% c("D8L4Y"))) %>%
  ggplot(aes(x=secondary_combination,y=Nuclei_intensity_mean_647,col=primary_combination)) +
  geom_boxplot() +
  coord_cartesian(ylim=c(0,300)) +
  theme(axis.text.x=element_text(angle=90,hjust=0.5,vjust=0.5))

```

## Covariance analysis

With clean single-cell data, we proceed to co-variance analysis to estimate
$\beta$.

### Data prep

Separate into sub-experiments with different detection channels and cell lines, and rename variables.

```{r slope_data_prep}

bkgrd_488_mClover <- tibble(CELLS="mAC-POLR2A",bkgrd_488=GFP_bkgrd,SECONDARY_488="None")
bkgrd_568_mCherry <- tibble(CELLS="mCherry-POLR2A",bkgrd_561=mCherry_bkgrd,SECONDARY_568="None")

dat_subset <- filtered_dat %>%
  select(-contains("haralick")) %>% 
  left_join(bind_rows(bkgrd_488,bkgrd_488_mClover)) %>%
  left_join(bind_rows(bkgrd_568,bkgrd_568_mCherry)) %>%
  left_join(bkgrd_647) %>%
  mutate(intensity_488=Nuclei_intensity_mean_488-bkgrd_488,
         intensity_561=Nuclei_intensity_mean_561-bkgrd_561,
         intensity_647=Nuclei_intensity_mean_647-bkgrd_647,
         intensity_dapi=Nuclei_intensity_mean_405-DAPI_bkgrd)

# Separate into 4 x pSer2 vs total experiments in the two cell lines with antibody channels switched

dat_mAC_pSer2_561_total_647 <- dat_subset %>%
  filter(CELLS=="mAC-POLR2A" & PRIMARY_RAT=="3E10" & PRIMARY_RABBIT=="D8L4Y" & SECONDARY_568=="Rat" & SECONDARY_647=="Rabbit") %>%
  select(well_name,
         FP=intensity_488,
         pSer2=intensity_561,
         D8L4Y=intensity_647,
         Nuclei_area,
         DAPI=intensity_dapi) %>%
  filter(FP > 0,
         pSer2 > 0,
         D8L4Y > 0,
         DAPI > 0,
         Nuclei_area > 0) %>%
  mutate(
    lFP      = log(FP),
    lpSer2   = log(pSer2),
    lD8L4Y   = log(D8L4Y),
    lArea    = log(Nuclei_area),
    lDAPI    = log(DAPI * Nuclei_area),
    well     = factor(well_name)
  )

dat_mAC_pSer2_647_total_561 <- dat_subset %>%
  filter(CELLS=="mAC-POLR2A" & PRIMARY_RAT=="3E10" & PRIMARY_RABBIT=="D8L4Y" & SECONDARY_568=="Rabbit" & SECONDARY_647=="Rat") %>%
  select(well_name,
         FP=intensity_488,
         pSer2=intensity_647,
         D8L4Y=intensity_561,
         Nuclei_area,
         DAPI=intensity_dapi) %>%
  filter(FP > 0,
         pSer2 > 0,
         D8L4Y > 0,
         DAPI > 0,
         Nuclei_area > 0) %>%
  mutate(
    lFP      = log(FP),
    lpSer2   = log(pSer2),
    lD8L4Y   = log(D8L4Y),
    lArea    = log(Nuclei_area),
    lDAPI    = log(DAPI * Nuclei_area),
    well     = factor(well_name)
  )

dat_mCh_pSer2_488_total_647 <- dat_subset %>%
  filter(CELLS=="mCherry-POLR2A" & PRIMARY_RAT=="3E10" & PRIMARY_RABBIT=="D8L4Y" & SECONDARY_488=="Rat" & SECONDARY_647=="Rabbit") %>%
  select(well_name,
         FP=intensity_561,
         pSer2=intensity_488,
         D8L4Y=intensity_647,
         Nuclei_area,
         DAPI=intensity_dapi) %>%
  filter(FP > 0,
         pSer2 > 0,
         D8L4Y > 0,
         DAPI > 0,
         Nuclei_area > 0) %>%
  mutate(
    lFP      = log(FP),
    lpSer2   = log(pSer2),
    lD8L4Y   = log(D8L4Y),
    lArea    = log(Nuclei_area),
    lDAPI    = log(DAPI * Nuclei_area),
    well     = factor(well_name)
  )

dat_mCh_pSer2_647_total_488 <- dat_subset %>%
  filter(CELLS=="mCherry-POLR2A" & PRIMARY_RAT=="3E10" & PRIMARY_RABBIT=="D8L4Y" & SECONDARY_488=="Rabbit" & SECONDARY_647=="Rat") %>%
  select(well_name,
         FP=intensity_561,
         pSer2=intensity_647,
         D8L4Y=intensity_488,
         Nuclei_area,
         DAPI=intensity_dapi) %>%
  filter(FP > 0,
         pSer2 > 0,
         D8L4Y > 0,
         DAPI > 0,
         Nuclei_area > 0) %>%
  mutate(
    lFP      = log(FP),
    lpSer2   = log(pSer2),
    lD8L4Y   = log(D8L4Y),
    lArea    = log(Nuclei_area),
    lDAPI    = log(DAPI * Nuclei_area),
    well     = factor(well_name)
  )

# Separate into 4 x pSer5 vs total experiments in the two cell lines with antibody channels switched

dat_mAC_pSer5_561_total_647 <- dat_subset %>%
  filter(CELLS=="mAC-POLR2A" & PRIMARY_RAT=="3E8" & PRIMARY_RABBIT=="D8L4Y" & SECONDARY_568=="Rat" & SECONDARY_647=="Rabbit") %>%
  select(well_name,
         FP=intensity_488,
         pSer5=intensity_561,
         D8L4Y=intensity_647,
         Nuclei_area,
         DAPI=intensity_dapi) %>%
  filter(FP > 0,
         pSer5 > 0,
         D8L4Y > 0,
         DAPI > 0,
         Nuclei_area > 0) %>%
  mutate(
    lFP      = log(FP),
    lpSer5   = log(pSer5),
    lD8L4Y   = log(D8L4Y),
    lArea    = log(Nuclei_area),
    lDAPI    = log(DAPI * Nuclei_area),
    well     = factor(well_name)
  )

dat_mAC_pSer5_647_total_561 <- dat_subset %>%
  filter(CELLS=="mAC-POLR2A" & PRIMARY_RAT=="3E8" & PRIMARY_RABBIT=="D8L4Y" & SECONDARY_568=="Rabbit" & SECONDARY_647=="Rat") %>%
  select(well_name,
         FP=intensity_488,
         pSer5=intensity_647,
         D8L4Y=intensity_561,
         Nuclei_area,
         DAPI=intensity_dapi) %>%
  filter(FP > 0,
         pSer5 > 0,
         D8L4Y > 0,
         DAPI > 0,
         Nuclei_area > 0) %>%
  mutate(
    lFP      = log(FP),
    lpSer5   = log(pSer5),
    lD8L4Y   = log(D8L4Y),
    lArea    = log(Nuclei_area),
    lDAPI    = log(DAPI * Nuclei_area),
    well     = factor(well_name)
  )

dat_mCh_pSer5_488_total_647 <- dat_subset %>%
  filter(CELLS=="mCherry-POLR2A" & PRIMARY_RAT=="3E8" & PRIMARY_RABBIT=="D8L4Y" & SECONDARY_488=="Rat" & SECONDARY_647=="Rabbit") %>%
  select(well_name,
         FP=intensity_561,
         pSer5=intensity_488,
         D8L4Y=intensity_647,
         Nuclei_area,
         DAPI=intensity_dapi) %>%
  filter(FP > 0,
         pSer5 > 0,
         D8L4Y > 0,
         DAPI > 0,
         Nuclei_area > 0) %>%
  mutate(
    lFP      = log(FP),
    lpSer5   = log(pSer5),
    lD8L4Y   = log(D8L4Y),
    lArea    = log(Nuclei_area),
    lDAPI    = log(DAPI * Nuclei_area),
    well     = factor(well_name)
  )

dat_mCh_pSer5_647_total_488 <- dat_subset %>%
  filter(CELLS=="mCherry-POLR2A" & PRIMARY_RAT=="3E8" & PRIMARY_RABBIT=="D8L4Y" & SECONDARY_488=="Rabbit" & SECONDARY_647=="Rat") %>%
  select(well_name,
         FP=intensity_561,
         pSer5=intensity_647,
         D8L4Y=intensity_488,
         Nuclei_area,
         DAPI=intensity_dapi) %>%
  filter(FP > 0,
         pSer5 > 0,
         D8L4Y > 0,
         DAPI > 0,
         Nuclei_area > 0) %>%
  mutate(
    lFP      = log(FP),
    lpSer5   = log(pSer5),
    lD8L4Y   = log(D8L4Y),
    lArea    = log(Nuclei_area),
    lDAPI    = log(DAPI * Nuclei_area),
    well     = factor(well_name)
  )

```

## Preliminary analysis

Before getting too deep into this, we plot the data to look for evidence of
buffering. Specifically, we bin cells by total Pol II level (using the 
antibody-free FP readout, with each cell expressed relative to its own well
median to remove well effects) and plot the average pSer2 and total Pol II
antibody (D8L4Y) in each bin.

If pSer2 tracks total Pol II (no buffering) it would rise proportionally with
D8L4Y, along the dashed line, but it actually rises more slowly, suggesting
buffering. 

```{r slope_analysis}

adj_well <- function(y,d) exp(resid(lm(reformulate("well", y), data = d)))

plot_IF_vs_FP <- function(df, title, readout_var = "pSer2") {
  
  intro <- tibble(
    FP = adj_well("lFP", df),
    !!readout_var := adj_well(paste0("l", readout_var), df),
    D8L4Y = adj_well("lD8L4Y", df)
  ) %>%
    mutate(bin = ntile(FP, 18)) %>%
    pivot_longer(all_of(c(readout_var, "D8L4Y")), names_to = "readout", values_to = "y") %>%
    group_by(readout, bin) %>%
    summarise(x  = exp(mean(log(FP))),
              ym = exp(mean(log(y))),
              lo = exp(mean(log(y)) - sd(log(y))),
              hi = exp(mean(log(y)) + sd(log(y))),
              .groups = "drop") %>%
    mutate(readout = factor(readout, levels = c("D8L4Y", readout_var)))
  
  if (readout_var=="pSer5") {
    colour_map <- setNames(c("#306fb7", "#5bba47"), c("D8L4Y", readout_var))
  } else {
     colour_map <- setNames(c("#306fb7", "#cf3827"), c("D8L4Y", readout_var))
  }

  p <- ggplot(intro, aes(x, ym, colour = readout, fill = readout)) +
    geom_abline(slope = 1, intercept = 0, linetype = 2, colour = "grey50") +
    geom_ribbon(aes(ymin = lo, ymax = hi), alpha = 0.1, colour = NA) +
    geom_line() + geom_point(size = 1.8) +
    scale_x_log10(breaks = c(0.7, 1, 1.5)) + 
    scale_y_log10(breaks = c(0.7, 1, 1.5)) +
    scale_colour_manual(values = colour_map) +
    scale_fill_manual(values   = colour_map) +
    coord_cartesian(xlim=c(-0.5,1.7), ylim=c(-0.5,1.7)) +
    labs(x = "total Pol II per cell (FP, relative to well median)",
         y = "signal (relative to well median)", colour = NULL, fill = NULL) +
    theme_bw() + theme(aspect.ratio = 1, panel.grid.minor = element_blank()) +
    ggtitle(title)
  
  return(p)
}

plot_IF_vs_FP(dat_mAC_pSer2_647_total_561,"mAC-POLR2A, pSer2 (647), D8L4Y (561)", readout_var = "pSer2")
plot_IF_vs_FP(dat_mAC_pSer2_561_total_647,"mAC-POLR2A, pSer2 (561), D8L4Y (647)", readout_var = "pSer2")
plot_IF_vs_FP(dat_mCh_pSer2_647_total_488,"mCherry-POLR2A, pSer2 (647), D8L4Y (488)", readout_var = "pSer2")
plot_IF_vs_FP(dat_mCh_pSer2_488_total_647,"mCherry-POLR2A, pSer2 (488), D8L4Y (647)", readout_var = "pSer2")

plot_IF_vs_FP(dat_mAC_pSer5_647_total_561,"mAC-POLR2A, pSer5 (647), D8L4Y (561)", readout_var = "pSer5")
plot_IF_vs_FP(dat_mAC_pSer5_561_total_647,"mAC-POLR2A, pSer5 (561), D8L4Y (647)", readout_var = "pSer5")
plot_IF_vs_FP(dat_mCh_pSer5_647_total_488,"mCherry-POLR2A, pSer5 (647), D8L4Y (488)", readout_var = "pSer5")
plot_IF_vs_FP(dat_mCh_pSer5_488_total_647,"mCherry-POLR2A, pSer5 (488), D8L4Y (647)", readout_var = "pSer5")


```

### Publication plot

Make plots of this for the main and supplementary data.

```{r pub_plots,fig.width=5,fig.height=2}

colour_map <- setNames(c("#cf3827", "#306fb7"), c("Total", "pSer2"))
df <- dat_mAC_pSer2_561_total_647

intro_mAC <- tibble(
  FP = adj_well("lFP", df),
  pSer2 = adj_well("lpSer2", df),
  Total = adj_well("lD8L4Y", df)
) %>%
  mutate(bin = ntile(FP, 20)) %>%
  pivot_longer(all_of(c("pSer2", "Total")), names_to = "readout", values_to = "y") %>%
  group_by(readout, bin) %>%
  summarise(x  = exp(mean(log(FP))),
            ym = exp(mean(log(y))),
            lo = exp(mean(log(y)) - sd(log(y))),
            hi = exp(mean(log(y)) + sd(log(y))),
            .groups = "drop") %>%
  mutate(readout = factor(readout, levels = c("Total", "pSer2"))) %>%
  mutate(cell_line="mAC-POLR2A")

tibble(
  FP = adj_well("lFP", df)) %>%
  mutate(bin = ntile(FP, 20)) %>%
  group_by(bin) %>%
  count() %>%
  ungroup() %>%
  slice_max(n,with_ties = F)

p_mClover <- ggplot(intro_mAC, aes(x, ym, colour = readout, fill = readout)) +
  geom_abline(slope = 1, intercept = 0, linetype = 2, colour = "grey50") +
  geom_ribbon(aes(ymin = lo, ymax = hi), alpha = 0.1, colour = NA) +
  geom_line() + 
  geom_point(size = 1.0) +
  scale_x_log10(breaks = c(0.5,1,2)) + 
  scale_y_log10(breaks = c(0.5,1,2)) +
  scale_colour_manual(values = colour_map, breaks=c("Total", "pSer2"), labels=c("Total\n(D8L4Y)", "pSer2\n(3E10)")) +
  scale_fill_manual(values   = colour_map, breaks=c("Total", "pSer2"), labels=c("Total\n(D8L4Y)", "pSer2\n(3E10)")) +
  coord_cartesian(xlim=c(0.5,2), ylim=c(0.5,2)) +
  labs(x = "mClover intensity",
       y = "IF intensity", colour = NULL, fill = NULL) +
  theme_bw(base_size = 8) + 
  theme(aspect.ratio = 1,
        panel.grid.minor = element_blank(),
        axis.text=element_text(size=8),
        axis.title=element_text(size=8),
        legend.key.size = unit(4,"mm"),
        legend.key.spacing.y = unit(2,"mm"))

p_mClover
ggsave(file.path(PLOTS_DIR, "pSer2_and_Total_vs_FP_mAC.pdf"),width=8,height=3.5,units="cm")


df <- dat_mCh_pSer2_488_total_647

intro_mCh <- tibble(
  FP = adj_well("lFP", df),
  pSer2 = adj_well("lpSer2", df),
  Total = adj_well("lD8L4Y", df)
) %>%
  mutate(bin = ntile(FP, 20)) %>%
  pivot_longer(all_of(c("pSer2", "Total")), names_to = "readout", values_to = "y") %>%
  group_by(readout, bin) %>%
  summarise(x  = exp(mean(log(FP))),
            ym = exp(mean(log(y))),
            lo = exp(mean(log(y)) - sd(log(y))),
            hi = exp(mean(log(y)) + sd(log(y))),
            .groups = "drop") %>%
  mutate(readout = factor(readout, levels = c("Total", "pSer2"))) %>%
  mutate(cell_line="mCherry-POLR2A")

tibble(
  FP = adj_well("lFP", df)) %>%
  mutate(bin = ntile(FP, 20)) %>%
  group_by(bin) %>%
  count() %>%
  ungroup() %>%
  slice_max(n,with_ties = F)

p_mCherry <- ggplot(intro_mCh, aes(x, ym, colour = readout, fill = readout)) +
  geom_abline(slope = 1, intercept = 0, linetype = 2, colour = "grey50") +
  geom_ribbon(aes(ymin = lo, ymax = hi), alpha = 0.1, colour = NA) +
  geom_line() + 
  geom_point(size = 1.0) +
  scale_x_log10(breaks = c(0.5,1,2)) + 
  scale_y_log10(breaks = c(0.5,1,2)) +
  scale_colour_manual(values = colour_map, breaks=c("Total", "pSer2"), labels=c("Total\n(D8L4Y)", "pSer2\n(3E10)")) +
  scale_fill_manual(values   = colour_map, breaks=c("Total", "pSer2"), labels=c("Total\n(D8L4Y)", "pSer2\n(3E10)")) +
  coord_cartesian(xlim=c(0.5,2), ylim=c(0.5,2)) +
  labs(x = "mCherry intensity",
       y = "IF intensity", colour = NULL, fill = NULL) +
  theme_bw(base_size = 8) + 
  theme(aspect.ratio = 1,
        panel.grid.minor = element_blank(),
        axis.text=element_text(size=8),
        axis.title=element_text(size=8),
        legend.key.size = unit(4,"mm"),
        legend.key.spacing.y = unit(2,"mm"))

p_mCherry
ggsave(file.path(PLOTS_DIR, "pSer2_and_Total_vs_FP_mCherry.pdf"),width=8,height=3.5,units="cm")

p_combined <- bind_rows(intro_mCh,intro_mAC) %>%
  mutate(cell_line = fct_relevel(cell_line, "mCherry-POLR2A", "mAC-POLR2A")) %>%   
  ggplot(aes(x, ym, colour = readout, fill = readout)) +
  geom_abline(slope = 1, intercept = 0, linetype = 2, colour = "grey50") +
  geom_ribbon(aes(ymin = lo, ymax = hi), alpha = 0.1, colour = NA) +
  geom_line() + 
  geom_point(size = 1.0) +
  scale_x_log10(breaks = c(0.5,1,2)) + 
  scale_y_log10(breaks = c(0.5,1,2)) +
  scale_colour_manual(values = colour_map, breaks=c("Total", "pSer2"), labels=c("Total\n(D8L4Y)", "pSer2\n(3E10)")) +
  scale_fill_manual(values   = colour_map, breaks=c("Total", "pSer2"), labels=c("Total\n(D8L4Y)", "pSer2\n(3E10)")) +
  coord_cartesian(xlim=c(0.5,2), ylim=c(0.5,2)) +
  labs(x = "Fluorescent protein intensity",
       y = "IF intensity", colour = NULL, fill = NULL) +
  facet_wrap(~cell_line) +
  theme_bw(base_size = 8) + 
  theme(aspect.ratio = 1,
        panel.grid.minor = element_blank(),
        strip.background = element_blank(),
        strip.text = element_text(size=8),
        legend.key.size = unit(4,"mm"),
        legend.key.spacing.y = unit(2,"mm"))

p_combined
ggsave(file.path(PLOTS_DIR, "pSer2_and_Total_vs_FP.pdf"),width=11,height=5,units="cm")

cell_counts <- bind_rows(
  mutate(dat_mCh_pSer2_488_total_647,cell_line="mCherry-POLR2A"),
  mutate(dat_mAC_pSer2_561_total_647,cell_line="mAC-POLR2A")) %>%
  group_by(cell_line,well_name) %>%
  count()

cell_counts %>%
  group_by(cell_line) %>%
  summarise(sum(n))

cell_counts %>%
  distinct(well_name) %>%
  group_by(cell_line) %>%
  count()

```


## Estimating the pSer2-POLR2A vs POLR2A slope from single-cell data

We use mClover or mCherry measured in the same cell as a reference POLR2A level, to
correct for the shared co-variation between the two antibody measurements.

We first correct for well-to-well variability from each log-transformed measurement: 
$P=\log\text{pSer2}$, $T=\log\text{D8L4Y}$, $C=\log\text{mClover}$ (or $\log\text{mCherry}$). We want the 
buffering slope $\beta = \mathrm{d}P/\mathrm{d}\hat T$ taken along genuine
changes in total Pol II ($\hat T$), with the antibody measurement noise removed.

We then regress $T$ on $C$ and keep the (noise-free) part,

$$\hat T = aC, \qquad a=\frac{\operatorname{cov}(T,C)}{\operatorname{var}(C)}.$$

The slope of $P$ on this cleaned axis, $\hat T$, is then an ordinary regression
slope:

$$
\beta=\frac{\operatorname{cov}(P,\hat T)}{\operatorname{var}(\hat T)} =\frac{\operatorname{cov}(P,aC)}{\operatorname{var}(aC)}=\frac{a\,\operatorname{cov}(P,C)}{a^{2}\,\operatorname{var}(C)}.
$$

Substituting in $a$ from above $\operatorname{var}(C)$ cancels, leaving the ratio of two covariances.

$$
\beta=\frac{\operatorname{cov}(P,C)}{\operatorname{cov}(T,C)}
$$

Equivalently, since $\hat T$ can be expressed as a function of $C$, we can use the chain rule,

$$
\beta=\frac{\mathrm{d}P}{\mathrm{d}\hat T}
=\frac{\mathrm{d}P/\mathrm{d}C}{\mathrm{d}\hat T/\mathrm{d}C}
=\frac{\operatorname{cov}(P,C)/\operatorname{var}(C)}{\operatorname{cov}(T,C)/\operatorname{var}(C)}
=\frac{\operatorname{cov}(P,C)}{\operatorname{cov}(T,C)} ,
$$

Our approach to finding $\beta$ is therefore to correct $P$ and $T$ for
well-to-well variation and then calculate the ratio
of their covariances.

```{r estimator}

adj   <- function(d, y, rhs) resid(lm(as.formula(paste(y, "~", rhs)), data = d))
betas <- function(d, rhs, var="lpSer2") {
  
  # adjust Pol II intensity values
  adjP <- adj(d, var, rhs)
  adjC <- adj(d, "lFP", rhs)
  adjT <- adj(d, "lD8L4Y", rhs)
  
  # get covariances for adjusted variables
  covPC <- cov(adjP, adjC)
  covPT <- cov(adjP, adjT)
  covCT <- cov(adjC, adjT)
  
  # return the ratio of covariances
  c(beta_FP = covPC / covCT,       # var slope, fluorescent protein as the reference (antibody-free)
    beta_D8L4Y  = covPT / covCT,   # var slope, D8L4Y as reference (shares IF noise with var)
    beta_naive = covPT / var(adjT),  # var slope from naive linear regression using D8L4Y
    covPC = covPC,
    covCT = covCT,
    covPT = covPT,
    excess = covPT - covPC,
    rel_D8L4Y = covCT / var(adjT),
    rel_FP = covCT / var(adjC))
}

current_data <- dat_mCh_pSer2_488_total_647 

RHS_MIN <- "well"
res  <- betas(current_data, RHS_MIN, var="lpSer2")

```

### What each of these slopes estimates

`betas()` returns three slopes that differ in how they treat measurement noise.

The **naive slope** (`beta_naive`, $\operatorname{cov}(P,T)/\operatorname{var}(T)$)
is an ordinary regression of pSer2 on the D8L4Y antibody. It carries two biases
that act in opposite directions: antibody noise on the $x$-axis dilutes it
downward (regression dilution), while the artefact that D8L4Y and pSer2 share as
co-stained antibodies inflates $\operatorname{cov}(P,T)$ upward. Because these
may partly cancel, `beta_naive` (`r sprintf("%.2f", res["beta_naive"])`) may not be
obviously wrong.

The **antibody-referenced slope** (`beta_D8L4Y`,
$\operatorname{cov}(P,T)/\operatorname{cov}(T,C)$) uses the FP only to remove the
regression dilution from the denominator, so it is the dilution-corrected slope
of pSer2 on D8L4Y. It still contains the shared antibody artefact in its
numerator, which is why it is the largest of the three
(`r sprintf("%.2f", res["beta_D8L4Y"])`).

The **FP-referenced slope** (`beta_FP`,
$\operatorname{cov}(P,C)/\operatorname{cov}(T,C)$) is the one reported. Because
the FP requires no antibody, it carries none of the antibody-borne artefact
shared by the two immunofluorescence channels, so the numerator
$\operatorname{cov}(P,C)$ is free of that artefact; and because the private noise
of the two total-Pol II readouts is uncorrelated, the denominator
$\operatorname{cov}(T,C)$ is free of regression-dilution.
Each channel therefore de-noises the other, and `beta_FP`
(`r sprintf("%.2f", res["beta_FP"])`) is the slope with both antibody-noise
problems removed. It is not free of the per-cell imaging factor common to all
fluorescence channels, which is treated separately below.
It is our estimate of $\beta$, the buffering slope defined
above; we report `beta_FP` throughout and refer to the quantity it estimates as
$\beta$. The gap between `beta_D8L4Y` and
`beta_FP` is the shared antibody artefact, and it corresponds to the *excess* covariance scaled by $\operatorname{cov}(T,C)$
 in the validation table below.

`rel_FP` and `rel_D8L4Y` are not slopes but the reproducible fractions
$R_C$ and $R_T$ of the two total readouts, defined in the reproducible-fraction
section below.

### Check correspondence of fluorescent protein and D8L4Y

On the log scale, a constant difference in absolute intensity
between the two channels is absorbed into the intercept and leaves the
slope unchanged. We therefore don't need to normalise intensities across
channels before calculating $\beta$.

A within-well robust regression of `lD8L4Y` on `lFP` gives slope
`r {lin <- MASS::rlm(lD8L4Y ~ lFP + well, data = current_data, method = "MM", maxit = 200); sprintf("%.3f", coef(lin)["lFP"])}`.
This is not a test of whether the two channels load equally on true Pol II:
regressing one noisy readout on another attenuates the slope by the noise in the
predictor, and this value is essentially $R_C$ (below). It shows only that the
two totals track each other closely and monotonically on the log scale. Equal
loading is not required in any case, because a constant loading factor cancels
from the ratio of covariances.

```{r proportionality}
ggplot(current_data, aes(lFP, lD8L4Y)) +
  geom_hex(bins = 70) +
  geom_smooth(method = MASS::rlm, method.args = list(method = "MM", maxit = 200), colour = "red") +
  labs(x = "log FP-POLR2A", y = "log D8L4Y (IF)") + theme_minimal()
```

### Validate that D8L4Y shares extra correlated measurement noise with pSer2

```{r diag-table, echo=FALSE}
tibble(
  quantity = c("slope (FP as reference)", "slope (D8L4Y as reference)",
               "slope (naive fit)", "excess D8L4Y-pSer2 covariance (shared noise)"),
  value = c(res["beta_FP"], res["beta_D8L4Y"], res["beta_naive"],
            res["excess"])
) %>% knitr::kable(digits = 3)
```


If the two antibodies shared no artefact, pSer2 should co-vary equally with FP
and with D8L4Y. As expected, D8L4Y shares extra covariance with pSer2 (greater
slope). This *excess* (`r sprintf("%.3f", res["excess"])`) accounts for
`r sprintf("%.0f%%", 100*res["excess"]/res["covPT"])` of the observed covariance
between pSer2 and D8L4Y (`r sprintf("%.3f", res["covPT"])`), likely because
D8L4Y and pSer2 are both POLR2A antibodies stained on the same fixed cells and
therefore share IF artefacts.

### Reproducible fraction of each total-Pol II readout

We can model the two readouts (fluorescent protein and immunofluorescence) as the
same shared signal $S$ (true total Pol II, common to both) plus noise specific to
each measurement channel, $e_C, e_T$,

$$
C = S + e_C,\qquad T = S + e_T,
\qquad e_C \perp e_T,\quad e_C,e_T \perp S .
$$

Independence is only required between the two total readouts, and it holds
because the FP is not measured using an antibody: $e_T$ still contains the antibody artefact
shared with pSer2 quantified above, but that confounding artefact is absent from $C$.

The *reproducible fraction* of a readout is the share of its variance
that is real shared signal, e.g. $R_C = \operatorname{var}(S)/\operatorname{var}(C)$.

Because the measurement noises are (by definition) independent, they disappear
from the covariance, while each variance keeps its own noise,

$$
\operatorname{cov}(C,T)=\operatorname{var}(S),\qquad
\operatorname{var}(C)=\operatorname{var}(S)+\operatorname{var}(e_C),\qquad
\operatorname{var}(T)=\operatorname{var}(S)+\operatorname{var}(e_T).
$$

Substituting gives the estimator used above,

$$
R_C=\frac{\operatorname{var}(S)}{\operatorname{var}(C)}=\frac{\operatorname{cov}(C,T)}{\operatorname{var}(C)},
\qquad
R_T=\frac{\operatorname{cov}(C,T)}{\operatorname{var}(T)} .
$$

The numerator is the same shared signal for both so the readout carrying less
channel-specific noise has the larger reproducible fraction.

In these data, $R_C =$ `r sprintf("%.3f", res["rel_FP"])` versus
$R_T =$ `r sprintf("%.3f", res["rel_D8L4Y"])`: `r sprintf("%.0f%%", 100*res["rel_FP"])`
of the FP's variance is the component it shares with D8L4Y, versus
`r sprintf("%.0f%%", 100*res["rel_D8L4Y"])` for D8L4Y. The FP is therefore the
cleaner of the two readouts, though only modestly so in this cell line. Neither
fraction affects $\beta$ — channel-private noise cancels from the covariances —
so these are reported as a measure of readout quality, not as a correction.

Because $S$ combines the shared imaging factor in with true Pol II, $R$ measures
reproducibility relative to that common signal, not the underlying biological
differences between cells.

### Confidence interval for $\beta$

We already estimated $\beta = \mathrm{d}P / \mathrm{d}\hat T$ above, and now use
bootstrapping to estimate a confidence interval.We sample all wells with
replacement and take the 2.5% - 97.5% range as a 95% confidence interval for 
$\beta$. We then examine whether $\beta$ estimates are reproducible across all
wells.

```{r boot}
bootstrap_reps <- 500
set.seed(1)
wells <- levels(current_data$well)
boot <- replicate(bootstrap_reps, {
  samp <- sample(wells, size = length(wells), replace = TRUE)
  d <- imap_dfr(samp, ~ filter(current_data, well == .x) %>%
                  mutate(well = factor(paste0(.x, "_", .y))))
  betas(d, RHS_MIN)["beta_FP"]
})
ci <- quantile(boot, c(.025, .975))
```

The buffering slope, corrected for antibody measurement noise using the FP
reference is

$$\hat\beta = `r sprintf("%.2f", res["beta_FP"])`
\quad (\text{95% CI } `r sprintf("%.2f", ci[1])`\text{–}`r sprintf("%.2f", ci[2])`,
\text{ well-cluster bootstrap}).$$

Because this slope is below 1, pSer2 is buffered against total Pol II variation
in unperturbed cells. Importantly, because the residual shared-imaging factor
biases this estimate upward, this is an upper bound.

### Decomposing $\beta$

Even with measurement noise removed, $\beta$ is not a single
mechanistic constant. Cell-to-cell variation in total Pol II has an *intrinsic*
component (fluctuations in Pol II abundance with the rest of the cell held
fixed) and an *extrinsic* component (e.g. cell size, cycle stage, global biosynthetic
capacity). pSer2 measures only the total, so the measured slope is a variance-weighted average of the
two responses,

$$\beta = w\,\beta_{\text{int}} + (1-w)\,\beta_{\text{ext}},$$

where $w$ is the fraction of true (non-measurement) Pol II variance that is
intrinsic. Extrinsic variation tends to be close to proportional
($\beta_{\text{ext}}\approx 1$, because everything scales together), so it pulls
$\beta$ upward and a sub-1 slope is a conservative reading of the
intrinsic buffering.

```{r conditioning}
# add cell size (log area) and cell cycle (log integrated DAPI) to the well adjustment
res_conditioned <- betas(current_data, "well + lArea + lDAPI", var = "lpSer2")

# Paired well-cluster bootstrap for the DIFFERENCE between the two slopes.
# Both are computed on the same resampled cells, so well-to-well variation
# common to the two models cancels and the difference is paired. 
set.seed(2)
bootstrap_reps <- 500
wells_c   <- levels(current_data$well)

dboot <- replicate(bootstrap_reps, {
  samp <- sample(wells_c, size = length(wells_c), replace = TRUE)
  d <- imap_dfr(samp, ~ filter(current_data, well == .x) %>%
                  mutate(well = factor(paste0(.x, "_", .y))))
  unname(betas(d, RHS_MIN)["beta_FP"] -
         betas(d, "well + lArea + lDAPI")["beta_FP"])
})

diff_hat <- unname(res["beta_FP"] - res_conditioned["beta_FP"])
ci_diff  <- quantile(dboot, c(.025, .975))

c(well_only       = unname(res["beta_FP"]),
  plus_size_cycle = unname(res_conditioned["beta_FP"]),
  difference      = diff_hat,
  diff_lo         = unname(ci_diff[1]),
  diff_hi         = unname(ci_diff[2]))
```

To test whether the sub-proportionality is this extrinsic scaling, we add cell
size ($\log$ area) and cell cycle ($\log$ integrated DAPI) to the adjustment
alongside `well`, bootstrapping both slopes on the same well resamples so that
the difference is paired.

Conditioning lowered the slope from `r sprintf("%.2f", res["beta_FP"])` to
`r sprintf("%.2f", res_conditioned["beta_FP"])` (difference
`r sprintf("%.3f", diff_hat)`, 95% CI `r sprintf("%.3f", ci_diff[1])`–`r sprintf("%.3f", ci_diff[2])`,
paired well bootstrap). The slope remains far below 1, so the buffering is not an
artefact of larger or cycling cells dragging both channels up together.

We do not interpret the size of the shift: we do not know how much of the
extrinsic variance these two proxies capture, and $\log$ DAPI is not a pure
extrinsic proxy, because POLR2A expression is itself cell-cycle-dependent, so
conditioning on DNA content also removes genuine Pol II variation.


### Robustness across wells, cell lines, and imaging setups

We can check on a per-well basis whether $\beta$ is consistently below 1 for the different imaging setups.

```{r per_well,fig.height=3,fig.width=6}

per_well_pSer2 <- bind_rows(
  mutate(dat_mAC_pSer2_647_total_561,cell_line="mAC-POLR2A",detection_channel="Far red"),
  mutate(dat_mAC_pSer2_561_total_647,cell_line="mAC-POLR2A",detection_channel="Red"),
  mutate(dat_mCh_pSer2_647_total_488,cell_line="mCherry-POLR2A",detection_channel="Far red"),
  mutate(dat_mCh_pSer2_488_total_647,cell_line="mCherry-POLR2A",detection_channel="Green")) %>%
  group_by(well,cell_line,detection_channel) %>%
  summarise(beta_pSer2 = betas(pick(everything()), "1", var="lpSer2")["beta_FP"],
            n = n(), .groups = "drop") 
per_well_pSer2 %>% knitr::kable(digits = 3)

per_well_pSer2 %>%
  ggplot(aes(x=cell_line,y=beta_pSer2,col=detection_channel)) +
  geom_boxplot() + 
  geom_jitter(position = position_jitterdodge()) + 
  geom_hline(yintercept = 1, linetype = 2) +
  scale_y_continuous(name = expression(beta), limits = c(0, 1.2), breaks = c(0, 0.5, 1)) +
  theme_bw() +
  theme(panel.grid.major.x = element_blank(),
        axis.title.x = element_blank())

per_well_pSer5 <- bind_rows(
  mutate(dat_mAC_pSer5_647_total_561,cell_line="mAC-POLR2A",detection_channel="Far red"),
  mutate(dat_mAC_pSer5_561_total_647,cell_line="mAC-POLR2A",detection_channel="Red"),
  mutate(dat_mCh_pSer5_647_total_488,cell_line="mCherry-POLR2A",detection_channel="Far red"),
  mutate(dat_mCh_pSer5_488_total_647,cell_line="mCherry-POLR2A",detection_channel="Green")) %>%
  group_by(well,cell_line,detection_channel) %>%
  summarise(beta_pSer5 = betas(pick(everything()), "1", var="lpSer5")["beta_FP"],
            n = n(), .groups = "drop") 
per_well_pSer5 %>% knitr::kable(digits = 3)

per_well_pSer5 %>%
  ggplot(aes(x=cell_line,y=beta_pSer5,col=detection_channel)) +
  geom_boxplot() + 
  geom_jitter(position = position_jitterdodge()) + 
  geom_hline(yintercept = 1, linetype = 2) +
  scale_y_continuous(name = expression(beta), limits = c(0, 1.2), breaks = c(0, 0.5, 1)) +
  theme_bw() +
  theme(panel.grid.major.x = element_blank(),
        axis.title.x = element_blank())

# Final plot omitting Far-Red 3E10 and 3E8 due to bleedthrough

pSer2_plot <- per_well_pSer2 %>%
  filter(!(cell_line=="mCherry-POLR2A" & detection_channel=="Far red")) %>%
  mutate(cell_line = fct_relevel(cell_line, "mCherry-POLR2A","mAC-POLR2A")) %>%
  ggplot(aes(x=cell_line,y=beta_pSer2)) +
  geom_boxplot(alpha=0.2,col="#306fb7",fill="#306fb7",linewidth=0.25, whisker.linewidth = 0.5,width=0.6) + 
  #geom_jitter(width=0.2,alpha=0.5,size=1,colour="#306fb7") + 
  geom_hline(yintercept = 1, linetype = 2) +
  scale_x_discrete(breaks=c("mCherry-POLR2A","mAC-POLR2A"),labels=c("mCherry-\nPOLR2A","mAC-\nPOLR2A")) +
  scale_y_continuous(name = expression(beta), limits = c(0, 1.1), breaks = c(0, 0.5, 1)) +
  theme_bw(base_size = 8) +
  theme(panel.grid.major.x = element_blank(),
        axis.title.x = element_blank(),
        axis.text.y = element_text(size=7),
        axis.text.x = element_text(size=7,angle=90,hjust=1,vjust=0.5),
        legend.position = "none")

pSer2_plot
ggsave(file.path(PLOTS_DIR, "beta_pSer2_by_well.pdf"),width=2.5,height=3.5,units="cm")


pSer5_pSer2_plot <- bind_rows(
  per_well_pSer5 %>% rename(beta=beta_pSer5) %>% mutate(antibody="pSer5"),
  per_well_pSer2 %>% rename(beta=beta_pSer2) %>% mutate(antibody="pSer2")) %>%
  mutate(antibody=factor(antibody,levels=c("pSer5","pSer2"))) %>%
  filter(!(cell_line=="mCherry-POLR2A" & detection_channel=="Far red")) %>%
  mutate(cell_line = fct_relevel(cell_line, "mCherry-POLR2A","mAC-POLR2A")) %>%
  ggplot(aes(x=cell_line,y=beta,fill=antibody,col=antibody)) +
  geom_boxplot(alpha=0.2,linewidth=0.25, whisker.linewidth = 0.5,width=0.6) + 
  #geom_jitter(width=0.2,alpha=0.5,size=0.5) + 
  geom_hline(yintercept = 1, linetype = 2) +
  scale_fill_manual(values=c("#5bba47","#306fb7")) +
  scale_colour_manual(values=c("#5bba47","#306fb7")) +
  scale_x_discrete(breaks=c("mAC-POLR2A","mCherry-POLR2A"),labels=c("mAC-\nPOLR2A","mCherry-\nPOLR2A")) +
  scale_y_continuous(name = expression(beta), limits = c(0, 1.1), breaks = c(0, 0.5, 1)) +
  theme_bw(base_size = 8) +
  theme(panel.grid.major.x = element_blank(),
        axis.title.x = element_blank(),
        axis.text.y = element_text(size=7),
        axis.text.x = element_text(size=7,angle=90,hjust=1,vjust=0.5),
        legend.title=element_blank(),
        legend.key.size=unit(3,"mm"),
        legend.key.spacing=unit(1,"mm"))

pSer5_pSer2_plot
ggsave(file.path(PLOTS_DIR, "beta_pSer2_pSer5_by_well.pdf"),width=5.0,height=3.5,units="cm")

pSer5_pSer2_plot_alt <- bind_rows(
  per_well_pSer5 %>% rename(beta=beta_pSer5) %>% mutate(antibody="pSer5"),
  per_well_pSer2 %>% rename(beta=beta_pSer2) %>% mutate(antibody="pSer2")) %>%
  mutate(antibody=factor(antibody,levels=c("pSer5","pSer2"))) %>%
  filter(!(cell_line=="mCherry-POLR2A" & detection_channel=="Far red")) %>%
  mutate(config = case_when(cell_line=="mAC-POLR2A" & detection_channel=="Red" ~ "mAC-POLR2A_total_647_phospho_568",
                            cell_line=="mAC-POLR2A" & detection_channel=="Far red" ~ "mAC-POLR2A_total_568_phospho_647",
                            cell_line=="mCherry-POLR2A" & detection_channel=="Green" ~ "mCherry-POLR2A_total_647_phospho_488",
                            TRUE ~ NA)) %>%
  mutate(config = fct_relevel(config, "mCherry-POLR2A_total_647_phospho_488",
                              "mAC-POLR2A_total_647_phospho_568",
                              "mAC-POLR2A_total_568_phospho_647")) %>%
  ggplot(aes(x=config,y=beta,fill=antibody,col=antibody)) +
  geom_boxplot(alpha=0.2,linewidth=0.25, whisker.linewidth = 0.5,width=0.6) + 
  geom_hline(yintercept = 1, linetype = 2) +
  scale_fill_manual(values=c("#5bba47","#306fb7")) +
  scale_colour_manual(values=c("#5bba47","#306fb7")) +
  scale_x_discrete(breaks=c("mCherry-POLR2A_total_647_phospho_488",
                            "mAC-POLR2A_total_647_phospho_568",
                            "mAC-POLR2A_total_568_phospho_647"),
                   labels=c("mCherry-POLR2A\n(total = Alexa647,\nphospho = Alexa488)",
                            "mAC-POLR2A\n(total = Alexa647,\nphospho = Alexa568)",
                            "mAC-POLR2A\n(total = Alexa568,\nphospho = Alexa647)")) +
  scale_y_continuous(name = expression(beta), limits = c(0, 1.1), breaks = c(0, 0.5, 1)) +
  theme_bw(base_size = 8) +
  theme(panel.grid.major.x = element_blank(),
        axis.title.x = element_blank(),
        axis.text.y = element_text(size=7),
        axis.text.x = element_text(size=7,angle=90,hjust=1,vjust=0.5),
        legend.title=element_blank(),
        legend.key.size=unit(3,"mm"),
        legend.key.spacing=unit(1,"mm"))

pSer5_pSer2_plot_alt
ggsave(file.path(PLOTS_DIR, "beta_pSer2_pSer5_by_well_by_config.pdf"),width=6.5,height=5,units="cm")

```

All wells fall below 1 range (`r sprintf("%.2f", min(per_well_pSer2[per_well_pSer2$detection_channel!="Far red",]$beta_pSer2))`–`r sprintf("%.2f", max(per_well_pSer2[per_well_pSer2$detection_channel!="Far red",]$beta_pSer2))`), so the result is reproducible across replicates and cell lines.

### Validate difference between pSer2 and pSer5

```{r beta_by_mark, message=FALSE}

mark_reps <- 500

beta_boot <- function(d, var, reps = mark_reps, seed = 1) {
  wells <- levels(droplevels(factor(d$well)))
  set.seed(seed)
  replicate(reps, {
    samp <- sample(wells, length(wells), replace = TRUE)
    dd <- imap_dfr(samp, ~ filter(d, well == .x) %>%
                     mutate(well = factor(paste0(.x, "_", .y))))
    unname(betas(dd, RHS_MIN, var = var)["beta_FP"])
  })
}

main_cfg <- list(
  `mCherry-POLR2A` = list(pSer2 = dat_mCh_pSer2_488_total_647,
                          pSer5 = dat_mCh_pSer5_488_total_647),
  `mAC-POLR2A`     = list(pSer2 = dat_mAC_pSer2_561_total_647,
                          pSer5 = dat_mAC_pSer5_561_total_647)
)

# pSer2 and pSer5 are stained in separate wells, so the two bootstraps are
# independent and their draws can be differenced directly to give a CI on
# beta_pSer5 - beta_pSer2.
beta_marks <- imap_dfr(main_cfg, function(cfg, line) {
  est2 <- unname(betas(cfg$pSer2, RHS_MIN, var = "lpSer2")["beta_FP"])
  est5 <- unname(betas(cfg$pSer5, RHS_MIN, var = "lpSer5")["beta_FP"])
  b2   <- beta_boot(cfg$pSer2, "lpSer2", seed = 1)
  b5   <- beta_boot(cfg$pSer5, "lpSer5", seed = 2)
  dd   <- b5 - b2
  tibble(
    cell_line  = line,
    beta_pSer2 = est2,
    pSer2_lo   = unname(quantile(b2, .025)), pSer2_hi = unname(quantile(b2, .975)),
    beta_pSer5 = est5,
    pSer5_lo   = unname(quantile(b5, .025)), pSer5_hi = unname(quantile(b5, .975)),
    difference = est5 - est2,
    diff_lo    = unname(quantile(dd, .025)), diff_hi = unname(quantile(dd, .975)),
    p_boot     = mean(dd <= 0)
  )
})

beta_marks %>% knitr::kable(digits = 3)
```

## Cell-to-cell variability

We have a look at raw cell-to-cell variability to see if this can be easily visualised as "buffering" in the sense of reducing variance.

```{r variability}

library(rstatix)

paired_long <- function(df, source_label, chans = c("D8L4Y", "pSer2")) {
  out <- df %>%
    group_by(well) %>%
    summarise(across(all_of(chans),
                     c(gcv = function(x) sqrt(exp(sd(log(x))^2) - 1))),
              .groups = "drop") %>%
    rename_with(~ sub("_gcv$", "", .x), ends_with("_gcv")) %>%
    drop_na(all_of(chans)) %>%                 # keep only wells with BOTH channels
    mutate(source = source_label)

  message(source_label, ": ", nrow(out), " wells with complete pairs")

  out %>%
    pivot_longer(cols = all_of(chans), names_to = "channel", values_to = "gcv") %>%
    arrange(source, channel, well)
}

gcv_all <- bind_rows(
  paired_long(dat_mCh_pSer2_488_total_647, "mCh_pSer2_488_total_647"),
  paired_long(dat_mAC_pSer2_647_total_561, "mAC_pSer2_647_total_561"),
  paired_long(dat_mAC_pSer2_561_total_647, "mAC_pSer2_561_total_647")
) %>%
  mutate(channel = factor(
    case_match(channel, "D8L4Y" ~ "Total", "pSer2" ~ "pSer2"),
    levels=c("Total","pSer2")))

# within each source, both channels must have identical well sets
stopifnot(
  gcv_all %>%
    group_by(source, channel) %>%
    summarise(wells = list(sort(well)), .groups = "drop") %>%
    group_by(source) %>%
    summarise(ok = n_distinct(wells) == 1, .groups = "drop") %>%
    pull(ok) %>% all()
)

# stats: D8L4Y vs pSer2 within each dataset
wilcox_res <- gcv_all %>%
  group_by(source) %>%
  pairwise_wilcox_test(gcv ~ channel, paired = TRUE)

ttest_res <- gcv_all %>%
  group_by(source) %>%
  pairwise_t_test(gcv ~ channel, paired = TRUE) %>%
  select(source, group1, group2, n1, n2, statistic, df, p)

effect_res <- gcv_all %>%
  group_by(source) %>%
  cohens_d(gcv ~ channel, paired = TRUE) %>%
  select(source, effsize, magnitude)

# combined summary: one row per dataset (t-test + effect size) 
summary_tbl <- left_join(ttest_res, effect_res, by = "source")

wilcox_res
summary_tbl
library(ggpubr)

colour_map <- setNames(c("#cf3827", "#306fb7"), c("Total", "pSer2"))

# Combined (all three, faceted)
gcv_stat <- gcv_all %>%
  group_by(source) %>%
  pairwise_wilcox_test(gcv ~ channel, paired = TRUE) %>%
  add_significance("p") %>%
  add_xy_position(x = "channel")

gcv_plot <- gcv_all %>%
  ggplot(aes(x = channel, y = gcv)) +
  geom_boxplot(aes(fill = channel), alpha = 0.2, colour = "black",
               width = 0.5, outlier.shape = NA) +
  geom_line(aes(group = well), colour = "grey70", linewidth = 0.3) +
  geom_point(aes(colour = channel), alpha = 0.7, size = 1) +
  stat_pvalue_manual(gcv_stat,
                     label = "p.signif",
                     tip.length = 0.01,
                     bracket.size = 0.4,
                     label.size = 3) +
  scale_fill_manual(values = colour_map) +
  scale_colour_manual(values = colour_map) +
  facet_wrap(~ source, labeller = as_labeller(c(
    "mCh_pSer2_488_total_647" = "mCherry-POLR2A",
    "mAC_pSer2_647_total_561" = "mAC-POLR2A 2",
    "mAC_pSer2_561_total_647" = "mAC-POLR2A 1"))) +
  scale_y_continuous(name = "Geometric CV",
                     expand = expansion(mult = c(0.03, 0.10))) +
  theme_bw(base_size = 8) +
  theme(panel.grid.minor = element_blank(),
        panel.grid.major.x = element_blank(),
        axis.title.x = element_blank(),
        legend.position = "none")
gcv_plot
ggsave(file.path(PLOTS_DIR, "gcv_plot_all.pdf"),width=6,height=4,units="cm")

# mCherry only (single panel, no facet)
mch_dat <- gcv_all %>% filter(source == "mCh_pSer2_488_total_647")

mch_stat <- mch_dat %>%
  pairwise_wilcox_test(gcv ~ channel, paired = TRUE) %>%
  add_significance("p") %>%
  add_xy_position(x = "channel")

gcv_plot_mCherry <- mch_dat %>%
  ggplot(aes(x = channel, y = gcv)) +
  geom_boxplot(aes(fill = channel), alpha = 0.2, colour = "black",
               width = 0.5, outlier.shape = NA) +
  geom_line(aes(group = well), colour = "grey70", linewidth = 0.3) +
  geom_point(aes(colour = channel), alpha = 0.7, size = 1) +
  stat_pvalue_manual(mch_stat,
                     label = "p.signif",
                     tip.length = 0.01,
                     bracket.size = 0.4,
                     label.size = 3) +
  scale_fill_manual(values = colour_map) +
  scale_colour_manual(values = colour_map) +
  scale_y_continuous(name = "Geometric C.V.",
                     expand = expansion(mult = c(0.1, 0.10))) +
  theme_bw(base_size = 8) +
  theme(panel.grid.minor = element_blank(),
        panel.grid.major.x = element_blank(),
        axis.title.x = element_blank(),
        axis.text = element_text(size=8),
        legend.position = "none")
gcv_plot_mCherry

ggsave(file.path(PLOTS_DIR, "gcv_plot_mCherry.pdf"),width=3,height=3.15,units="cm")

# mAC only (two panels, faceted) 
mac_dat <- gcv_all %>% filter(str_starts(source, "mAC"))

mac_stat <- mac_dat %>%
  group_by(source) %>%
  pairwise_wilcox_test(gcv ~ channel, paired = TRUE) %>%
  add_significance("p") %>%
  add_xy_position(x = "channel")

gcv_plot_mAC <- mac_dat %>%
  ggplot(aes(x = channel, y = gcv)) +
  geom_boxplot(aes(fill = channel), alpha = 0.2, colour = "black",
               width = 0.5, outlier.shape = NA) +
  geom_line(aes(group = well), colour = "grey70", linewidth = 0.3) +
  geom_point(aes(colour = channel), alpha = 0.7, size = 1) +
  stat_pvalue_manual(mac_stat,
                     label = "p.signif",
                     tip.length = 0.01,
                     bracket.size = 0.4,
                     label.size = 3) +
  scale_fill_manual(values = colour_map) +
  scale_colour_manual(values = colour_map) +
  facet_wrap(~ source, labeller = as_labeller(c(
    "mAC_pSer2_647_total_561" = "mAC-POLR2A 2",
    "mAC_pSer2_561_total_647" = "mAC-POLR2A 1"))) +
  scale_y_continuous(name = "Geometric C.V.",
                     expand = expansion(mult = c(0.03, 0.10))) +
  theme_bw(base_size = 8) +
  theme(panel.grid.minor = element_blank(),
        panel.grid.major.x = element_blank(),
        axis.title.x = element_blank(),
        legend.position = "none")
gcv_plot_mAC
ggsave(file.path(PLOTS_DIR, "gcv_plot_mAC.pdf"),width=5,height=4,units="cm")


```

Choose a representative well and plot the distribution

```{r check_distribution}

gcv_all %>%
  filter(channel %in% c("Total", "pSer2")) %>%
  group_by(source, channel) %>%
  mutate(channel_median = median(gcv),
         channel_mad = mad(gcv)) %>%
  ungroup() %>%
  mutate(z = abs(gcv - channel_median) / channel_mad) %>%
  select(well, source, channel, z) %>%
  pivot_wider(names_from = channel, values_from = z, names_prefix = "z_") %>%
  filter(!is.na(z_Total), !is.na(z_pSer2)) %>%
  mutate(combined_dev = z_Total + z_pSer2) %>%
  group_by(source) %>%
  slice_min(combined_dev, n = 1) %>%
  ungroup()

highlight_well <- "L07" 

mch_dat %>%
  ggplot(aes(x = channel, y = gcv)) +
  geom_boxplot(aes(fill = channel), alpha = 0.2, colour = "black",
               width = 0.5, outlier.shape = NA) +
  geom_line(aes(group = well), colour = "grey70", linewidth = 0.3) +
  geom_point(aes(colour = channel), alpha = 0.7, size = 1) +
  # highlighted well on top
  geom_line(data = mch_dat %>% filter(well == highlight_well),
            aes(group = well), colour = "black", linewidth = 0.8) +
  geom_point(data = mch_dat %>% filter(well == highlight_well),
             aes(colour = channel), size = 2.2, shape = 21, fill = "white", stroke = 1) +
  stat_pvalue_manual(mch_stat,
                     label = "p.signif",
                     tip.length = 0.01,
                     bracket.size = 0.4,
                     label.size = 3) +
  scale_fill_manual(values = colour_map) +
  scale_colour_manual(values = colour_map) +
  scale_y_continuous(name = "Geometric C.V.",
                     expand = expansion(mult = c(0.1, 0.10))) +
  theme_bw(base_size = 8) +
  theme(panel.grid.minor = element_blank(),
        panel.grid.major.x = element_blank(),
        axis.title.x = element_blank(),
        legend.position = "none")


p <- dat_mCh_pSer2_488_total_647 %>%
  group_by(well_name) %>%
  mutate(pSer2 = pSer2/median(pSer2), Total = D8L4Y/median(D8L4Y)) %>%
  select(well_name, pSer2, Total) %>%
  pivot_longer(-well_name) %>%
  ggplot(aes(x = value, fill = name, color = name)) +
  geom_histogram(bins = 300, alpha = 0.4, position = "identity", color = NA) +
  geom_step(stat = "bin", bins = 300, position = "identity",
            direction = "mid", pad = TRUE, linewidth = 0.3) +
  scale_colour_manual(values = colour_map) +
  scale_fill_manual(values = colour_map) +
  theme_bw(8) +
  scale_y_continuous(name = "Number of cells", breaks = c(0, 1000, 2000)) +
  scale_x_log10(name = "Nuclear intensity") +
  coord_cartesian(xlim = c(0.5, 2)) +
  theme(panel.grid = element_blank(),
        legend.position = "none",
        axis.text=element_text(size=8))

p
ggsave(file.path(PLOTS_DIR, "pSer2_and_D8L4Y_single_cell_distribution.pdf"),
       plot = p, width = 3, height = 3.5, units = "cm")
  
```

## Summary and caveats

- pSer2 scales sub-proportionally with total Pol II in unperturbed cells:
$\hat\beta \approx `r sprintf("%.2f", res["beta_FP"])`$
(CI `r sprintf("%.2f", ci[1])`–`r sprintf("%.2f", ci[2])`), consistent across all wells.

- The genetic mClover or mCherry readout is essential: about
`r sprintf("%.0f%%", 100*res["excess"]/res["covPT"])` of the D8L4Y antibody's covariance with
pSer2 is technical artifact shared between the two antibodies, so any analysis resting on the
two antibodies alone (a naive straight-line fit, or treating total Pol II as a hidden quantity
measured only by the two antibodies) is confounded and would over- or under-state buffering.
- $\beta$ is a variance-weighted mix of an intrinsic and an extrinsic response,
$\beta = w\,\beta_{\text{int}} + (1-w)\,\beta_{\text{ext}}$. Conditioning on
measured cell-size and cell-cycle proxies only lowered $\beta$ slightly,
so the sub-proportionality is not entirely a
size/cycle artefact. The intrinsic slope is not identified from unperturbed data: the value
should be read as an upper bound on intrinsic buffering.

- The estimate is conservative: the shared-imaging factor and the extrinsic mixing both bias
$\beta$ toward 1, and the scale is validated by the near slope-1 mClover–D8L4Y
relationship. The reported interval is sampling precision from the well bootstrap on one
cell line and channel configuration; it does not include these one-sided systematic biases,
which all imply the true buffering is at least this strong.

