---
title: "mAC_mACmChPOLR2A_TotalPolII"
output: html_document
---

```{r setup_r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls())
library(tidyverse)
library(patchwork)

cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
```

```{r}

read_quantification <- function(dir_name) {
  csv_files <- Sys.glob(file.path(dir_name,"*.csv"))
  dt_list <- lapply(csv_files, data.table::fread)
  df <- tibble(path = csv_files,
               feature_values = dt_list) %>%
    mutate(dir_name = dir_name,
           file_name = basename(path),
           feature_values = purrr::map(feature_values,tibble)) %>%
    separate(file_name,into = c("nuclei","well_name","ch1","ch2","ch3","ch4","sample","id"),sep="[,_.]", extra = "drop") %>%
    mutate(well_name = str_replace(well_name, "^Well", ""),
           field_id = as.integer(id))
  return(df)
}
```


### Load data

```{r}
ana_dir <- "/srv/scratch/z3532965/src/publications/2026_POLR2A_homeostasis/Imaging/NikonSpinningDisk/mAC_mACmCh_TotalPolII"
data_dir <- "/srv/scratch/berrylab/z3536241/NikonSpinningDisk"
acquisition_dirs <- purrr::map_chr(
  c("260518_mAC_mACmCh_TotalPolII/20260524_154125_815",
    "260601_mAC_mACmChPOLR2A_TotalPolII/20260609_153400_658"),~file.path(data_dir,.))
intensity_image_dirs <- map_chr(acquisition_dirs,~file.path(.,"OME-TIFF"))
intensity_image_mip_dirs <- map_chr(acquisition_dirs,~file.path(.,"OME-TIFF-MIP"))
label_image_dirs <- map_chr(acquisition_dirs,~file.path(.,"SEGMENTATION"))
quantification_dirs <- map_chr(acquisition_dirs,~file.path(.,"QUANTIFICATION"))

read_quantification <- function(dir_name) {
  csv_files <- Sys.glob(file.path(dir_name,"*.csv"))
  dt_list <- lapply(csv_files, data.table::fread)
  df <- tibble(path = csv_files,
               feature_values = dt_list) %>%
    mutate(dir_name = dir_name,
           file_name = basename(path),
           feature_values = purrr::map(feature_values,tibble)) %>%
    separate(file_name,into = c("nuclei","well_name","ch1","ch2","ch3","ch4","sample","id"),sep="[,_.]", extra = "drop") %>%
    mutate(well_name = str_replace(well_name, "^Well", ""),
           field_id = as.integer(id))
  return(df)
}

pad_well_names <- function(names) {
  gsub("(\\D)(\\d)$", "\\10\\2", names)
}

# features
f <- lapply(quantification_dirs,read_quantification) %>%
  bind_rows() %>%
  mutate(acquisition_dir = dirname(dirname(path)),
         imaging_date = basename(acquisition_dir) %>% str_extract("\\d{8}(?=_)")) %>%
  select(-c(ch1, ch2, ch3, ch4, sample))

# layouts
plate_layouts <- tibble(layout_filename=c("260518_mAC_mACmCh_TotalPolII/METADATA/260518_mAC_mACmCh_TotalPolII.csv",
                                          "260601_mAC_mACmCh_TotalPolII/METADATA/260601_mAC_mACmCh_TotalPolII.csv"),
                        plating_date=c("260518","260601"),
                        imaging_date=c("20260524","20260609"),
                        metadata=purrr::map(layout_filename,~read_csv(file.path(ana_dir,.),show_col_types = FALSE, name_repair = ~sub(" ", "_", .)))) %>%
  unnest(metadata) %>%
  mutate(well_name=pad_well_names(WELL)) %>%
  select(-WELL) %>%
  left_join(distinct(f,acquisition_dir,well_name,imaging_date),by = join_by(imaging_date, well_name))

```

Load metadata and join with quantified image data.

```{r load metadata}
# read image metadata and join to plate metadata
m <- tibble(acquisition_dir = acquisition_dirs) %>%
  mutate(image_dir = purrr::map_chr(acquisition_dirs,~file.path(.,"OME-TIFF-MIP"))) %>%
  mutate(metadata_file = purrr::map(image_dir,~list.files(.))) %>%
  unnest(metadata_file) %>%
  filter(grepl("metadata.csv$",metadata_file)) %>%
  mutate(dat = purrr::map(file.path(image_dir,metadata_file),~read_csv(.,show_col_types = FALSE))) %>%
  separate(metadata_file,into = c("well_name","ch1","ch2","ch3","ch4","sample"),sep="[,_.]", extra = "drop") %>%
  unnest(dat) %>%
  mutate(well_name = str_replace(well_name, "^Well", ""),
           field_id = as.integer(field_id)) %>%
  select(-c(ch1, ch2, ch3, ch4, sample)) %>%
  left_join(filter(plate_layouts,!is.na(acquisition_dir)),by=join_by(acquisition_dir,well_name))

# attach image and plate metadata 
dat <- right_join(m,f,by = join_by(acquisition_dir, well_name, field_id, imaging_date)) %>%
  unnest(feature_values) %>%
  select(-matches("None$|XXX$")) %>%
  rename_with(~ str_replace(., "Channel:0:0_", "Nuclei_")) %>%
  mutate(
    well_row = str_extract(well_name, "^[A-Z]"),
    well_col = as.integer(str_extract(well_name, "[0-9]+$"))
  ) %>%
  filter(Nuclei_is_border==F)

```

### Check plate layouts

```{r}

source("/srv/scratch/z3532965/src/blana/Scott/plot_plate_layout.R")

plate_layout_1 <- filter(plate_layouts,plating_date=="260518")
plate_layout_2 <- filter(plate_layouts,plating_date=="260601")

plate_layout_1 %>% group_by(CELL,DRUG_CONCENTRATION,DRUG_TIME_MINUTES,PRIMARY,SECONDARY) %>%
  count()

plate_layout_2 %>% group_by(CELL,DRUG_CONCENTRATION,DRUG_TIME_MINUTES,PRIMARY,SECONDARY) %>%
  count()


```

#### Replicate 1

```{r plate_layout_1}
plot_plate_layout(plate_layout_1, color_by = "CELL", title = "Cell type")
plot_plate_layout(plate_layout_1, color_by = "DRUG", title = "Drug")
plot_plate_layout(plate_layout_1, color_by = "DRUG_CONCENTRATION", title = "Drug (µM)", log10_scale = TRUE)
plot_plate_layout(plate_layout_1, color_by = "DRUG_TIME_MINUTES", title = "Time (min)")
plot_plate_layout(plate_layout_1, color_by = "PRIMARY", title = "Primary antibody")
plot_plate_layout(plate_layout_1, color_by = "SECONDARY", title = "Secondary antibody")
```

#### Replicate 2

```{r plate_layout_2}
plot_plate_layout(plate_layout_2, color_by = "CELL", title = "Cell type")
plot_plate_layout(plate_layout_2, color_by = "DRUG", title = "Drug")
plot_plate_layout(plate_layout_2, color_by = "DRUG_CONCENTRATION", title = "Drug (µM)", log10_scale = TRUE)
plot_plate_layout(plate_layout_2, color_by = "DRUG_TIME_MINUTES", title = "Time (min)")
plot_plate_layout(plate_layout_2, color_by = "PRIMARY", title = "Primary antibody")
plot_plate_layout(plate_layout_2, color_by = "SECONDARY", title = "Secondary antibody")
```
### Background subtraction

```{r subtract_background}

E1_IF_Zero <- dat %>% 
  filter(plating_date=="260518" & PRIMARY == "No primary") %>% 
  group_by(well_name) %>% 
  summarise(zero=mean(Nuclei_intensity_mean_647)) %>% 
  summarise(zero=mean(zero)) %>% pull(zero)
E2_IF_Zero <- 161.8488

IF_Zero <- tibble(plating_date=c("260518","260601"),
                  bkgrd_647=c(E1_IF_Zero,E2_IF_Zero))

GFP_Zero <- dat %>% 
  filter(CELL == "HCT116 OsTIRF74G") %>% 
  group_by(plating_date,well_name) %>% 
  summarise(zero=mean(Nuclei_intensity_mean_488)) %>% 
  summarise(bkgrd_488=mean(zero))

RFP_Zero <- dat %>% 
  filter(CELL == "HCT116 OsTIRF74G") %>% 
  group_by(plating_date,well_name) %>% 
  summarise(zero=mean(Nuclei_intensity_mean_561)) %>% 
  summarise(bkgrd_561=mean(zero))

DAPI_Zero <- tibble(plating_date=c("260518","260601"),
                  bkgrd_405=c(170,170))

dat <- dat %>%
  left_join(IF_Zero) %>%
  left_join(GFP_Zero) %>%
  left_join(RFP_Zero) %>%
  left_join(DAPI_Zero) %>%
  mutate(Nuclei_intensity_mean_405_minus_bkgrd = Nuclei_intensity_mean_405 - bkgrd_405,
         Nuclei_intensity_mean_488_minus_bkgrd = Nuclei_intensity_mean_488 - bkgrd_488,
         Nuclei_intensity_mean_561_minus_bkgrd = Nuclei_intensity_mean_561 - bkgrd_561,
         Nuclei_intensity_mean_647_minus_bkgrd = Nuclei_intensity_mean_647 - bkgrd_647)

```

### Check EU effect on pSer2

```{r EU_on_pSer2}

dat %>%
  filter(PRIMARY=="pSer2 Pol II (3E10)") %>%
  ggplot(aes(x=Nuclei_intensity_mean_647_minus_bkgrd,col=DRUG,group=well_name)) +
  geom_density()

dat %>%
  filter(PRIMARY=="pSer2 Pol II (3E10)") %>%
  group_by(DRUG,well_name) %>%
  summarise(IF_intensity_well = mean(Nuclei_intensity_mean_647_minus_bkgrd)) %>%
  summarise(IF_intensity_well_mean = mean(IF_intensity_well),
            IF_intensity_well_sd = sd(IF_intensity_well)) %>%
  ungroup() %>%
  mutate(IF_intensity_well_mean_norm = IF_intensity_well_mean/min(IF_intensity_well_mean),
         IF_intensity_well_sd_norm = IF_intensity_well_sd/min(IF_intensity_well_mean),
         DRUG = factor(if_else(DRUG=="Media","Media","1 mM\n5-EU"),levels=c("Media","1 mM\n5-EU"))) %>%
  ggplot(aes(y=IF_intensity_well_mean_norm,x=DRUG,fill=DRUG)) +
  geom_col() +
  geom_errorbar(aes(ymin=IF_intensity_well_mean_norm-IF_intensity_well_sd_norm,
                    ymax=IF_intensity_well_mean_norm+IF_intensity_well_sd_norm),width=0.25) +
  scale_fill_manual(values = c("grey50","#FF0000")) +
  scale_y_continuous(name="pSer2 POLR2A (3E10)",expand = expansion(mult = c(0, 0.05)),breaks=c(0,0.5,1)) +
  theme_bw(8) +
  theme(legend.position = "none",
        axis.title.x=element_blank(),
        panel.grid = element_blank())
ggsave("pSer2_with_1mMEU.pdf",width=3,height=4,units="cm")

```

### mCherry and GFP levels over time

```{r mCherry_and_GFP}

dat %>%
  filter(CELL=="HCT116 OsTIRF74G mAC-/mChPOLR2A Clone 1") %>%
  group_by(imaging_date,well_name,DRUG,DRUG_CONCENTRATION,DRUG_TIME_MINUTES) %>%
  summarise(mean_mCherry=mean(Nuclei_intensity_mean_561_minus_bkgrd),
            mean_mClover=mean(Nuclei_intensity_mean_488_minus_bkgrd)) %>%
  group_by(imaging_date,DRUG,DRUG_CONCENTRATION,DRUG_TIME_MINUTES) %>%
  summarise(expt_mean_mCherry=mean(mean_mCherry),
            expt_sd_mCherry=sd(mean_mCherry),
            expt_mean_mClover=mean(mean_mClover),
            expt_sd_mClover=sd(mean_mClover)) %>%
  ggplot(aes(x=DRUG_TIME_MINUTES,y=expt_mean_mCherry,colour = DRUG)) +
  geom_point() +
  geom_line() +
  geom_errorbar(aes(ymin=expt_mean_mCherry-expt_sd_mCherry,ymax=expt_mean_mCherry+expt_sd_mCherry)) +
  facet_wrap(~imaging_date)
    

```


### QC

We see pronounced row effects which are correlated with imaging time for the 405 and 647 lasers. Likely a reduction in power during sample acquisition. Present in both replicates to different degrees.

```{r qc}

# Check stability of control wells over well_col
control_wells_over_time <- dat %>%
  filter(DRUG=="Vehicle" & PRIMARY=="Total Pol II (F12)") %>%
  group_by(CELL,plating_date,imaging_date,well_name,well_col,well_row) %>%
  summarise(across(c(Nuclei_intensity_mean_647_minus_bkgrd,
                     Nuclei_intensity_mean_488_minus_bkgrd,
                     Nuclei_intensity_mean_561_minus_bkgrd,
                     Nuclei_intensity_mean_405_minus_bkgrd,
                     acquisition_time_rel),c(mean=~mean(.)))) %>%
  pivot_longer(matches("Nuclei_intensity"),names_prefix = "Nuclei_intensity_mean_",names_to = "channel", values_to = "mean_intensity")

initial_value <- control_wells_over_time %>%
  group_by(channel,imaging_date,CELL) %>%
  filter(acquisition_time_rel_mean==min(acquisition_time_rel_mean)) %>%
  ungroup() %>%
  rename(initial_mean_intensity=mean_intensity) %>%
  select(channel,imaging_date,CELL,initial_mean_intensity)

control_wells_over_time %>%
  left_join(initial_value) %>%
  filter(!(CELL=="HCT116 OsTIRF74G" & channel %in% c("488_minus_bkgrd_mean","561_minus_bkgrd_mean"))) %>%
  filter(!(CELL=="HCT116 OsTIRF74G mAC-POLR2A Clone 5" & channel %in% c("561_minus_bkgrd_mean"))) %>%
  ggplot(aes(x=acquisition_time_rel_mean,col=CELL,y=mean_intensity/initial_mean_intensity)) +
  geom_point() +
  facet_grid(imaging_date~channel) +
  scale_y_continuous(limits=c(0,1.1),breaks=c(0,0.5,1))

# well_col
control_wells_over_time %>%
  left_join(initial_value) %>%
  ggplot(aes(x=well_col,col=channel,y=mean_intensity/initial_mean_intensity)) +
  geom_point() +
  facet_wrap(~imaging_date) +
  scale_y_continuous(limits=c(0,NA))

# well_row
control_wells_over_time %>%
  left_join(initial_value) %>%
  ggplot(aes(x=well_row,col=channel,y=mean_intensity/initial_mean_intensity)) +
  geom_point() +
  facet_wrap(~imaging_date) +
  scale_y_continuous(limits=c(0,NA))

```

### Acquisition decay correction

Signal in the 647 (IF) and 405 (DAPI) channels declines over acquisition time,
likely due to laser power reduction during the run. Cell lines occupy distinct
plate regions, so they are acquired at different times and enter the decay curve
at different points.

We fit a model in which the **relative** decay shape is *shared* across cell
lines, and cell lines differ only by an overall multiplicative baseline:

$$I_{ij}(t) = S_i \cdot \big[(1 - c)\,e^{-k t_{ij}} + c\big]$$

where $i$ indexes cell line, $j$ indexes well, $k$ (rate) and $c$ (asymptote
*fraction*, $0 \le c \le 1$) are shared, and $S_i$ is the per-cell-line baseline
(the expected signal at $t = 0$). The bracketed term $f(t) = (1-c)e^{-kt} + c$
is identical for every cell line and satisfies $f(0) = 1$.

Because the decay shape is shared and relative, the **correction factor depends
only on the shared parameters**, not on $S_i$:

$$\mathrm{cf}(t) = \frac{1}{(1 - c)\,e^{-k t} + c}$$

This is the same for all cell lines at a given $t$, and is bounded above by
$1/c$. A cell line measured only late therefore cannot produce a runaway
correction — its baseline $S_i$ is absorbed and drops out entirely.

```{r decay-correction-setup}
library(minpack.lm)   # nlsLM(): robust Levenberg-Marquardt NLS
```

```{r decay-calibration-data}
# Well-level means for vehicle-treated, Total-Pol-II-stained control wells.
# Fit on raw background-subtracted intensities; the per-cell-line baseline
# S_i is estimated jointly with the shared decay shape.
calibration_data <- dat %>%
  filter(DRUG == "Vehicle" & PRIMARY == "Total Pol II (F12)") %>%
  group_by(imaging_date, CELL, well_name, well_col, well_row) %>%
  summarise(
    across(
      c(Nuclei_intensity_mean_647_minus_bkgrd,
        Nuclei_intensity_mean_405_minus_bkgrd,
        acquisition_time_rel),
      mean
    ),
    .groups = "drop"
  ) %>%
  pivot_longer(
    cols         = c(Nuclei_intensity_mean_647_minus_bkgrd,
                     Nuclei_intensity_mean_405_minus_bkgrd),
    names_to     = "channel",
    values_to    = "mean_intensity",
    names_prefix = "Nuclei_intensity_mean_"
  )
```

```{r decay-fit-function}
# Fit shared relative-decay shape with per-cell-line multiplicative baseline:
#   I = S_i * ((1 - cc) * exp(-k * t) + cc)
#
# Shared parameters: k (rate), cc (asymptote fraction in [0, 1]).
# Per-cell-line parameters: S_i (baseline = signal at t = 0).
#
# nlsLM has no vector parameters, so the S vector is unrolled into the
# formula dynamically: c(S1, S2, ...)[idx]. The relative shape f(t) is
# identical across cell lines, so the correction factor 1/f(t) is
# independent of S_i by construction.

fit_decay <- function(df) {

  cells   <- unique(df$CELL)
  n_cells <- length(cells)

  df <- df %>%
    mutate(cell_idx = as.integer(factor(CELL, levels = cells)))

  t   <- df$acquisition_time_rel
  y   <- df$mean_intensity
  idx <- df$cell_idx

  # Start values for S_i: per-cell-line mean intensity
  S_start <- df %>%
    group_by(cell_idx) %>%
    summarise(s = mean(mean_intensity), .groups = "drop") %>%
    pull(s)
  names(S_start) <- paste0("S", seq_along(S_start))

  S_expr <- paste0("c(", paste(names(S_start), collapse = ", "), ")")

  data_for_nls <- data.frame(t = t, y = y, idx = idx)

  # Full model: shared k and cc, per-cell-line S_i
  fit_full <- tryCatch(
    nlsLM(
      as.formula(paste0("y ~ ", S_expr, "[idx] * ((1 - cc) * exp(-k * t) + cc)")),
      data    = data_for_nls,
      start   = as.list(c(S_start, k = 1e-4, cc = 0.7)),
      lower   = c(rep(0,   n_cells), k = 0,   cc = 0),
      upper   = c(rep(Inf, n_cells), k = Inf, cc = 1),
      control = nls.lm.control(maxiter = 500)
    ),
    error = function(e) { message("Full fit failed: ", e$message); NULL }
  )

  # Fallback: pure shared-k exponential (cc fixed at 0)
  fit_simple <- if (is.null(fit_full)) tryCatch(
    nlsLM(
      as.formula(paste0("y ~ ", S_expr, "[idx] * exp(-k * t)")),
      data    = data_for_nls,
      start   = as.list(c(S_start, k = 1e-4)),
      lower   = c(rep(0, n_cells), k = 0),
      control = nls.lm.control(maxiter = 500)
    ),
    error = function(e) { message("Simple fit failed: ", e$message); NULL }
  ) else NULL

  if (!is.null(fit_full)) {
    cf     <- coef(fit_full)
    params <- list(k = cf[["k"]], cc = cf[["cc"]], model = "shared_asymptotic")
    S_vals <- cf[names(S_start)]
  } else if (!is.null(fit_simple)) {
    cf     <- coef(fit_simple)
    params <- list(k = cf[["k"]], cc = 0, model = "shared_simple")
    S_vals <- cf[names(S_start)]
  } else {
    warning("All fits failed — using identity correction")
    params <- list(k = 0, cc = 1, model = "identity")  # cc=1 -> cf(t)=1
    S_vals <- S_start
  }

  params$cell_S <- setNames(as.numeric(S_vals), cells)
  list(params = params)
}
```

```{r decay-fit-models}
decay_fits <- calibration_data %>%
  group_by(imaging_date, channel) %>%
  nest() %>%
  mutate(fit_result = map(data, fit_decay))

# Shared parameters (k, cc) — these alone define the correction
decay_shared <- decay_fits %>%
  mutate(
    k     = map_dbl(fit_result, ~ .x$params$k),
    cc    = map_dbl(fit_result, ~ .x$params$cc),
    model = map_chr(fit_result, ~ .x$params$model)
  ) %>%
  select(imaging_date, channel, k, cc, model)

# Per-cell-line baselines (for plotting / diagnostics only)
decay_baselines <- decay_fits %>%
  mutate(cell_S = map(fit_result, ~ enframe(.x$params$cell_S,
                                            name = "CELL", value = "S"))) %>%
  select(imaging_date, channel, cell_S) %>%
  unnest(cell_S)

decay_shared %>%
  mutate(
    t_max  = map_dbl(imaging_date, ~ max(
      calibration_data$acquisition_time_rel[calibration_data$imaging_date == .x],
      na.rm = TRUE)),
    max_cf = 1 / ((1 - cc) * exp(-k * t_max) + cc)
  ) %>%
  knitr::kable(
    digits  = 5,
    caption = "Shared decay parameters and max correction factor (cell-line independent)"
  )

decay_baselines %>%
  knitr::kable(digits = 1,
               caption = "Per-cell-line baseline S_i (does not affect correction)")
```

```{r decay-fit-plot, fig.cap="Decay fits. Points are per-well means; curves show the fitted model per cell line (shared shape, scaled by S_i)."}
fit_curves <- decay_fits %>%
  left_join(
    calibration_data %>%
      group_by(imaging_date) %>%
      summarise(t_max = max(acquisition_time_rel, na.rm = TRUE), .groups = "drop"),
    by = "imaging_date"
  ) %>%
  mutate(
    curves = map2(fit_result, t_max, function(fr, tmax) {
      t_grid <- seq(0, tmax, length.out = 300)
      k  <- fr$params$k
      cc <- fr$params$cc
      imap_dfr(fr$params$cell_S, function(S, cell) {
        tibble(CELL = cell,
               acquisition_time_rel = t_grid,
               fitted = S * ((1 - cc) * exp(-k * t_grid) + cc))
      })
    })
  ) %>%
  select(imaging_date, channel, curves) %>%
  unnest(curves)

calibration_data %>%
  ggplot(aes(x = acquisition_time_rel, colour = CELL)) +
  geom_point(aes(y = mean_intensity), alpha = 0.6) +
  geom_line(data = fit_curves, aes(y = fitted), linewidth = 1) +
  facet_grid(imaging_date ~ channel, scales = "free_y") +
  labs(
    title    = "Exponential decay fit — 647 & 405 channels",
    subtitle = "Shared relative shape (1-c)·exp(-k·t)+c; per-cell-line baseline S_i",
    x        = "Relative acquisition time",
    y        = "Mean background-subtracted intensity",
    colour   = NULL
  ) +
  theme_bw() +
  theme(legend.position = "bottom")
```

```{r decay-apply-correction}
# Correction factor uses ONLY the shared parameters k and cc.
# cf(t) = 1 / ((1 - cc) * exp(-k * t) + cc),  bounded above by 1/cc.
correction_factor <- function(t, k, cc) {
  1 / ((1 - cc) * exp(-k * t) + cc)
}

cf_647 <- decay_shared %>%
  filter(channel == "647_minus_bkgrd") %>%
  select(imaging_date, k_647 = k, cc_647 = cc)

cf_405 <- decay_shared %>%
  filter(channel == "405_minus_bkgrd") %>%
  select(imaging_date, k_405 = k, cc_405 = cc)

dat_corrected <- dat %>%
  left_join(cf_647, by = "imaging_date") %>%
  mutate(
    cf_647 = correction_factor(acquisition_time_rel, k_647, cc_647),
    Nuclei_intensity_mean_647_minus_bkgrd_corrected =
      Nuclei_intensity_mean_647_minus_bkgrd * cf_647
  ) %>%
  left_join(cf_405, by = "imaging_date") %>%
  mutate(
    cf_405 = correction_factor(acquisition_time_rel, k_405, cc_405),
    Nuclei_intensity_mean_405_minus_bkgrd_corrected =
      Nuclei_intensity_mean_405_minus_bkgrd * cf_405
  )
```

```{r decay-qc-plot, fig.cap="Signal stability in control wells before (faded, dashed) and after (solid) correction, coloured by cell line."}
check_corrected <- dat_corrected %>%
  filter(DRUG == "Vehicle" & PRIMARY == "Total Pol II (F12)") %>%
  group_by(imaging_date, CELL, well_name, well_col, well_row) %>%
  summarise(
    across(
      c(Nuclei_intensity_mean_647_minus_bkgrd,
        Nuclei_intensity_mean_647_minus_bkgrd_corrected,
        Nuclei_intensity_mean_405_minus_bkgrd,
        Nuclei_intensity_mean_405_minus_bkgrd_corrected,
        acquisition_time_rel),
      mean
    ),
    .groups = "drop"
  ) %>%
  pivot_longer(
    cols         = matches("Nuclei_intensity_mean_(647|405)"),
    names_to     = "channel",
    values_to    = "mean_intensity",
    names_prefix = "Nuclei_intensity_mean_"
  ) %>%
  mutate(
    corrected = if_else(str_detect(channel, "corrected"),
                        "After correction", "Before correction"),
    channel   = str_remove(channel, "_corrected")
  )

check_corrected %>%
  ggplot(aes(x        = acquisition_time_rel,
             y        = mean_intensity,
             colour   = CELL,
             alpha    = corrected,
             linetype = corrected)) +
  geom_point() +
  geom_smooth(aes(group = interaction(CELL, corrected)),
              method = "lm", se = FALSE, linewidth = 0.7) +
  facet_wrap(channel ~ imaging_date , scales = "free_y") +
  scale_alpha_manual(values    = c("Before correction" = 0.3,
                                   "After correction"  = 0.8)) +
  scale_linetype_manual(values = c("Before correction" = "dashed",
                                   "After correction"  = "solid")) +
  ylim(c(0,NA)) +
  labs(
    title    = "Signal stability before/after decay correction",
    x        = "Relative acquisition time",
    y        = "Mean background-subtracted intensity",
    colour   = NULL, alpha = NULL, linetype = NULL
  ) +
  theme_bw() +
  theme(legend.position = "bottom")
```

```{r investigate_corrected}

dat_corrected_well_summary <- dat_corrected %>%
  group_by(plating_date,imaging_date,well_name,CELL,DRUG,DRUG_CONCENTRATION,DRUG_TIME_MINUTES,PRIMARY) %>%
  summarise(across(matches("Nuclei_intensity_mean"),c(mean=~mean(.))),.groups = 'drop')

zero_point <- dat_corrected_well_summary %>%
  filter(DRUG=="Vehicle") %>%
  mutate(DRUG_TIME_MINUTES=0) %>%
  select(-DRUG_CONCENTRATION,-DRUG) %>%
  expand_grid(distinct(dat_corrected_well_summary,DRUG,DRUG_CONCENTRATION))

dat_corrected_plate_summary <- dat_corrected_well_summary %>%
  group_by(plating_date,imaging_date,CELL,DRUG,DRUG_CONCENTRATION,DRUG_TIME_MINUTES,PRIMARY) %>%
  summarise(across(matches("Nuclei_intensity_mean"),c(mean=~mean(.),sd=~sd(.))),.groups = 'drop')

zero_point_plate <- zero_point %>%
  group_by(plating_date,imaging_date,CELL,PRIMARY,DRUG,DRUG_CONCENTRATION) %>%
  summarise(across(matches("Nuclei_intensity_mean"),c(mean=~mean(.),sd=~sd(.))),.groups = 'drop')

# compare before/after correction
dat_corrected_plate_summary %>%
  bind_rows(zero_point_plate) %>%
  filter(CELL == "HCT116 OsTIRF74G mAC-POLR2A Clone 5" &
         PRIMARY == "Total Pol II (F12)") %>%
  pivot_longer(
    cols = c(
      Nuclei_intensity_mean_647_minus_bkgrd_corrected_mean_mean,
      Nuclei_intensity_mean_647_minus_bkgrd_corrected_mean_sd,
      Nuclei_intensity_mean_647_minus_bkgrd_mean_mean,
      Nuclei_intensity_mean_647_minus_bkgrd_mean_sd
    ),
    names_to      = c("name", ".value"),
    names_pattern = "(.*)_(mean|sd)$"
  ) %>%
  ggplot(aes(x = DRUG_TIME_MINUTES, y = mean, col = name)) +
  geom_point() +
  geom_line() +
  geom_errorbar(aes(ymin = mean - sd, ymax = mean + sd), width = 0) +
  facet_wrap(DRUG_CONCENTRATION ~ plating_date)

normalisers <- zero_point %>%
  group_by(plating_date,imaging_date,CELL,PRIMARY) %>%
  summarise(across(matches("Nuclei_intensity_mean"),c(normaliser=~mean(.))),.groups = 'drop')

dat_corrected_overall_summary <- dat_corrected_well_summary %>%
  bind_rows(zero_point) %>%
  left_join(normalisers) %>%
  group_by(CELL,DRUG,DRUG_CONCENTRATION,DRUG_TIME_MINUTES,PRIMARY) %>%
  mutate(normalised_Nuclei_intensity_mean_647_minus_bkgrd_mean = 
           Nuclei_intensity_mean_647_minus_bkgrd_mean / Nuclei_intensity_mean_647_minus_bkgrd_mean_normaliser,
         normalised_Nuclei_intensity_mean_647_minus_bkgrd_corrected_mean = 
           Nuclei_intensity_mean_647_minus_bkgrd_corrected_mean / Nuclei_intensity_mean_647_minus_bkgrd_corrected_mean_normaliser) %>%
  summarise(across(matches("normalised_Nuclei_intensity_mean"),c(mean=~mean(.),sd=~sd(.))),.groups = 'drop')

uncorrected <- dat_corrected_overall_summary %>%
  filter(CELL != "HCT116 OsTIRF74G" &
         PRIMARY == "Total Pol II (F12)" &
           DRUG=="5PhIAA") %>%
  filter(DRUG_CONCENTRATION %in% c(1,3,10) | CELL=="HCT116 OsTIRF74G mAC-/mChPOLR2A Clone 1") %>%
  ggplot(aes(x=DRUG_TIME_MINUTES,
             y=normalised_Nuclei_intensity_mean_647_minus_bkgrd_mean_mean,
             col=factor(DRUG_CONCENTRATION))) +
  geom_point() +
  geom_line() +
  geom_errorbar(aes(ymin=normalised_Nuclei_intensity_mean_647_minus_bkgrd_mean_mean -
                      normalised_Nuclei_intensity_mean_647_minus_bkgrd_mean_sd,
                    ymax=normalised_Nuclei_intensity_mean_647_minus_bkgrd_mean_mean +
                      normalised_Nuclei_intensity_mean_647_minus_bkgrd_mean_sd)) +
  coord_cartesian(xlim=c(0,NA),ylim=c(0,1.1)) +
  facet_wrap(~CELL,ncol=1)

corrected <- dat_corrected_overall_summary %>%
  filter(CELL != "HCT116 OsTIRF74G" &
           PRIMARY == "Total Pol II (F12)" &
           DRUG=="5PhIAA") %>%
  filter(DRUG_CONCENTRATION %in% c(1,3,10) | CELL=="HCT116 OsTIRF74G mAC-/mChPOLR2A Clone 1") %>%
  ggplot(aes(x=DRUG_TIME_MINUTES,
             y=normalised_Nuclei_intensity_mean_647_minus_bkgrd_corrected_mean_mean,
             col=factor(DRUG_CONCENTRATION))) +
  geom_point() +
  geom_line() +
  geom_errorbar(aes(ymin=normalised_Nuclei_intensity_mean_647_minus_bkgrd_corrected_mean_mean -
                      normalised_Nuclei_intensity_mean_647_minus_bkgrd_corrected_mean_sd,
                    ymax=normalised_Nuclei_intensity_mean_647_minus_bkgrd_corrected_mean_mean +
                      normalised_Nuclei_intensity_mean_647_minus_bkgrd_corrected_mean_sd)) +
  coord_cartesian(xlim=c(0,NA),ylim=c(0,1.1)) +
  facet_wrap(~CELL,ncol=1)

library(patchwork)
corrected + uncorrected + plot_layout(guides='collect')

dat_corrected_overall_summary %>%
  filter(CELL != "HCT116 OsTIRF74G" &
           PRIMARY == "Total Pol II (F12)" &
           DRUG=="5PhIAA") %>%
  filter(DRUG_TIME_MINUTES!=360) %>%
  filter(DRUG_CONCENTRATION %in% c(10000) | CELL!="HCT116 OsTIRF74G mAC-/mChPOLR2A Clone 1") %>%
  pivot_longer(
    cols = c(
      normalised_Nuclei_intensity_mean_647_minus_bkgrd_corrected_mean_mean,
      normalised_Nuclei_intensity_mean_647_minus_bkgrd_corrected_mean_sd,
      normalised_Nuclei_intensity_mean_647_minus_bkgrd_mean_mean,
      normalised_Nuclei_intensity_mean_647_minus_bkgrd_mean_sd
    ),
    names_to      = c("name", ".value"),
    names_pattern = "(.*)_(mean|sd)$"
  ) %>%
  ggplot(aes(x = DRUG_TIME_MINUTES, y = mean, col = name)) +
  geom_point() +
  geom_line() +
  scale_y_continuous(limits = c(0,1.2)) +
  geom_errorbar(aes(ymin = mean - sd, ymax = mean + sd), width = 0) +
  facet_wrap(CELL~DRUG_CONCENTRATION,nrow=1) +
  theme(legend.position = "bottom",
        legend.direction = "vertical")
  

```

```{r check_nucl_area}

dat %>%
  filter(DRUG_CONCENTRATION==10000) %>%
  distinct(CELL,DRUG_TIME_MINUTES)

dat_corrected %>%
  filter(DRUG_CONCENTRATION==10000 | CELL=="HCT116 OsTIRF74G") %>%
  ggplot() +
  geom_density(aes(x=Nuclei_area*Nuclei_intensity_mean_405_minus_bkgrd_corrected,col=DRUG_TIME_MINUTES,group=interaction(DRUG_TIME_MINUTES,well_name))) +
  facet_grid(CELL~imaging_date) +
  xlim(c(0,1.5e6))

```

