---
title: "Stats on Pol II log~2~ fold-changes"
subtitle: "Deleted-effect contrasts with multivariate-*t* adjustment"
date: "`r format(Sys.Date(), '%d %B %Y')`"
params:
  csv_path: "data/metadata_boxplot_data_all_figureS7_ptype_Sep2026.csv"
  out_dir: "results"
  seed: 20260914
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    number_sections: true
    code_folding: show
    df_print: paged
    fig_width: 11
    fig_height: 3.6
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE, message = FALSE, warning = FALSE,
  dpi = 150, fig.align = "center"
)
```

```{r libraries}
suppressPackageStartupMessages({
  library(readr)
  library(dplyr)
  library(tidyr)
  library(purrr)
  library(emmeans)
  library(ggplot2)
  library(RColorBrewer)
  library(knitr)
})

set.seed(params$seed)   # mvt p-values are simulation-based 
dir.create(params$out_dir, showWarnings = FALSE, recursive = TRUE)
```

# Definitions

We are interested in whether the stratified groups result in statistically
different log2 fold-changes in Pol II ChIP-seq from the other groups.

We want to highlight each group against the unweighted mean of the
remaining groups of the same variable. In `emmeans` this is called a 
deleted-effect contrast (`method = "del.eff"`).

Four response columns are tested (TSS + Gene body) x (90min + 8h):

| response column | figure sub-panel |
|---|---|
| `TSS_log2FC_90min_0h` | TSS 1.5 h / 0 h |
| `TSS_log2FC_8h_0h` | TSS 8 h / 0 h |
| `GeneBody_log2FC_90min_0h` | Gene Body 1.5 h / 0 h |
| `GeneBody_log2FC_8h_0h` | Gene Body 8 h / 0 h |

Note that the column ``promoter`` (promoter class) is not categorical because
each promoter can be in multiple classes.

```{r config}
responses <- c(
  "TSS_log2FC_90min_0h",
  "TSS_log2FC_8h_0h",
  "GeneBody_log2FC_90min_0h",
  "GeneBody_log2FC_8h_0h"
)

# Pretty facet labels, in figure order
resp_labels <- c(
  TSS_log2FC_90min_0h      = "TSS\n1.5 h / 0 h",
  TSS_log2FC_8h_0h         = "TSS\n8 h / 0 h",
  GeneBody_log2FC_90min_0h = "Gene Body\n1.5 h / 0 h",
  GeneBody_log2FC_8h_0h    = "Gene Body\n8 h / 0 h"
)

id_cols <- c("gene_id", "gene_name", "transcript_id", "symbol")

special_cols <- c("promoter")

# Axis order for the known bins
level_orders <- list(
  length_category = c("1\u20135 kb", "5\u201310 kb", "10\u201325 kb",
                      "25\u201350 kb", "50\u201375 kb", ">75 kb"),
  baseMean_bin    = paste0("Q", 1:8),
  pausing_bin     = c("<-2", "-2 to -1", "-1 to 0", "0 to 1",
                      "1 to 2", "2 to 3", "3 to 4", "> 4"),
  promoter_type   = c("CCAAT", "GC box", "Initiator", "TATA", "no information")
)

# Which figure row each grouping column corresponds to
panel_meta <- tibble::tribble(
  ~variable,         ~panel, ~title,             ~xlab,                          ~palette,
  "pausing_bin",     "S7B",  "Pausing index",    "log2 pausing index (0 h)",     "pausing",
  "baseMean_bin",    "S7C",  "Pol II occupancy", "log2(baseMean) quantile (n=8)", "occupancy",
  "length_category", "S7D",  "Gene length",      "Gene length category (kb)",    "length"
)

group_vars <- c("length_category","baseMean_bin","pausing_bin")

```

# Input data

```{r read-data}
stopifnot(file.exists(params$csv_path))
dat <- read_csv(params$csv_path, show_col_types = FALSE, progress = FALSE)

missing_resp <- setdiff(responses, names(dat))
if (length(missing_resp)) {
  stop("missing response column(s): ", paste(missing_resp, collapse = ", "))
}
```

```{r level-order}
order_levels <- function(x, var) {
  obs  <- sort(unique(as.character(x[!is.na(x)])))
  want <- level_orders[[var]]
  if (is.null(want)) return(obs)
  extra <- setdiff(obs, want)
  if (length(extra)) {
    warning("levels of '", var, "' not in the configured order, appended: ",
            paste(extra, collapse = ", "), call. = FALSE)
  }
  c(intersect(want, obs), extra)
}
```

Group sizes differ substantially for non-quartile binning. Counts below are the `n` used by
each test, after dropping genes with a missing response value.

```{r group-sizes}
size_tbl <- map_dfr(group_vars, function(gv) {
  map_dfr(responses, function(rsp) {
    dat |>
      transmute(level = as.character(.data[[gv]]), y = .data[[rsp]]) |>
      filter(!is.na(y), !is.na(level)) |>
      count(level, name = "n") |>
      mutate(variable = gv, response = rsp)
  })
})

size_tbl |>
  mutate(level = factor(level, levels = unlist(map(group_vars, ~ order_levels(dat[[.x]], .x))))) |>
  select(variable, level, response, n) |>
  pivot_wider(names_from = response, values_from = n, values_fill = 0) |>
  arrange(variable, level) |>
  kable(caption = "Genes contributing to each test")
```

Important to note that TSS responses are defined for 9,849 genes, while 
gene-body responses are defined for 8,983 genes (short genes excluded). The
1–5 kb class is absent from the gene-body panels, so gene length is tested over
six levels for TSS and five for gene body.

# Statistical approach

Each family is a one-way linear model on the gene-level log~2~ fold-changes:

```r
mod <- lm(response ~ group)
em  <- emmeans(mod, ~ group)
contrast(em, method = "del.eff", adjust = "mvt")
```

**`lm(response ~ group)`** — a one-way model whose fitted values are the group
means.

**`emmeans(mod, ~ group)`** — the estimated marginal means. With a single categorical
predictor these are the plain group means, reported with a standard error pooled
across groups.

**`method = "del.eff"`** — deleted-effect contrasts. For level *i* of a variable with
*k* levels, the contrast is

$$\hat\delta_i \;=\; \bar{y}_i \;-\; \frac{1}{k-1}\sum_{j \neq i} \bar{y}_j$$

i.e. that level minus the **unweighted** mean of the others.

**`adjust = "mvt"`** — the multivariate-*t* adjustment. The *k* deleted-effect
contrasts within one family are strongly correlated. `mvt` integrates the
exact multivariate-*t* distribution of
the whole set rather than assuming independence. It controls the family-wise error
rate across the *k* contrasts of that family. It is computed by quasi-Monte-Carlo
integration, which is why we fix the seed.

**One-way ANOVA** — alongside each family, `anova(mod)` gives the overall test that
all group means are equal, i.e. that the grouping variable explains nothing.

**Which multiplicity correction is primary** — `mvt` operates within one family, but 
we later apply a Benjamini–Hochberg correction across all
106 class-vs-rest contrasts (`p_BH_fig`).

## Fitting

```{r fit-function}
fit_family <- function(resp, gv) {
  d <- dat |>
    transmute(y = .data[[resp]], group = as.character(.data[[gv]])) |>
    filter(!is.na(y), !is.na(group)) |>
    mutate(group = factor(group, levels = order_levels(group, gv))) |>
    droplevels()
  if (nlevels(d$group) < 2) return(NULL)

  mod <- lm(y ~ group, data = d)
  em  <- emmeans(mod, ~ group)
  ct  <- contrast(em, method = "del.eff", adjust = "mvt")

  omnibus <- anova(mod)

  p_unadj <- as.data.frame(summary(contrast(em, method = "del.eff", adjust = "none")))$p.value
  ct_df <- as.data.frame(summary(ct, infer = c(TRUE, TRUE)))
  # del.eff returns exactly one contrast per level, in level order. Take names
  # from the factor rather than parsing emmeans' labels, which parenthesise
  # levels containing special characters (e.g. "<-2" becomes "(<-2)").
  stopifnot(nrow(ct_df) == nlevels(d$group))

  means <- as.data.frame(summary(em)) |>
    mutate(response = resp, variable = gv, .before = 1) |>
    left_join(count(d, group, name = "n"), by = "group")

  contrasts <- ct_df |>
    mutate(response       = resp,
           variable       = gv,
           level          = levels(d$group),
           contrast_label = as.character(contrast),
           .before        = 1) |>
    left_join(count(d, group, name = "n_level"), by = c("level" = "group")) |>
    mutate(p_raw       = p_unadj,
           n_total     = nrow(d),
           omnibus_F   = omnibus[["F value"]][1],
           omnibus_df1 = omnibus[["Df"]][1],
           omnibus_df2 = omnibus[["Df"]][2],
           omnibus_p   = omnibus[["Pr(>F)"]][1])

  list(means = means, contrasts = contrasts)
}
```

```{r fit-all}
grid <- expand_grid(response = responses, variable = group_vars)
res  <- compact(pmap(grid, function(response, variable) fit_family(response, variable)))

emm_tbl <- bind_rows(map(res, "means"))

con_tbl <- bind_rows(map(res, "contrasts")) |>
  mutate(p_BH_BD_panels_only = p.adjust(p.value, method = "BH")) |>
  rename(estimate_vs_rest = estimate, p_mvt = p.value) |>
  select(response, variable, level, contrast_label, n_level, n_total,
         estimate_vs_rest, SE, df, lower.CL, upper.CL, t.ratio,
         p_raw, p_mvt, p_BH_BD_panels_only,
         omnibus_F, omnibus_df1, omnibus_df2, omnibus_p)

write_csv(con_tbl, file.path(params$out_dir, "del_eff_contrasts.csv"))
write_csv(emm_tbl, file.path(params$out_dir, "group_emmeans.csv"))
```

# One-way ANOVA

```{r omnibus}
omni <- con_tbl |>
  distinct(response, variable, omnibus_F, omnibus_df1, omnibus_df2, omnibus_p, n_total) |>
  mutate(response = factor(response, levels = responses)) |>
  arrange(variable, response) |>
  transmute(variable, response,
            n = n_total,
            `F` = sprintf("%.2f", omnibus_F),
            df = sprintf("%d, %d", omnibus_df1, omnibus_df2),
            p = format.pval(omnibus_p, digits = 3, eps = 1e-16))

kable(omni, caption = "Overall test that all group means are equal")
```


- Only ``length_category`` is not-significant for TSS at 90 min. For all others, the grouping is significant.
- ``baseMean_bin`` is borderline for gene body at 8 h.

```{r stars-helper}
stars <- function(p) {
  cut(p,
      breaks = c(-Inf, 0.001, 0.01, 0.05, Inf),
      labels = c("***", "**", "*", ""))
}
```

# Promoter class

The `promoter` column stores motif *combinations*: `GC_box`, `CCAAT; GC_box`,
`CCAAT; GC_box; Initiator; TATA`, etc.

```{r promoter-classes}
motifs        <- c("CCAAT", "GC_box", "Initiator", "TATA")
motif_labels  <- c(CCAAT = "CCAAT", GC_box = "GC box", Initiator = "Initiator",
                   TATA = "TATA", none = "no information")
no_info_level <- "no promoter information"

has_promoter <- "promoter" %in% names(dat)

if (has_promoter) {
  memb <- dat |>
    select(gene_id, promoter, all_of(responses)) |>
    filter(!is.na(promoter))
  for (m in motifs) memb[[m]] <- grepl(m, memb$promoter, fixed = TRUE)
  memb$none <- memb$promoter == no_info_level

  tibble(class = unname(motif_labels[c(motifs, "none")]),
         genes = vapply(c(motifs, "none"), function(m) sum(memb[[m]]), integer(1))) |>
    mutate(`% of genes` = sprintf("%.1f%%", 100 * genes / nrow(memb))) |>
    kable(caption = paste0("Class sizes (total genes = ", nrow(memb), ")"))
}
```

```{r promoter-overlap}
if (has_promoter) {
  n_motif <- rowSums(memb[motifs])
  tibble(`motifs per gene` = c("0 (no information)", "1", "2", "3", "4"),
         genes = as.integer(table(factor(n_motif, levels = 0:4)))) |>
    kable()
}
```

We code the 16 observed combinations as levels of a single factor
factor and calculate the gene-weighted contrast. The estimate is then the 
marginal difference in mean log~2~ fold-change between genes carrying that 
motif and genes not carrying it. Because all five contrasts come from one model, `mvt`
still applies.

## Class vs rest, and class vs zero

```{r promoter-contrasts}
# Which of the 16 motif combinations belong to a given promoter class?
# "none" is the single "no promoter information" combination; the others are any
# combination whose label contains that motif.
combos_in_class <- function(cls, combos) {
  if (cls == "none") combos == no_info_level else grepl(cls, combos, fixed = TRUE)
}

# Weights over the 16 combination means.
#   vs_rest = FALSE -> gene-weighted average of the class's own combinations.
#                      Weights sum to 1, so the contrast estimates the class mean
#                      and emmeans tests it against zero.
#   vs_rest = TRUE  -> that average minus the gene-weighted average of every
#                      combination outside the class. Weights sum to 0, giving the
#                      exact marginal difference (motif present - motif absent).
# Weighting each combination by its gene count is what makes these averages equal
# the mean over genes, rather than the mean over combinations.
class_weights <- function(in_class, n_genes, vs_rest) {
  stopifnot(any(in_class), !all(in_class))
  w <- ifelse(in_class, n_genes / sum(n_genes[in_class]), 0)
  if (vs_rest) w <- w - ifelse(in_class, 0, n_genes / sum(n_genes[!in_class]))
  w
}

# Summarise one set of weights, keeping both the unadjusted p-value (which the
# figure-wide BH correction later uses) and the within-panel mvt value.
contrast_summary <- function(em, weights) {
  unadjusted <- as.data.frame(summary(contrast(em, weights, adjust = "none")))
  adjusted   <- as.data.frame(summary(contrast(em, weights, adjust = "mvt"),
                                      infer = c(TRUE, TRUE)))
  adjusted$p_raw <- unadjusted$p.value
  adjusted$class <- as.character(adjusted$contrast)
  adjusted
}

promoter_fit <- function(resp) {
  genes <- memb |> filter(!is.na(.data[[resp]]))
  model_data <- genes |>
    transmute(y = .data[[resp]], combo = factor(promoter)) |>
    droplevels()

  em      <- emmeans(lm(y ~ combo, data = model_data), ~ combo)
  combos  <- levels(model_data$combo)
  n_genes <- as.numeric(table(model_data$combo)[combos])

  # one row per promoter class, carrying its display name and its gene count
  # (these sum to more than the number of genes, because the classes overlap)
  classes <- tibble(key = c(motifs, "none")) |>
    mutate(class   = unname(motif_labels[key]),
           n_class = vapply(key, function(k) sum(genes[[k]]), integer(1)))

  membership <- lapply(classes$key, combos_in_class, combos = combos)
  named <- function(x) setNames(x, classes$class)

  vs_rest <- contrast_summary(em, named(lapply(membership, class_weights,
                                               n_genes = n_genes, vs_rest = TRUE))) |>
    transmute(class, estimate_vs_rest = estimate, lower.CL, upper.CL, t.ratio,
              p_raw, p_mvt = p.value)

  vs_zero <- contrast_summary(em, named(lapply(membership, class_weights,
                                               n_genes = n_genes, vs_rest = FALSE))) |>
    transmute(class, mean_log2FC = estimate, mean_lower = lower.CL,
              mean_upper = upper.CL, p_vs_zero = p.value)

  classes |>
    select(class, n_class) |>
    left_join(vs_rest, by = "class") |>
    left_join(vs_zero, by = "class") |>
    mutate(response = resp, n_total = nrow(model_data)) |>
    select(response, class, estimate_vs_rest, lower.CL, upper.CL, t.ratio,
           p_raw, p_mvt, mean_log2FC, mean_lower, mean_upper, p_vs_zero,
           n_class, n_total)
}

if (has_promoter) {
  promoter_tbl <- map_dfr(responses, promoter_fit) |>
    mutate(class = factor(class, levels = unname(motif_labels[c(motifs, "none")])))
}
```

Results shown after the multiplicity correction

## Does promoter class matter at all?

The equivalent of the one-way ANOVA above is a 
joint Wald test of the five class-vs-rest contrasts.

```{r promoter-joint}
promoter_joint <- purrr::map_dfr(responses, function(resp) {
  d <- memb |>
    filter(!is.na(.data[[resp]])) |>
    transmute(y = .data[[resp]], combo = factor(promoter)) |>
    droplevels()
  em      <- emmeans(lm(y ~ combo, data = d), ~ combo)
  combos  <- levels(d$combo)
  n_genes <- as.numeric(table(d$combo)[combos])

  weights <- setNames(
    lapply(lapply(c(motifs, "none"), combos_in_class, combos = combos),
           class_weights, n_genes = n_genes, vs_rest = TRUE),
    unname(motif_labels[c(motifs, "none")]))

  joint <- test(contrast(em, weights), joint = TRUE)
  tibble(response = resp, df1 = joint$df1, df2 = joint$df2,
         F = joint$F.ratio, p = joint$p.value)
})
write_csv(promoter_joint, file.path(params$out_dir, "promoter_omnibus.csv"))

promoter_joint |>
  mutate(response = factor(response, levels = responses)) |>
  arrange(response) |>
  transmute(response,
            test = sprintf("F(%d, %d) = %.2f", df1, df2, F),
            p = format.pval(p, digits = 3, eps = 1e-16)) |>
  kable(caption = "Joint Wald test of the five promoter classes")
```

At 1.5 h the test is significant but only just, and no individual class survives
correction, so there is quite weak evidence that promoter class influences response.

## FDR control

```{r figure-wide-bh}
all_contrasts <- bind_rows(
  con_tbl      |> transmute(panel = variable, response, item = level, p_raw),
  promoter_tbl |> transmute(panel = "promoter", response, item = as.character(class), p_raw)
) |>
  mutate(p_BH_fig = p.adjust(p_raw, method = "BH"))

con_tbl <- con_tbl |>
  left_join(all_contrasts |> filter(panel != "promoter") |>
              select(panel, response, item, p_BH_fig),
            by = c("variable" = "panel", "response", "level" = "item"))
promoter_tbl <- promoter_tbl |>
  mutate(item = as.character(class)) |>
  left_join(all_contrasts |> filter(panel == "promoter") |>
              select(response, item, p_BH_fig), by = c("response", "item")) |>
  select(-item)

write_csv(con_tbl, file.path(params$out_dir, "del_eff_contrasts.csv"))
write_csv(promoter_tbl, file.path(params$out_dir, "promoter_contrasts.csv"))

tibble(
  family = c("all contrasts in Figure S7",
             "B-D only (mvt, within panel)",
             "B-D only (BH, within panel)",
             "A only (mvt, within panel)"),
  n = c(nrow(all_contrasts), nrow(con_tbl), nrow(con_tbl), nrow(promoter_tbl)),
  significant = c(sum(all_contrasts$p_BH_fig < 0.05),
                  sum(con_tbl$p_mvt < 0.05),
                  sum(p.adjust(con_tbl$p_raw, "BH") < 0.05),
                  sum(promoter_tbl$p_mvt < 0.05))
) |>
  kable(caption = "Significant contrasts at 0.05 under each choice of family and method")
```

```{r bh-changes}
bind_rows(
  con_tbl |> transmute(panel = variable, response, item = level,
                       p_within = p_mvt, p_BH_fig),
  promoter_tbl |> transmute(panel = "promoter (A)", response,
                            item = as.character(class), p_within = p_mvt, p_BH_fig)
) |>
  filter((p_within < 0.05) != (p_BH_fig < 0.05)) |>
  mutate(response = factor(response, levels = responses)) |>
  arrange(panel, response) |>
  transmute(panel, response, item,
            `p (mvt, within panel)` = format.pval(p_within, digits = 3, eps = 1e-16),
            `p (BH, whole figure)`  = format.pval(p_BH_fig, digits = 3, eps = 1e-16),
            change = ifelse(p_BH_fig < 0.05, "becomes significant", "loses significance")) |>
  kable(caption = "Annotations that move when the figure becomes one BH family")
```

```{r promoter-table}
if (has_promoter) {
  write_csv(promoter_tbl, file.path(params$out_dir, "figureS7A_promoter_contrasts.csv"))
  promoter_tbl |>
    mutate(response = factor(response, levels = responses)) |>
    arrange(response, class) |>
    transmute(response, class, n = n_class,
              `mean log2FC` = sprintf("%+.4f", mean_log2FC),
              `p vs 0` = format.pval(p_vs_zero, digits = 3, eps = 1e-16),
              `vs rest` = sprintf("%+.4f", estimate_vs_rest),
              `95% CI` = sprintf("[%+.3f, %+.3f]", lower.CL, upper.CL),
              `p vs rest (BH, figure)` = format.pval(p_BH_fig, digits = 3, eps = 1e-16),
              sig = as.character(stars(p_BH_fig)),
              `p vs rest (mvt, panel)` = format.pval(p_mvt, digits = 3, eps = 1e-16)) |>
    kable(caption = "Promoter classes: mean change, and class vs all genes outside the class")
}
```

## Adjusting for co-occurrence

Nearly two thirds of TATA-containing promoters also carry a GC box, so a marginal TATA
effect may be carried by its partners. The model below estimates each motif's effect
with the others held constant; the reference is a promoter carrying none of the four.

```{r promoter-adjusted}
if (has_promoter) {
  adj <- map_dfr(responses, function(resp) {
    d <- memb |> filter(!is.na(.data[[resp]]))
    mod <- lm(reformulate(motifs, response = resp), data = d)
    ci  <- confint(mod)
    tibble(response = resp,
           motif    = rownames(coef(summary(mod))),
           estimate = coef(summary(mod))[, "Estimate"],
           lower    = ci[, 1], upper = ci[, 2],
           t        = coef(summary(mod))[, "t value"],
           p        = coef(summary(mod))[, "Pr(>|t|)"]) |>
      filter(motif != "(Intercept)")
  }) |>
    mutate(motif = sub("TRUE$", "", motif))
  write_csv(adj, file.path(params$out_dir, "figureS7A_promoter_adjusted.csv"))

  adj |>
    mutate(response = factor(response, levels = responses)) |>
    arrange(response, motif) |>
    transmute(response, motif,
              `adjusted effect` = sprintf("%+.4f", estimate),
              `95% CI` = sprintf("[%+.3f, %+.3f]", lower, upper),
              t = sprintf("%.2f", t),
              p = format.pval(p, digits = 3, eps = 1e-16)) |>
    kable(caption = "Each motif adjusted for the other three (reference: none of the four)")
}
```


## Figure

```{r fig-s7a, fig.height = 3.6, fig.width = 11}
if (has_promoter) {
  pal_A <- c("CCAAT" = "#FBB4AE", "GC box" = "#B3CDE3", "Initiator" = "#DECBE4",
             "TATA" = "#FB8072", "no information" = "#CCEBC5")

  # genes are duplicated into every class they belong to — class sizes sum to more
  # than the number of genes, by construction
  d_long <- map_dfr(c(motifs, "none"), function(m) {
    memb |> filter(.data[[m]]) |>
      select(all_of(responses)) |>
      mutate(class = unname(motif_labels[m]))
  }) |>
    pivot_longer(all_of(responses), names_to = "response", values_to = "y") |>
    filter(!is.na(y)) |>
    mutate(class    = factor(class, levels = names(pal_A)),
           response = factor(response, levels = responses))

  yr    <- range(d_long$y, na.rm = TRUE)
  ann_A <- promoter_tbl |>
    mutate(class    = factor(as.character(class), levels = names(pal_A)),
           response = factor(response, levels = responses),
           label    = as.character(stars(p_BH_fig)),
           y        = yr[2] + 0.13 * diff(yr))

  p_A <- ggplot(d_long, aes(x = class, y = y, fill = class)) +
    geom_hline(yintercept = 0, linetype = "dashed", colour = "red", linewidth = 0.4) +
    geom_boxplot(outlier.size = 0.35, outlier.alpha = 0.45, outlier.stroke = 0,
                 linewidth = 0.3, colour = "grey20") +
    geom_text(data = ann_A, aes(x = class, y = y, label = label),
              inherit.aes = FALSE, size = 2.5) +
    facet_wrap(~ response, nrow = 1, labeller = labeller(response = resp_labels)) +
    scale_fill_manual(values = pal_A, guide = "none") +
    coord_cartesian(ylim = c(yr[1] - 0.02 * diff(yr), yr[2] + 0.22 * diff(yr))) +
    labs(title = "S7A  Promoter type", x = "Promoter type",
         y = expression(log[2]~"fold-change")) +
    theme_bw(base_size = 9) +
    theme(plot.title = element_text(face = "bold", size = 10),
          panel.grid = element_blank(),
          strip.background = element_blank(),
          strip.text = element_text(size = 8, lineheight = 1.05),
          axis.text.x = element_text(angle = 45, hjust = 1, size = 7),
          panel.spacing = unit(0.35, "lines"))
  print(p_A)
  ggsave(file.path(params$out_dir, "promoter_type.png"),
         plot = p_A, width = 11, height = 3.6, dpi = 300)
}
```

# Deleted-effects contrast results

Estimates are on the log~2~ scale: the difference between that bin's mean and the
unweighted mean of the other bins. `sig` and `p (BH, figure)` are the primary
inference (the false-discovery rate across all contrasts).

```{r contrast-tables, results = "asis"}
for (gv in panel_meta$variable) {
  pm <- panel_meta |> filter(variable == gv)
  cat("\n\n## ", pm$panel, " — ", pm$title, " (`", gv, "`)\n\n", sep = "")
  con_tbl |>
    filter(variable == gv) |>
    mutate(response = factor(response, levels = responses),
           sig      = as.character(stars(p_BH_fig))) |>
    arrange(response, level) |>
    transmute(response,
              level,
              n = n_level,
              estimate = sprintf("%+.4f", estimate_vs_rest),
              `95% CI` = sprintf("[%+.3f, %+.3f]", lower.CL, upper.CL),
              t = sprintf("%.2f", t.ratio),
              `p (BH, figure)` = format.pval(p_BH_fig, digits = 3, eps = 1e-16),
              sig,
              `p (mvt, panel)` = format.pval(p_mvt, digits = 3, eps = 1e-16)) |>
    kable(caption = paste0("Deleted-effect contrasts — ", pm$title)) |>
    print()
  cat("\n\n")
}
```

# Figures

```{r palettes}
palettes <- list(
  pausing   = rev(brewer.pal(8, "RdYlBu")),
  occupancy = colorRampPalette(c("#F7FBFF", "#08519C"))(8),
  length    = colorRampPalette(c("#FFF5EB", "#D94801"))(6)
)
```

```{r plot-function}
make_panel <- function(gv) {
  pm   <- panel_meta |> filter(variable == gv)
  levs <- order_levels(dat[[gv]], gv)
  pal  <- palettes[[pm$palette]]
  names(pal) <- levs[seq_along(pal)]

  d <- dat |>
    select(level = all_of(gv), all_of(responses)) |>
    pivot_longer(all_of(responses), names_to = "response", values_to = "y") |>
    filter(!is.na(y), !is.na(level)) |>
    mutate(level    = factor(level, levels = levs),
           response = factor(response, levels = responses))

  # Common y range across the four sub-panels
  yr    <- range(d$y, na.rm = TRUE)
  y_ann <- yr[2] + 0.13 * diff(yr)
  ylim  <- c(yr[1] - 0.02 * diff(yr), yr[2] + 0.22 * diff(yr))

  ann <- con_tbl |>
    filter(variable == gv) |>
    mutate(level    = factor(level, levels = levs),
           response = factor(response, levels = responses),
           label    = as.character(stars(p_BH_fig)),
           y        = y_ann)

  ggplot(d, aes(x = level, y = y, fill = level)) +
    geom_hline(yintercept = 0, linetype = "dashed", colour = "red", linewidth = 0.4) +
    geom_boxplot(outlier.size = 0.35, outlier.alpha = 0.45, outlier.stroke = 0,
                 linewidth = 0.3, colour = "grey20") +
    geom_text(data = ann, aes(x = level, y = y, label = label),
              inherit.aes = FALSE, size = 2.5, vjust = 0.5) +
    facet_wrap(~ response, nrow = 1,
               labeller = labeller(response = resp_labels)) +
    scale_fill_manual(values = pal, guide = "none") +
    coord_cartesian(ylim = ylim) +
    labs(title = paste0(pm$panel, "  ", pm$title),
         x = pm$xlab,
         y = expression(log[2]~"fold-change")) +
    theme_bw(base_size = 9) +
    theme(
      plot.title      = element_text(face = "bold", size = 10),
      panel.grid      = element_blank(),
      strip.background = element_blank(),
      strip.text      = element_text(size = 8, lineheight = 1.05),
      axis.text.x     = element_text(angle = 45, hjust = 1, size = 7),
      panel.spacing   = unit(0.35, "lines")
    )
}
```

## Pausing index

Genes are binned by their log~2~ pausing index at 0 h. The bins run from
de-paused (`<-2`) to strongly paused (`> 4`).

```{r fig-pausing-bin, fig.height = 3.6, fig.width = 11}
p_B <- make_panel("pausing_bin")
p_B
ggsave(file.path(params$out_dir, "pausing_index.png"),
       plot = p_B, width = 11, height = 3.6, dpi = 300)
```

## Baseline Pol II occupancy

Genes are binned into eight equal-count quantiles of log~2~ baseMean Pol II signal,
Q1 (lowest) to Q8 (highest).

```{r fig-s7c, fig.height = 3.6, fig.width = 11}
p_C <- make_panel("baseMean_bin")
p_C
ggsave(file.path(params$out_dir, "polII_occupancy.png"),
       plot = p_C, width = 11, height = 3.6, dpi = 300)
```

## Gene length

The 1–5 kb class has no gene-body measurements, so that x position is empty in the
two gene-body facets and carries no annotation.

```{r fig-s7d, fig.height = 3.6, fig.width = 11}
p_D <- make_panel("length_category")
p_D
ggsave(file.path(params$out_dir, "gene_length.png"),
       plot = p_D, width = 11, height = 3.6, dpi = 300)
```

# Continuous predictors — gene-level regression

Pausing index and gene length are continuous, and given that we can observe trends across 
ordered bins, we consider an analysis without binning to see if 
results hold up. We regress each log~2~ fold-change on the underlying continuous 
variable across all genes. We use heteroskedasticity-consistent standard error (HC3),
because the varaince is not constant along either predictor.

```{r continuous-fit}
dat <- dat |> mutate(log10_length = log10(length))

cont_preds <- tibble::tribble(
  ~var,                 ~label,                             ~axis,
  "log2_pausing_ratio", "Baseline pausing index (log2)",    "log2 pausing index (0 h)",
  "log10_length",       "Gene length (log10 bp)",           "Gene length (bp, log10 scale)"
)

continuous_fit <- function(resp, pred) {
  d2 <- dat |>
    transmute(y = .data[[resp]], x = .data[[pred]]) |>
    filter(!is.na(y), !is.na(x))
  m  <- lm(y ~ x, data = d2)
  V  <- sandwich::vcovHC(m, type = "HC3")
  b  <- unname(coef(m)[2]); se <- sqrt(V[2, 2]); df <- m$df.residual
  tstat <- b / se
  tibble(response = resp, predictor = pred, n = nrow(d2),
         slope = b, lower.CL = b - qt(0.975, df) * se, upper.CL = b + qt(0.975, df) * se,
         t.ratio = tstat, p_value = 2 * pt(-abs(tstat), df),
         spearman_rho = cor(d2$x, d2$y, method = "spearman"),
         r_squared = summary(m)$r.squared)
}

cont_tbl <- tidyr::expand_grid(response = responses, predictor = cont_preds$var) |>
  purrr::pmap_dfr(function(response, predictor) continuous_fit(response, predictor)) |>
  left_join(cont_preds, by = c("predictor" = "var"))
write_csv(cont_tbl, file.path(params$out_dir, "continuous_regression.csv"))

cont_tbl |>
  mutate(response = factor(response, levels = responses)) |>
  arrange(predictor, response) |>
  transmute(predictor = label, response, n,
            slope = sprintf("%+.4f", slope),
            `95% CI` = sprintf("[%+.4f, %+.4f]", lower.CL, upper.CL),
            t = sprintf("%.2f", t.ratio),
            p = format.pval(p_value, digits = 3, eps = 1e-16),
            `Spearman rho` = sprintf("%+.3f", spearman_rho),
            `R^2` = sprintf("%.4f", r_squared)) |>
  kable(caption = "Gene-level regression of each log2 fold-change on the continuous predictor (HC3 standard errors)")
```

```{r continuous-panel-fn}
make_cont_panel <- function(pred, title, xlab, log_x = FALSE) {
  d_long <- dat |>
    select(x = all_of(pred), all_of(responses)) |>
    pivot_longer(all_of(responses), names_to = "response", values_to = "y") |>
    filter(!is.na(x), !is.na(y)) |>
    mutate(response = factor(response, levels = responses))

  ann <- cont_tbl |>
    filter(predictor == pred) |>
    mutate(response = factor(response, levels = responses),
           label = sprintf("slope %+.3f\n%s\nrho %+.2f",
                           slope,
                           ifelse(p_value < 2e-16, "p < 2e-16",
                                  paste0("p = ", sprintf("%.3g", p_value))),
                           spearman_rho))

  yr <- quantile(d_long$y, c(0.001, 0.999), na.rm = TRUE)
  xr <- range(d_long$x, na.rm = TRUE)

  p <- ggplot(d_long, aes(x = x, y = y)) +
    geom_hline(yintercept = 0, linetype = "dashed", colour = "red", linewidth = 0.4) +
    geom_hex(bins = 55) +
    scale_fill_gradient(low = "grey88", high = "grey15", trans = "log10",
                        name = "genes", guide = guide_colourbar(barwidth = 0.4,
                                                                barheight = 4)) +
    geom_smooth(method = "gam", formula = y ~ s(x, bs = "cs"), se = FALSE,
                colour = "#1f78b4", linetype = "dashed", linewidth = 0.6) +
    geom_smooth(method = "lm", formula = y ~ x, se = FALSE,
                colour = "black", linewidth = 0.6) +
    geom_text(data = ann, aes(x = xr[1], y = yr[2], label = label), inherit.aes = FALSE,
              hjust = 0, vjust = 1, size = 2.1, lineheight = 1.05) +
    facet_wrap(~ response, nrow = 1, labeller = labeller(response = resp_labels)) +
    coord_cartesian(ylim = yr) +
    labs(title = title, x = xlab, y = expression(log[2]~"fold-change")) +
    theme_bw(base_size = 9) +
    theme(plot.title = element_text(face = "bold", size = 10),
          panel.grid = element_blank(),
          strip.background = element_blank(),
          strip.text = element_text(size = 8, lineheight = 1.05),
          panel.spacing = unit(0.35, "lines"))
  if (log_x) p <- p + scale_x_continuous(labels = function(v) scales::comma(10^v))
  p
}
```

## Pausing index as a continuous variable

```{r fig-s7e, fig.height = 3.6, fig.width = 11}
p_E <- make_cont_panel("log2_pausing_ratio",
                       "Baseline pausing index (continuous)",
                       expression(log[2]~"pausing index (0 h)"))
p_E
ggsave(file.path(params$out_dir, "pausing_continuous.png"),
       plot = p_E, width = 11, height = 3.6, dpi = 300)
```

## Gene length as a continuous variable

```{r fig-s7f, fig.height = 3.6, fig.width = 11}
p_F <- make_cont_panel("log10_length",
                       "Gene length (continuous)",
                       "Gene length (bp)", log_x = TRUE)
p_F
ggsave(file.path(params$out_dir, "length_continuous.png"),
       plot = p_F, width = 11, height = 3.6, dpi = 300)
```


```{r continuous-vs-binned}
bind_rows(
  cont_tbl |> transmute(response, predictor = label,
                        analysis = "continuous (gene-level)",
                        direction = ifelse(slope > 0, "positive", "negative"),
                        p = p_value),
  tidyr::expand_grid(response = responses,
                     variable = c("pausing_bin", "length_category")) |>
    purrr::pmap_dfr(function(response, variable) {
      d2 <- dat |>
        transmute(y = .data[[response]], group = as.character(.data[[variable]])) |>
        filter(!is.na(y), !is.na(group)) |>
        mutate(group = factor(group, levels = order_levels(group, variable))) |>
        droplevels()
      ct <- contrast(emmeans(lm(y ~ group, data = d2), ~ group), "poly", max.degree = 1)
      as.data.frame(summary(ct)) |>
        filter(contrast == "linear") |>
        transmute(response,
                  predictor = ifelse(variable == "pausing_bin",
                                     "Baseline pausing index (log2)", "Gene length (log10 bp)"),
                  analysis  = "linear trend across bins",
                  direction = ifelse(estimate > 0, "positive", "negative"),
                  p = p.value)
    })
) |>
  mutate(response = factor(response, levels = responses)) |>
  arrange(predictor, response, analysis) |>
  transmute(predictor, response, analysis, direction,
            p = format.pval(p, digits = 3, eps = 1e-16)) |>
  kable(caption = "Continuous gene-level slope against the binned linear trend, for the same data")
```

# Statistics for the Results text

The deleted-effect contrasts answer "does this bin differ from the other bins", 
which is appropriate for determining which levels of the stratification are 
especially effected, if any.

Several sentences in the Results make different claims, which need different tests:

| sentence in the Results | what it claims | test needed |
|---|---|---|
| "Pol II reduction at proximal promoters was observed across all gene lengths, expression levels…" | every bin is *below zero* | each bin mean vs 0 |
| "genes with a log~2~ pausing index > 2 lost more TSS-bound Pol II than genes with lower pausing indices" | one *set of bins* vs the complementary set, at a threshold read off the figure | pooled two-set contrast, corrected for the threshold search |
| "shorter genes tended to show reduced gene-body Pol II" | a *gradient* across the ordered bins | linear trend |

We now compute all three are computed below on the same models.

## Is each bin's mean different from zero?

```{r vs-zero}
vs_zero <- expand_grid(response = responses, variable = group_vars) |>
  pmap_dfr(function(response, variable) {
    d  <- dat |>
      transmute(y = .data[[response]], group = as.character(.data[[variable]])) |>
      filter(!is.na(y), !is.na(group)) |>
      mutate(group = factor(group, levels = order_levels(group, variable))) |>
      droplevels()
    em <- emmeans(lm(y ~ group, data = d), ~ group)
    as.data.frame(summary(em, infer = c(TRUE, TRUE), null = 0, adjust = "mvt")) |>
      transmute(response, variable, level = as.character(group),
                n = as.integer(table(d$group)[as.character(group)]),
                mean_log2FC = emmean, lower.CL, upper.CL, t.ratio, p_mvt = p.value)
  })
write_csv(vs_zero, file.path(params$out_dir, "level_vs_zero.csv"))

vs_zero |>
  mutate(response = factor(response, levels = responses)) |>
  group_by(response, variable) |>
  summarise(bins = n(),
            `below 0 (p < 0.05)` = sum(mean_log2FC < 0 & p_mvt < 0.05),
            `above 0 (p < 0.05)` = sum(mean_log2FC > 0 & p_mvt < 0.05),
            `not distinguishable from 0` = sum(p_mvt >= 0.05),
            .groups = "drop") |>
  arrange(response, variable) |>
  kable(caption = "How many bins differ from zero, per family (mvt-adjusted within family)")
```

All groups individually are below zero for TSS at 90 min, which is the statement in results.

## Pooled two-set contrasts

Coefficients are weighted by the number of genes in each bin, so the estimate
is the difference in mean log~2~ fold-change between the two gene sets.

```{r pooled-contrasts}
pooled_one <- function(resp, gv, set_a, label) {
  d <- dat |>
    transmute(y = .data[[resp]], group = as.character(.data[[gv]])) |>
    filter(!is.na(y), !is.na(group)) |>
    mutate(group = factor(group, levels = order_levels(group, gv))) |>
    droplevels()
  lv <- levels(d$group); a <- lv %in% set_a
  if (!any(a) || all(a)) return(NULL)
  n <- as.numeric(table(d$group)[lv])
  w <- ifelse(a, n / sum(n[a]), -n / sum(n[!a]))
  ct <- contrast(emmeans(lm(y ~ group, data = d), ~ group), setNames(list(w), label))
  as.data.frame(summary(ct, infer = c(TRUE, TRUE))) |>
    transmute(response = resp, variable = gv, comparison = label,
              levels_in_set = paste(lv[a], collapse = " + "),
              n_set = sum(n[a]), n_rest = sum(n[!a]),
              difference = estimate, lower.CL, upper.CL, df, t.ratio, p_value = p.value)
}

specs <- list(
  list(gv = "pausing_bin",     set = c("2 to 3", "3 to 4", "> 4"),
       lab = "pausing index > 2 vs <= 2"),
  list(gv = "pausing_bin",     set = "> 4",
       lab = "pausing index > 4 vs <= 4"),
  list(gv = "length_category", set = c("1\u20135 kb", "5\u201310 kb"),
       lab = "genes < 10 kb vs >= 10 kb"),
  list(gv = "baseMean_bin",    set = c("Q1", "Q2"),
       lab = "lowest-expressed quartiles (Q1-Q2) vs rest")
)
pooled_tbl <- map_dfr(responses, function(r) map_dfr(specs, ~ pooled_one(r, .x$gv, .x$set, .x$lab)))
write_csv(pooled_tbl, file.path(params$out_dir, "pooled_contrasts.csv"))

pooled_tbl |>
  mutate(response = factor(response, levels = responses)) |>
  arrange(comparison, response) |>
  transmute(comparison, response,
            `set / rest` = sprintf("%d / %d", n_set, n_rest),
            difference = sprintf("%+.4f", difference),
            `95% CI` = sprintf("[%+.3f, %+.3f]", lower.CL, upper.CL),
            t = sprintf("%.2f", t.ratio),
            p = format.pval(p_value, digits = 3, eps = 1e-16)) |>
  kable(caption = "Pooled, gene-weighted contrasts matching the Results sentences")
```


## Correcting for multiple possible thresholds

These thresholds are post-hoc (decided after examining the boxplots), not 
pre-specified, so we need to correct the p-values for many possible choices of
threshold.

An ordered factor with *k* levels admits *k*−1 splits into a low and a high set,
and all of them are contrasts on the same model — so `mvt` across those *k*−1 splits
corrects for having searched all thresholds, and is appropriate when there is
a strong correlation between nested splits.

```{r threshold-search}
split_search <- expand_grid(response = responses, variable = group_vars) |>
  pmap_dfr(function(response, variable) {
    d <- dat |>
      transmute(y = .data[[response]], group = as.character(.data[[variable]])) |>
      filter(!is.na(y), !is.na(group)) |>
      mutate(group = factor(group, levels = order_levels(group, variable))) |>
      droplevels()
    mod <- lm(y ~ group, data = d)
    lv <- levels(d$group); k <- length(lv); n <- as.numeric(table(d$group)[lv])
    W <- lapply(seq_len(k - 1), function(i) {
      a <- seq_along(lv) > i; ifelse(a, n / sum(n[a]), -n / sum(n[!a]))
    })
    names(W) <- paste0("above ", lv[seq_len(k - 1)])
    em <- emmeans(mod, ~ group)
    raw <- as.data.frame(summary(contrast(em, W, adjust = "none"), infer = c(TRUE, TRUE)))
    adj <- as.data.frame(summary(contrast(em, W, adjust = "mvt")))
    tibble(response, variable, split = names(W), n_high = vapply(W, function(w) sum(n[w > 0]), numeric(1)),
           difference = raw$estimate, lower.CL = raw$lower.CL, upper.CL = raw$upper.CL,
           p_raw = raw$p.value, p_splits_corrected = adj$p.value)
  })
write_csv(split_search, file.path(params$out_dir, "threshold_search_corrected.csv"))

# the splits the Results text names
used_splits <- tibble::tribble(
  ~response,                  ~variable,         ~split,
  "TSS_log2FC_90min_0h",      "pausing_bin",     "above 1 to 2",
  "TSS_log2FC_8h_0h",         "pausing_bin",     "above 3 to 4",
  "GeneBody_log2FC_90min_0h", "pausing_bin",     "above 1 to 2",
  "GeneBody_log2FC_8h_0h",    "pausing_bin",     "above 1 to 2",
  "GeneBody_log2FC_90min_0h", "length_category", "above 5\u201310 kb",
  "GeneBody_log2FC_8h_0h",    "length_category", "above 5\u201310 kb"
)

used_splits |>
  left_join(split_search, by = c("response", "variable", "split")) |>
  transmute(response, variable, split,
            difference = sprintf("%+.4f", difference),
            `p (raw)` = format.pval(p_raw, digits = 3, eps = 1e-16),
            `p (corrected)` =
              format.pval(p_splits_corrected, digits = 3, eps = 1e-16)) |>
  kable()
```

## Linear trend across the ordered bins

```{r trends}
trend <- expand_grid(response = responses, variable = group_vars) |>
  pmap_dfr(function(response, variable) {
    d <- dat |>
      transmute(y = .data[[response]], group = as.character(.data[[variable]])) |>
      filter(!is.na(y), !is.na(group)) |>
      mutate(group = factor(group, levels = order_levels(group, variable))) |>
      droplevels()
    ct <- contrast(emmeans(lm(y ~ group, data = d), ~ group), "poly", max.degree = 1)
    as.data.frame(summary(ct, infer = c(TRUE, TRUE))) |>
      filter(contrast == "linear") |>
      transmute(response, variable, k_levels = nlevels(d$group),
                linear_estimate = estimate, lower.CL, upper.CL, t.ratio, p_value = p.value)
  })
write_csv(trend, file.path(params$out_dir, "linear_trends.csv"))

trend |>
  mutate(response = factor(response, levels = responses)) |>
  arrange(variable, response) |>
  transmute(variable, response, k = k_levels,
            `linear contrast` = sprintf("%+.3f", linear_estimate),
            `95% CI` = sprintf("[%+.3f, %+.3f]", lower.CL, upper.CL),
            t = sprintf("%.2f", t.ratio),
            p = format.pval(p_value, digits = 3, eps = 1e-16)) |>
  kable(caption = "Linear component across the ordered bins")
```


```{r session}
sessionInfo()
```
