---
title: "Pol II IF with ARMC5 siRNA and transcriptional inhibitors"
author: "Scott Berry"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup_r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

rm(list=ls())
library(tidyverse)
library(patchwork)

wt_col <- "#184cee"
armc5_col <- "#9d24ec"

cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
```

# Experiment

# Data handling

Original data (Perkin-Elmer operetta `.tiff` files) are stored on UNSW's data archive under Berry Lab project D0419427. Data were maximum-intensity projected and converted to OME-TIFF using the `blimp` tools.

# Analysis

## Image analysis 

Image analysis performed using the following series of python scripts (found on github for [2023-10-06](https://github.com/berrygroup/publications/tree/main/2024_ARMC5/hct116_inhibitors_IF/))

- `illumination_correction_IF.ipynb`
- `save_example_images_IF.ipynb` (used to generate example images and training data for the cellpose model)
- `run_quantification_IF.py` (parallelised using e.g. `PBS/batch_quantify_230821.pbs`)

## Data analysis

Quantitative data analysis performed in this document.

### Load data

Load data from `scratch/berrylab` on UNSW's Katana HPC.

```{r load_data}

ana_dir <- "/srv/scratch/z3532965/src/publications/2024_ARMC5/hct116_inhibitors_IF"
data_dir <- "/srv/scratch/berrylab/z3532965/operetta_cls/ARMC5_TransInhib"
acquisition_dirs <- purrr::map_chr(
  c("230821_ARMC5_TransInhib/230821_ARMC5_TransInhib_TotalRNAPol__2023-08-25T14_29_20-Measurement 1",
    "231110_ARMC5_TransInhib/231110_ARMC5_TransInhib__2023-11-14T18_12_24-Measurement 1",
    "240322_ARMC5_TransInhib/240322_ARMC5_TransInhib__2024-03-26T13_50_23-Measurement 1"),~file.path(data_dir,.))
intensity_image_dirs <- map_chr(acquisition_dirs,~file.path(.,"OME-TIFF"))
intensity_image_mip_dirs <- map_chr(acquisition_dirs,~file.path(.,"OME-TIFF-MIP"))
label_image_dirs <- map_chr(acquisition_dirs,~file.path(.,"SEGMENTATION"))
quantification_dirs <- map_chr(acquisition_dirs,~file.path(.,"QUANTIFICATION"))

dir_name <- quantification_dirs[1]

read_quantification <- function(dir_name) {
  csv_files <- Sys.glob(file.path(dir_name,"*.csv"))
  dt_list <- lapply(csv_files, data.table::fread)
  df <- tibble(path = csv_files,
               feature_values = dt_list) %>%
    mutate(dir_name = dir_name,
           file_name = basename(path),
           feature_values = purrr::map(feature_values,tibble))
  return(df)
}

# features
f <- lapply(quantification_dirs,read_quantification) %>%
  bind_rows() %>%
  mutate(acquisition_dir = dirname(dirname(path)),
         imaging_date = basename(acquisition_dir) %>% str_extract("\\d{6}(?=_)"))
```

Load metadata and join with quantified image data.

```{r load metadata}

pad_well_names <- function(names) {
  gsub("(\\D)(\\d)$", "\\10\\2", names)
}

# layouts
plate_layouts <- tibble(layout_filename=c("230821_ARMC5_TransInhib.csv","231110_ARMC_TransInhib.csv","240322_ARMC5_TransInhib.csv"),
                        plating_date=c("230818","231107","240319"),
                        imaging_date=c("230821","231110","240322"),
                        metadata=purrr::map(layout_filename,~read_csv(file.path(ana_dir,"METADATA",.),show_col_types = FALSE, name_repair = ~sub(" ", "_", .)))) %>%
  unnest(metadata) %>%
  mutate(well_name=pad_well_names(WELL)) %>%
  select(-WELL) %>%
  rename_with(~gsub(" ", "_",str_to_lower(.))) %>%
  rename(siRNA=sirna)

# read image metadata and join to plate metadata
m <- tibble(acquisition_dir = acquisition_dirs) %>%
  mutate(image_dir = purrr::map_chr(acquisition_dirs,~file.path(.,"OME-TIFF-MIP"))) %>%
  mutate(metadata_file = purrr::map(image_dir,~list.files(.))) %>%
  unnest(metadata_file) %>%
  filter(grepl("metadata.csv$",metadata_file)) %>%
  mutate(dat = purrr::map(file.path(image_dir,metadata_file),~read_csv(.,show_col_types = FALSE))) %>%
  unnest(dat)

# attach image and plate metadata 
dat <- right_join(mutate(m,file_id = str_extract(URL, ".+?(?=\\.)")),
                  mutate(f,file_id = str_extract(file_name, ".+?(?=\\.)")),
                  by = join_by(acquisition_dir, file_id)) %>%
  rename(well_name=WellName) %>%
  left_join(filter(plate_layouts,!is.na(imaging_date)),by=join_by(imaging_date,well_name)) %>%
  select(-TimepointID) %>%
  unnest(feature_values)

```

Check that the data are complete.

```{r check_complete}

# check completeness (count wells)
dat %>% 
  distinct(acquisition_dir,well_name) %>%
  group_by(acquisition_dir) %>%
  count()

# check completeness (count sites per well)
dat %>% 
  distinct(acquisition_dir,well_name,FieldID) %>%
  group_by(acquisition_dir,well_name) %>%
  count() %>%
  filter(!(n %in% c(18,23,25)))

```

Some basic QC regarding number of cells per well, etc.

```{r cell_counts, fig.width=8,fig.height=12}

dat %>% 
  distinct(plating_date,imaging_date,well_name,siRNA,primary,secondary) %>%
  group_by(plating_date,imaging_date,siRNA,primary,secondary) %>%
  count() %>%
  ungroup() %>%
  ggplot(aes(x=siRNA,y=n)) + geom_col(position = "dodge") + facet_grid(interaction(primary,secondary)~plating_date) +
  coord_flip() +
  ggtitle("Number of wells of each condition")

dat %>% 
  group_by(plating_date,imaging_date,siRNA,primary,secondary,drug) %>%
  count() %>%
  ungroup() %>%
  ggplot(aes(x=siRNA,y=n,fill=drug)) + geom_col(position = "dodge") + facet_grid(interaction(primary,secondary)~plating_date) +
  coord_flip() +
  ggtitle("Total number of cells in each condition")

```

### Subtract background

Calculate background by maximum intensity projecting, illumination-correcting, and getting the mean intensity in a region containing no cells. See `calculate_background.ipynb` in the corresponding github repository. For the IF channel, we use the no-primary controls on each plate as the background.

```{r subtract_background}

# updated 23/6/24 

# 230821
bg1 <- tibble(
  imaging_date = "230821",
  bg_568 = mean(c(139.9,140.4,139.4)),
  bg_DAPI = mean(c(137.5,138.3,135.3)))

# 231110
bg2 <- tibble(
  imaging_date = "231110",
  bg_568 = mean(c(144.7,149.4,145.3)),
  bg_DAPI = mean(c(166.6,151.36,163.1)))

# 240322
bg3 <- tibble(
  imaging_date = "240322",
  bg_568 = mean(c(142.4,142.5,143.6)),
  bg_DAPI = mean(c(144.6,141.8,141.0)))

bg_manual <- bind_rows(bg1,bg2,bg3)

# for Alexa 568, we use the no primary control to subtract background

bg_no_primary_all <- dat %>%
  filter(primary=="No primary") %>%
  group_by(imaging_date,well_name,siRNA) %>%
  summarise(bg_568 = mean(`Nuclei_intensity_mean_Alexa 568`),.groups = 'drop')

bg_no_primary_all %>%
  ggplot(aes(x=siRNA,y=bg_568,col=imaging_date)) + geom_jitter()

bg_no_primary <- bg_no_primary_all %>%
  group_by(imaging_date) %>%
  summarise(bg_568_no_primary = mean(bg_568),.groups = 'drop')

dat_minus_bg <- left_join(dat,bg_manual,by="imaging_date") %>%
  left_join(bg_no_primary,by="imaging_date") %>%
  mutate(`Nuclei_intensity_mean_DAPI` = `Nuclei_intensity_mean_DAPI` - bg_DAPI,
         `Nuclei_intensity_mean_Alexa 568` = `Nuclei_intensity_mean_Alexa 568` - bg_568_no_primary,
         
         `Nuclei_intensity_median_DAPI` = `Nuclei_intensity_median_DAPI` - bg_DAPI,
         `Nuclei_intensity_median_Alexa 568` = `Nuclei_intensity_median_Alexa 568` - bg_568_no_primary)

```

### Sum intensities

Compute sum intensities by multiplying the background-subtracted means by the corresponding area

```{r sum_intensities}

dat_minus_bg <- mutate(dat_minus_bg,
                       `Nuclei_intensity_sum_DAPI` = `Nuclei_intensity_mean_DAPI` * Nuclei_area,
                       `Nuclei_intensity_sum_Alexa 568` = `Nuclei_intensity_mean_Alexa 568` * Nuclei_area)

```

### Data cleanup

Remove border cells 

```{r remove_variables}

dat_non_border <- filter(dat_minus_bg,!Nuclei_is_border)

```

# Check illumination correction

After correction of illumination biases across the field of view, there should be no bias for mean intensities across the field-of-view.

```{r test_illum_corr}

dat_non_border_sample_longer <- dat_non_border %>%
  filter(imaging_date=="240322") %>%
  filter(well_name=="F06") %>%
  select(Nuclei_centroid_0,Nuclei_centroid_1,Nuclei_intensity_mean_DAPI,`Nuclei_intensity_mean_Alexa 568`) %>%
  pivot_longer(-c(Nuclei_centroid_0,Nuclei_centroid_1)) 

rescaling_limits <- dat_non_border_sample_longer %>%
  group_by(name) %>%
  summarise(upper_limit = quantile(value,0.99))

dat_non_border_sample_longer <- left_join(dat_non_border_sample_longer,rescaling_limits,by="name")

channels <- unique(dat_non_border_sample_longer$name)

p <- list(length=length(channels))
for (i in 1:length(channels)) {
  p[[i]] <- dat_non_border_sample_longer %>%
    filter(name==channels[i]) %>%
    ggplot(aes(x=Nuclei_centroid_0,y=Nuclei_centroid_1,col=value/upper_limit)) +
    geom_point(size=0.5) +
    scale_color_viridis_c(limits=c(0,1)) +
    facet_wrap(~name) +
    coord_fixed() +
    theme(legend.position = "none")
}

wrap_plots(p) & ggtitle("Illumination biases")

```

## Check DAPI distributions

```{r check_DAPI}

dat_non_border %>%
  ggplot(aes(x=Nuclei_intensity_sum_DAPI,col=as.factor(well_name))) +
  geom_density() +
  facet_wrap(siRNA~imaging_date,scales = "free") +
  theme(legend.position = "none") +
  scale_x_continuous(limits = c(0,8e6))

```
## Nuclear Area

```{r nuclear_area_qc}

dat_non_border %>%
  ggplot(aes(x=Nuclei_area,col=as.factor(well_name))) +
  geom_density() +
  facet_wrap(imaging_date~siRNA,scales = "free") +
  theme(legend.position = "none") +
  scale_x_continuous(limits = c(0,2e4)) +
  geom_vline(xintercept = 700) +
  geom_vline(xintercept = 20000)

dat_non_border <- filter(dat_non_border,Nuclei_area>700 & Nuclei_area<20000)

```

```{r plot_time_courses}

dat_non_border <- dat_non_border %>% 
  mutate(siRNA=case_when(siRNA=="Scramble" ~ "Scrambled",
                         siRNA=="ARMC5 pool" ~ "ARMC5",
                         TRUE ~ siRNA)) %>%
  mutate(siRNA=factor(siRNA,levels=c("Scrambled","ARMC5")),
         drug=factor(drug,levels=c("Vehicle","Triptolide 1 µM","LDC4297 10 µM","THZ1 1 µM","AZD4573 1 µM","DRB 100 µM","0")))

well_summaries <- dat_non_border %>%
  group_by(imaging_date,siRNA,drug,drug_time_minutes,primary,secondary,well_name) %>%
  summarise(Nuclei_intensity_mean_Alexa_568=mean(`Nuclei_intensity_mean_Alexa 568`),.groups="drop") 

normalisers <- well_summaries %>%
  filter(siRNA=="Scrambled" & drug=='0' & primary=="Total Pol") %>%
  group_by(imaging_date) %>%
  summarise(day_mean_scrambled=mean(Nuclei_intensity_mean_Alexa_568))

well_summaries_normalised <- well_summaries %>%
  left_join(normalisers,by = join_by(imaging_date)) %>%
  mutate(Nuclei_intensity_mean_Alexa_568_plate_normalised=Nuclei_intensity_mean_Alexa_568/day_mean_scrambled)

starting_level_well_summaries_normalised <- well_summaries_normalised %>% 
  filter(drug=='0' & primary=="Total Pol") %>%
  select(-drug) %>%
  expand_grid(drug=unique(well_summaries$drug))

all_well_summaries_normalised <- bind_rows(well_summaries_normalised,starting_level_well_summaries_normalised)

all_overall_summaries_normalised <- all_well_summaries_normalised %>%
  group_by(siRNA,drug,drug_time_minutes,primary,secondary) %>%
  summarise(sd_Nuclei_intensity_mean_Alexa_568_plate_normalised=sd(Nuclei_intensity_mean_Alexa_568_plate_normalised),
            Nuclei_intensity_mean_Alexa_568_plate_normalised=mean(Nuclei_intensity_mean_Alexa_568_plate_normalised),
            .groups = "drop") %>%
  filter(drug!='0')

all_well_summaries_normalised %>%
  filter(drug!='0' & primary=="Total Pol") %>%
  ggplot(aes(x=drug_time_minutes/60,y=Nuclei_intensity_mean_Alexa_568_plate_normalised,col=siRNA)) +
  facet_wrap(~drug,nrow=1) +
  geom_point(pch=4,alpha=0.4,size=0.75) +
  geom_ribbon(data=all_overall_summaries_normalised,
                aes(fill=siRNA,
                    ymin=Nuclei_intensity_mean_Alexa_568_plate_normalised-sd_Nuclei_intensity_mean_Alexa_568_plate_normalised,
                    ymax=Nuclei_intensity_mean_Alexa_568_plate_normalised+sd_Nuclei_intensity_mean_Alexa_568_plate_normalised),
              col=NA,alpha=0.2) +
  geom_line(data=all_overall_summaries_normalised) +
  geom_point(data=all_overall_summaries_normalised,size=1) +
  scale_x_continuous(name="Time (h)",breaks=pretty) +
  scale_y_continuous(name="Mean nuclear\nRPB1 intensity (F12)",limits = c(0,1.9),expand = c(0,0),breaks=pretty) +
  scale_color_manual(values = c(wt_col,armc5_col)) +
  scale_fill_manual(values = c(wt_col,armc5_col)) +
  theme_bw(base_size = 8) +
  theme(panel.grid.minor=element_blank(),
        legend.key.size = unit(3,"mm"))
ggsave("PLOTS/HCT116_wt_ARMC5_kd_timecourse_total.pdf",width=5.5,height=1.5)

all_well_summaries_normalised %>%
  filter(drug!='0' & primary=="Total Pol" & siRNA=="Scrambled") %>%
  ggplot(aes(x=drug_time_minutes/60,y=Nuclei_intensity_mean_Alexa_568_plate_normalised,col=siRNA)) +
  facet_wrap(~drug,nrow=1) +
  geom_point(pch=4,alpha=0.4,size=0.75) +
  geom_ribbon(data=filter(all_overall_summaries_normalised,siRNA=="Scrambled"),
                aes(fill=siRNA,
                    ymin=Nuclei_intensity_mean_Alexa_568_plate_normalised-sd_Nuclei_intensity_mean_Alexa_568_plate_normalised,
                    ymax=Nuclei_intensity_mean_Alexa_568_plate_normalised+sd_Nuclei_intensity_mean_Alexa_568_plate_normalised),
              col=NA,alpha=0.2) +
  geom_line(data=filter(all_overall_summaries_normalised,siRNA=="Scrambled")) +
  geom_point(data=filter(all_overall_summaries_normalised,siRNA=="Scrambled"),size=1) +
  scale_x_continuous(name="Time (h)",breaks=pretty) +
  scale_y_continuous(name="Mean nuclear\nRPB1 intensity (F12)",limits = c(0,1.9),expand = c(0,0),breaks=pretty) +
  scale_color_manual(values = c(wt_col,armc5_col)) +
  scale_fill_manual(values = c(wt_col,armc5_col)) +
  theme_bw(base_size = 8) +
  theme(panel.grid.minor=element_blank(),
        legend.key.size = unit(3,"mm"))
ggsave("PLOTS/HCT116_wt_ARMC5_kd_timecourse_total_noARMC5.pdf",width=5.5,height=1.5)


```

```{r save_summaries}

write_csv(x = all_well_summaries_normalised,file = "SUMMARIES/Drug_IF_well_summaries.csv")
write_csv(x = all_overall_summaries_normalised,file = "SUMMARIES/Drug_IF_overall_summaries.csv")

```

```{r session_info}

sessionInfo()

```
