---
title: "Total Pol II IF in HCT116 cells with ARMC5 KD (EU experiment)"
author: "Scott Berry"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup_r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

rm(list=ls())
library(tidyverse)
library(patchwork)
library(lme4)
library(emmeans)

wt_col <- "#184cee"
armc5_col <- "#9d24ec"

cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
```

# Experiment

# Data handling

Original data (Nikon `.nd2` files) are stored on UNSW's data archive under Berry Lab project D0419427. Data were maximum-intensity projected and converted to OME-TIFF using the `blimp` tools.

# Analysis

## Image analysis 

Image analysis performed using the following series of python scripts (found on github for [hct116_EU](https://github.com/berrygroup/publications/tree/main/2024_ARMC5/hct116_EU))

- `illumination_correction_EU_IF.ipynb`
- `save_example_images_EU_IF.ipynb` (used to generate example images and training data for the cellpose model)
- `run_quantification_EU.py` (parallelised using e.g. `PBS/batch_quantify_IF_240628.pbs`)

## Data analysis

Quantitative data analysis performed in this document.

### Load data

Load data from `scratch/berrylab` on UNSW's Katana HPC.

```{r load_data}

ana_dir <- "/srv/scratch/z3532965/src/publications/2024_ARMC5/hct116_EU"
data_dir <- "/srv/scratch/berrylab/z3532965/NikonSpinningDisk"
acquisition_dirs <- purrr::map_chr(
  c("ARMC5_KD_EU/240628_ARMC5_EU/Cycle1_IF/20240701_151828_786",
    "ARMC5_KD_EU/240705_ARMC5_EU/Cycle1_IF/20240708_135644_856",
    "ARMC5_KD_EU/240713_ARMC5_EU/Cycle1_IF/20240715_144144_254"),~file.path(data_dir,.))
intensity_image_dirs <- map_chr(acquisition_dirs,~file.path(.,"OME-TIFF"))
intensity_image_mip_dirs <- map_chr(acquisition_dirs,~file.path(.,"OME-TIFF-MIP"))
label_image_dirs <- map_chr(acquisition_dirs,~file.path(.,"SEGMENTATION"))
quantification_dirs <- map_chr(acquisition_dirs,~file.path(.,"QUANTIFICATION"))

read_quantification <- function(dir_name) {
  csv_files <- Sys.glob(file.path(dir_name,"*.csv"))
  dt_list <- lapply(csv_files, data.table::fread)
  df <- tibble(path = csv_files,
               feature_values = dt_list) %>%
    mutate(dir_name = dir_name,
           file_name = basename(path),
           
           # acq3 has an extra two channels!
           #file_name = str_remove(file_name,",561,405"),
           
           feature_values = purrr::map(feature_values,tibble)) %>%
    separate(file_name,into = c("well_name","ch1","ch2","sample","id"),sep="[,_.]", extra = "drop") %>%
    mutate(well_name = str_replace(well_name, "^Well", ""),
           field_id = as.integer(id))
  return(df)
}

pad_well_names <- function(names) {
  gsub("(\\D)(\\d)$", "\\10\\2", names)
}

# features
f <- lapply(quantification_dirs,read_quantification) %>%
  bind_rows() %>%
  mutate(acquisition_dir = dirname(dirname(path)),
         imaging_date = basename(acquisition_dir) %>% str_extract("\\d{8}(?=_)")) %>%
  select(-c(ch1, ch2, sample))

# layouts
plate_layouts <- tibble(layout_filename=c("240628_ARMC5KD_EU.csv","240705_ARMC5KD_EU.csv","240713_ARMC5KD_EU.csv"),
                        plating_date=c("240628","240705","240713"),
                        imaging_date=c("20240701","20240708","20240715"),
                        metadata=purrr::map(layout_filename,~read_csv(file.path(ana_dir,"METADATA",.),show_col_types = FALSE, name_repair = ~sub(" ", "_", .)))) %>%
  unnest(metadata) %>%
  mutate(well_name=pad_well_names(WELL)) %>%
  select(-WELL) %>%
  left_join(distinct(f,acquisition_dir,imaging_date),by = join_by(imaging_date)) %>%
  rename(siRNA=SIRNA,primary=PRIMARY,secondary=SECONDARY,cell=CELL)

```

Check that the data are complete.

```{r check_complete}

# check completeness (count wells)
f %>% 
  distinct(acquisition_dir,well_name) %>%
  group_by(acquisition_dir) %>%
  count()

# check completeness (count sites per well)
f %>% 
  group_by(acquisition_dir,well_name) %>%
  count() %>%
  filter(!(n %in% c(9,16,24)))

```

Load metadata and join with quantified image data.

```{r load metadata}
# read image metadata and join to plate metadata
m <- tibble(acquisition_dir = acquisition_dirs) %>%
  mutate(image_dir = purrr::map_chr(acquisition_dirs,~file.path(.,"OME-TIFF-MIP"))) %>%
  mutate(metadata_file = purrr::map(image_dir,~list.files(.))) %>%
  unnest(metadata_file) %>%
  filter(grepl("metadata.csv$",metadata_file)) %>%
  mutate(dat = purrr::map(file.path(image_dir,metadata_file),~read_csv(.,show_col_types = FALSE)),
         # acq3 has an extra two channels!
         #metadata_file = str_remove(metadata_file,",561,405")
         ) %>%
  separate(metadata_file,into = c("well_name","ch1","ch2","sample"),sep="[,_.]", extra = "drop") %>%
  unnest(dat) %>%
  mutate(well_name = str_replace(well_name, "^Well", ""),
         field_id = as.integer(field_id)) %>%
  select(-c(ch1, ch2, sample)) %>%
  left_join(plate_layouts,by=join_by(acquisition_dir,well_name))

# attach image and plate metadata 
dat <- right_join(m,f,by = join_by(acquisition_dir, well_name, field_id, imaging_date)) %>%
  unnest(feature_values) %>%
  select(-matches("^Nuclei_x"))

```

Some basic QC regarding number of cells per well, etc.

```{r cell_counts, fig.width=8,fig.height=8}

dat %>% 
  distinct(cell,plating_date,imaging_date,well_name,siRNA,primary) %>%
  group_by(cell,plating_date,imaging_date,siRNA,primary) %>%
  count() %>%
  ungroup() %>%
  ggplot(aes(x=cell,y=n,fill=siRNA)) + geom_col(position = "dodge") + facet_grid(primary~imaging_date) +
  coord_flip() +
  ggtitle("Number of wells of each condition")

dat %>% 
  group_by(cell,plating_date,imaging_date,siRNA,primary) %>%
  count() %>%
  ungroup() %>%
  ggplot(aes(x=cell,y=n,fill=siRNA)) + geom_col(position = "dodge") + facet_grid(primary~imaging_date) +
  coord_flip() +
  ggtitle("Total number of cells in each condition")

```

### Subtract background

Calculate background by maximum intensity projecting, illumination-correcting, and getting the mean intensity in a region containing no cells.


```{r subtract_background}

# update 3/9/24 for EU-IF in HCT116

bg_561 <- mean(c(166.8,169.4,161.4))
bg_405 <-mean(c(152.0,153.7,150.5))

dat_minus_bg <- mutate(dat,
                       Nuclei_intensity_mean_561 = Nuclei_intensity_mean_561 - bg_561,
                       Nuclei_intensity_mean_405 = Nuclei_intensity_mean_405 - bg_405,
                       
                       Nuclei_intensity_median_561 = Nuclei_intensity_median_561 - bg_561,
                       Nuclei_intensity_median_405 = Nuclei_intensity_median_405 - bg_405)

```

### Sum intensities

Compute sum intensities by multiplying the background-subtracted means by the corresponding area

```{r sum_intensities}

dat_minus_bg <- mutate(dat_minus_bg,
                       Nuclei_intensity_sum_561 = Nuclei_intensity_mean_561 * Nuclei_area,
                       Nuclei_intensity_sum_405 = Nuclei_intensity_mean_405 * Nuclei_area)

```

## Data cleanup

Remove border cells 

```{r remove_variables}

dat_non_border <- filter(dat_minus_bg,!Nuclei_is_border)

```

### Check illumination correction

After correction of illumination biases across the field of view, there should be no bias for mean intensities across the field-of-view.

```{r test_illum_corr}

dat_non_border_sample_longer <- dat_non_border %>%
  filter(imaging_date=="20240701") %>%
  filter(well_name=="D03") %>%
  select(Nuclei_centroid_0,Nuclei_centroid_1,Nuclei_intensity_mean_405,Nuclei_intensity_mean_561) %>%
  pivot_longer(-c(Nuclei_centroid_0,Nuclei_centroid_1)) 

rescaling_limits <- dat_non_border_sample_longer %>%
  group_by(name) %>%
  summarise(upper_limit = quantile(value,0.99))

dat_non_border_sample_longer <- left_join(dat_non_border_sample_longer,rescaling_limits,by="name")

channels <- unique(dat_non_border_sample_longer$name)

p <- list(length=length(channels))
for (i in 1:length(channels)) {
  p[[i]] <- dat_non_border_sample_longer %>%
    filter(name==channels[i]) %>%
    ggplot(aes(x=Nuclei_centroid_0,y=Nuclei_centroid_1,col=value/upper_limit)) +
    geom_point(size=0.5) +
    scale_color_viridis_c(limits=c(0,1)) +
    facet_wrap(~name) +
    coord_fixed() +
    theme(legend.position = "none")
}

wrap_plots(p) & ggtitle("Illumination biases")

```

### Check DAPI distributions

```{r check_DAPI}

dat_non_border %>%
  ggplot(aes(x=Nuclei_intensity_sum_405,col=as.factor(well_name))) +
  geom_density() +
  facet_wrap(imaging_date~siRNA,scales = "free") +
  theme(legend.position = "none") +
  scale_x_continuous(limits = c(0,6e6)) +
  scale_y_continuous(labels = NULL)

```

### Nuclear Area

```{r nuclear_area_qc}

dat_non_border %>%
  ggplot(aes(x=Nuclei_area,col=as.factor(well_name))) +
  geom_density() +
  facet_wrap(imaging_date~siRNA,scales = "free") +
  theme(legend.position = "none") +
  geom_vline(xintercept = 150) +
  scale_y_continuous(labels = NULL)

dat_non_border <- filter(dat_non_border,Nuclei_area>150)

```

## Quantify stains of interest

### Check IF distributions

```{r IF_by_cell, fig.width=14,fig.height=10}

dat_non_border <- mutate(dat_non_border,
                         siRNA = case_when(siRNA=="Scramble" ~ "Scrambled",
                                           siRNA=="ARMC5 pool" ~ "ARMC5",
                                           siRNA=="No reaction" ~ "None",
                                           TRUE ~ NA_character_),
                         siRNA=factor(siRNA,levels=c("None","Scrambled","ARMC5"))) 

dat_non_border %>% 
  ggplot(aes(x=siRNA,y=Nuclei_intensity_mean_561,col=siRNA,group=interaction(well_name,siRNA))) + 
  #geom_boxplot(outlier.shape=NA) + 
  geom_boxplot() + 
  facet_grid(primary~imaging_date) +
  #coord_cartesian(ylim = c(0,1000)) +
  theme_bw(base_size = 8) +
  theme(axis.text.x = element_text(angle=90,hjust=1,vjust=0.5))

```

Seems to be some very bright intensities in some wells. Check for outliers using boxplot rule.

```{r outlier_sites_detection}

field_averages <- dat_non_border %>%
  group_by(image_dir,filename_ome_tiff,primary,siRNA,imaging_date,well_name,field_id) %>%
  summarise(Nuclei_intensity_mean_561_mean=mean(Nuclei_intensity_mean_561),.groups = 'drop') 

outlier_fields <- field_averages %>%
  group_by(image_dir,primary,siRNA,imaging_date) %>%
  summarise(across(Nuclei_intensity_mean_561_mean,c(p25=~quantile(.,0.25),p75=~quantile(.,0.75))),.groups = 'drop') %>%
  left_join(field_averages,by = join_by(image_dir, primary, siRNA, imaging_date)) %>%
  mutate(Nuclei_intensity_mean_561_mean_IQR = Nuclei_intensity_mean_561_mean_p75 - Nuclei_intensity_mean_561_mean_p25,
         outlier = case_when(
           Nuclei_intensity_mean_561_mean > Nuclei_intensity_mean_561_mean_p75 + 1.5*Nuclei_intensity_mean_561_mean_IQR ~ TRUE,
           Nuclei_intensity_mean_561_mean < Nuclei_intensity_mean_561_mean_p25 - 1.5*Nuclei_intensity_mean_561_mean_IQR ~ TRUE,
           TRUE ~ FALSE))

outlier_fields %>%
  ggplot(aes(y=Nuclei_intensity_mean_561_mean,x=siRNA,col=outlier)) +
  geom_jitter() +
  scale_y_log10() +
  facet_wrap(primary~imaging_date)

outlier_fields %>%
  count(imaging_date,primary,siRNA,outlier) %>%
  group_by(imaging_date,primary,siRNA) %>%
  mutate(freq=n/sum(n)) %>%
  filter(outlier=="TRUE")

```

Remove outlier fields-of-view

```{r outlier_sites_removal}

dat_non_border <- dat_non_border %>%
  anti_join(filter(outlier_fields,outlier==TRUE),by = join_by(image_dir, well_name, field_id, filename_ome_tiff, imaging_date, siRNA, primary))

```

### Check plate layouts

Visualise plate layouts and examine the well-means across the plates for each experiment.

```{r check_plate_layouts, fig.width=9,fig.height=3}

dat_non_border %>%
  mutate(well_row = factor(str_extract(well_name, "[A-Z]"),levels=LETTERS),
         well_column = as.integer(as.character(str_extract(well_name, "[0-9]+")))) %>%
  distinct(imaging_date,well_name,well_row,well_column,siRNA) %>%
  ggplot(aes(x=well_column,y=well_row,fill=siRNA)) +
  geom_tile() +
  scale_y_discrete(limits=rev) +
  facet_wrap(~imaging_date) +
  ggtitle("siRNA")

dat_non_border %>%
  mutate(well_row = factor(str_extract(well_name, "[A-Z]"),levels=LETTERS),
         well_column = as.integer(as.character(str_extract(well_name, "[0-9]+")))) %>%
  distinct(imaging_date,well_name,well_row,well_column,primary) %>%
  ggplot(aes(x=well_column,y=well_row,fill=primary)) +
  geom_tile() +
  scale_y_discrete(limits=rev) +
  facet_wrap(~imaging_date) +
  ggtitle("primary")


# compute fold-change from mean for each well
dat_non_border %>%
  mutate(well_row = factor(str_extract(well_name, "[A-Z]"),levels=LETTERS),
         well_column = as.integer(as.character(str_extract(well_name, "[0-9]+")))) %>%
  group_by(imaging_date,well_name,well_row,well_column,primary,siRNA) %>%
  summarise(well_mean_IF = mean(Nuclei_intensity_mean_561),.groups="drop") %>%
  left_join(dat_non_border %>%
              group_by(imaging_date,well_name,primary,siRNA) %>%
              summarise(well_mean_IF = mean(Nuclei_intensity_mean_561),.groups="drop") %>%
              group_by(imaging_date,primary,siRNA) %>%
              summarise(plate_mean_IF = mean(well_mean_IF),.groups="drop"),
            by = join_by(imaging_date, primary,siRNA)) %>%
  mutate(log2FC_from_mean = log2(well_mean_IF/plate_mean_IF)) %>%
  ggplot(aes(x=well_column,y=well_row,fill=log2FC_from_mean)) +
  geom_tile() +
  scale_fill_distiller(type = "div", palette = "RdBu", limits = c(-2,2)) +
  scale_y_discrete(limits=rev) +
  facet_wrap(~imaging_date) +
  ggtitle("Within plate difference from mean per well")

# compute fold-change from mean for each well
dat_non_border %>%
  mutate(well_row = factor(str_extract(well_name, "[A-Z]"),levels=LETTERS),
         well_column = as.integer(as.character(str_extract(well_name, "[0-9]+")))) %>%
  group_by(imaging_date,well_name,well_row,well_column,primary,siRNA) %>%
  summarise(well_mean_IF = mean(Nuclei_intensity_mean_561),.groups="drop") %>%
  left_join(dat_non_border %>%
              group_by(imaging_date,well_name,primary,siRNA) %>%
              summarise(well_mean_IF = mean(Nuclei_intensity_mean_561),.groups="drop") %>%
              group_by(primary,siRNA) %>%
              summarise(overall_mean_IF = mean(well_mean_IF),.groups="drop"),
            by = join_by(primary,siRNA)) %>%
  mutate(log2FC_from_mean = log2(well_mean_IF/overall_mean_IF)) %>%
  ggplot(aes(x=well_column,y=well_row,fill=log2FC_from_mean)) +
  geom_tile() +
  scale_fill_distiller(type = "div", palette = "RdBu", limits = c(-2,2)) +
  scale_y_discrete(limits=rev) +
  facet_wrap(~imaging_date) +
  ggtitle("Experiment difference from mean per well")

```

Correct for plate-to-plate variation in IF staining for plotting (don't use normalisation for statistics, just include plate and well as random effects.)

```{r plate_normalisation}

normalisation_data <- dat_non_border %>%
  filter(primary=="Total Pol II" & siRNA!="None") %>%
  group_by(imaging_date,well_name,siRNA) %>%
  summarise(well_mean_IF = mean(Nuclei_intensity_mean_561),.groups="drop") %>%
  group_by(imaging_date,siRNA) %>%
  summarise(plate_mean_IF = mean(well_mean_IF),.groups="drop")

normalisation_factors <- normalisation_data %>%
  filter(imaging_date=="20240701") %>%
  select(-imaging_date) %>%
  rename(normalise_to = plate_mean_IF) %>%
  left_join(normalisation_data,by="siRNA") %>%
  mutate(normalisation_factor=normalise_to/plate_mean_IF) %>%
  group_by(imaging_date) %>%
  summarise(plate_normaliser = mean(normalisation_factor),.groups = "drop")

normalisation_factors

```

Now average over replicates, including plate-to-plate normalisation factors.

```{r average_over_replicates,fig.height=4,fig.width=12}

well_summaries <- dat_non_border %>%
  filter(primary=="Total Pol II" & siRNA!="None") %>%
  left_join(normalisation_factors, by = join_by(imaging_date)) %>%
  mutate(Nuclei_intensity_mean_561_plate_normalised = Nuclei_intensity_mean_561 * plate_normaliser) %>%
  group_by(imaging_date,well_name,siRNA) %>%
  summarise(Nuclei_intensity_mean_561 = mean(Nuclei_intensity_mean_561),
            Nuclei_intensity_mean_561_plate_normalised = mean(Nuclei_intensity_mean_561_plate_normalised),
            .groups='drop')

overall_summaries <- well_summaries %>%
  group_by(siRNA) %>%
  summarise(mean_Nuclei_intensity_mean_561 = mean(Nuclei_intensity_mean_561),
            sd_Nuclei_intensity_mean_561 = sd(Nuclei_intensity_mean_561),
            mean_Nuclei_intensity_mean_561_plate_normalised = mean(Nuclei_intensity_mean_561_plate_normalised),
            sd_Nuclei_intensity_mean_561_plate_normalised = sd(Nuclei_intensity_mean_561_plate_normalised),
            .groups='drop')

axis_rescaler <- filter(overall_summaries,siRNA=="Scrambled") %>% pull(mean_Nuclei_intensity_mean_561_plate_normalised)

no_primary_well_summaries <- dat_non_border %>%
  filter(primary=="No primary") %>%
  left_join(normalisation_factors, by = join_by(imaging_date)) %>%
  mutate(Nuclei_intensity_mean_561_plate_normalised = Nuclei_intensity_mean_561 * plate_normaliser) %>%
  group_by(imaging_date,well_name,siRNA) %>%
  summarise(Nuclei_intensity_mean_561 = mean(Nuclei_intensity_mean_561),
            Nuclei_intensity_mean_561_plate_normalised = mean(Nuclei_intensity_mean_561_plate_normalised),
            .groups='drop')

no_primary_overall_summaries <- no_primary_well_summaries %>%
  group_by(siRNA) %>%
  summarise(mean_Nuclei_intensity_mean_561 = mean(Nuclei_intensity_mean_561),
            sd_Nuclei_intensity_mean_561 = sd(Nuclei_intensity_mean_561),
            mean_Nuclei_intensity_mean_561_plate_normalised = mean(Nuclei_intensity_mean_561_plate_normalised),
            sd_Nuclei_intensity_mean_561_plate_normalised = sd(Nuclei_intensity_mean_561_plate_normalised),
            .groups='drop')

with_primary <- well_summaries %>%
  ggplot(aes(x=siRNA,y=Nuclei_intensity_mean_561_plate_normalised/axis_rescaler,col=siRNA)) +
  geom_col(data = overall_summaries, aes(y=mean_Nuclei_intensity_mean_561_plate_normalised/axis_rescaler,fill=siRNA),col=NA,alpha=0.4,width=0.8) +
  geom_point(data = overall_summaries, aes(y=mean_Nuclei_intensity_mean_561_plate_normalised/axis_rescaler)) +
  geom_errorbar(data = overall_summaries, aes(y=mean_Nuclei_intensity_mean_561_plate_normalised/axis_rescaler,
                                              ymin=(mean_Nuclei_intensity_mean_561_plate_normalised-sd_Nuclei_intensity_mean_561_plate_normalised)/axis_rescaler,
                                              ymax=(mean_Nuclei_intensity_mean_561_plate_normalised+sd_Nuclei_intensity_mean_561_plate_normalised)/axis_rescaler),width=0.25) +
  geom_jitter(width=0.25,pch=4,alpha=0.3) +
  scale_y_continuous(limits = c(0,1.8),name="Mean nuclear\nRPB1 intensity",expand = c(0,0)) +
  theme_bw(base_size = 8) + 
  scale_color_manual(values = c(wt_col,armc5_col)) +
  scale_fill_manual(values = c(wt_col,armc5_col)) +
  theme(axis.text.x = element_text(angle=45,hjust=1,vjust=1),
        axis.title.x = element_blank(),
        panel.grid = element_blank(),
        legend.key.size = unit(2.5,"mm"))

no_primary <- no_primary_well_summaries %>%
  ggplot(aes(x="No primary",y=Nuclei_intensity_mean_561_plate_normalised/axis_rescaler,col=siRNA)) +
  geom_col(data = no_primary_overall_summaries, aes(y=mean_Nuclei_intensity_mean_561_plate_normalised/axis_rescaler,fill=siRNA),col=NA,alpha=0.4,width=0.8) +
  geom_point(data = no_primary_overall_summaries, aes(y=mean_Nuclei_intensity_mean_561_plate_normalised/axis_rescaler)) +
  geom_errorbar(data = no_primary_overall_summaries, aes(y=mean_Nuclei_intensity_mean_561_plate_normalised/axis_rescaler,
                                              ymin=(mean_Nuclei_intensity_mean_561_plate_normalised-sd_Nuclei_intensity_mean_561_plate_normalised)/axis_rescaler,
                                              ymax=(mean_Nuclei_intensity_mean_561_plate_normalised+sd_Nuclei_intensity_mean_561_plate_normalised)/axis_rescaler),width=0.25) +
  geom_jitter(width=0.25,pch=4,alpha=0.3) +
  scale_y_continuous(limits = c(0,1.8),name="Mean nuclear\nRPB1 intensity",expand = c(0,0)) +
  theme_bw(base_size = 8) + 
  scale_color_manual(values = c("grey")) +
  scale_fill_manual(values = c("grey")) +
  theme(axis.text.x = element_text(angle=45,hjust=1,vjust=1),
        axis.title.x = element_blank(),
        axis.title.y=element_blank(),
        panel.grid = element_blank(),
        legend.key.size = unit(2.5,"mm"))

with_primary + no_primary + plot_layout(widths=c(2,1),guides="collect")

ggsave("PLOTS/IF_barplots_by_siRNA.pdf",width=6,height=4,units="cm")

```

```{r well_numbers}

well_summaries %>%
  group_by(imaging_date,siRNA) %>%
  count()

well_summaries %>%
  group_by(siRNA) %>%
  count()

```

```{r save_summaries}

write_csv(x = well_summaries,file = "SUMMARIES/siRNA_EU_IF_well_summaries.csv")
write_csv(x = overall_summaries,file = "SUMMARIES/siRNA_EU_IF_overall_summaries.csv")

```


```{r session_info}

sessionInfo()

```
