---
title: "Poly(A) FISH in ARMC5 KO cells"
author: "Scott Berry"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup_r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

rm(list=ls())
library(tidyverse)
library(patchwork)

wt_col <- "#184cee"
armc5_col <- "#9d24ec"

cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
```

# Experiment

# Data handling

Original data (Nikon `.nd2` files) are stored on UNSW's data archive under Berry Lab project D0419427. Data were maximum-intensity projected and converted to OME-TIFF using the `blimp` tools.

# Analysis

## Image analysis 

Image analysis performed using the following series of python scripts (found on github for [hek_polyA](https://github.com/berrygroup/publications/tree/main/2024_ARMC5/hek_polyA))

- `illumination_correction_polyA_1.ipynb` and `illumination_correction_polyA_2.ipynb`
- `save_example_images_polyA.ipynb` (used to generate example images and training data for the cellpose model)
- `run_quantification_polyA_1.py` (parallelised using `PBS/batch_quantify_polyA_240315.pbs`)
- `run_quantification_polyA_2.py` (parallelised using `PBS/batch_quantify_polyA_240322.pbs` and `PBS/batch_quantify_polyA_240329.pbs`)

## Data analysis

Quantitative data analysis performed in this document.

### Load data

Load data from `scratch/berrylab` on UNSW's Katana HPC.

```{r load_data}

ana_dir <- "/srv/scratch/z3532965/src/publications/2024_ARMC5/hek_polyA"
data_dir <- "/srv/scratch/berrylab/z3532965/NikonSpinningDisk"
acquisition_dirs <- purrr::map_chr(
  c("240315/240314_ARMC5KO_PolyA/20240315_142116_487",
    "240322/240322_ARMC5KO_PolyA/20240322_143358_441",
    "240403/240329_ARMC5KO_PolyA/20240403_160700_583"),~file.path(data_dir,.))
intensity_image_dirs <- map_chr(acquisition_dirs,~file.path(.,"OME-TIFF"))
intensity_image_mip_dirs <- map_chr(acquisition_dirs,~file.path(.,"OME-TIFF-MIP"))
label_image_dirs <- map_chr(acquisition_dirs,~file.path(.,"SEGMENTATION"))
quantification_dirs <- map_chr(acquisition_dirs,~file.path(.,"QUANTIFICATION"))

read_quantification <- function(dir_name) {
  csv_files <- Sys.glob(file.path(dir_name,"*.csv"))
  dt_list <- lapply(csv_files, data.table::fread)
  df <- tibble(path = csv_files,
               feature_values = dt_list) %>%
    mutate(dir_name = dir_name,
           file_name = basename(path),
           feature_values = purrr::map(feature_values,tibble)) %>%
    separate(file_name,into = c("well_name","ch1","ch2","ch3","ch4","sample","id"),sep="[,_.]", extra = "drop") %>%
    mutate(well_name = str_replace(well_name, "^Well", ""),
           field_id = as.integer(id))
  return(df)
}

pad_well_names <- function(names) {
  gsub("(\\D)(\\d)$", "\\10\\2", names)
}

# features
f <- lapply(quantification_dirs,read_quantification) %>%
  bind_rows() %>%
  mutate(acquisition_dir = dirname(dirname(path)),
         imaging_date = basename(acquisition_dir) %>% str_extract("\\d{8}(?=_)")) %>%
  select(-c(ch1, ch2, ch3, ch4, sample))

# layouts
plate_layouts <- tibble(layout_filename=c("240311_ARMC5KO_PolyA_longlayout.csv","240311_ARMC5KO_PolyA_longlayout.csv","240311_ARMC5KO_PolyA_longlayout.csv"),
                        plating_date=c("240311","240319","240329"),
                        imaging_date=c("20240315","20240322","20240403"),
                        metadata=purrr::map(layout_filename,~read_csv(file.path(ana_dir,"METADATA",.),show_col_types = FALSE, name_repair = ~sub(" ", "_", .)))) %>%
  unnest(metadata) %>%
  mutate(well_name=pad_well_names(Well_name)) %>%
  select(-Well_name) %>%
  left_join(distinct(f,acquisition_dir,well_name,imaging_date),by = join_by(imaging_date, well_name))

```

Check that the data are complete.

```{r check_complete}

# check completeness (count wells)
f %>% 
  distinct(acquisition_dir,well_name) %>%
  group_by(acquisition_dir) %>%
  count()

# check completeness (count sites per well)
f %>% 
  group_by(acquisition_dir,well_name) %>%
  count() %>%
  filter(!(n %in% c(16)))

```

Load metadata and join with quantified image data.

```{r load metadata}
# read image metadata and join to plate metadata
m <- tibble(acquisition_dir = acquisition_dirs) %>%
  mutate(image_dir = purrr::map_chr(acquisition_dirs,~file.path(.,"OME-TIFF-MIP"))) %>%
  mutate(metadata_file = purrr::map(image_dir,~list.files(.))) %>%
  unnest(metadata_file) %>%
  filter(grepl("metadata.csv$",metadata_file)) %>%
  mutate(dat = purrr::map(file.path(image_dir,metadata_file),~read_csv(.,show_col_types = FALSE))) %>%
  separate(metadata_file,into = c("well_name","ch1","ch2","ch3","ch4","sample"),sep="[,_.]", extra = "drop") %>%
  unnest(dat) %>%
  mutate(well_name = str_replace(well_name, "^Well", ""),
           field_id = as.integer(field_id)) %>%
  select(-c(ch1, ch2, ch3, ch4, sample)) %>%
  left_join(filter(plate_layouts,!is.na(acquisition_dir)),by=join_by(acquisition_dir,well_name))

# attach image and plate metadata 
dat <- right_join(m,f,by = join_by(acquisition_dir, well_name, field_id, imaging_date)) %>%
  unnest(feature_values) %>%
  select(-matches("None$"))

```

Some basic QC regarding number of cells per well, etc.

```{r cell_counts, fig.width=8,fig.height=12}

dat %>% 
  distinct(Cell,plating_date,imaging_date,well_name,Hybridisation) %>%
  group_by(Cell,plating_date,imaging_date,Hybridisation) %>%
  count() %>%
  ungroup() %>%
  ggplot(aes(x=Cell,y=n)) + geom_col(position = "dodge") + facet_grid(Hybridisation~imaging_date) +
  coord_flip() +
  ggtitle("Number of wells of each condition")

dat %>% 
  group_by(Cell,plating_date,imaging_date,Hybridisation) %>%
  count() %>%
  ungroup() %>%
  ggplot(aes(x=Cell,y=n)) + geom_col(position = "dodge") + facet_grid(Hybridisation~imaging_date) +
  coord_flip() +
  ggtitle("Total number of cells in each condition")

```

### Subtract background

Calculate background by maximum intensity projecting, illumination-correcting, and getting the mean intensity in a region containing no cells. See `calculate_background.ipynb` in the corresponding github repository.


```{r subtract_background}

bg_488 <-mean(c(163.671,158.321,161.861))
bg_647 <- mean(c(131.045,129.943,133.541))
bg_405 <-mean(c(193.651,193.498,192.183))

dat_minus_bg <- mutate(dat,
                       Nuclei_intensity_mean_647 = Nuclei_intensity_mean_647 - bg_647,
                       Nuclei_intensity_mean_405 = Nuclei_intensity_mean_405 - bg_405,
                       Nuclei_intensity_mean_488 = Nuclei_intensity_mean_488 - bg_488,
                       
                       Cell_intensity_mean_647 = Cell_intensity_mean_647 - bg_647,
                       Cell_intensity_mean_405 = Cell_intensity_mean_405 - bg_405,
                       Cell_intensity_mean_488 = Cell_intensity_mean_488 - bg_488,
                       
                       Nuclei_intensity_median_647 = Nuclei_intensity_median_647 - bg_647,
                       Nuclei_intensity_median_405 = Nuclei_intensity_median_405 - bg_405,
                       Nuclei_intensity_median_488 = Nuclei_intensity_median_488 - bg_488,
                       
                       Cell_intensity_median_647 = Cell_intensity_median_647 - bg_647,
                       Cell_intensity_median_405 = Cell_intensity_median_405 - bg_405,
                       Cell_intensity_median_488 = Cell_intensity_median_488 - bg_488)

```

### Sum intensities

Compute sum intensities by multiplying the background-subtracted means by the corresponding area

```{r sum_intensities}

dat_minus_bg <- mutate(dat_minus_bg,
                       Nuclei_intensity_sum_647 = Nuclei_intensity_mean_647 * Nuclei_area,
                       Nuclei_intensity_sum_405 = Nuclei_intensity_mean_405 * Nuclei_area,
                       Nuclei_intensity_sum_488 = Nuclei_intensity_mean_488 * Nuclei_area,
                       
                       Cell_intensity_sum_647 = Cell_intensity_mean_647 * Cell_area,
                       Cell_intensity_sum_405 = Cell_intensity_mean_405 * Cell_area,
                       Cell_intensity_sum_488 = Cell_intensity_mean_488 * Cell_area,
                       
                       Cytoplasm_intensity_sum_647 = Cell_intensity_sum_647 - Nuclei_intensity_sum_647,
                       Cytoplasm_intensity_sum_405 = Cell_intensity_sum_405 - Nuclei_intensity_sum_405,
                       Cytoplasm_intensity_sum_488 = Cell_intensity_sum_488 - Nuclei_intensity_sum_488,
                       
                       Cytoplasm_area = Cell_area - Nuclei_area,
                       
                       Cytoplasm_intensity_mean_647 = Cytoplasm_intensity_sum_647 / Cytoplasm_area,
                       Cytoplasm_intensity_mean_405 = Cytoplasm_intensity_sum_405 / Cytoplasm_area,
                       Cytoplasm_intensity_mean_488 = Cytoplasm_intensity_sum_488 / Cytoplasm_area)

```

### Data cleanup

Remove border cells 

```{r remove_variables}

dat_non_border <- filter(dat_minus_bg,!Nuclei_is_border)

```

# Check illumination correction

After correction of illumination biases across the field of view, there should be no bias for mean intensities across the field-of-view.

```{r test_illum_corr}

dat_non_border_sample_longer <- dat_non_border %>%
  filter(imaging_date=="20240322") %>%
  filter(well_name=="F04") %>%
  select(Nuclei_centroid_0,Nuclei_centroid_1,Nuclei_intensity_mean_405,Nuclei_intensity_mean_488,Nuclei_intensity_mean_647) %>%
  pivot_longer(-c(Nuclei_centroid_0,Nuclei_centroid_1)) 

rescaling_limits <- dat_non_border_sample_longer %>%
  group_by(name) %>%
  summarise(upper_limit = quantile(value,0.99))

dat_non_border_sample_longer <- left_join(dat_non_border_sample_longer,rescaling_limits,by="name")

channels <- unique(dat_non_border_sample_longer$name)

p <- list(length=length(channels))
for (i in 1:length(channels)) {
  p[[i]] <- dat_non_border_sample_longer %>%
    filter(name==channels[i]) %>%
    ggplot(aes(x=Nuclei_centroid_0,y=Nuclei_centroid_1,col=value/upper_limit)) +
    geom_point(size=0.5) +
    scale_color_viridis_c(limits=c(0,1)) +
    facet_wrap(~name) +
    coord_fixed() +
    theme(legend.position = "none")
}

wrap_plots(p) & ggtitle("Illumination biases")

```

## Check DAPI distributions

```{r check_DAPI}

dat_non_border %>%
  ggplot(aes(x=Nuclei_intensity_sum_405,col=as.factor(well_name))) +
  geom_density() +
  facet_wrap(imaging_date~Cell,scales = "free") +
  theme(legend.position = "none") +
  scale_x_continuous(limits = c(0,8e6))

dat_non_border %>%
  ggplot(aes(x=Nuclei_area,col=as.factor(well_name))) +
  geom_density() +
  facet_wrap(imaging_date~Cell,scales = "free") +
  theme(legend.position = "none") +
  scale_x_continuous()

```

## Check poly(A) distributions

```{r FISH_by_treatment, fig.width=8,fig.height=4}

dat_non_border <- filter(dat_non_border)

dat_non_border %>%
  ggplot(aes(x=Cell,y=Nuclei_intensity_mean_647,group=interaction(well_name,Hybridisation))) + 
  geom_boxplot(aes(fill=Hybridisation),outlier.shape=NA) + facet_grid(.~imaging_date) +
  coord_cartesian(ylim = c(0,1500)) +
  theme_bw(base_size = 8) +
  theme(axis.text.x = element_text(angle = 90,hjust=1,vjust=0.5))

ggsave("PLOTS/FISH_boxplot_per_well_by_cell.pdf",width=14,height=10)

```

```{r check_plate_layouts, fig.width=9,fig.height=3}

dat_non_border %>%
  mutate(well_row = factor(str_extract(well_name, "[A-Z]"),levels=LETTERS),
         well_column = as.integer(as.character(str_extract(well_name, "[0-9]+")))) %>%
  distinct(imaging_date,well_name,well_row,well_column,Cell) %>%
  ggplot(aes(x=well_column,y=well_row,fill=Cell)) +
  geom_tile() +
  scale_y_discrete(limits=rev) +
  facet_wrap(~imaging_date)

dat_non_border %>%
  mutate(well_row = factor(str_extract(well_name, "[A-Z]"),levels=LETTERS),
         well_column = as.integer(as.character(str_extract(well_name, "[0-9]+")))) %>%
  distinct(imaging_date,well_name,well_row,well_column,Hybridisation) %>%
  ggplot(aes(x=well_column,y=well_row,fill=Hybridisation)) +
  geom_tile() +
  scale_y_discrete(limits=rev) +
  facet_wrap(~imaging_date)

# compute fold-change from mean for each well
dat_non_border %>%
  filter(Hybridisation!="Negative") %>%
  mutate(well_row = factor(str_extract(well_name, "[A-Z]"),levels=LETTERS),
         well_column = as.integer(as.character(str_extract(well_name, "[0-9]+")))) %>%
  group_by(imaging_date,well_name,well_row,well_column,Hybridisation,Cell) %>%
  summarise(well_mean_IF = mean(Nuclei_intensity_mean_647),.groups="drop") %>%
  left_join(dat_non_border %>%
              group_by(imaging_date,well_name,Hybridisation,Cell) %>%
              summarise(well_mean_IF = mean(Nuclei_intensity_mean_647),.groups="drop") %>%
              group_by(imaging_date,Hybridisation,Cell) %>%
              summarise(plate_mean_IF = mean(well_mean_IF),.groups="drop"),
            by = join_by(imaging_date,Hybridisation,Cell)) %>%
  mutate(log2FC_from_mean = log2(well_mean_IF/plate_mean_IF)) %>%
  ggplot(aes(x=well_column,y=well_row,fill=log2FC_from_mean)) +
  geom_tile() +
  scale_fill_distiller(type = "div", palette = "RdBu", limits = c(-2,2)) +
  scale_y_discrete(limits=rev) +
  facet_wrap(~imaging_date) +
  ggtitle("Within plate difference from mean per well")

# compute fold-change from mean for each well
dat_non_border %>%
  filter(Hybridisation!="Negative") %>%
  mutate(well_row = factor(str_extract(well_name, "[A-Z]"),levels=LETTERS),
         well_column = as.integer(as.character(str_extract(well_name, "[0-9]+")))) %>%
  group_by(imaging_date,well_name,well_row,well_column,Hybridisation,Cell) %>%
  summarise(well_mean_IF = mean(Nuclei_intensity_mean_647),.groups="drop") %>%
  left_join(dat_non_border %>%
              group_by(imaging_date,well_name,Hybridisation,Cell) %>%
              summarise(well_mean_IF = mean(Nuclei_intensity_mean_647),.groups="drop") %>%
              group_by(Hybridisation,Cell) %>%
              summarise(overall_mean_IF = mean(well_mean_IF),.groups="drop"),
            by = join_by(Hybridisation, Cell)) %>%
  mutate(log2FC_from_mean = log2(well_mean_IF/overall_mean_IF)) %>%
  ggplot(aes(x=well_column,y=well_row,fill=log2FC_from_mean)) +
  geom_tile() +
  scale_fill_distiller(type = "div", palette = "RdBu", limits = c(-2,2)) +
  scale_y_discrete(limits=rev) +
  facet_wrap(~imaging_date) +
  ggtitle("Experiment difference from mean per well")

```

No extreme outliers other than plate-to-plate variation in absolute intensities. We can compute a rescaling factor across plates using HEK293 parental wells.

```{r plate_normalisation}

normalisation_data <- dat_non_border %>%
  filter(Hybridisation!="Negative" & Cell=="HEK293 parental") %>%
  select(imaging_date,well_name,Cell,Nucleus=Nuclei_intensity_mean_647,`Whole cell`=Cell_intensity_mean_647,Cytoplasm=Cytoplasm_intensity_mean_647) %>%
  pivot_longer(-c(imaging_date,well_name,Cell)) %>%
  group_by(imaging_date,well_name,Cell,name) %>%
  summarise(well_mean_IF = mean(value,na.rm=T),.groups="drop") %>%
  group_by(imaging_date,Cell,name) %>%
  summarise(plate_mean_IF = mean(well_mean_IF),.groups="drop")

# compute a normalisation factor for plates on imaging_date 20240322

normalisation_factors <- normalisation_data %>%
  filter(imaging_date=="20240315") %>%
  select(-imaging_date) %>%
  rename(normalise_to = plate_mean_IF) %>%
  left_join(normalisation_data,by = join_by(Cell,name)) %>%
  mutate(normalisation_factor=normalise_to/plate_mean_IF) %>%
  group_by(imaging_date,name) %>%
  summarise(plate_normaliser = mean(normalisation_factor),.groups = "drop")

```


```{r average_over_replicates,fig.height=3.5,fig.width=3}

well_summaries <- dat_non_border %>%
  filter(Hybridisation!="Negative") %>%
  select(imaging_date,well_name,Cell,Nucleus=Nuclei_intensity_mean_647,`Whole cell`=Cell_intensity_mean_647,Cytoplasm=Cytoplasm_intensity_mean_647) %>%
  pivot_longer(-c(imaging_date,well_name,Cell)) %>%
  left_join(normalisation_factors, by = join_by(imaging_date,name)) %>%
  mutate(value_normalised = value * plate_normaliser) %>%
  group_by(imaging_date,well_name,Cell,name) %>%
  summarise(value = mean(value,na.rm=T),
            value_plate_normalised = mean(value_normalised,na.rm=T),
            .groups='drop') %>%
  mutate(cell_labels=case_when(Cell=="HEK293 parental" ~ "HEK293",
                               Cell=="HEK293 ARMC5 KO 1A3" ~ "ARMC5 K.O. 1",
                               Cell=="HEK293 ARMC5 KO 2B6" ~ "ARMC5 K.O. 2",
                               TRUE ~ NA_character_),
         cell_labels=factor(cell_labels,levels=c("HEK293","ARMC5 K.O. 1","ARMC5 K.O. 2")),
         name=factor(name,levels=c("Nucleus","Cytoplasm","Whole cell")))

overall_summaries <- well_summaries %>%
  group_by(Cell,cell_labels,name) %>%
  summarise(mean = mean(value),
            sd = sd(value),
            mean_plate_normalised = mean(value_plate_normalised),
            sd_plate_normalised = sd(value_plate_normalised),
            .groups='drop')

axis_rescalers <- filter(overall_summaries,Cell=="HEK293 parental") %>% select(name,axis_rescaler=mean_plate_normalised)

well_summaries <- left_join(well_summaries,axis_rescalers,by = join_by(name)) 
overall_summaries <- left_join(overall_summaries,axis_rescalers,by = join_by(name)) 

well_summaries %>%
  ggplot(aes(x=cell_labels,y=value_plate_normalised/axis_rescaler,col=cell_labels)) +
  geom_col(data = overall_summaries, aes(y=mean_plate_normalised/axis_rescaler,fill=cell_labels),col=NA,alpha=0.4,position=position_dodge(0.8),width=0.8) +
  geom_point(data = overall_summaries, aes(y=mean_plate_normalised/axis_rescaler),position=position_dodge(0.8)) +
  geom_errorbar(data = overall_summaries, aes(y=mean_plate_normalised/axis_rescaler,
                                              ymin=mean_plate_normalised/axis_rescaler-sd_plate_normalised/axis_rescaler,
                                              ymax=mean_plate_normalised/axis_rescaler+sd_plate_normalised/axis_rescaler),
                position=position_dodge(0.8),width=0.4) +
  geom_jitter(width=0.3,pch=4,alpha=0.5,size=0.7) +
  scale_y_continuous(limits = c(0,1.25),name="Mean Poly(A)\nFISH intensity",breaks = pretty, expand = c(0,0)) +
  scale_color_manual(values = c(wt_col,armc5_col,armc5_col)) +
  scale_fill_manual(values = c(wt_col,armc5_col,armc5_col)) +
  facet_wrap(~name) +
  theme_bw(8) + 
  theme(axis.text.x = element_text(angle=90,hjust=1,vjust=0.5),
        axis.title.x = element_blank(),
        panel.grid = element_blank(),
        legend.position = "none")
ggsave("PLOTS/polyA_barplots_by_cell.pdf",width=2,height=2)


well_summaries %>%
  filter(name!="Cytoplasm") %>%
  ggplot(aes(x=cell_labels,y=value_plate_normalised/axis_rescaler,col=cell_labels)) +
  geom_col(data = filter(overall_summaries,name!="Cytoplasm"), aes(y=mean_plate_normalised/axis_rescaler,fill=cell_labels),col=NA,alpha=0.4,position=position_dodge(0.8),width=0.8) +
  geom_point(data = filter(overall_summaries,name!="Cytoplasm"), aes(y=mean_plate_normalised/axis_rescaler),position=position_dodge(0.8)) +
  geom_errorbar(data = filter(overall_summaries,name!="Cytoplasm"), aes(y=mean_plate_normalised/axis_rescaler,
                                              ymin=mean_plate_normalised/axis_rescaler-sd_plate_normalised/axis_rescaler,
                                              ymax=mean_plate_normalised/axis_rescaler+sd_plate_normalised/axis_rescaler),
                position=position_dodge(0.8),width=0.4) +
  geom_jitter(width=0.3,pch=4,alpha=0.4,size=0.7) +
  scale_y_continuous(limits = c(0,1.25),name="Mean Poly(A)\nFISH intensity",breaks = pretty, expand = c(0,0)) +
  scale_color_manual(values = c(wt_col,armc5_col,armc5_col)) +
  scale_fill_manual(values = c(wt_col,armc5_col,armc5_col)) +
  facet_wrap(~name) +
  theme_bw(8) + 
  theme(axis.text.x = element_text(angle=90,hjust=1,vjust=0.5),
        axis.title.x = element_blank(),
        panel.grid = element_blank(),
        legend.position = "none")
ggsave("PLOTS/polyA_barplots_by_cell_no_cytoplasm.pdf",width=1.5,height=2)



```


Suspect that these small differences arise due to cell desnsity changes, check this by plotting as a function of cell area.


```{r area_plots}

# compare area distributions
dat_non_border %>%
  ggplot(aes(x=Cell_area,col=Cell)) + 
  facet_wrap(~imaging_date) + 
  geom_density() +
  scale_x_continuous(limits = c(0,40000))

area_features <- dat_non_border %>%
  filter(Hybridisation!="Negative") %>%
  select(imaging_date,well_name,field_id,label,Cell,
         Nucleus=Nuclei_area,
         `Whole cell`=Cell_area,
         Cytoplasm=Cytoplasm_area) %>%
    pivot_longer(-c(imaging_date,well_name,field_id,label,Cell),values_to = "area")

cut_borders <- function(x){
  pattern <- "(\\(|\\[)(-*[0-9]+\\.*[0-9]*),(-*[0-9]+\\.*[0-9]*)(\\)|\\])"
  
  start <- as.numeric(gsub(pattern,"\\2", x))
  end <- as.numeric(gsub(pattern,"\\3", x))
  
  data.frame(start, end)
}

single_cell_intensities_areas <- dat_non_border %>%
  filter(Hybridisation!="Negative") %>%
  select(imaging_date,well_name,field_id,label,Cell,
         Nucleus=Nuclei_intensity_mean_647,
         `Whole cell`=Cell_intensity_mean_647,
         Cytoplasm=Cytoplasm_intensity_mean_647) %>%
  pivot_longer(-c(imaging_date,well_name,field_id,label,Cell)) %>%
  left_join(normalisation_factors, by = join_by(imaging_date,name)) %>%
  mutate(value_normalised = value * plate_normaliser) %>%
  left_join(area_features,by = join_by(imaging_date, well_name, field_id, label, Cell, name)) %>%
  filter(name=="Whole cell") %>%
  mutate(cell_labels=case_when(Cell=="HEK293 parental" ~ "HEK293",
                               Cell=="HEK293 ARMC5 KO 1A3" ~ "ARMC5 K.O. 1",
                               Cell=="HEK293 ARMC5 KO 2B6" ~ "ARMC5 K.O. 2",
                               TRUE ~ NA_character_),
         cell_labels=factor(cell_labels,levels=c("HEK293","ARMC5 K.O. 1","ARMC5 K.O. 2")),
         name=factor(name,levels=c("Nucleus","Cytoplasm","Whole cell")),
         area_sq_micron=area*0.1625*0.1625,
         area_bin=cut_width(area_sq_micron,width=50)) 

get_bin_center <- function(bin_name) {
  clean_bin <- substr(bin_name, 2, nchar(bin_name) - 1)
  parts <- unlist(strsplit(clean_bin, "[,]"))
  numeric_values <- as.numeric(parts)
  center <- (numeric_values[2] - numeric_values[1]) / 2 + numeric_values[1]
  return(center)
}

cut_borders_lut <- distinct(single_cell_intensities_areas,area_bin) %>%
  mutate(area_bin=as.character(area_bin)) %>%
  mutate(area_bin_centre = map(area_bin,~get_bin_center(.))) %>%
  unnest(cols = c(area_bin_centre))

well_summaries_intensities_areas <- single_cell_intensities_areas %>%
  left_join(cut_borders_lut,by = join_by(area_bin)) %>%
  group_by(area_bin_centre,imaging_date,well_name,Cell,name) %>%
  summarise(value_normalised=mean(value_normalised,na.rm = T),
            n_cells=n(),.groups = "drop") %>% 
  filter(area_bin_centre >= 50 & area_bin_centre <= 1000)

overall_summaries_intensities_by_area_bin <- well_summaries_intensities_areas %>% 
  group_by(area_bin_centre,Cell,name) %>%
  summarise(value_normalised_mean=mean(value_normalised,na.rm = T),
            value_normalised_sd=sd(value_normalised,na.rm = T),
            n_cells=mean(n_cells),.groups = "drop") %>%
  group_by(Cell,name) %>%
  mutate(frac_cells = n_cells/sum(n_cells)) %>%
  ungroup()

cell_count_plot <- overall_summaries_intensities_by_area_bin %>%
  ggplot(aes(x=area_bin_centre,y=frac_cells,col=Cell)) + 
  geom_line() + 
  scale_x_continuous(limits = c(0,1100)) +
  scale_y_continuous(name="Fraction of cells") +
  theme_bw(8) +
  theme(legend.position = "none",
        axis.title.x = element_blank())

intensity_plot <- well_summaries_intensities_areas %>% 
  ggplot(aes(x=area_bin_centre,y=value_normalised,col=Cell)) + 
  geom_jitter(size=0.1,alpha=0.2) +
  geom_line(data=overall_summaries_intensities_by_area_bin,
             aes(x=area_bin_centre,y=value_normalised_mean)) + 
  geom_point(data=overall_summaries_intensities_by_area_bin,
             aes(x=area_bin_centre,y=value_normalised_mean),size=0.5) + 
  geom_ribbon(data=overall_summaries_intensities_by_area_bin,
             aes(x=area_bin_centre,
                 y=value_normalised_mean,
                 ymin=value_normalised_mean-value_normalised_sd,
                 ymax=value_normalised_mean+value_normalised_sd,
                 fill=Cell),col=NA,alpha=0.1) + 
  scale_x_continuous(name="Cell area (µm<sup>2</sup>)",limits = c(0,1100)) +
  scale_y_continuous(name="Mean poly(A) intensity",limits = c(0,750)) +
  theme_bw(8) +
  theme(axis.title.x = ggtext::element_markdown()) 

cell_count_plot / intensity_plot + plot_layout(heights = c(1,3),guides = "collect")

ggsave("PLOTS/polyA_by_area_whole_cell.pdf",width=4,height=3)

```


```{r well_numbers}

dat_non_border %>%
  filter(Hybridisation!="Negative") %>%
  group_by(imaging_date,well_name,Cell) %>%
  count()

well_summaries %>%
  group_by(imaging_date,Cell,name) %>%
  count()

well_summaries %>%
  group_by(Cell,name) %>%
  count()

```


```{r save_summaries}

write_csv(x = well_summaries,file = "SUMMARIES/polyA_well_summaries.csv")
write_csv(x = overall_summaries,file = "SUMMARIES/polyA_overall_summaries.csv")

```



```{r session_info}

sessionInfo()

```
