---
title: "POLR2A IF in ARMC5 KO cells"
author: "Scott Berry"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup_r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

rm(list=ls())
library(tidyverse)
library(patchwork)

wt_col <- "#184cee"
armc5_col <- "#9d24ec"

cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
```

# Experiment

# Data handling

Original data (Nikon `.nd2` files) are stored on UNSW's data archive under Berry Lab project D0419427. Data were maximum-intensity projected and converted to OME-TIFF using the `blimp` tools.

# Analysis

## Image analysis 

Image analysis performed using the following python scripts (found on github for [hek_inhibitors_IF](https://github.com/berrygroup/publications/tree/main/2024_ARMC5/hek_inhibitors_IF))

- `illumination_correction_EU.ipynb`
- `save_example_images_EU.ipynb` (used to generate example images and training data for the cellpose model)
- `run_quantification_IF.py` (parallelised using `PBS/batch_quantify_EU_240305.pbs`)

## Data analysis

Quantitative data analysis performed in this document.

### Load data

Load data from `scratch/berrylab` on UNSW's Katana HPC.

```{r load_data}

ana_dir <- "/srv/scratch/z3532965/src/publications/2024_ARMC5/hek_inhibitors_IF"
data_dir <- "/srv/scratch/berrylab/z3532965/NikonSpinningDisk"
acquisition_dirs <- purrr::map_chr(
  c("240305/240301_ARMC5KO_IF/20240305_184055_331",
    "240308/240304_ARMC5KO_IF/20240308_165103_250",
    "240314/240311_ARMC5KO_IF/20240314_185145_903"),~file.path(data_dir,.))
intensity_image_dirs <- map_chr(acquisition_dirs,~file.path(.,"OME-TIFF"))
intensity_image_mip_dirs <- map_chr(acquisition_dirs,~file.path(.,"OME-TIFF-MIP"))
label_image_dirs <- map_chr(acquisition_dirs,~file.path(.,"SEGMENTATION"))
quantification_dirs <- map_chr(acquisition_dirs,~file.path(.,"QUANTIFICATION"))

read_quantification <- function(dir_name) {
  csv_files <- Sys.glob(file.path(dir_name,"*.csv"))
  dt_list <- lapply(csv_files, data.table::fread)
  df <- tibble(path = csv_files,
               feature_values = dt_list) %>%
    mutate(dir_name = dir_name,
           file_name = basename(path),
           feature_values = purrr::map(feature_values,tibble)) %>%
    separate(file_name,into = c("well_name","ch1","ch2","ch3","ch4","sample","id"),sep="[,_.]", extra = "drop") %>%
    mutate(well_name = str_replace(well_name, "^Well", ""),
           field_id = as.integer(id))
  return(df)
}

pad_well_names <- function(names) {
  gsub("(\\D)(\\d)$", "\\10\\2", names)
}

# features
f <- lapply(quantification_dirs,read_quantification) %>%
  bind_rows() %>%
  mutate(acquisition_dir = dirname(dirname(path)),
         imaging_date = basename(acquisition_dir) %>% str_extract("\\d{8}(?=_)")) %>%
  select(-c(ch1, ch2, ch3, ch4, sample))

# layouts
plate_layouts <- tibble(layout_filename=c("240301_ARMC5KO_IF_longlayout.csv","240304_ARMC5KO_IF_longlayout.csv","240311_ARMC5KO_IF_longlayout.csv"),
                        plating_date=c("240301","240304","240311"),
                        imaging_date=c("20240305","20240308","20240314"),
                        metadata=purrr::map(layout_filename,~read_csv(file.path(ana_dir,"METADATA",.),show_col_types = FALSE, name_repair = ~sub(" ", "_", .)))) %>%
  unnest(metadata) %>%
  mutate(well_name=pad_well_names(Well_name)) %>%
  select(-Well_name) %>%
  left_join(distinct(f,acquisition_dir,well_name,imaging_date),by = join_by(imaging_date, well_name))

```

Check that data are complete.

```{r check_complete}

# check completeness (count wells)
f %>% 
  distinct(acquisition_dir,well_name) %>%
  group_by(acquisition_dir) %>%
  count()

# check completeness (count sites per well)
f %>% 
  group_by(acquisition_dir,well_name) %>%
  count() %>%
  filter(!(n %in% c(16,25)))

```

Load metadata and join with quantified image data.

```{r load metadata}
# read image metadata and join to plate metadata
m <- tibble(acquisition_dir = acquisition_dirs) %>%
  mutate(image_dir = purrr::map_chr(acquisition_dirs,~file.path(.,"OME-TIFF-MIP"))) %>%
  mutate(metadata_file = purrr::map(image_dir,~list.files(.))) %>%
  unnest(metadata_file) %>%
  filter(grepl("metadata.csv$",metadata_file)) %>%
  mutate(dat = purrr::map(file.path(image_dir,metadata_file),~read_csv(.,show_col_types = FALSE))) %>%
  separate(metadata_file,into = c("well_name","ch1","ch2","ch3","ch4","sample"),sep="[,_.]", extra = "drop") %>%
  unnest(dat) %>%
  mutate(well_name = str_replace(well_name, "^Well", ""),
           field_id = as.integer(field_id)) %>%
  select(-c(ch1, ch2, ch3, ch4, sample)) %>%
  left_join(filter(plate_layouts,!is.na(acquisition_dir)),by=join_by(acquisition_dir,well_name))

# attach image and plate metadata 
dat <- right_join(m,f,by = join_by(acquisition_dir, well_name, field_id, imaging_date)) %>%
  unnest(feature_values) %>%
  select(-matches("XXX$"))

```

Some basic QC regarding number of cells per well, etc.

```{r cell_counts, fig.width=8,fig.height=12}

dat %>% 
  distinct(Cell,plating_date,imaging_date,well_name,Drug,Drug_time,Primary,Secondary) %>%
  group_by(Cell,plating_date,imaging_date,Drug,Drug_time,Primary,Secondary) %>%
  count() %>%
  ungroup() %>%
  ggplot(aes(x=Cell,y=n,fill=interaction(Drug,Drug_time))) + geom_col(position = "dodge") + facet_grid(interaction(Primary,Secondary)~imaging_date) +
  coord_flip() +
  ggtitle("Number of wells of each condition")

dat %>% 
  group_by(Cell,plating_date,imaging_date,Drug,Drug_time,Primary,Secondary) %>%
  count() %>%
  ungroup() %>%
  ggplot(aes(x=Cell,y=n,fill=Drug)) + geom_col(position = "dodge") + facet_grid(interaction(Primary,Secondary)~imaging_date) +
  coord_flip() +
  ggtitle("Total number of cells in each condition")

```

### Subtract background

Calculate background by maximum intensity projecting, illumination-correcting, and getting the mean intensity in a region containing no cells.


```{r subtract_background}

# calculated manually across three different illumination corrected images
bg_488 <-mean(c(150.269, 143.756, 148.334))
bg_647 <- mean(c(117.062, 113.934, 116.316))
bg_DAPI <-mean(c(147.770, 142.327, 145.496))

dat_minus_bg <- mutate(dat,
                       Nuclei_intensity_mean_647 = Nuclei_intensity_mean_647 - bg_647,
                       Nuclei_intensity_mean_DAPI = Nuclei_intensity_mean_DAPI - bg_DAPI,
                       Nuclei_intensity_mean_488 = Nuclei_intensity_mean_488 - bg_488,
                       
                       Nuclei_intensity_median_647 = Nuclei_intensity_median_647 - bg_647,
                       Nuclei_intensity_median_DAPI = Nuclei_intensity_median_DAPI - bg_DAPI,
                       Nuclei_intensity_median_488 = Nuclei_intensity_median_488 - bg_488)

```

### Sum intensities

Compute sum intensities by multiplying the background-subtracted means by the corresponding area.

```{r sum_intensities}

dat_minus_bg <- mutate(dat_minus_bg,
                       Nuclei_intensity_sum_647 = Nuclei_intensity_mean_647 * Nuclei_area,
                       Nuclei_intensity_sum_DAPI = Nuclei_intensity_mean_DAPI * Nuclei_area,
                       Nuclei_intensity_sum_488 = Nuclei_intensity_mean_488 * Nuclei_area)

```

### Data cleanup

Remove border cells 

```{r remove_variables}

dat_non_border <- filter(dat_minus_bg,!Nuclei_is_border)

```

### Check illumination correction

After correction of illumination biases across the field of view, there should be no bias for mean intensities across the field-of-view.

```{r test_illum_corr}

dat_non_border_sample_longer <- dat_non_border %>%
  filter(imaging_date=="20240308") %>%
  filter(well_name=="F06") %>%
  select(Nuclei_centroid_0,Nuclei_centroid_1,Nuclei_intensity_mean_DAPI,Nuclei_intensity_mean_488,Nuclei_intensity_mean_647) %>%
  pivot_longer(-c(Nuclei_centroid_0,Nuclei_centroid_1)) 

rescaling_limits <- dat_non_border_sample_longer %>%
  group_by(name) %>%
  summarise(upper_limit = quantile(value,0.99))

dat_non_border_sample_longer <- left_join(dat_non_border_sample_longer,rescaling_limits,by="name")

channels <- unique(dat_non_border_sample_longer$name)

p <- list(length=length(channels))
for (i in 1:length(channels)) {
  p[[i]] <- dat_non_border_sample_longer %>%
    filter(name==channels[i]) %>%
    ggplot(aes(x=Nuclei_centroid_0,y=Nuclei_centroid_1,col=value/upper_limit)) +
    geom_point(size=0.5) +
    scale_color_viridis_c(limits=c(0,1)) +
    facet_wrap(~name) +
    coord_fixed() +
    theme(legend.position = "none")
}

wrap_plots(p) & ggtitle("Illumination biases")

```

### Check DAPI distributions

These look relatively reproducible within experiments, with one exception removed below.

```{r check_DAPI}

dat_non_border %>%
  ggplot(aes(x=Nuclei_intensity_sum_DAPI,col=as.factor(well_name))) +
  geom_density() +
  facet_wrap(Cell~imaging_date,scales = "free") +
  theme(legend.position = "none") +
  scale_x_continuous(limits = c(0,5e6)) +
  scale_y_continuous(labels = NULL)

```

```{r remove_dapi_outlier_well}

dat_non_border %>%
  filter(Cell=="HEK293 ARMC5 KO 2B6" & imaging_date=="20240314") %>%
  group_by(well_name) %>%
  summarise(mean_sum_dapi = mean(Nuclei_intensity_sum_DAPI)) %>%
  arrange(-mean_sum_dapi) %>%
  head(5)

dat_non_border <- filter(dat_non_border,!(Cell=="HEK293 ARMC5 KO 2B6" & imaging_date=="20240314" & well_name=="H10"))

```


### Check IF distributions

We now visualise the IF staining distributions in each well, separating by treatment.

```{r IF_by_treatment, fig.width=14,fig.height=10}

dat_non_border <- mutate(dat_non_border,Drug=factor(Drug,levels=c("None","Vehicle","Triptolide 300 nM","LDC4297 10 µM","THZ1 1 µM","AZD4573 1 µM","DRB 100 µM"))) 

pTotal <- dat_non_border %>%
  filter(Primary=="Total Pol II F12") %>%
  ggplot(aes(x=Drug,y=Nuclei_intensity_mean_488,col=Cell,group=interaction(well_name,Cell))) + 
  geom_boxplot(outlier.shape=NA) + facet_grid(Primary~imaging_date) +
  coord_cartesian(ylim = c(0,750)) +
  theme_bw(base_size = 8) +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank())

pS5P <- dat_non_border %>%
  filter(Primary=="pSer5 Pol II 3E8") %>%
  ggplot(aes(x=Drug,y=Nuclei_intensity_mean_488,col=Cell,group=interaction(well_name,Cell))) + 
  geom_boxplot(outlier.shape=NA) + facet_grid(Primary~imaging_date) +
  coord_cartesian(ylim = c(0,4500)) +
  theme_bw(base_size = 8) +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank())

pS2P <- dat_non_border %>%
  filter(Primary=="pSer2 Pol II 3E10") %>%
  ggplot(aes(x=Drug,y=Nuclei_intensity_mean_488,col=Cell,group=interaction(well_name,Cell))) + 
  geom_boxplot(outlier.shape=NA) + facet_grid(Primary~imaging_date) +
  coord_cartesian(ylim = c(0,750)) +
  theme_bw(base_size = 8) +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_text(angle=90,hjust=1,vjust=0.5))


pTotal / pS5P / pS2P + plot_layout(guides="collect")

ggsave("PLOTS/IF_boxplot_per_well_by_drug.pdf",width=14,height=10)

```

As above, but separating by cell type

```{r IF_by_cell, fig.width=14,fig.height=10}

pTotal <- dat_non_border %>%
  filter(Primary=="Total Pol II F12") %>%
  ggplot(aes(x=Cell,y=Nuclei_intensity_mean_488,col=Drug,group=interaction(well_name,Drug))) + 
  geom_boxplot(outlier.shape=NA) + facet_grid(Primary~imaging_date) +
  coord_cartesian(ylim = c(0,750)) +
  theme_bw(base_size = 8) +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank())

pS5P <- dat_non_border %>%
  filter(Primary=="pSer5 Pol II 3E8") %>%
  ggplot(aes(x=Cell,y=Nuclei_intensity_mean_488,col=Drug,group=interaction(well_name,Drug))) + 
  geom_boxplot(outlier.shape=NA) + facet_grid(Primary~imaging_date) +
  coord_cartesian(ylim = c(0,4500)) +
  theme_bw(base_size = 8) +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank())

pS2P <- dat_non_border %>%
  filter(Primary=="pSer2 Pol II 3E10") %>%
  ggplot(aes(x=Cell,y=Nuclei_intensity_mean_488,col=Drug,group=interaction(well_name,Drug))) + 
  geom_boxplot(outlier.shape=NA) + facet_grid(Primary~imaging_date) +
  coord_cartesian(ylim = c(0,750)) +
  theme_bw(base_size = 8) +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_text(angle=90,hjust=1,vjust=0.5))


pTotal / pS5P / pS2P + plot_layout(guides="collect")

ggsave("PLOTS/IF_boxplot_per_well_by_cell.pdf",width=14,height=10)

```

Visualise plate layouts and examine the well-means across the plates for each experiment.

```{r check_plate_layouts, fig.width=9,fig.height=3}

dat_non_border %>%
  mutate(well_row = factor(str_extract(well_name, "[A-Z]"),levels=LETTERS),
         well_column = as.integer(as.character(str_extract(well_name, "[0-9]+")))) %>%
  distinct(imaging_date,well_name,well_row,well_column,Drug) %>%
  ggplot(aes(x=well_column,y=well_row,fill=Drug)) +
  geom_tile() +
  scale_y_discrete(limits=rev) +
  facet_wrap(~imaging_date) +
  ggtitle("Drug")

dat_non_border %>%
  mutate(well_row = factor(str_extract(well_name, "[A-Z]"),levels=LETTERS),
         well_column = as.integer(as.character(str_extract(well_name, "[0-9]+")))) %>%
  distinct(imaging_date,well_name,well_row,well_column,Cell) %>%
  ggplot(aes(x=well_column,y=well_row,fill=Cell)) +
  geom_tile() +
  scale_y_discrete(limits=rev) +
  facet_wrap(~imaging_date) +
  ggtitle("Cell type")

dat_non_border %>%
  mutate(well_row = factor(str_extract(well_name, "[A-Z]"),levels=LETTERS),
         well_column = as.integer(as.character(str_extract(well_name, "[0-9]+")))) %>%
  distinct(imaging_date,well_name,well_row,well_column,Primary) %>%
  ggplot(aes(x=well_column,y=well_row,fill=Primary)) +
  geom_tile() +
  scale_y_discrete(limits=rev) +
  facet_wrap(~imaging_date) +
  ggtitle("Primary antibody")

dat_non_border %>%
  mutate(well_row = factor(str_extract(well_name, "[A-Z]"),levels=LETTERS),
         well_column = as.integer(as.character(str_extract(well_name, "[0-9]+")))) %>%
  count(imaging_date,well_name,well_row,well_column,Secondary) %>%
  ggplot(aes(x=well_column,y=well_row,fill=n)) +
  geom_tile() +
  scale_y_discrete(limits=rev) +
  facet_wrap(~imaging_date) +
  scale_fill_viridis_c() +
  ggtitle("Cell number")


# compute fold-change from mean for each well
dat_non_border %>%
  filter(Primary!="None") %>%
  mutate(well_row = factor(str_extract(well_name, "[A-Z]"),levels=LETTERS),
         well_column = as.integer(as.character(str_extract(well_name, "[0-9]+")))) %>%
  group_by(imaging_date,well_name,well_row,well_column,Primary,Cell,Drug) %>%
  summarise(well_mean_IF = mean(Nuclei_intensity_mean_488),.groups="drop") %>%
  left_join(dat_non_border %>%
              group_by(imaging_date,well_name,Primary,Cell,Drug) %>%
              summarise(well_mean_IF = mean(Nuclei_intensity_mean_488),.groups="drop") %>%
              group_by(imaging_date,Primary,Cell,Drug) %>%
              summarise(plate_mean_IF = mean(well_mean_IF),.groups="drop"),
            by = join_by(imaging_date,Primary,Cell,Drug)) %>%
  mutate(log2FC_from_mean = log2(well_mean_IF/plate_mean_IF)) %>%
  ggplot(aes(x=well_column,y=well_row,fill=log2FC_from_mean)) +
  geom_tile() +
  scale_fill_distiller(type = "div", palette = "RdBu", limits = c(-2,2)) +
  scale_y_discrete(limits=rev) +
  facet_wrap(~imaging_date) +
  ggtitle("Within plate difference from mean per well")

# compute fold-change from mean for each well
dat_non_border %>%
  filter(Primary!="None") %>%
  mutate(well_row = factor(str_extract(well_name, "[A-Z]"),levels=LETTERS),
         well_column = as.integer(as.character(str_extract(well_name, "[0-9]+")))) %>%
  group_by(imaging_date,well_name,well_row,well_column,Secondary,Primary,Cell,Drug) %>%
  summarise(well_mean_IF = mean(Nuclei_intensity_mean_488),.groups="drop") %>%
  left_join(dat_non_border %>%
              group_by(imaging_date,well_name,Secondary,Primary,Cell,Drug) %>%
              summarise(well_mean_IF = mean(Nuclei_intensity_mean_488),.groups="drop") %>%
              group_by(Secondary,Primary,Cell,Drug) %>%
              summarise(overall_mean_IF = mean(well_mean_IF),.groups="drop"),
            by = join_by(Secondary,Primary, Cell, Drug)) %>%
  mutate(log2FC_from_mean = log2(well_mean_IF/overall_mean_IF)) %>%
  ggplot(aes(x=well_column,y=well_row,fill=log2FC_from_mean)) +
  geom_tile() +
  scale_fill_distiller(type = "div", palette = "RdBu", limits = c(-2,2)) +
  scale_y_discrete(limits=rev) +
  facet_wrap(~imaging_date) +
  ggtitle("Experiment difference from mean per well")

```

No extreme outliers other than plate-to-plate variation in absolute intensities. We can compute a rescaling factor across plates using vehicle-treated wells. 20240314 seems to be in the middle for all stains so normalise to this.

```{r plate_normalisation}

normalisation_data <- dat_non_border %>%
  filter(Primary!="None" & Drug=="Vehicle") %>%
  group_by(imaging_date,well_name,Primary,Cell) %>%
  summarise(well_mean_IF = mean(Nuclei_intensity_mean_488),.groups="drop") %>%
  group_by(imaging_date,Primary,Cell) %>%
  summarise(plate_mean_IF = mean(well_mean_IF),.groups="drop")

normalisation_data %>%
  pivot_wider(names_from = imaging_date, values_from=plate_mean_IF) %>%
  ggplot(aes(x=`20240314`,y=`20240305`)) + geom_point() + geom_abline() + facet_wrap(~Primary)

normalisation_data %>%
  pivot_wider(names_from = imaging_date, values_from=plate_mean_IF) %>%
  ggplot(aes(x=`20240314`,y=`20240308`)) + geom_point() + geom_abline() + facet_wrap(~Primary)

# compute a normalisation factor for plates on imaging_date 20240305 and 20240308

normalisation_factors <- normalisation_data %>%
  filter(imaging_date=="20240314") %>%
  select(-imaging_date) %>%
  rename(normalise_to = plate_mean_IF) %>%
  left_join(normalisation_data,by = join_by(Primary, Cell)) %>%
  mutate(normalisation_factor=normalise_to/plate_mean_IF) %>%
  group_by(Primary,imaging_date) %>%
  summarise(plate_normaliser = mean(normalisation_factor),.groups = "drop")

```

After normalising data, we can calculate average and standard deviations across replicates.

```{r average_over_replicates,fig.height=4,fig.width=12}

well_summaries <- dat_non_border %>%
  filter(Primary!="None") %>%
  left_join(normalisation_factors, by = join_by(imaging_date, Primary)) %>%
  mutate(Nuclei_intensity_mean_488_plate_normalised = Nuclei_intensity_mean_488 * plate_normaliser) %>%
  group_by(imaging_date,well_name,Primary,Cell,Drug) %>%
  summarise(Nuclei_intensity_mean_488 = mean(Nuclei_intensity_mean_488),
            Nuclei_intensity_mean_488_plate_normalised = mean(Nuclei_intensity_mean_488_plate_normalised),
            .groups='drop') %>%
  mutate(Primary,factor(Primary,levels=c("Total Pol II F12","pSer5 Pol II 3E8","pSer2 Pol II 3E10")))

overall_summaries <- well_summaries %>%
  group_by(Primary,Cell,Drug) %>%
  summarise(mean_Nuclei_intensity_mean_488 = mean(Nuclei_intensity_mean_488),
            sd_Nuclei_intensity_mean_488 = sd(Nuclei_intensity_mean_488),
            mean_Nuclei_intensity_mean_488_plate_normalised = mean(Nuclei_intensity_mean_488_plate_normalised),
            sd_Nuclei_intensity_mean_488_plate_normalised = sd(Nuclei_intensity_mean_488_plate_normalised),
            .groups='drop')

well_summaries %>%
  ggplot(aes(x=Drug,y=Nuclei_intensity_mean_488_plate_normalised,col=Cell)) +
  facet_wrap(~factor(Primary, levels=c("Total Pol II F12","pSer5 Pol II 3E8","pSer2 Pol II 3E10")), nrow=1,scales="free_y") +
  geom_col(data = overall_summaries, aes(y=mean_Nuclei_intensity_mean_488_plate_normalised,fill=Cell),col=NA,alpha=0.4,position=position_dodge(0.8),width=0.8) +
  geom_point(data = overall_summaries, aes(y=mean_Nuclei_intensity_mean_488_plate_normalised),position=position_dodge(0.8)) +
  geom_errorbar(data = overall_summaries, aes(y=mean_Nuclei_intensity_mean_488_plate_normalised,
                                              ymin=mean_Nuclei_intensity_mean_488_plate_normalised-sd_Nuclei_intensity_mean_488_plate_normalised,
                                              ymax=mean_Nuclei_intensity_mean_488_plate_normalised+sd_Nuclei_intensity_mean_488_plate_normalised),
                position=position_dodge(0.8),width=0.1) +
  geom_jitter(position=position_jitterdodge(jitter.width=0.15,dodge.width=0.8),pch=4,alpha=0.5) +
  scale_y_continuous(limits = c(0,NA),name="Nuclear intensity (A.U.)") +
  theme_bw() + 
  theme(axis.text.x = element_text(angle=90,hjust=1,vjust=0.5),
        axis.title.x = element_blank(),
        panel.grid = element_blank())
ggsave("PLOTS/IF_barplots_by_cell.pdf",width=12,height=4)

well_summaries %>%
  ggplot(aes(x=factor(Cell, levels=c("HEK293 parental","HEK293 ARMC5 KO 1A3","HEK293 ARMC5 KO 2B6")),
                      y=Nuclei_intensity_mean_488_plate_normalised,col=Drug)) +
  facet_wrap(~factor(Primary, levels=c("Total Pol II F12","pSer5 Pol II 3E8","pSer2 Pol II 3E10")), nrow=1,scales="free_y") +
  geom_col(data = overall_summaries, aes(y=mean_Nuclei_intensity_mean_488_plate_normalised,fill=Drug),col=NA,alpha=0.4,position=position_dodge(0.8),width=0.8) +
  geom_point(data = overall_summaries, aes(y=mean_Nuclei_intensity_mean_488_plate_normalised),position=position_dodge(0.8)) +
  geom_errorbar(data = overall_summaries, aes(y=mean_Nuclei_intensity_mean_488_plate_normalised,
                                              ymin=mean_Nuclei_intensity_mean_488_plate_normalised-sd_Nuclei_intensity_mean_488_plate_normalised,
                                              ymax=mean_Nuclei_intensity_mean_488_plate_normalised+sd_Nuclei_intensity_mean_488_plate_normalised),
                position=position_dodge(0.8),width=0.1) +
  geom_jitter(position=position_jitterdodge(jitter.width=0.2,dodge.width=0.8),pch=4,alpha=0.8) +
  scale_y_continuous(limits = c(0,NA),name="Nuclear intensity (A.U.)") +
  theme_bw() + 
  theme(axis.text.x = element_text(angle=90,hjust=1,vjust=0.5),
        axis.title.x = element_blank(),
        panel.grid = element_blank())

ggsave("PLOTS/IF_barplots_by_drug.pdf",width=12,height=4)

```

And make a cleaner plot of vehicle-treated wells for publication.

```{r barplot_vehicle_only,fig.height=2,fig.width=1.1}

well_summaries_vehicle <- filter(well_summaries,Drug %in% c("Vehicle")) %>%
  mutate(cell_labels=case_when(Cell=="HEK293 parental" ~ "HEK293",
                               Cell=="HEK293 ARMC5 KO 1A3" ~ "ARMC5 K.O. #1",
                               Cell=="HEK293 ARMC5 KO 2B6" ~ "ARMC5 K.O. #2",
                               TRUE ~ NA_character_),
         cell_labels=factor(cell_labels,levels=c("HEK293","ARMC5 K.O. #1","ARMC5 K.O. #2")))
overall_summaries_vehicle <- filter(overall_summaries,Drug %in% c("Vehicle")) %>%
  mutate(cell_labels=case_when(Cell=="HEK293 parental" ~ "HEK293",
                               Cell=="HEK293 ARMC5 KO 1A3" ~ "ARMC5 K.O. #1",  
                               Cell=="HEK293 ARMC5 KO 2B6" ~ "ARMC5 K.O. #2",
                               TRUE ~ NA_character_),
         cell_labels=factor(cell_labels,levels=c("HEK293","ARMC5 K.O. #1","ARMC5 K.O. #2")))

norm_axis <- filter(overall_summaries_vehicle,Cell=="HEK293 parental") %>% select(Primary,mean_Nuclei_intensity_mean_488_plate_normalised)
norm_axis_total <- norm_axis %>% filter(Primary=="Total Pol II F12") %>% pull(mean_Nuclei_intensity_mean_488_plate_normalised)
norm_axis_S5P <- norm_axis %>% filter(Primary=="pSer5 Pol II 3E8") %>% pull(mean_Nuclei_intensity_mean_488_plate_normalised)
norm_axis_S2P <- norm_axis %>% filter(Primary=="pSer2 Pol II 3E10") %>% pull(mean_Nuclei_intensity_mean_488_plate_normalised)

well_summaries_vehicle %>%
  filter(Primary=="Total Pol II F12") %>%
  ggplot(aes(x=cell_labels,
             y=Nuclei_intensity_mean_488_plate_normalised/norm_axis_total)) +
  geom_col(data = filter(overall_summaries_vehicle,Primary=="Total Pol II F12"),
           aes(y=mean_Nuclei_intensity_mean_488_plate_normalised/norm_axis_total,fill=Cell),col=NA,alpha=0.4,position=position_dodge(0.8),width=0.8) +
  geom_point(data = filter(overall_summaries_vehicle,Primary=="Total Pol II F12"),
             aes(y=mean_Nuclei_intensity_mean_488_plate_normalised/norm_axis_total,col=Cell),position=position_dodge(0.8),size=0.75) +
  geom_errorbar(data = filter(overall_summaries_vehicle,Primary=="Total Pol II F12"),
                aes(y=mean_Nuclei_intensity_mean_488_plate_normalised/norm_axis_total,
                    ymin=mean_Nuclei_intensity_mean_488_plate_normalised/norm_axis_total-sd_Nuclei_intensity_mean_488_plate_normalised/norm_axis_total,
                    ymax=mean_Nuclei_intensity_mean_488_plate_normalised/norm_axis_total+sd_Nuclei_intensity_mean_488_plate_normalised/norm_axis_total,
                    col=Cell),
                position=position_dodge(0.8),width=0.1) +
  geom_jitter(aes(col=Cell),pch=4,alpha=0.4,position = position_jitter(width = 0.2)) +
  scale_y_continuous(limits = c(0,2.2),name="Mean nuclear \n RPB1 intensity",breaks=c(0,1,2), expand = c(0,0)) +
  scale_color_manual(values = c(armc5_col,armc5_col,wt_col)) +
  scale_fill_manual(values = c(armc5_col,armc5_col,wt_col)) +
  theme_bw(8) + 
  theme(axis.text.x = element_text(angle=90,hjust=1,vjust=0.5),
        axis.title.x = element_blank(),
        panel.grid = element_blank(),
        legend.position = "none")

ggsave("PLOTS/IF_barplot_vehicle_total.pdf",width=1.1,height=1.75)

pS5 <- well_summaries_vehicle %>%
  filter(Primary %in% c("pSer5 Pol II 3E8")) %>%
  ggplot(aes(x=cell_labels,
             y=Nuclei_intensity_mean_488_plate_normalised/norm_axis_S5P)) +
  geom_col(data = filter(overall_summaries_vehicle,Primary=="pSer5 Pol II 3E8"),
           aes(y=mean_Nuclei_intensity_mean_488_plate_normalised/norm_axis_S5P,fill=Cell),col=NA,alpha=0.4,position=position_dodge(0.8),width=0.8) +
  geom_point(data = filter(overall_summaries_vehicle,Primary=="pSer5 Pol II 3E8"),
             aes(y=mean_Nuclei_intensity_mean_488_plate_normalised/norm_axis_S5P,col=Cell),position=position_dodge(0.8),size=0.75) +
  geom_errorbar(data = filter(overall_summaries_vehicle,Primary=="pSer5 Pol II 3E8"),
                aes(y=mean_Nuclei_intensity_mean_488_plate_normalised/norm_axis_S5P,
                    ymin=mean_Nuclei_intensity_mean_488_plate_normalised/norm_axis_S5P-sd_Nuclei_intensity_mean_488_plate_normalised/norm_axis_S5P,
                    ymax=mean_Nuclei_intensity_mean_488_plate_normalised/norm_axis_S5P+sd_Nuclei_intensity_mean_488_plate_normalised/norm_axis_S5P,
                    col=Cell),
                position=position_dodge(0.8),width=0.1) +
  geom_jitter(aes(col=Cell),pch=4,alpha=0.4,position = position_jitter(width = 0.2)) +
  scale_y_continuous(limits = c(0,2.2),name="Mean nuclear\nintensity",breaks=c(0,1,2), expand = c(0,0)) +
  scale_color_manual(values = c(armc5_col,armc5_col,wt_col)) +
  scale_fill_manual(values = c(armc5_col,armc5_col,wt_col)) +
  theme_bw(8) + 
  theme(axis.text.x = element_text(angle=90,hjust=1,vjust=0.5),
        axis.title.x = element_blank(),
        panel.grid = element_blank(),
        legend.position = "none")

pS2 <- well_summaries_vehicle %>%
  filter(Primary %in% c("pSer2 Pol II 3E10")) %>%
  ggplot(aes(x=cell_labels,
             y=Nuclei_intensity_mean_488_plate_normalised/norm_axis_S2P)) +
  geom_col(data = filter(overall_summaries_vehicle,Primary=="pSer2 Pol II 3E10"),
           aes(y=mean_Nuclei_intensity_mean_488_plate_normalised/norm_axis_S2P,fill=Cell),col=NA,alpha=0.4,position=position_dodge(0.8),width=0.8) +
  geom_point(data = filter(overall_summaries_vehicle,Primary=="pSer2 Pol II 3E10"),
             aes(y=mean_Nuclei_intensity_mean_488_plate_normalised/norm_axis_S2P,col=Cell),position=position_dodge(0.8),size=0.75) +
  geom_errorbar(data = filter(overall_summaries_vehicle,Primary=="pSer2 Pol II 3E10"),
                aes(y=mean_Nuclei_intensity_mean_488_plate_normalised/norm_axis_S2P,
                    ymin=mean_Nuclei_intensity_mean_488_plate_normalised/norm_axis_S2P-sd_Nuclei_intensity_mean_488_plate_normalised/norm_axis_S2P,
                    ymax=mean_Nuclei_intensity_mean_488_plate_normalised/norm_axis_S2P+sd_Nuclei_intensity_mean_488_plate_normalised/norm_axis_S2P,
                    col=Cell),
                position=position_dodge(0.8),width=0.1) +
  geom_jitter(aes(col=Cell),pch=4,alpha=0.4,position = position_jitter(width = 0.2)) +
  scale_y_continuous(limits = c(0,2.2),name="Mean nuclear\nintensity",breaks=c(0,1,2), expand = c(0,0)) +
  scale_color_manual(values = c(armc5_col,armc5_col,wt_col)) +
  scale_fill_manual(values = c(armc5_col,armc5_col,wt_col)) +
  theme_bw(8) + 
  theme(axis.text.x = element_text(angle=90,hjust=1,vjust=0.5),
        axis.title.x = element_blank(),
        panel.grid = element_blank(),
        legend.position = "none")

pS5 / pS2 + plot_layout(axes = "collect")

ggsave("PLOTS/IF_barplot_vehicle_phos.pdf",width=1.1,height=2.5)

```

```{r save_summaries}

write_csv(x = well_summaries,file = "SUMMARIES/Drug_IF_well_summaries.csv")
write_csv(x = overall_summaries,file = "SUMMARIES/Drug_IF_overall_summaries.csv")

```

```{r session_info}

sessionInfo()

```
