---
title: "Pol II IF with ARMC5 and INTS8 siRNA"
author: "Scott Berry"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup_r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

rm(list=ls())
library(tidyverse)
library(patchwork)

wt_col <- "#184cee"
armc5_col <- "#9d24ec"
ints8_col <- "#5ba274"
armc5_ints8_col <- "#f3a031"

cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
```

# Experiment

# Data handling

Original data (Nikon `.nd2` files) are stored on UNSW's data archive under Berry Lab project D0419427. Data were maximum-intensity projected and converted to OME-TIFF using the `blimp` tools.

# Analysis

## Image analysis 

Image analysis performed using the following series of python scripts (found on github for [hct116_IF](https://github.com/berrygroup/publications/tree/main/2024_ARMC5/hct116_IF))

- `illumination_correction_IF.ipynb`
- `save_example_images_IF.ipynb` (used to generate example images and training data for the cellpose model)
- `run_quantification_IF.py` (parallelised using `PBS/batch_quantify_IF_240328.pbs`)

## Data analysis

Quantitative data analysis performed in this document.

### Load data

Load data from `scratch/berrylab` on UNSW's Katana HPC.

```{r load_data}

ana_dir <- "/srv/scratch/z3532965/src/publications/2024_ARMC5/hct116_IF"
data_dir <- "/srv/scratch/berrylab/z3532965/NikonSpinningDisk/"
acquisition_dirs <- purrr::map_chr(
  c("240328/240325_siRNA_ARMC5_INTS8/20240328_172603_573",
    "240404/240401_HCT116_ARMC5_INTS8/20240404_153226_053",
    "240701/240628_siRNA_ARMC5_INTS8/20240701_193044_047"),~file.path(data_dir,.))
intensity_image_dirs <- map_chr(acquisition_dirs,~file.path(.,"OME-TIFF"))
intensity_image_mip_dirs <- map_chr(acquisition_dirs,~file.path(.,"OME-TIFF-MIP"))
label_image_dirs <- map_chr(acquisition_dirs,~file.path(.,"SEGMENTATION"))
quantification_dirs <- map_chr(acquisition_dirs,~file.path(.,"QUANTIFICATION"))

read_quantification <- function(dir_name) {
  csv_files <- Sys.glob(file.path(dir_name,"*.csv"))
  dt_list <- lapply(csv_files, data.table::fread)
  df <- tibble(path = csv_files,
               feature_values = dt_list) %>%
    mutate(dir_name = dir_name,
           file_name = basename(path),
           feature_values = purrr::map(feature_values,tibble)) %>%
    separate(file_name,into = c("well_name","ch1","ch2","ch3","ch4","sample","id"),sep="[,_.]", extra = "drop") %>%
    mutate(well_name = str_replace(well_name, "^Well", ""),
           field_id = as.integer(id))
  return(df)
}

pad_well_names <- function(names) {
  gsub("(\\D)(\\d)$", "\\10\\2", names)
}

# features
f <- lapply(quantification_dirs,read_quantification) %>%
  bind_rows() %>%
  mutate(acquisition_dir = dirname(dirname(path)),
         imaging_date = basename(acquisition_dir) %>% str_extract("\\d{8}(?=_)")) %>%
  select(-c(ch1, ch2, ch3, ch4, sample))

# layouts
plate_layouts <- tibble(layout_filename=c("240325_siRNA_ARMC5_INTS8_long.csv",
                                          "240401_siRNA_ARM_INTS_longlayout.csv",
                                          "240628_siRNA_ARMC5_INTS8.csv"),
                        plating_date=c("240325","240401","240628"),
                        imaging_date=c("20240328","20240404","20240701"),
                        metadata=purrr::map(layout_filename,~read_csv(file.path(ana_dir,"METADATA",.),show_col_types = FALSE, name_repair = ~sub(" ", "_", .)))) %>%
  unnest(metadata) %>%
  mutate(well_name=pad_well_names(Well)) %>%
  select(-Well) %>%
  left_join(distinct(f,acquisition_dir,well_name,imaging_date),by = join_by(imaging_date, well_name)) %>%
  filter(!is.na(acquisition_dir)) %>%
  mutate(siRNA=if_else(siRNA=="Scramble","Scrambled",siRNA))

```

Check that the data are complete.

```{r check_complete}

# check completeness (count wells)
f %>% 
  distinct(acquisition_dir,well_name) %>%
  group_by(acquisition_dir) %>%
  count()

# check completeness (count sites per well)
f %>% 
  group_by(acquisition_dir,well_name) %>%
  count() %>%
  filter(!(n %in% c(16)))

```

Load metadata and join with quantified image data.

```{r load metadata}
# read image metadata and join to plate metadata
m <- tibble(acquisition_dir = acquisition_dirs) %>%
  mutate(image_dir = purrr::map_chr(acquisition_dirs,~file.path(.,"OME-TIFF-MIP"))) %>%
  mutate(metadata_file = purrr::map(image_dir,~list.files(.))) %>%
  unnest(metadata_file) %>%
  filter(grepl("metadata.csv$",metadata_file)) %>%
  mutate(dat = purrr::map(file.path(image_dir,metadata_file),~read_csv(.,show_col_types = FALSE))) %>%
  separate(metadata_file,into = c("well_name","ch1","ch2","ch3","ch4","sample"),sep="[,_.]", extra = "drop") %>%
  unnest(dat) %>%
  mutate(well_name = str_replace(well_name, "^Well", ""),
           field_id = as.integer(field_id)) %>%
  select(-c(ch1, ch2, ch3, ch4, sample)) %>%
  left_join(filter(plate_layouts,!is.na(acquisition_dir)),by=join_by(acquisition_dir,well_name))

# attach image and plate metadata 
dat <- right_join(m,f,by = join_by(acquisition_dir, well_name, field_id, imaging_date)) %>%
  unnest(feature_values) %>%
  select(-matches("None$"))

```

Some basic QC regarding number of cells per well, etc.

```{r cell_counts, fig.width=8,fig.height=12}

dat %>% 
  distinct(plating_date,imaging_date,well_name,siRNA,Primary,Secondary) %>%
  group_by(plating_date,imaging_date,siRNA,Primary,Secondary) %>%
  count() %>%
  ungroup() %>%
  ggplot(aes(x=siRNA,y=n)) + geom_col(position = "dodge") + facet_grid(interaction(Primary,Secondary)~plating_date) +
  coord_flip() +
  ggtitle("Number of wells of each condition")

dat %>% 
  group_by(plating_date,imaging_date,siRNA,Primary,Secondary) %>%
  count() %>%
  ungroup() %>%
  ggplot(aes(x=siRNA,y=n)) + geom_col(position = "dodge") + facet_grid(interaction(Primary,Secondary)~plating_date) +
  coord_flip() +
  ggtitle("Total number of cells in each condition")

```

### Subtract background

Calculate background by maximum intensity projecting, illumination-correcting, and getting the mean intensity in a region containing no cells. See `calculate_background.ipynb` in the corresponding github repository.


```{r subtract_background}

# updated 6/4/24 using 240328/240325_siRNA_ARMC5_INTS8/20240328_172603_573 dataset

bg_488 <- mean(c(174.621,173.198,174.548))
bg_561 <- mean(c(147.854,149.994,149.375))
bg_405 <- mean(c(165.801,161.261,162.563))

dat_minus_bg <- mutate(dat,
                       Nuclei_intensity_mean_561 = Nuclei_intensity_mean_561 - bg_561,
                       Nuclei_intensity_mean_405 = Nuclei_intensity_mean_405 - bg_405,
                       Nuclei_intensity_mean_488 = Nuclei_intensity_mean_488 - bg_488,
                       
                       Nuclei_intensity_median_561 = Nuclei_intensity_median_561 - bg_561,
                       Nuclei_intensity_median_405 = Nuclei_intensity_median_405 - bg_405,
                       Nuclei_intensity_median_488 = Nuclei_intensity_median_488 - bg_488)

```

### Sum intensities

Compute sum intensities by multiplying the background-subtracted means by the corresponding area

```{r sum_intensities}

dat_minus_bg <- mutate(dat_minus_bg,
                       Nuclei_intensity_sum_561 = Nuclei_intensity_mean_561 * Nuclei_area,
                       Nuclei_intensity_sum_405 = Nuclei_intensity_mean_405 * Nuclei_area,
                       Nuclei_intensity_sum_488 = Nuclei_intensity_mean_488 * Nuclei_area)

```

### Data cleanup

Remove border cells 

```{r remove_variables}

dat_non_border <- filter(dat_minus_bg,!Nuclei_is_border)

```

# Check illumination correction

After correction of illumination biases across the field of view, there should be no bias for mean intensities across the field-of-view.

```{r test_illum_corr}

dat_non_border_sample_longer <- dat_non_border %>%
  filter(imaging_date=="20240328") %>%
  filter(well_name=="F06") %>%
  select(Nuclei_centroid_0,Nuclei_centroid_1,Nuclei_intensity_mean_405,Nuclei_intensity_mean_488,Nuclei_intensity_mean_561) %>%
  pivot_longer(-c(Nuclei_centroid_0,Nuclei_centroid_1)) 

rescaling_limits <- dat_non_border_sample_longer %>%
  group_by(name) %>%
  summarise(upper_limit = quantile(value,0.99))

dat_non_border_sample_longer <- left_join(dat_non_border_sample_longer,rescaling_limits,by="name")

channels <- unique(dat_non_border_sample_longer$name)

p <- list(length=length(channels))
for (i in 1:length(channels)) {
  p[[i]] <- dat_non_border_sample_longer %>%
    filter(name==channels[i]) %>%
    ggplot(aes(x=Nuclei_centroid_0,y=Nuclei_centroid_1,col=value/upper_limit)) +
    geom_point(size=0.5) +
    scale_color_viridis_c(limits=c(0,1)) +
    facet_wrap(~name) +
    coord_fixed() +
    theme(legend.position = "none")
}

wrap_plots(p) & ggtitle("Illumination biases")

```

## Check DAPI distributions

```{r check_DAPI}

dat_non_border %>%
  ggplot(aes(x=Nuclei_intensity_sum_405,col=as.factor(well_name))) +
  geom_density() +
  facet_wrap(imaging_date~interaction(siRNA,Cell),scales = "free") +
  theme(legend.position = "none") +
  scale_x_continuous(limits = c(0,4e6))

```
## Check "No reaction" wells on 2024/03/28

Some strange-looking results in the histograms of DAPI for some "no reaction" wells on 2024/03/28. Removing low cell number wells seems to fix this issue.

```{r check_no_reaction}

qc <- dat_non_border %>%
  filter(siRNA=="No reaction") %>%
  group_by(well_name,imaging_date,Cell) %>%
  summarise(n_cells = n(), mean_sum_DAPI = mean(Nuclei_intensity_sum_405),.groups = "drop")

qc %>% arrange(imaging_date,well_name)

low_cell_number_wells <- filter(qc,n_cells < 500) %>% select(well_name,imaging_date)

dat_non_border %>%
  anti_join(low_cell_number_wells,by = join_by(well_name, imaging_date)) %>%
  ggplot(aes(x=Nuclei_intensity_sum_405,col=as.factor(well_name))) +
  geom_density() +
  facet_wrap(imaging_date~interaction(siRNA,Cell),scales = "free") +
  theme(legend.position = "none") +
  scale_x_continuous(limits = c(0,4e6))

dat_non_border <- dat_non_border %>% anti_join(low_cell_number_wells,by = join_by(well_name, imaging_date))

```
## Nuclear Area

```{r nuclear_area_qc}

dat_non_border %>%
  ggplot(aes(x=Nuclei_area,col=as.factor(well_name))) +
  geom_density() +
  facet_wrap(imaging_date~interaction(siRNA,Cell),scales = "free") +
  theme(legend.position = "none") +
  geom_vline(xintercept = 750)

dat_non_border <- filter(dat_non_border,Nuclei_area>750)

```

## mCherry distributions

```{r mCherry_by_siRNA, fig.width=8,fig.height=4}

dat_non_border %>%
  ggplot(aes(x=interaction(siRNA,Cell),y=Nuclei_intensity_mean_561,group=interaction(siRNA,Cell,well_name))) + 
  geom_boxplot(outlier.shape = NA) +
  coord_cartesian(ylim = c(0,250)) +
  facet_wrap(~imaging_date) +
  theme_bw(base_size = 8) +
  theme(axis.text.x = element_text(angle=90,hjust=1,vjust=0.5))
ggsave("PLOTS/mCherry-POLR2A_siRNA_boxplots_per_well.pdf",width=8,height=6)

dat_non_border %>%
  group_by(well_name,siRNA,imaging_date,Cell) %>%
  summarise(Nuclei_intensity_mean_561=mean(Nuclei_intensity_mean_561),.groups = "drop") %>%
  ggplot(aes(x=siRNA,y=Nuclei_intensity_mean_561)) + 
  geom_boxplot() +
  geom_jitter(width = 0.2) +
  coord_cartesian(ylim = c(0,250)) +
  facet_wrap(Cell~imaging_date) +
  theme_bw(base_size = 8) +
  theme(axis.text.x = element_text(angle=90,hjust=1,vjust=0.5))

ggsave("PLOTS/mCherry-POLR2A_siRNA_well_averages.pdf",width=4,height=3)


```

```{r plot_mCherry}

background_values_from_HCT116 <- dat_non_border %>%
  filter(Cell=="HCT116") %>%
  group_by(well_name,imaging_date) %>%
  summarise(Nuclei_intensity_mean_561=mean(Nuclei_intensity_mean_561),.groups="drop") %>%
  group_by(imaging_date) %>%
  summarise(Nuclei_intensity_mean_561_bkgrd = mean(Nuclei_intensity_mean_561),.groups="drop")

dat_non_border %>%
  group_by(well_name,siRNA,imaging_date,Cell) %>%
  summarise(Nuclei_intensity_mean_561=mean(Nuclei_intensity_mean_561),.groups="drop") %>%
  mutate(siRNA = if_else(Cell=="HCT116","No mCherry-RPB1",siRNA)) %>%
  filter(!(Cell=="HCT116 mCherryPOLR2A" & siRNA=="No reaction")) %>%
  ggplot(aes(x=siRNA,y=Nuclei_intensity_mean_561)) + 
  geom_boxplot() +
  geom_jitter(width = 0.2) +
  coord_cartesian(ylim = c(0,250)) +
  facet_wrap(~imaging_date) +
  theme_bw(base_size = 8) +
  theme(axis.text.x = element_text(angle=90,hjust=1,vjust=0.5))
  
dat_non_border_summarised_minus_mCherry_bkgrd <- dat_non_border %>%
  filter(Cell!="HCT116") %>%
  filter(siRNA!="No reaction") %>%
  group_by(well_name,siRNA,imaging_date,Cell) %>%
  summarise(Nuclei_intensity_mean_561=mean(Nuclei_intensity_mean_561),.groups="drop") %>%
  left_join(background_values_from_HCT116,by = join_by(imaging_date)) %>%
  mutate(Nuclei_intensity_mean_561_minus_bkgrd=Nuclei_intensity_mean_561-Nuclei_intensity_mean_561_bkgrd)

dat_non_border_summarised_minus_mCherry_bkgrd %>%
  ggplot(aes(x=siRNA,y=Nuclei_intensity_mean_561_minus_bkgrd)) + 
  geom_boxplot() +
  geom_jitter(width = 0.2) +
  coord_cartesian(ylim = c(0,150)) +
  facet_wrap(~imaging_date) +
  theme_bw(base_size = 8) +
  theme(axis.text.x = element_text(angle=90,hjust=1,vjust=0.5))

plate_normalisers <- dat_non_border_summarised_minus_mCherry_bkgrd %>%
  filter(siRNA=="Scramble") %>%
  group_by(imaging_date) %>%
  summarise(plate_mean_scrambled = mean(Nuclei_intensity_mean_561_minus_bkgrd))

dat_non_border_summarised_minus_mCherry_bkgrd %>%
  left_join(plate_normalisers) %>%
  mutate(Nuclei_intensity_mean_561_minus_bkgrd_plate_normalised = Nuclei_intensity_mean_561_minus_bkgrd / plate_mean_scrambled) %>%
  ggplot(aes(x=siRNA,y=Nuclei_intensity_mean_561_minus_bkgrd_plate_normalised,group=interaction(imaging_date,siRNA),fill=imaging_date)) + 
  geom_boxplot() +
  coord_cartesian(ylim = c(0,2)) +
  theme_bw(base_size = 8) +
  theme(axis.text.x = element_text(angle=90,hjust=1,vjust=0.5))

to_plot <- dat_non_border_summarised_minus_mCherry_bkgrd %>%
  left_join(plate_normalisers) %>%
  mutate(Nuclei_intensity_mean_561_minus_bkgrd_plate_normalised = Nuclei_intensity_mean_561_minus_bkgrd / plate_mean_scrambled) %>%
  mutate(siRNA = if_else(siRNA=="Scramble","Scrambled",siRNA),
         siRNA = if_else(siRNA=="ARMC5+INTS8","ARMC5 + INTS8 (pools)",siRNA),
         siRNA = factor(siRNA,levels=c(
           "Scrambled",
           "ARMC5 #1","ARMC5 #2","ARMC5 #3","ARMC5 pool",
           "INTS8 #1","INTS8 #2","INTS8 #3","INTS8 pool",
           "ARMC5 + INTS8 (pools)"))) %>%
  mutate(siRNA_colour = case_when(
    siRNA == "Scrambled" ~ "Scrambled",
    siRNA %in% c("ARMC5 #1","ARMC5 #2","ARMC5 #3","ARMC5 pool") ~ "ARMC5",
    siRNA %in% c("INTS8 #1","INTS8 #2","INTS8 #3","INTS8 pool") ~ "INTS8",
    siRNA == "ARMC5 + INTS8 (pools)" ~ "ARMC5 + INTS8",
    TRUE ~ NA_character_),
    siRNA_colour=factor(siRNA_colour,levels=c("Scrambled","ARMC5","INTS8","ARMC5 + INTS8")))
  
to_plot_summary <- group_by(to_plot,siRNA,siRNA_colour) %>%
  summarise(mean_Nuclei_intensity_mean_561_minus_bkgrd_plate_normalised = mean(Nuclei_intensity_mean_561_minus_bkgrd_plate_normalised),
            sd_Nuclei_intensity_mean_561_minus_bkgrd_plate_normalised = sd(Nuclei_intensity_mean_561_minus_bkgrd_plate_normalised))

to_plot %>%
  ggplot(aes(x=siRNA,y=Nuclei_intensity_mean_561_minus_bkgrd_plate_normalised,fill=siRNA_colour,col=siRNA_colour)) + 
  geom_col(data=to_plot_summary,aes(y=mean_Nuclei_intensity_mean_561_minus_bkgrd_plate_normalised),
           col=NA,alpha=0.4,width=0.8) +
  geom_errorbar(data=to_plot_summary,aes(
    y=mean_Nuclei_intensity_mean_561_minus_bkgrd_plate_normalised,
    ymin=mean_Nuclei_intensity_mean_561_minus_bkgrd_plate_normalised-sd_Nuclei_intensity_mean_561_minus_bkgrd_plate_normalised,
    ymax=mean_Nuclei_intensity_mean_561_minus_bkgrd_plate_normalised+sd_Nuclei_intensity_mean_561_minus_bkgrd_plate_normalised),
    width=0.4) +
  geom_point(data=to_plot_summary, aes(y=mean_Nuclei_intensity_mean_561_minus_bkgrd_plate_normalised),size=0.75) +
  scale_y_continuous(name="Mean nuclear\nmCherry-RPB1\nintensity",limits = c(0,2),expand = c(0,0)) +
  scale_fill_manual(values = c(wt_col,armc5_col,ints8_col,armc5_ints8_col)) +
  scale_colour_manual(values = c(wt_col,armc5_col,ints8_col,armc5_ints8_col)) +
  geom_jitter(pch=4,alpha=0.4,position = position_jitter(width = 0.2),size=0.4) +
  theme_bw(base_size = 8) +
  theme(axis.text.x = element_text(angle=90,hjust=1,vjust=0.5),
        axis.title.x = element_blank(),
        panel.grid = element_blank(),
        legend.title = element_blank(),
        legend.key.size = unit(3,"mm"))
ggsave(filename = "PLOTS/mCherry_barplots_by_siRNA_all.pdf",width=3,height=2)

```

## Check IF distributions

```{r IF_by_treatment, fig.width=6,fig.height=8}

pTotal <- dat_non_border %>%
  filter(Primary=="Total Pol II") %>%
  ggplot(aes(x=siRNA,y=Nuclei_intensity_mean_488,group=interaction(well_name,siRNA))) + 
  geom_boxplot(outlier.shape=NA,aes(fill=imaging_date)) + facet_grid(Primary~plating_date) +
  coord_cartesian(ylim = c(0,1500)) +
  theme_bw(base_size = 8) +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank())

pS5P <- dat_non_border %>%
  filter(Primary=="pSer5 Pol II") %>% 
  ggplot(aes(x=siRNA,y=Nuclei_intensity_mean_488,group=interaction(well_name,siRNA))) + 
  geom_boxplot(outlier.shape=NA,aes(fill=imaging_date)) + facet_grid(Primary~plating_date) +
  coord_cartesian(ylim = c(0,4000)) +
  theme_bw(base_size = 8) +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank())

pS2P <- dat_non_border %>%
  filter(Primary=="pSer2 Pol II") %>%
  ggplot(aes(x=siRNA,y=Nuclei_intensity_mean_488,group=interaction(well_name,siRNA))) + 
  geom_boxplot(outlier.shape=NA,aes(fill=imaging_date)) + facet_grid(Primary~plating_date) +
  coord_cartesian(ylim = c(0,1500)) +
  theme_bw(base_size = 8) +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_text(angle=90,hjust=1,vjust=0.5))


pTotal / pS5P / pS2P + plot_layout(guides="collect")

ggsave("PLOTS/mCherry-POLR2A_IF_siRNA_boxplot_per_well.pdf",width=6,height=10)

```

pSer2 results for ARMC5 are inconsistent across the ARMC5 conditions and are inconsistent with previous results showing limited effect of ARMC5 on Ser2P also in both HCT with siRNA and HEK KO vs wt cells, so exclude these for now. Include Ser5P results (which are consistent across the ARMC5 siRNAs and with previous data). Repeat these Ser2P experiments later if necessary.

```{r plot_IF}

dat_non_border_summarised_IF <- dat_non_border %>%
  filter(siRNA!="No reaction") %>%
  filter(Primary!="None") %>%
  group_by(well_name,siRNA,imaging_date,Cell,Primary,Secondary) %>%
  summarise(Nuclei_intensity_mean_488=mean(Nuclei_intensity_mean_488),.groups="drop") 

plate_normalisers_IF <- dat_non_border_summarised_IF %>%
  filter(siRNA=="Scramble") %>%
  group_by(imaging_date,Primary,Secondary) %>%
  summarise(plate_mean_scrambled = mean(Nuclei_intensity_mean_488),.groups="drop")

to_plot_IF <- dat_non_border_summarised_IF %>%
  left_join(plate_normalisers_IF) %>%
  mutate(Nuclei_intensity_mean_488_plate_normalised = Nuclei_intensity_mean_488 / plate_mean_scrambled) %>%
  mutate(siRNA = if_else(siRNA=="Scramble","Scrambled",siRNA),
         siRNA = if_else(siRNA=="ARMC5+INTS8","ARMC5 + INTS8 (pools)",siRNA),
         siRNA = factor(siRNA,levels=c(
           "Scrambled",
           "ARMC5 #1","ARMC5 #2","ARMC5 #3","ARMC5 pool",
           "INTS8 #1","INTS8 #2","INTS8 #3","INTS8 pool",
           "ARMC5 + INTS8 (pools)"))) %>%
  mutate(siRNA_colour = case_when(
    siRNA == "Scrambled" ~ "Scrambled",
    siRNA %in% c("ARMC5 #1","ARMC5 #2","ARMC5 #3","ARMC5 pool") ~ "ARMC5",
    siRNA %in% c("INTS8 #1","INTS8 #2","INTS8 #3","INTS8 pool") ~ "INTS8",
    siRNA == "ARMC5 + INTS8 (pools)" ~ "ARMC5 + INTS8",
    TRUE ~ NA_character_),
    siRNA_colour=factor(siRNA_colour,levels=c("Scrambled","ARMC5","INTS8","ARMC5 + INTS8"))) %>%
  mutate(Primary_label = case_when(
    Primary=="Total Pol II" ~ "Total RPB1",
    Primary=="pSer5 Pol II" ~ "Ser5 phos. RPB1",
    Primary=="pSer2 Pol II" ~ "Ser2 phos. RPB1",
    Primary=="POLR2B" ~ "Total RPB2",
    TRUE ~ NA_character_),
    Primary_label=factor(Primary_label,levels=c("Total RPB1","Ser5 phos. RPB1","Ser2 phos. RPB1","Total RPB2")))

to_plot_IF_summary <- group_by(to_plot_IF,siRNA,siRNA_colour,Primary,Primary_label,Secondary) %>%
  summarise(mean_Nuclei_intensity_mean_488_plate_normalised = mean(Nuclei_intensity_mean_488_plate_normalised),
            sd_Nuclei_intensity_mean_488_plate_normalised = sd(Nuclei_intensity_mean_488_plate_normalised))

to_plot_IF %>%
  ggplot(aes(x=siRNA,y=Nuclei_intensity_mean_488_plate_normalised,fill=siRNA_colour,col=siRNA_colour)) + 
  geom_col(data=to_plot_IF_summary,aes(y=mean_Nuclei_intensity_mean_488_plate_normalised),
           col=NA,alpha=0.4,width=0.8) +
  geom_errorbar(data=to_plot_IF_summary,aes(
    y=mean_Nuclei_intensity_mean_488_plate_normalised,
    ymin=mean_Nuclei_intensity_mean_488_plate_normalised-sd_Nuclei_intensity_mean_488_plate_normalised,
    ymax=mean_Nuclei_intensity_mean_488_plate_normalised+sd_Nuclei_intensity_mean_488_plate_normalised),
    width=0.4) +
  geom_point(data=to_plot_IF_summary, aes(y=mean_Nuclei_intensity_mean_488_plate_normalised),size=0.75) +
  scale_y_continuous(name="Mean nuclear\nmCherry-RPB1\nintensity",limits = c(0,2),expand = c(0,0)) +
  scale_fill_manual(values = c(wt_col,armc5_col,ints8_col,armc5_ints8_col)) +
  scale_colour_manual(values = c(wt_col,armc5_col,ints8_col,armc5_ints8_col)) +
  geom_jitter(pch=4,alpha=0.4,position = position_jitter(width = 0.2),size=0.4) +
  facet_wrap(~Primary) +
  theme_bw(base_size = 8) +
  theme(axis.text.x = element_text(angle=90,hjust=1,vjust=0.5),
        axis.title.x = element_blank(),
        panel.grid = element_blank(),
        legend.title = element_blank(),
        legend.key.size = unit(2,"mm"))

filter(to_plot_IF,
       siRNA %in% c("Scrambled","ARMC5 pool","INTS8 pool","ARMC5 + INTS8 (pools)") & Primary %in% c("Total Pol II","pSer5 Pol II")) %>%
  ggplot(aes(x=siRNA,y=Nuclei_intensity_mean_488_plate_normalised,fill=siRNA_colour,col=siRNA_colour)) + 
  geom_col(data=filter(to_plot_IF_summary,siRNA %in% c("Scrambled","ARMC5 pool","INTS8 pool","ARMC5 + INTS8 (pools)") & Primary %in% c("Total Pol II","pSer5 Pol II")),
           aes(y=mean_Nuclei_intensity_mean_488_plate_normalised),
           col=NA,alpha=0.4,width=0.8) +
  geom_errorbar(data=filter(to_plot_IF_summary,siRNA %in% c("Scrambled","ARMC5 pool","INTS8 pool","ARMC5 + INTS8 (pools)") & Primary %in% c("Total Pol II","pSer5 Pol II")),
                aes(y=mean_Nuclei_intensity_mean_488_plate_normalised,
                    ymin=mean_Nuclei_intensity_mean_488_plate_normalised-sd_Nuclei_intensity_mean_488_plate_normalised,
                    ymax=mean_Nuclei_intensity_mean_488_plate_normalised+sd_Nuclei_intensity_mean_488_plate_normalised),
                width=0.4) +
  geom_point(data=filter(to_plot_IF_summary,siRNA %in% c("Scrambled","ARMC5 pool","INTS8 pool","ARMC5 + INTS8 (pools)") & Primary %in% c("Total Pol II","pSer5 Pol II")),
             aes(y=mean_Nuclei_intensity_mean_488_plate_normalised),size=0.75) +
  scale_y_continuous(name="Mean nuclear\nintensity",limits = c(0,2.2),expand = c(0,0)) +
  scale_fill_manual(values = c(wt_col,armc5_col,ints8_col,armc5_ints8_col)) +
  scale_colour_manual(values = c(wt_col,armc5_col,ints8_col,armc5_ints8_col)) +
  geom_jitter(pch=4,alpha=0.4,position = position_jitter(width = 0.2),size=0.4) +
  facet_grid(.~Primary_label) +
  theme_bw(base_size = 8) +
  theme(axis.text.x = element_text(angle=90,hjust=1,vjust=0.5),
        axis.title.x = element_blank(),
        panel.grid = element_blank(),
        legend.title = element_blank(),
        legend.key.size = unit(3,"mm"))
ggsave(filename = "PLOTS/mCherry_barplots_IF_by_siRNA_subset.pdf",width=3,height=2)

```


```{r check_plate_layouts, fig.width=9,fig.height=3}

dat_non_border %>%
  mutate(well_row = factor(str_extract(well_name, "[A-Z]"),levels=LETTERS),
         well_column = as.integer(as.character(str_extract(well_name, "[0-9]+")))) %>%
  distinct(imaging_date,well_name,well_row,well_column,siRNA) %>%
  ggplot(aes(x=well_column,y=well_row,fill=siRNA)) +
  geom_tile() +
  scale_y_discrete(limits=rev) +
  facet_wrap(~imaging_date)

dat_non_border %>%
  mutate(well_row = factor(str_extract(well_name, "[A-Z]"),levels=LETTERS),
         well_column = as.integer(as.character(str_extract(well_name, "[0-9]+")))) %>%
  distinct(imaging_date,well_name,well_row,well_column,Cell) %>%
  ggplot(aes(x=well_column,y=well_row,fill=Cell)) +
  geom_tile() +
  scale_y_discrete(limits=rev) +
  facet_wrap(~imaging_date)

dat_non_border %>%
  mutate(well_row = factor(str_extract(well_name, "[A-Z]"),levels=LETTERS),
         well_column = as.integer(as.character(str_extract(well_name, "[0-9]+")))) %>%
  distinct(imaging_date,well_name,well_row,well_column,Primary) %>%
  ggplot(aes(x=well_column,y=well_row,fill=Primary)) +
  geom_tile() +
  scale_y_discrete(limits=rev) +
  facet_wrap(~imaging_date)

dat_non_border %>%
  mutate(well_row = factor(str_extract(well_name, "[A-Z]"),levels=LETTERS),
         well_column = as.integer(as.character(str_extract(well_name, "[0-9]+")))) %>%
  count(imaging_date,well_name,well_row,well_column,Secondary) %>%
  ggplot(aes(x=well_column,y=well_row,fill=n)) +
  geom_tile() +
  scale_y_discrete(limits=rev) +
  facet_wrap(~imaging_date) +
  scale_fill_viridis_c()


# compute fold-change from mean for each well
dat_non_border %>%
  filter(Primary!="None") %>%
  mutate(well_row = factor(str_extract(well_name, "[A-Z]"),levels=LETTERS),
         well_column = as.integer(as.character(str_extract(well_name, "[0-9]+")))) %>%
  group_by(imaging_date,well_name,well_row,well_column,Primary,Cell,siRNA) %>%
  summarise(well_mean_IF = mean(Nuclei_intensity_mean_488),.groups="drop") %>%
  left_join(dat_non_border %>%
              group_by(imaging_date,well_name,Primary,Cell,siRNA) %>%
              summarise(well_mean_IF = mean(Nuclei_intensity_mean_488),.groups="drop") %>%
              group_by(imaging_date,Primary,Cell,siRNA) %>%
              summarise(plate_mean_IF = mean(well_mean_IF),.groups="drop"),
            by = join_by(imaging_date,Primary,Cell,siRNA)) %>%
  mutate(log2FC_from_mean = log2(well_mean_IF/plate_mean_IF)) %>%
  ggplot(aes(x=well_column,y=well_row,fill=log2FC_from_mean)) +
  geom_tile() +
  scale_fill_distiller(type = "div", palette = "RdBu", limits = c(-2,2)) +
  scale_y_discrete(limits=rev) +
  facet_wrap(~imaging_date) +
  ggtitle("Within plate difference from mean per well")

# compute fold-change from mean for each well
dat_non_border %>%
  filter(Primary!="None") %>%
  mutate(well_row = factor(str_extract(well_name, "[A-Z]"),levels=LETTERS),
         well_column = as.integer(as.character(str_extract(well_name, "[0-9]+")))) %>%
  group_by(imaging_date,well_name,well_row,well_column,Secondary,Primary,Cell,siRNA) %>%
  summarise(well_mean_IF = mean(Nuclei_intensity_mean_488),.groups="drop") %>%
  left_join(dat_non_border %>%
              group_by(imaging_date,well_name,Secondary,Primary,Cell,siRNA) %>%
              summarise(well_mean_IF = mean(Nuclei_intensity_mean_488),.groups="drop") %>%
              group_by(Secondary,Primary,Cell,siRNA) %>%
              summarise(overall_mean_IF = mean(well_mean_IF),.groups="drop"),
            by = join_by(Secondary,Primary, Cell, siRNA)) %>%
  mutate(log2FC_from_mean = log2(well_mean_IF/overall_mean_IF)) %>%
  ggplot(aes(x=well_column,y=well_row,fill=log2FC_from_mean)) +
  geom_tile() +
  scale_fill_distiller(type = "div", palette = "RdBu", limits = c(-2,2)) +
  scale_y_discrete(limits=rev) +
  facet_wrap(~imaging_date) +
  ggtitle("Experiment difference from mean per well")

```


```{r well_numbers}

dat_non_border %>%
  filter(Primary!="None") %>%
  group_by(imaging_date,Primary,Cell,siRNA,well_name) %>%
  count() 

to_plot %>%
  group_by(imaging_date,Cell,siRNA) %>%
  count()

to_plot_IF %>%
  group_by(imaging_date,Primary,Cell,siRNA) %>%
  count()

to_plot_IF %>%
  group_by(Primary,Cell,siRNA) %>%
  count()


```


```{r session_info}

sessionInfo()

```
